### 2016-08-02 old_style.tex
First commit

% Template for the submission to:
%   The Annals of Probability           [aop]
%   The Annals of Applied Probability   [aap]
%   The Annals of Statistics            [aos]
%   The Annals of Applied Statistics    [aoas]
%   Stochastic Systems                  [ssy]
%
%Author: In this template, the places where you need to add information
%        (or delete line) are indicated by {???}.  Mostly the information
%        required is obvious, but some explanations are given in lines starting
%Author:
%All other lines should be ignored.  After editing, there should be
%no instances of ??? after this line.

% use option [preprint] to remove info line at bottom
% journal options: aop,aap,aos,aoas,ssy
% natbib option: authoryear
\documentclass[aoas,preprint,authoryear]{imsart}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

\newcommand{\glnote}[1]{\textcolor{red}{[GL: #1]}}

\begin{document}

\begin{frontmatter}

\title{Learning to Transfer with Semi-Supervised Triply Adversarial Nets}

% indicate corresponding author with \corref{}
% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
% \thankstext{t1}{Thanks to somebody}
% \address{line 1\\ line 2\\ printead{e1}}
% \affiliation{Some University}

\author{\fnms{Gilles} \snm{Louppe}}
\affiliation{New York University}

\begin{abstract}

In classification, transfer learning (or its variants known as covariate shift
or domain adaptation) arises whenever target instances are governed by a
distribution that may be arbitrarily different from the distribution of the
source instances. This problem has traditionally been solved by reweighting
approaches or by learning robust representations over domains. In this work, we
propose a new paradigm based on the assumption that the covariate shift is
caused by the use of a different representation of the same underlying objects.
Accordingly, we propose to learn how to transform source instances into target
instances, possibly across input spaces of distinct dimensions, structures or
supports. For this purpose, we extend the generative adversarial networks
framework of \cite{goodfellow2014generative} to a triply adversarial process: a
transformer network $T$ for generating target instances from source instances, a
discriminative network $D$ for separating transformed source instances from
actual target instances, and a classifier network $C \circ T$ for classifying source
instances in the projected space. This 3-player game results in a network $T$
capable of transforming source into target instances, while preserving
separation between classes as enabled by $C$ in the adversarial setup.
Experiments demonstrate the potential of this novative approach, with promising
results when the construction of $C$ can be bootstrapped in a semi-supervised
way  from a few labeled instances from the target space.

\end{abstract}

\end{frontmatter}

\section{Introduction}

% - transfer learning has traditionally been solved through density ratio reweighting
% - works fine, but has the big issue of not working if the support of the source density does not cover the target density support\
%   support is needed for reweighting, but not here (Gretton et al) we can even have different space

% - in this work, we propose instead a new paradigm for solving transfer learning
% - build a generative model for transforming samples from p1 to reproduce p0
% - adversarial learning from Goodfellow, with the practical difference that we often have only a finite set of samples (we need to avoid learning a lookup table)

% other view point: start from dual adversaries, then add to the mix C is a regularizer
% - ensure p(y|t(x)) ~ p(y|x) by having a third classifier in the mix
% issue: T is not unique, how make sure samples are mapped correctly?

% - application: fix simulated data

% Ganin: domains share the same input space, we dont necessarily have to
%        seek a common robust representation, here we seek instead to transform training into test

\section{Method}

% framework
% doubly adversarial network
% T is not unique => triply adversarial + regularization
% proofs?

% we should state that it makes sense only when the same object (\in Omega) can exists in both space, or when there is a suitable transformation from one to the other
% semi-supervised transfer learning: use seeds from the target space to bootstrap C

\section{Experiments}

% - regularization
% - network architecture (propagate x, similar to highway networks)

\section{Related work}

% also related to multi-view learning

\section{Conclusions}

\bibliographystyle{apalike}
\bibliography{bibliography.bib}

\end{document}

### 2016-08-02 paper.tex
First commit

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
\PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2016}

\usepackage[final]{nips_2016}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{color}

\newcommand{\glnote}[1]{\textcolor{red}{[GL: #1]}}

\title{Adversarial Training of Neural Networks\\
against Systematic Uncertainty}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Gilles Louppe \\
  New York University\\
  \texttt{g.louppe@nyu.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}

\glnote{Distinction between statistical and systematic uncertainty.}
\glnote{Define nuisance parameters.}
\glnote{We want to build an accurate classifier whose output remains invariant with
respect to systematic uncertainties.}
% https://www.slac.stanford.edu/econf/C030908/papers/TUAT004.pdf

\section{Problem statement}
\label{sec:problem}

Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
Let consider the multivariate random variables $X_\lambda: \Omega \mapsto \mathbb{R}^p$
and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a nuisance
parameter $\lambda$ encoding for systematic uncertainties in its probability
distribution. That is, $X_\lambda$ and $Y$ induce together the joint probability
distribution $p(X_\lambda,Y)$. For training, let further assume a finite set $\{
 \lambda_i, x_i, y_i, \}_{i=1}^N$ of realizations $X_{\lambda_i}(\omega_i),
Y(\omega_i)$, for $\omega_i \in \Omega$ and known values $\lambda_i$ of the
nuisance parameter. Our goal is to learn a function $f : \mathbb{R}^p \mapsto
{\cal Y}$ (e.g., a classifier if ${\cal Y}$ is a finite set of classes)
minimizing the expected value of a loss $L(Y, f(X_{\lambda}))$, with the constraint that
$f(X_\lambda)$ should be robust to the value of the nuisance parameter
$\lambda$ -- which remains unknown at test time. More specifically, we aim at building $f$ such that in the ideal case
\begin{equation}\label{eqn1}
f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega))
\end{equation}
for any sample $\omega \in
\Omega$ and any $\lambda_i, \lambda_j$ pair of values of the nuisance parameter.

Since we do not have training tuples $(X_{\lambda_i}(\omega),
X_{\lambda_j}(\omega))$ (for the same unknown $\omega$), we propose instead to
solve the closely related problem of finding a predictive function $f$ such that
\begin{equation}\label{eqn2}
    P(\{ \omega | f(X_{\lambda_i}(\omega)) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega')) = y\}) \text{ for all $y \in {\cal Y}$}.
\end{equation}
In words, we are looking for a predictive function $f$ such that  the
distribution of $f(X_\lambda)$ is invariant with respect to the nuisance
parameter $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn1} is
true necessarily satisfies Eqn.~\ref{eqn2}. The converse is however in
general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega)) = y \}$
and $\{ \omega' | f(X_{\lambda_j}(\omega)) = y \}$ do not need to be the same
for the equality to hold. \glnote{Yet, in practice, this criterion is often good enough.}

\section{Method}
\label{sec:method}

\section{Experiments}

\section{Related work}

\glnote{Similar to domain adaptation, but with infinitely many domains,
as parameterized by $\lambda$.}

\section{Conclusions}

\subsubsection*{Acknowledgments}

\bibliographystyle{ieeetr}
{\small
\bibliography{bibliography.bib}}

\end{document}

### 2016-08-02 paper.tex
save

@@
 sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
-Let consider the multivariate random variables $X_\lambda: \Omega \mapsto \mathbb{R}^p$
-and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a nuisance
-parameter $\lambda$ encoding for systematic uncertainties in its probability
-distribution. That is, $X_\lambda$ and $Y$ induce together the joint probability
-distribution $p(X_\lambda,Y)$. For training, let further assume a finite set $\{
- \lambda_i, x_i, y_i, \}_{i=1}^N$ of realizations $X_{\lambda_i}(\omega_i),
-Y(\omega_i)$, for $\omega_i \in \Omega$ and known values $\lambda_i$ of the
-nuisance parameter. Our goal is to learn a function $f : \mathbb{R}^p \mapsto
-{\cal Y}$ (e.g., a classifier if ${\cal Y}$ is a finite set of classes)
-minimizing the expected value of a loss $L(Y, f(X_{\lambda}))$, with the constraint that
-$f(X_\lambda)$ should be robust to the value of the nuisance parameter
-$\lambda$ -- which remains unknown at test time. More specifically, we aim at building $f$ such that in the ideal case
-\begin{equation}\label{eqn1}
-f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega))
-\end{equation}
-for any sample $\omega \in
-\Omega$ and any $\lambda_i, \lambda_j$ pair of values of the nuisance parameter.
+Let consider the multivariate random variables $X_\lambda: \Omega \mapsto
+\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
+nuisance parameter $\lambda$ whose values define the family of its systematic uncertainties.
+That is, $X_\lambda$ and $Y$ induce together
+a joint probability distribution written as $p(X,Y|\lambda)$, where the
+conditional on $\lambda$ denotes $X_\lambda$. For training, let further assume a
+finite set $\{ \lambda_i, x_i, y_i, \}_{i=1}^N$ of realizations
+$X_{\lambda_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
+values $\lambda_i$ of the nuisance parameter. Our goal is to learn a function $f :
+\mathbb{R}^p \mapsto {\cal Y}$ (e.g., a classifier if ${\cal Y}$ is a finite set
+of classes) minimizing the expected value of a loss $L(Y, f(X_{\lambda}))$, with
+the constraint that $f(X_\lambda)$ should be robust to the value of the nuisance
+parameter $\lambda$ -- which remains unknown at test time. More specifically, we
+aim at building $f$ such that in the ideal case \begin{equation}\label{eqn1}
+f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega)) \end{equation} for any
+sample $\omega \in \Omega$ and any $\lambda_i, \lambda_j$ pair of values of the
+nuisance parameter.
 
@@
 general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega)) = y \}$
-and $\{ \omega' | f(X_{\lambda_j}(\omega)) = y \}$ do not need to be the same
-for the equality to hold. \glnote{Yet, in practice, this criterion is often good enough.}
+and $\{ \omega' | f(X_{\lambda_j}(\omega')) = y \}$ do not need to be the same
+for the equality to hold. \glnote{This criterion is still adequate for most purposes.}
 

### 2016-08-02 paper.tex
save
method

@@
 % https://www.slac.stanford.edu/econf/C030908/papers/TUAT004.pdf
+\glnote{Motivate the criterion: allow to derive guarantees no matter $\lambda$,
+while $f$ may otherwise be good or very bad depending on the nuisance if trained
+on the mixture. E.g. to form hypothesis to be later confirmed by data.}
 
@@
 That is, $X_\lambda$ and $Y$ induce together
-a joint probability distribution written as $p(X,Y|\lambda)$, where the
+a joint probability distribution $p(X,Y|\lambda)$, where the
 conditional on $\lambda$ denotes $X_\lambda$. For training, let further assume a
-finite set $\{ \lambda_i, x_i, y_i, \}_{i=1}^N$ of realizations
+finite set $\{ x_i, y_i, \lambda_i \}_{i=1}^N$ of realizations
 $X_{\lambda_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
@@
 parameter $\lambda$ -- which remains unknown at test time. More specifically, we
-aim at building $f$ such that in the ideal case \begin{equation}\label{eqn1}
-f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega)) \end{equation} for any
+aim at building $f$ such that in the ideal case
+\begin{equation}\label{eqn:criterion-true}
+f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega))
+\end{equation} for any
 sample $\omega \in \Omega$ and any $\lambda_i, \lambda_j$ pair of values of the
@@
 solve the closely related problem of finding a predictive function $f$ such that
-\begin{equation}\label{eqn2}
+\begin{equation}\label{eqn:criterion}
     P(\{ \omega | f(X_{\lambda_i}(\omega)) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega')) = y\}) \text{ for all $y \in {\cal Y}$}.
@@
 distribution of $f(X_\lambda)$ is invariant with respect to the nuisance
-parameter $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn1} is
-true necessarily satisfies Eqn.~\ref{eqn2}. The converse is however in
+parameter $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn:criterion-true} is
+true necessarily satisfies Eqn.~\ref{eqn:criterion}. The converse is however in
 general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega)) = y \}$
 and $\{ \omega' | f(X_{\lambda_j}(\omega')) = y \}$ do not need to be the same
-for the equality to hold. \glnote{This criterion is still adequate for most purposes.}
+for the equality to hold.
 
@@
 \label{sec:method}
+
+Adversarial training was first proposed by \cite{goodfellow2014generative} as a
+way to build a generative model capable of producing samples from random noise
+$z \sim p_Z$. More specifically, the authors pit a generative model $G$ against
+an adversary classifier $D$ whose repelling objective is to recognize real from
+generated data. Both models $G$ and $D$ are trained simultaneously, in such a
+way that $G$ learns to produce samples that are difficult to identify by $D$,
+while $D$ incrementally adapts to changes in $G$. At the equilibrium, $G$ models
+a distribution whose samples can be identified by $D$ only by chance. In other
+words, assuming enough capacity in $D$ and  $G$, the distribution $p_{G(Z)}$
+eventually converges towards the real distribution $p_X$.
+
+In this work, we repurpose adversarial training for regularizing the
+construction of the predictive model $f$ so as to satisfy
+Eqn.~\ref{eqn:criterion}.
+
+\glnote{describe baseline}
+\glnote{describe adversarial approach}
+\glnote{proof that it solves Eqn. 2}
 

### 2016-08-02 paper.tex
method

@@
 \mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
-nuisance parameter $\lambda$ whose values define the family of its systematic uncertainties.
-That is, $X_\lambda$ and $Y$ induce together
-a joint probability distribution $p(X,Y|\lambda)$, where the
+nuisance parameter $\lambda \in \mathbb{R}$ whose values define a continuously
+parameterized family of its systematic uncertainties. That is, $X_\lambda$ and
+$Y$ induce together a joint probability distribution $p(X,Y|\lambda)$, where the
 conditional on $\lambda$ denotes $X_\lambda$. For training, let further assume a
@@
 $X_{\lambda_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
-values $\lambda_i$ of the nuisance parameter. Our goal is to learn a function $f :
-\mathbb{R}^p \mapsto {\cal Y}$ (e.g., a classifier if ${\cal Y}$ is a finite set
-of classes) minimizing the expected value of a loss $L(Y, f(X_{\lambda}))$, with
-the constraint that $f(X_\lambda)$ should be robust to the value of the nuisance
-parameter $\lambda$ -- which remains unknown at test time. More specifically, we
-aim at building $f$ such that in the ideal case
+values $\lambda_i$ of the nuisance parameter. Our goal is to learn a function
+$f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal Y}$ of parameters $\theta_f$
+(e.g., a neural network-based classifier if ${\cal Y}$ is a finite set of
+classes) minimizing  a loss ${\cal L}_f(\theta_f)$, with the constraint that $f(X_\lambda ;
+\theta_f)$ should be robust to the value of the nuisance parameter $\lambda$ --
+which remains unknown at test time. More specifically, we aim at building $f$
+such that in the ideal case
 \begin{equation}\label{eqn:criterion-true}
-f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega))
+f(X_{\lambda_i}(\omega) ; \theta_f) = f(X_{\lambda_j}(\omega) ; \theta_f)
 \end{equation} for any
@@
 \begin{equation}\label{eqn:criterion}
-    P(\{ \omega | f(X_{\lambda_i}(\omega)) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega')) = y\}) \text{ for all $y \in {\cal Y}$}.
+    P(\{ \omega | f(X_{\lambda_i}(\omega) ; \theta_f) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega') ; \theta_f) = y\}) \text{ for all $y \in {\cal Y}$}.
 \end{equation}
 In words, we are looking for a predictive function $f$ such that  the
-distribution of $f(X_\lambda)$ is invariant with respect to the nuisance
+distribution of $f(X_\lambda; \theta_f)$ is invariant with respect to the nuisance
 parameter $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn:criterion-true} is
 true necessarily satisfies Eqn.~\ref{eqn:criterion}. The converse is however in
-general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega)) = y \}$
-and $\{ \omega' | f(X_{\lambda_j}(\omega')) = y \}$ do not need to be the same
+general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega); \theta_f) = y \}$
+and $\{ \omega' | f(X_{\lambda_j}(\omega'); \theta_f) = y \}$ do not need to be the same
 for the equality to hold.
@@
 way to build a generative model capable of producing samples from random noise
-$z \sim p_Z$. More specifically, the authors pit a generative model $G$ against
-an adversary classifier $D$ whose repelling objective is to recognize real from
-generated data. Both models $G$ and $D$ are trained simultaneously, in such a
-way that $G$ learns to produce samples that are difficult to identify by $D$,
-while $D$ incrementally adapts to changes in $G$. At the equilibrium, $G$ models
-a distribution whose samples can be identified by $D$ only by chance. In other
-words, assuming enough capacity in $D$ and  $G$, the distribution $p_{G(Z)}$
-eventually converges towards the real distribution $p_X$.
+$z \sim p_Z$. More specifically, the authors pit a generative model $g: {\cal Z}
+\mapsto \mathbb{R}^p$ against an adversary classifier $d : \mathbb{R}^p \mapsto \{
+0, 1\}$ whose antagonistic objective is to recognize real data $X$ from generated data $g(Z)$. Both
+models $g$ and $d$ are trained simultaneously, in such a way that $g$ learns to
+produce samples that are difficult to identify by $d$, while $d$ incrementally
+adapts to changes in $g$. At the equilibrium, $g$ models a distribution whose
+samples can be identified by $d$ only by chance. In other words, assuming enough
+capacity in $d$ and  $g$, the distribution $p_{g(Z)}$ eventually converges
+towards the real distribution $p_X$.
 
-In this work, we repurpose adversarial training for regularizing the
-construction of the predictive model $f$ so as to satisfy
-Eqn.~\ref{eqn:criterion}.
+In this work, we repurpose adversarial training as a means to regularize the
+predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. In
+particular, we pit $f$ against an adversary regression model $r(\cdot ;
+\theta_r) : \mathbb{R} \mapsto \mathbb{R}$ of parameters $\theta_r$ and
+associated loss ${\cal L}_r(\theta_r)$, taking as input realizations of $f(X_\lambda; \theta_f)$
+and producing as output a prediction $\smash{\hat\lambda} = r(f(X_\lambda); \theta_f)$
+of the nuisance parameter. If $f(X_\lambda; \theta_f)$ varies with $\lambda$,
+then the corresponding correlation can be captured by $r$. By contrast, if
+$f(X_\lambda; \theta_f)$ is invariant with $\lambda$ as we require, then $r$
+should perform poorly. Training $f$ such that it minimizes the performance of $r$
+should therefore acts as a regularization towards Eqn.~\ref{eqn:criterion}.
 
-\glnote{describe baseline}
-\glnote{describe adversarial approach}
-\glnote{proof that it solves Eqn. 2}
+As for generative adversarial networks, we propose to
+train $f$ and $r$ simultaneously, which can be carried out by considering
+the value function
+\begin{equation}
+    E(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - {\cal L}_r(\theta_r)
+\end{equation}
+and finding the saddle point $\smash{\hat\theta_f}, \smash{\hat\theta_r}$ such that
+\begin{align}
+    \smash{\hat\theta_f} &= \arg \min_{\theta_f} E(\theta_f, \smash{\hat\theta_r}), \\
+    \smash{\hat\theta_r} &= \arg \max_{\theta_r} E(\smash{\hat\theta_f}, \theta_r).
+\end{align}
+
+\glnote{What are the necessary conditions on ${\cal L}_r$ to imply
+Eqn.~\ref{eqn:criterion} at the saddle point? We should clarify and prove that
+formally! There may also be assumptions required on the learning problem itself.}
 

### 2016-08-02 paper.tex
save

@@
 (e.g., a neural network-based classifier if ${\cal Y}$ is a finite set of
-classes) minimizing  a loss ${\cal L}_f(\theta_f)$, with the constraint that $f(X_\lambda ;
+classes) and minimizing  a loss ${\cal L}_f(\theta_f)$. In addition, we require
+the constraint that $f(X_\lambda ;
 \theta_f)$ should be robust to the value of the nuisance parameter $\lambda$ --
@@
 associated loss ${\cal L}_r(\theta_r)$, taking as input realizations of $f(X_\lambda; \theta_f)$
-and producing as output a prediction $\smash{\hat\lambda} = r(f(X_\lambda); \theta_f)$
-of the nuisance parameter. If $f(X_\lambda; \theta_f)$ varies with $\lambda$,
+and producing as output predictions $\smash{\hat\lambda} = r(f(X_\lambda); \theta_f)$
+of the nuisance parameter. If $f(X_\lambda(\omega); \theta_f)$ varies with $\lambda$,
 then the corresponding correlation can be captured by $r$. By contrast, if
-$f(X_\lambda; \theta_f)$ is invariant with $\lambda$ as we require, then $r$
+$f(X_\lambda(\omega); \theta_f)$ is invariant with $\lambda$ as we require, then $r$
 should perform poorly. Training $f$ such that it minimizes the performance of $r$
@@
 
+\glnote{Explain this can be done simply using SGD.}
 \glnote{What are the necessary conditions on ${\cal L}_r$ to imply

### 2016-08-03 paper.tex
save

@@
 classes) and minimizing  a loss ${\cal L}_f(\theta_f)$. In addition, we require
-the constraint that $f(X_\lambda ;
-\theta_f)$ should be robust to the value of the nuisance parameter $\lambda$ --
+that $f(X_\lambda ; \theta_f)$ should be robust to the value of the nuisance parameter $\lambda$ --
 which remains unknown at test time. More specifically, we aim at building $f$
@@
 
+\glnote{An alternative objective, but which leads to something
+significantly different, is to aim at getting the best model adapted
+to the unknown $\lambda$, possibly using predictive information from
+regions of the input space affected by the nuisance parameter.}
+
 \section{Method}
@@
 adapts to changes in $g$. At the equilibrium, $g$ models a distribution whose
-samples can be identified by $d$ only by chance. In other words, assuming enough
+samples can be identified by $d$ only by chance. That is, assuming enough
 capacity in $d$ and  $g$, the distribution $p_{g(Z)}$ eventually converges
@@
 \theta_r) : \mathbb{R} \mapsto \mathbb{R}$ of parameters $\theta_r$ and
-associated loss ${\cal L}_r(\theta_r)$, taking as input realizations of $f(X_\lambda; \theta_f)$
-and producing as output predictions $\smash{\hat\lambda} = r(f(X_\lambda); \theta_f)$
-of the nuisance parameter. If $f(X_\lambda(\omega); \theta_f)$ varies with $\lambda$,
+associated loss ${\cal L}_r(\theta_r)$. This regressor takes as input realizations of $f(X_\lambda; \theta_f)$
+and produces as output predictions $\smash{\hat\lambda} = r(f(X_\lambda; \theta_f))$
+of the nuisance parameter. If $p(f(X_\lambda; \theta_f))$ varies with $\lambda$,
 then the corresponding correlation can be captured by $r$. By contrast, if
-$f(X_\lambda(\omega); \theta_f)$ is invariant with $\lambda$ as we require, then $r$
-should perform poorly. Training $f$ such that it minimizes the performance of $r$
-should therefore acts as a regularization towards Eqn.~\ref{eqn:criterion}.
+$p(f(X_\lambda; \theta_f))$ is invariant with $\lambda$ as we require, then $r$
+should perform poorly. Training $f$ such that it additionally minimizes the performance of $r$
+therefore acts as a regularization towards Eqn.~\ref{eqn:criterion}.
 
 As for generative adversarial networks, we propose to
-train $f$ and $r$ simultaneously, which can be carried out by considering
+train $f$ and $r$ simultaneously, which we carry out by considering
 the value function
@@
 \end{equation}
-and finding the saddle point $\smash{\hat\theta_f}, \smash{\hat\theta_r}$ such that
+that we optimize by finding the saddle point $(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ such that
 \begin{align}
@@
 \glnote{Similar to domain adaptation, but with infinitely many domains,
-as parameterized by $\lambda$.}
+as parameterized by $\lambda$, also related to transfer learning.}
 

### 2016-08-08 paper.tex
added some notes

@@
 % https://www.slac.stanford.edu/econf/C030908/papers/TUAT004.pdf
-\glnote{Motivate the criterion: allow to derive guarantees no matter $\lambda$,
-while $f$ may otherwise be good or very bad depending on the nuisance if trained
-on the mixture. E.g. to form hypothesis to be later confirmed by data.}
+\glnote{Motivate the criterion (which may not be obvious for the ML crowd).}
 
@@
 Since we do not have training tuples $(X_{\lambda_i}(\omega),
-X_{\lambda_j}(\omega))$ (for the same unknown $\omega$), we propose instead to
+X_{\lambda_j}(\omega))$ (for the same unknown $\omega$) \glnote{Actually we do, as discussed with Michael!}, we propose instead to
 solve the closely related problem of finding a predictive function $f$ such that
@@
 
+\glnote{Expand the equation to make it clearer that $f$ is plugged into $r$.}
 \glnote{Explain this can be done simply using SGD.}

### 2016-08-22 paper.tex
save

--- 2016-08-03 paper.tex
@@
 % https://www.slac.stanford.edu/econf/C030908/papers/TUAT004.pdf
-\glnote{Motivate the criterion: allow to derive guarantees no matter $\lambda$,
-while $f$ may otherwise be good or very bad depending on the nuisance if trained
-on the mixture. E.g. to form hypothesis to be later confirmed by data.}
+\glnote{Motivate the criterion (which may not be obvious for the ML crowd). See pivotal quantity motivation.}
 
@@
 \mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
-nuisance parameter $\lambda \in \mathbb{R}$ whose values define a continuously
+nuisance parameter $\lambda \in \Lambda \subseteq \mathbb{R}$ whose values define a
 parameterized family of its systematic uncertainties. That is, $X_\lambda$ and
@@
 \end{equation}
-In words, we are looking for a predictive function $f$ such that  the
-distribution of $f(X_\lambda; \theta_f)$ is invariant with respect to the nuisance
-parameter $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn:criterion-true} is
+In words, we are looking for a predictive function $f$ which is a pivotal quantity with respect to the nuisance parameter.
+That is, such that  the
+distribution of $f(X_\lambda; \theta_f)$ is invariant with respect to $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn:criterion-true} is
 true necessarily satisfies Eqn.~\ref{eqn:criterion}. The converse is however in
@@
 for the equality to hold.
-
-\glnote{An alternative objective, but which leads to something
-significantly different, is to aim at getting the best model adapted
-to the unknown $\lambda$, possibly using predictive information from
-regions of the input space affected by the nuisance parameter.}
 
@@
 predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. In
-particular, we pit $f$ against an adversary regression model $r(\cdot ;
-\theta_r) : \mathbb{R} \mapsto \mathbb{R}$ of parameters $\theta_r$ and
-associated loss ${\cal L}_r(\theta_r)$. This regressor takes as input realizations of $f(X_\lambda; \theta_f)$
+particular, we pit $f$ against an adversary model $r(\cdot ;
+\theta_r) : \mathbb{R} \mapsto \Lambda$ of parameters $\theta_r$ and
+associated loss ${\cal L}_r(\theta_r)$. This model takes as input realizations of $f(X_\lambda; \theta_f)$
 and produces as output predictions $\smash{\hat\lambda} = r(f(X_\lambda; \theta_f))$

### 2016-08-28 tex/paper.tex
save

--- 2016-08-22 paper.tex
@@
 \mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
-nuisance parameter $\lambda \in \Lambda \subseteq \mathbb{R}$ whose values define a
+nuisance parameter $\lambda \in \Lambda$ whose values define a
 parameterized family of its systematic uncertainties. That is, $X_\lambda$ and
@@
 solve the closely related problem of finding a predictive function $f$ such that
-\begin{equation}\label{eqn:criterion}
+\begin{equation}\label{eqn:criterion-measure}
     P(\{ \omega | f(X_{\lambda_i}(\omega) ; \theta_f) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega') ; \theta_f) = y\}) \text{ for all $y \in {\cal Y}$}.
 \end{equation}
-In words, we are looking for a predictive function $f$ which is a pivotal quantity with respect to the nuisance parameter.
-That is, such that  the
-distribution of $f(X_\lambda; \theta_f)$ is invariant with respect to $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn:criterion-true} is
-true necessarily satisfies Eqn.~\ref{eqn:criterion}. The converse is however in
-general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega); \theta_f) = y \}$
-and $\{ \omega' | f(X_{\lambda_j}(\omega'); \theta_f) = y \}$ do not need to be the same
+In words, we are looking for a predictive function $f$ which is a pivotal
+quantity \citep{degroot1986probability} with respect to the nuisance parameter.
+That is, such that  the distribution of $f(X_\lambda; \theta_f)$ is invariant
+with respect to $\lambda$. Note that a function $f$ for which
+Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
+Eqn.~\ref{eqn:criterion-measure}. The converse is however in general not true, since the
+sets of samples $\{ \omega | f(X_{\lambda_i}(\omega); \theta_f) = y \}$ and $\{
+\omega' | f(X_{\lambda_j}(\omega'); \theta_f) = y \}$ do not need to be the same
 for the equality to hold.
+With this distinction, but as only Eqn.~\ref{eqn:criterion-measure} is
+of direct interest in this work, we further simplify notations by denoting
+the pivotal quantity criterion as
+\begin{equation}\label{eqn:criterion}
+    p(f(X ; \theta_f) | \lambda_i ) = p(f(X ; \theta_f) | \lambda_j ) \text{ for all $\lambda_i,\lambda_j \in  \Lambda$}
+\end{equation}.
 
@@
 
-In this work, we repurpose adversarial training as a means to regularize the
+In this work, we repurpose adversarial training as a means to constraint the
 predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. In
@@
 \theta_r) : \mathbb{R} \mapsto \Lambda$ of parameters $\theta_r$ and
-associated loss ${\cal L}_r(\theta_r)$. This model takes as input realizations of $f(X_\lambda; \theta_f)$
-and produces as output predictions $\smash{\hat\lambda} = r(f(X_\lambda; \theta_f))$
-of the nuisance parameter. If $p(f(X_\lambda; \theta_f))$ varies with $\lambda$,
+associated loss ${\cal L}_r(\theta_r)$. This model takes as input realizations of $f(X; \theta_f)$
+and produces as output estimates $r(f(X; \theta_f)) = \hat{p}(\lambda | f(X; \theta_f)) $
+that $f(X; \theta_f)$ is generated from the nuisance value $\lambda$. If $p(f(X; \theta_f)|\lambda)$ varies with $\lambda$,
 then the corresponding correlation can be captured by $r$. By contrast, if
-$p(f(X_\lambda; \theta_f))$ is invariant with $\lambda$ as we require, then $r$
+$p(f(X; \theta_f)|\lambda)$ is invariant with $\lambda$ as we require, then $r$
 should perform poorly. Training $f$ such that it additionally minimizes the performance of $r$
@@
 
-\glnote{Expand the equation to make it clearer that $f$ is plugged into $r$.}
-\glnote{Explain this can be done simply using SGD.}
-\glnote{What are the necessary conditions on ${\cal L}_r$ to imply
-Eqn.~\ref{eqn:criterion} at the saddle point? We should clarify and prove that
-formally! There may also be assumptions required on the learning problem itself.}
+\glnote{Sketch the optimization algoritm}
+
+\section{Theoretical results}
+
+\glnote{proof 1: if r optimal and r(f(x)) = 1/N then f is pivotal. [OK]}
+\glnote{proof 2: if Lf can be minimized under the pivotal constraint, then at the saddle point r is such that r(f(x)) = 1/N. [OK]}
 

### 2016-08-29 tex/paper.tex
added algorithm

@@
 \usepackage{color}
+
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 \begin{equation}\label{eqn:criterion-true}
-f(X_{\lambda_i}(\omega) ; \theta_f) = f(X_{\lambda_j}(\omega) ; \theta_f)
-\end{equation} for any
-sample $\omega \in \Omega$ and any $\lambda_i, \lambda_j$ pair of values of the
+f(X_{\lambda}(\omega) ; \theta_f) = f(X_{\lambda^\prime}(\omega) ; \theta_f)
+\end{equation} for all
+samples $\omega \in \Omega$ and all $\lambda, \lambda^\prime$ pairs of values of the
 nuisance parameter.
 
-Since we do not have training tuples $(X_{\lambda_i}(\omega),
-X_{\lambda_j}(\omega))$ (for the same unknown $\omega$), we propose instead to
+Since we do not have training tuples $(X_{\lambda}(\omega),
+X_{\lambda^\prime}(\omega))$ (for the same unknown $\omega$), we propose instead to
 solve the closely related problem of finding a predictive function $f$ such that
 \begin{equation}\label{eqn:criterion-measure}
-    P(\{ \omega | f(X_{\lambda_i}(\omega) ; \theta_f) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega') ; \theta_f) = y\}) \text{ for all $y \in {\cal Y}$}.
+    P(\{ \omega | f(X_{\lambda}(\omega) ; \theta_f) = y \} ) = P( \{ \omega' | f(X_{\lambda^\prime}(\omega') ; \theta_f) = y\}) \text{ for all $y \in {\cal Y}$}.
 \end{equation}
@@
 That is, such that  the distribution of $f(X_\lambda; \theta_f)$ is invariant
-with respect to $\lambda$. Note that a function $f$ for which
+with respect to the value of $\lambda$. Note that a function $f$ for which
 Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
 Eqn.~\ref{eqn:criterion-measure}. The converse is however in general not true, since the
-sets of samples $\{ \omega | f(X_{\lambda_i}(\omega); \theta_f) = y \}$ and $\{
-\omega' | f(X_{\lambda_j}(\omega'); \theta_f) = y \}$ do not need to be the same
+sets of samples $\{ \omega | f(X_{\lambda}(\omega); \theta_f) = y \}$ and $\{
+\omega' | f(X_{\lambda^\prime}(\omega'); \theta_f) = y \}$ do not need to be the same
 for the equality to hold.
-With this distinction, but as only Eqn.~\ref{eqn:criterion-measure} is
-of direct interest in this work, we further simplify notations by denoting
+In order to simplify notations,
+and as only Eqn.~\ref{eqn:criterion-measure} is
+of direct interest in this work, we denote from here on
 the pivotal quantity criterion as
 \begin{equation}\label{eqn:criterion}
-    p(f(X ; \theta_f) | \lambda_i ) = p(f(X ; \theta_f) | \lambda_j ) \text{ for all $\lambda_i,\lambda_j \in  \Lambda$}
-\end{equation}.
+    p(f(X ; \theta_f) | \lambda ) = p(f(X ; \theta_f) | \lambda^\prime ) \text{ for all $\lambda,\lambda^\prime \in  \Lambda$.}
+\end{equation}
 
@@
 predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. In
-particular, we pit $f$ against an adversary model $r(\cdot ;
+particular, we pit $f$ against an adversary classifier $r(\cdot ;
 \theta_r) : \mathbb{R} \mapsto \Lambda$ of parameters $\theta_r$ and
-associated loss ${\cal L}_r(\theta_r)$. This model takes as input realizations of $f(X; \theta_f)$
-and produces as output estimates $r(f(X; \theta_f)) = \hat{p}(\lambda | f(X; \theta_f)) $
-that $f(X; \theta_f)$ is generated from the nuisance value $\lambda$. If $p(f(X; \theta_f)|\lambda)$ varies with $\lambda$,
+associated loss ${\cal L}_r(\theta_r)$.
+Assuming that $\Lambda$ defines a finite family of nuisance values $\lambda_l$ (for $l=1, \dots, |\Lambda|$),
+this model takes as input realizations of $f(X; \theta_f)$
+and produces as output probability estimates $r(f(X; \theta_f); \theta_r)_l = \hat{p}(\lambda_l| f(X; \theta_f)) $
+that $f(X; \theta_f)$ is generated from the nuisance value $\lambda_l$. If $p(f(X; \theta_f)|\lambda)$ varies with $\lambda$,
 then the corresponding correlation can be captured by $r$. By contrast, if
-$p(f(X; \theta_f)|\lambda)$ is invariant with $\lambda$ as we require, then $r$
-should perform poorly. Training $f$ such that it additionally minimizes the performance of $r$
+$p(f(X; \theta_f)|\lambda)$ is invariant with $\lambda$, as we require, then $r$
+should perform poorly and be close to random guessing. Training $f$ such that it additionally minimizes the performance of $r$
 therefore acts as a regularization towards Eqn.~\ref{eqn:criterion}.
@@
 \begin{align}
-    \smash{\hat\theta_f} &= \arg \min_{\theta_f} E(\theta_f, \smash{\hat\theta_r}), \\
-    \smash{\hat\theta_r} &= \arg \max_{\theta_r} E(\smash{\hat\theta_f}, \theta_r).
+    \smash{\hat\theta_f} &= \arg \min_{\theta_f} E(\theta_f, \smash{\hat\theta_r}), \label{eqn:min_thetaf} \\
+    \smash{\hat\theta_r} &= \arg \max_{\theta_r} E(\smash{\hat\theta_f}, \theta_r) \label{eqn:max_thetar}.
 \end{align}
+The adversarial training procedure to obtain $(\smash{\hat\theta_f},
+\smash{\hat\theta_r})$ is formally presented in
+Algorithm~\ref{alg:adversarial-training} in the case of $f$ being a classifier
+and of using the log-loss for both ${\cal L}_f$ and ${\cal L}_r$. The algorithm
+consists in using stochastic gradient descent alternatively to optimize
+Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.
 
-\glnote{Sketch the optimization algoritm}
+\begin{algorithm}[t]
+\caption{Adversarial training of a classifier $f$ against an adversary $r$.\\
+{\it Inputs:} training data $\{ x_i, y_i, \lambda_i \}_{i=1}^N$\\
+{\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$\\
+{\it Hyper-parameters:} Number $T$ of training iterations, Number $K$ of gradient steps to update $r$.}
+\label{alg:adversarial-training}
+\begin{algorithmic}[1]
+    \For{$t=1$ to $T$}
+        \For{$k=1$ to $K$} \Comment{Update $r$}
+            \State{Sample minibatch $\{x_m, \lambda_m \}_{m=1}^M$} of size $M$;
+            \State{With $\theta_f$ fixed, update the adversary $r$ by ascending its stochastic gradient
+            $$\nabla_{\theta_r} \sum_{m=1}^M \left[ -\sum_{l=1}^{|\Lambda|} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+        \EndFor
+        \State{Sample minibatch $\{x_m, y_m, \lambda_m \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
+        \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient
+        $$\nabla_{\theta_f} \sum_{m=1}^M \left[ \sum_{c=1}^{|{\cal Y}|} 1(y_m = y_c) \log f(x_m;\theta_f)_c -\sum_{l=1}^{|\Lambda|} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+    \EndFor
+\end{algorithmic}
+\end{algorithm}
 
 \section{Theoretical results}
+
+In this section, we show that Algorithm~\ref{alg:adversarial-training} converges
+to a classifier $f$ which is a pivotal quantity in the sense of Eqn.~\ref{eqn:criterion}.
 

### 2016-08-29 tex/paper.tex
save (work in progress)

@@
             \State{Sample minibatch $\{x_m, \lambda_m \}_{m=1}^M$} of size $M$;
-            \State{With $\theta_f$ fixed, update the adversary $r$ by ascending its stochastic gradient
-            $$\nabla_{\theta_r} \sum_{m=1}^M \left[ -\sum_{l=1}^{|\Lambda|} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+            \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
+            $$\nabla_{\theta_r} \sum_{m=1}^M \left[ -\sum_{\lambda_l \in \Lambda} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
         \EndFor
         \State{Sample minibatch $\{x_m, y_m, \lambda_m \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
-        \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient
-        $$\nabla_{\theta_f} \sum_{m=1}^M \left[ \sum_{c=1}^{|{\cal Y}|} 1(y_m = y_c) \log f(x_m;\theta_f)_c -\sum_{l=1}^{|\Lambda|} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+        \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
+        $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ \sum_{y_c \in {\cal Y}} 1(y_m = y_c) \log f(x_m;\theta_f)_c -\sum_{\lambda_l \in \Lambda} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
     \EndFor
@@
 to a classifier $f$ which is a pivotal quantity in the sense of Eqn.~\ref{eqn:criterion}.
+Results below are derived in a non-parametric setting, e.g. by assuming
+that both $f$ and $r$ have enough capacity. To simplify the presentation,
+we also assume the uniform prior
+$p(\lambda) = \frac{1}{|\Lambda|}$ for all $\lambda \in \Lambda$, e.g. by
+having the same number of training samples for each modality $\lambda$ of
+the nuisance parameter.
+
+\begin{proposition}
+
+Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
+\theta_r)$. If $r(f(x;\theta_f) ; \hat\theta_r )_l = p(\lambda_l| f(X; \theta_f)) = \frac{1}{|\Lambda|}$ for all
+$x$ and all $\lambda_l$, then $f$ is a pivotal quantity.
+
+\end{proposition}
 

### 2016-08-29 tex/paper.tex
fix signs

--- 2016-08-29 tex/paper.tex
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}
 \caption{Adversarial training of a classifier $f$ against an adversary $r$.\\
@@
             \State{Sample minibatch $\{x_m, \lambda_m \}_{m=1}^M$} of size $M$;
-            \State{With $\theta_f$ fixed, update the adversary $r$ by ascending its stochastic gradient
-            $$\nabla_{\theta_r} \sum_{m=1}^M \left[ -\sum_{l=1}^{|\Lambda|} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+            \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
+            $$\nabla_{\theta_r} \sum_{m=1}^M \left[\sum_{\lambda_l \in \Lambda} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
         \EndFor
         \State{Sample minibatch $\{x_m, y_m, \lambda_m \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
-        \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient
-        $$\nabla_{\theta_f} \sum_{m=1}^M \left[ \sum_{c=1}^{|{\cal Y}|} 1(y_m = y_c) \log f(x_m;\theta_f)_c -\sum_{l=1}^{|\Lambda|} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+        \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
+        $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\sum_{y_c \in {\cal Y}} 1(y_m = y_c) \log f(x_m;\theta_f)_c  +\sum_{\lambda_l \in \Lambda} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
     \EndFor
@@
 
-In this section, we show that Algorithm~\ref{alg:adversarial-training} converges
-to a classifier $f$ which is a pivotal quantity in the sense of Eqn.~\ref{eqn:criterion}.
+In this section, we show that in the setting of
+Algorithm~\ref{alg:adversarial-training}, the procedure converges to a
+classifier $f$ which is a pivotal quantity in the sense of
+Eqn.~\ref{eqn:criterion}. Results below are derived in a non-parametric setting,
+by assuming that both $f$ and $r$ have enough capacity. To simplify the
+presentation, we also assume the uniform prior $p(\lambda) =
+\frac{1}{|\Lambda|}$ for all $\lambda \in \Lambda$, e.g. by having the same
+number of training samples for each modality $\lambda$ of the nuisance
+parameter.
+
+\begin{proposition}
+
+Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
+\theta_r)$. If $r(f(X;\theta_f) ; \hat\theta_r )_l = p(\lambda_l| f(X; \theta_f)) = \frac{1}{|\Lambda|}$
+for all $\lambda_l$, then $f$ is a pivotal quantity.
+
+\end{proposition}
 

### 2016-08-29 tex/paper.tex
L_r also depends on theta_f
save

@@
 \theta_r) : \mathbb{R} \mapsto \Lambda$ of parameters $\theta_r$ and
-associated loss ${\cal L}_r(\theta_r)$.
+associated loss ${\cal L}_r(\theta_f, \theta_r)$.
 Assuming that $\Lambda$ defines a finite family of nuisance values $\lambda_l$ (for $l=1, \dots, |\Lambda|$),
-this model takes as input realizations of $f(X; \theta_f)$
+this classifier takes as input realizations of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$ parameters,
 and produces as output probability estimates $r(f(X; \theta_f); \theta_r)_l = \hat{p}(\lambda_l| f(X; \theta_f)) $
@@
 \begin{equation}
-    E(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - {\cal L}_r(\theta_r)
+    E(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - {\cal L}_r(\theta_f, \theta_r)
 \end{equation}
@@
 Algorithm~\ref{alg:adversarial-training} in the case of $f$ being a classifier
-and of using the log-loss for both ${\cal L}_f$ and ${\cal L}_r$. The algorithm
+and of the cross-entropy loss for both ${\cal L}_f$ and ${\cal L}_r$. The algorithm
 consists in using stochastic gradient descent alternatively to optimize

### 2016-08-30 tex/paper.tex
save

@@
 Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
-\theta_r)$. If $r(f(X;\theta_f) ; \hat\theta_r )_l = p(\lambda_l| f(X; \theta_f)) = \frac{1}{|\Lambda|}$
+\theta_r)$. If $r(f(X;\theta_f) ; \hat\theta_r )_l = \frac{1}{|\Lambda|}$
 for all $\lambda_l$, then $f$ is a pivotal quantity.
@@
 
-\glnote{proof 1: if r optimal and r(f(x)) = 1/N then f is pivotal. [OK]}
+\begin{proof}
+
+Let us first recall that the cross-entropy for distributions $p$ and $q$ is
+minimized when $p=q$. For ${\cal L}_r$ defined as the cross-entropy between the
+true conditional distribution of the nuisance $p_{\lambda|f(X;\theta_f)}$ and
+the approximate conditional distribution of the nuisance
+$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\hat\theta_r =
+\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
+\theta_r)$ are therefore such that $p_{\lambda|f(X;\theta_f)} =
+p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$.
+
+In other words, for all $\lambda_l \in  \Lambda$, we have $r(f(X;\theta_f);\hat\theta_r)_l = p(\lambda_l|f(X;\theta_f))$.
+By assumption, $r(f(X;\theta_f);\hat\theta_r)_l = \frac{1}{|\Lambda|}$,
+and therefore $p(\lambda_l|f(X;\theta_f)) = \frac{1}{|\Lambda|}$.
+Using the Bayes' rule, we write
+\begin{align*}
+    p(f(X;\theta_f)|\lambda_l) &= \frac{ p(\lambda_l|f(X;\theta_f)) p(f(X;\theta_f)) } { p(\lambda_l)} \\
+                               &= \frac{ \tfrac{1}{|\Lambda|} p(f(X;\theta_f)) } { \tfrac{1}{|\Lambda|} } \\
+                               &= p(f(X;\theta_f)),
+\end{align*}
+which holds for all $\lambda_l \in \Lambda$ and implies that $f$ is a pivotal quantity.
+\end{proof}
+
+\begin{proposition}
 \glnote{proof 2: if Lf can be minimized under the pivotal constraint, then at the saddle point r is such that r(f(x)) = 1/N. [OK]}
+\end{proposition}
 

### 2016-08-30 tex/paper.tex
rename lambda to z

--- 2016-08-29 tex/paper.tex
@@
 sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
-Let consider the multivariate random variables $X_\lambda: \Omega \mapsto
-\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
-nuisance parameter $\lambda \in \Lambda$ whose values define a
-parameterized family of its systematic uncertainties. That is, $X_\lambda$ and
-$Y$ induce together a joint probability distribution $p(X,Y|\lambda)$, where the
-conditional on $\lambda$ denotes $X_\lambda$. For training, let further assume a
-finite set $\{ x_i, y_i, \lambda_i \}_{i=1}^N$ of realizations
-$X_{\lambda_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
-values $\lambda_i$ of the nuisance parameter. Our goal is to learn a function
+Let consider the multivariate random variables $X_z: \Omega \mapsto
+\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence on a
+nuisance parameter $Z$ whose values $z \in {\cal Z}$  define a
+parameterized family of its systematic uncertainties. That is, $X_z$ and
+$Y$ induce together a joint probability distribution $p(X,Y|z)$, where the
+conditional on $z$ denotes $X_z$. For training, let further assume a
+finite set $\{ x_i, y_i, z_i \}_{i=1}^N$ of realizations
+$X_{z_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
+values $z_i$ of the nuisance parameter. Our goal is to learn a function
 $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal Y}$ of parameters $\theta_f$
@@
 classes) and minimizing  a loss ${\cal L}_f(\theta_f)$. In addition, we require
-that $f(X_\lambda ; \theta_f)$ should be robust to the value of the nuisance parameter $\lambda$ --
+that $f(X_z ; \theta_f)$ should be robust to the value $z$ of the nuisance parameter  --
 which remains unknown at test time. More specifically, we aim at building $f$
@@
 \begin{equation}\label{eqn:criterion-true}
-f(X_{\lambda}(\omega) ; \theta_f) = f(X_{\lambda^\prime}(\omega) ; \theta_f)
+f(X_{z}(\omega) ; \theta_f) = f(X_{z^\prime}(\omega) ; \theta_f)
 \end{equation} for all
-samples $\omega \in \Omega$ and all $\lambda, \lambda^\prime$ pairs of values of the
+samples $\omega \in \Omega$ and all $z, z^\prime$ pairs of values of the
 nuisance parameter.
 
-Since we do not have training tuples $(X_{\lambda}(\omega),
-X_{\lambda^\prime}(\omega))$ (for the same unknown $\omega$), we propose instead to
+Since we do not have training tuples $(X_{z}(\omega),
+X_{z^\prime}(\omega))$ (for the same unknown $\omega$), we propose instead to
 solve the closely related problem of finding a predictive function $f$ such that
 \begin{equation}\label{eqn:criterion-measure}
-    P(\{ \omega | f(X_{\lambda}(\omega) ; \theta_f) = y \} ) = P( \{ \omega' | f(X_{\lambda^\prime}(\omega') ; \theta_f) = y\}) \text{ for all $y \in {\cal Y}$}.
+    P(\{ \omega | f(X_{z}(\omega) ; \theta_f) = y \} ) = P( \{ \omega' | f(X_{z^\prime}(\omega') ; \theta_f) = y\}) \text{ for all $y \in {\cal Y}$}.
 \end{equation}
@@
 quantity \citep{degroot1986probability} with respect to the nuisance parameter.
-That is, such that  the distribution of $f(X_\lambda; \theta_f)$ is invariant
-with respect to the value of $\lambda$. Note that a function $f$ for which
+That is, such that  the distribution of $f(X_z; \theta_f)$ is invariant
+with respect to the value $z$ of the nuisance. Note that a function $f$ for which
 Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
 Eqn.~\ref{eqn:criterion-measure}. The converse is however in general not true, since the
-sets of samples $\{ \omega | f(X_{\lambda}(\omega); \theta_f) = y \}$ and $\{
-\omega' | f(X_{\lambda^\prime}(\omega'); \theta_f) = y \}$ do not need to be the same
+sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = y \}$ and $\{
+\omega' | f(X_{z^\prime}(\omega'); \theta_f) = y \}$ do not need to be the same
 for the equality to hold.
@@
 \begin{equation}\label{eqn:criterion}
-    p(f(X ; \theta_f) | \lambda ) = p(f(X ; \theta_f) | \lambda^\prime ) \text{ for all $\lambda,\lambda^\prime \in  \Lambda$.}
+    p(f(X ; \theta_f) | z ) = p(f(X ; \theta_f) | z^\prime ) \text{ for all $z,z^\prime \in  {\cal Z}$.}
 \end{equation}
@@
 particular, we pit $f$ against an adversary classifier $r(\cdot ;
-\theta_r) : \mathbb{R} \mapsto \Lambda$ of parameters $\theta_r$ and
+\theta_r) : \mathbb{R} \mapsto {\cal Z}$ of parameters $\theta_r$ and
 associated loss ${\cal L}_r(\theta_f, \theta_r)$.
-Assuming that $\Lambda$ defines a finite family of nuisance values $\lambda_l$ (for $l=1, \dots, |\Lambda|$),
+Assuming that ${\cal Z}$ defines a finite family of nuisance values $z_l$ (for $l=1, \dots, |{\cal Z}|$),
 this classifier takes as input realizations of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$ parameters,
-and produces as output probability estimates $r(f(X; \theta_f); \theta_r)_l = \hat{p}(\lambda_l| f(X; \theta_f)) $
-that $f(X; \theta_f)$ is generated from the nuisance value $\lambda_l$. If $p(f(X; \theta_f)|\lambda)$ varies with $\lambda$,
+and produces as output probability estimates $r(f(X; \theta_f); \theta_r)_l = \hat{p}(z_l| f(X; \theta_f)) $
+that $f(X; \theta_f)$ is generated from the nuisance value $z_l$. If $p(f(X; \theta_f)|z)$ varies with $z$,
 then the corresponding correlation can be captured by $r$. By contrast, if
-$p(f(X; \theta_f)|\lambda)$ is invariant with $\lambda$, as we require, then $r$
+$p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then $r$
 should perform poorly and be close to random guessing. Training $f$ such that it additionally minimizes the performance of $r$
@@
 \caption{Adversarial training of a classifier $f$ against an adversary $r$.\\
-{\it Inputs:} training data $\{ x_i, y_i, \lambda_i \}_{i=1}^N$\\
+{\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$\\
 {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$\\
@@
         \For{$k=1$ to $K$} \Comment{Update $r$}
-            \State{Sample minibatch $\{x_m, \lambda_m \}_{m=1}^M$} of size $M$;
+            \State{Sample minibatch $\{x_m, z_m \}_{m=1}^M$} of size $M$;
             \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
-            $$\nabla_{\theta_r} \sum_{m=1}^M \left[\sum_{\lambda_l \in \Lambda} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+            $$\nabla_{\theta_r} \sum_{m=1}^M \left[\sum_{z_l \in {\cal Z}} 1(z_m = z_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
         \EndFor
-        \State{Sample minibatch $\{x_m, y_m, \lambda_m \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
+        \State{Sample minibatch $\{x_m, y_m, z_m \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
         \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
-        $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\sum_{y_c \in {\cal Y}} 1(y_m = y_c) \log f(x_m;\theta_f)_c  +\sum_{\lambda_l \in \Lambda} 1(\lambda_m = \lambda_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+        $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\sum_{y_c \in {\cal Y}} 1(y_m = y_c) \log f(x_m;\theta_f)_c  +\sum_{z_l \in {\cal Z}} 1(z_m = z_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
     \EndFor
@@
 by assuming that both $f$ and $r$ have enough capacity. To simplify the
-presentation, we also assume the uniform prior $p(\lambda) =
-\frac{1}{|\Lambda|}$ for all $\lambda \in \Lambda$, e.g. by having the same
-number of training samples for each modality $\lambda$ of the nuisance
+presentation, we also assume the uniform prior $p(z) =
+\frac{1}{|{\cal Z}|}$ for all $z \in {\cal Z}$, e.g. by having the same
+number of training samples for each modality $z$ of the nuisance
 parameter.
@@
 Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
-\theta_r)$. If $r(f(X;\theta_f) ; \hat\theta_r )_l = p(\lambda_l| f(X; \theta_f)) = \frac{1}{|\Lambda|}$
-for all $\lambda_l$, then $f$ is a pivotal quantity.
+\theta_r)$. If $r(f(X;\theta_f) ; \hat\theta_r )_l = \frac{1}{|{\cal Z}|}$
+for all $z_l$, then $f$ is a pivotal quantity.
 
@@
 
-\glnote{proof 1: if r optimal and r(f(x)) = 1/N then f is pivotal. [OK]}
+\begin{proof}
+
+Let us first recall that the cross-entropy for distributions $p$ and $q$ is
+minimized when $p=q$. For ${\cal L}_r$ defined as the cross-entropy between the
+true conditional distribution of the nuisance $p_{Z|f(X;\theta_f)}$ and
+the approximate conditional distribution of the nuisance
+$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\hat\theta_r =
+\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
+\theta_r)$ are therefore such that $p_{Z|f(X;\theta_f)} =
+p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$.
+
+In other words, for all $z_l \in {\cal Z}$, we have $r(f(X;\theta_f);\hat\theta_r)_l = p(z_l|f(X;\theta_f))$.
+By assumption, $r(f(X;\theta_f);\hat\theta_r)_l = \frac{1}{|{\cal Z}|}$,
+and therefore $p(z_l|f(X;\theta_f)) = \frac{1}{|{\cal Z}|}$.
+Using the Bayes' rule, we write
+\begin{align*}
+    p(f(X;\theta_f)|z_l) &= \frac{ p(z_l|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z_l)} \\
+                         &= \frac{ \tfrac{1}{|{\cal Z}|} p(f(X;\theta_f)) } { \tfrac{1}{|{\cal Z}|} } \\
+                         &= p(f(X;\theta_f)),
+\end{align*}
+which holds for all $z_l \in {\cal Z}$ and implies that $f$ is a pivotal quantity.
+\end{proof}
+
+\begin{proposition}
 \glnote{proof 2: if Lf can be minimized under the pivotal constraint, then at the saddle point r is such that r(f(x)) = 1/N. [OK]}
+\end{proposition}
 
@@
 \glnote{Similar to domain adaptation, but with infinitely many domains,
-as parameterized by $\lambda$, also related to transfer learning.}
+as parameterized by $Z$, also related to transfer learning.}
 

### 2016-08-30 tex/paper.tex
Proposition 2

@@
 
-\begin{proposition}
+\begin{proposition}\label{prop:1}
 
@@
 the approximate conditional distribution of the nuisance
-$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\hat\theta_r =
+$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\smash{\hat\theta}_r =
 \arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
@@
 
-In other words, for all $z_l \in {\cal Z}$, we have $r(f(X;\theta_f);\hat\theta_r)_l = p(z_l|f(X;\theta_f))$.
-By assumption, $r(f(X;\theta_f);\hat\theta_r)_l = \frac{1}{|{\cal Z}|}$,
+In other words, for all $z_l \in {\cal Z}$, we have $r(f(X;\theta_f);\smash{\hat\theta}_r)_l = p(z_l|f(X;\theta_f))$.
+By assumption, $r(f(X;\theta_f);\smash{\hat\theta}_r)_l = \frac{1}{|{\cal Z}|}$,
 and therefore $p(z_l|f(X;\theta_f)) = \frac{1}{|{\cal Z}|}$.
@@
 
+\begin{proposition}\label{prop:2}
+
+If there exists a saddle point $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$
+for Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar} such that
+$E(\hat\theta_f, \hat\theta_r) = H(p_{Y|X}) - \log |{\cal Z}|$, then
+$f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier with respect to
+$L_f$ and a pivotal quantity.
+
+\end{proposition}
+
+\begin{proof}
+
+For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
+\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
+\theta_r)$, in which case $p_{Z|f(X;\theta_f)} =
+p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$ and $L_r$ reduces to the entropy
+$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
+value function $E$ can therefore be rewritten as $$E'(\theta_f) = L_f(\theta_f) -
+H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
+\log |{\cal Z}| \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
+holds at $\smash{\hat\theta}_f$ when
+\begin{itemize}
+    \item $\smash{\hat\theta}_f$ corresponds to
+    the parameters of an optimal classifier, in which case the log-loss $L_f$
+    reduces to $H(p_{Y|X})$,
+    \item all
+   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
+   $p(z_l|f(X;\smash{\hat\theta}_f)) = \frac{1}{|{\cal Z}|}$ for all $z_l \in {\cal
+   Z}$ and $H(p_{Z|f(X;\smash{\hat\theta}_f)}) =
+   -\sum_{z_l \in {\cal Z}} p(z|\smash{\hat\theta}_f) \log  p(z|\smash{\hat\theta}_f) =  -\sum_{z_l \in {\cal Z}} \frac{1}{|{\cal Z}|} \log \frac{1}{|{\cal Z}|} = \log |{\cal Z}|$.
+\end{itemize}
+Accordingly, the second condition implies that $r(f(X;\smash{\hat\theta}_f) ;
+\smash{\hat\theta}_r )_l = \frac{1}{|{\cal Z}|}$ and therefore that at this
+point, because of Proposition~\ref{prop:1}, $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$, $f$ is both an optimal
+classifier and a pivotal quantity.
+\end{proof}
+
+\glnote{We should further discuss that in practice, the equality may never hold.
+We should discuss in which circumstances. In such case, the pivotal quantity
+constraint can however be enforced by outweighting the $L_r$, resulting in a
+trade-off between classifier optimality and pivotality.}
+
 \begin{proposition}
-\glnote{proof 2: if Lf can be minimized under the pivotal constraint, then at the saddle point r is such that r(f(x)) = 1/N. [OK]}
+    \glnote{It remains to prove that the procedure of Algorithm 1
+    converges towards that saddle point. The proof should be similar to the proof of convergence in the GAN paper.}
 \end{proposition}

### 2016-08-30 tex/paper.tex
save

@@
 \arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
-\theta_r)$ are therefore such that $p_{Z|f(X;\theta_f)} =
-p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$.
+\theta_r)$ are therefore such that $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$.
 
@@
 \max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
-\theta_r)$, in which case $p_{Z|f(X;\theta_f)} =
-p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$ and $L_r$ reduces to the entropy
+\theta_r)$, in which case $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$ and $L_r$ reduces to the entropy
 $H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
@@
 \log |{\cal Z}| \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
-holds at $\smash{\hat\theta}_f$ when
+holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
 \begin{itemize}
@@
 \smash{\hat\theta}_r )_l = \frac{1}{|{\cal Z}|}$ and therefore that at this
-point, because of Proposition~\ref{prop:1}, $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$, $f$ is both an optimal
-classifier and a pivotal quantity.
+point, because of Proposition~\ref{prop:1}, the optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
 \end{proof}

### 2016-08-30 tex/paper.tex
save

@@
 We should discuss in which circumstances. In such case, the pivotal quantity
-constraint can however be enforced by outweighting the $L_r$, resulting in a
+constraint can however be enforced by outweighting the $L_r$ term, resulting in a
 trade-off between classifier optimality and pivotality.}

### 2016-08-30 tex/paper.tex
added note
fix typos

@@
 $E(\hat\theta_f, \hat\theta_r) = H(p_{Y|X}) - \log |{\cal Z}|$, then
-$f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier with respect to
-$L_f$ and a pivotal quantity.
+$f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier  and a pivotal quantity.
 
@@
     the parameters of an optimal classifier, in which case the log-loss $L_f$
-    reduces to $H(p_{Y|X})$,
+    reduces to its minimum value $H(p_{Y|X})$,
     \item all
@@
    Z}$ and $H(p_{Z|f(X;\smash{\hat\theta}_f)}) =
-   -\sum_{z_l \in {\cal Z}} p(z|\smash{\hat\theta}_f) \log  p(z|\smash{\hat\theta}_f) =  -\sum_{z_l \in {\cal Z}} \frac{1}{|{\cal Z}|} \log \frac{1}{|{\cal Z}|} = \log |{\cal Z}|$.
+   -\sum_{z_l \in {\cal Z}} p(z_l|f(X;\smash{\hat\theta}_f)) \log  p(z_l|f(X;\smash{\hat\theta}_f)) =  -\sum_{z_l \in {\cal Z}} \frac{1}{|{\cal Z}|} \log \frac{1}{|{\cal Z}|} = \log |{\cal Z}|$.
 \end{itemize}
@@
 \end{proposition}
+
+\glnote{This seems to naturally extend to the case where $Z$ takes continuous
+value, where $L_r$ could be a continuous version of the cross-entropy?}
 

### 2016-08-30 tex/paper.tex
added note
save
some tex cleanup

@@
 
-% if you need to pass options to natbib, use, e.g.:
 \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading nips_2016
-%
-% to avoid loading the natbib package, add option nonatbib:
-% \usepackage[nonatbib]{nips_2016}
 
 \usepackage[final]{nips_2016}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2016}
-
 \usepackage[utf8]{inputenc} % allow utf-8 input
@@
 \usepackage{microtype}      % microtypography
-
 \usepackage{amssymb}
@@
 \usepackage{color}
-
 \usepackage{algorithm}
@@
 
-% The \author macro works with any number of authors. There are two
-% commands used to separate the names and addresses of multiple
-% authors: \And and \AND.
-%
 % Using \And between authors leaves it to LaTeX to determine where to
@@
   \texttt{g.louppe@nyu.edu} \\
-  %% examples of more authors
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
   %% \texttt{email} \\
@@
 \begin{document}
-% \nipsfinalcopy is no longer used
 
@@
 
+% ==============================================================================
+
 \begin{abstract}
 \end{abstract}
+
+% ==============================================================================
 
@@
 \glnote{Motivate the criterion (which may not be obvious for the ML crowd). See pivotal quantity motivation.}
+
+% ==============================================================================
 
@@
 \end{equation}
+
+% ==============================================================================
 
@@
 
+% ==============================================================================
+
 \section{Theoretical results}
@@
 \begin{proposition}\label{prop:1}
-
 Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
@@
 for all $z_l$, then $f$ is a pivotal quantity.
-
 \end{proposition}
@@
 \begin{proof}
-
 Let us first recall that the cross-entropy for distributions $p$ and $q$ is
@@
 \begin{proposition}\label{prop:2}
-
 If there exists a saddle point $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$
@@
 $f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier  and a pivotal quantity.
-
 \end{proposition}
@@
 \begin{proof}
-
 For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
@@
 
-\glnote{We should further discuss that in practice, the equality may never hold.
-We should discuss in which circumstances. In such case, the pivotal quantity
-constraint can however be enforced by outweighting the $L_r$ term, resulting in a
-trade-off between classifier optimality and pivotality.}
-
 \begin{proposition}
@@
 
+\glnote{We should further discuss that in practice, the equality in Prop 2 may never hold.
+We should discuss in which circumstances. In such case, the pivotal quantity
+constraint can however be enforced by outweighting the $L_r$ term, resulting in a
+trade-off between classifier optimality and pivotality.}
+
 \glnote{This seems to naturally extend to the case where $Z$ takes continuous
@@
 
+% ==============================================================================
+
 \section{Experiments}
+
+% ==============================================================================
 
@@
 
+\glnote{Other applications: removing implicit bias in data (e.g. gender bias).}
+
+% ==============================================================================
+
 \section{Conclusions}
+
+% ==============================================================================
 
@@
 \bibliographystyle{ieeetr}
-{\small
-\bibliography{bibliography.bib}}
+\bibliography{bibliography.bib}
 

### 2016-08-31 tex/paper.tex
save
added note for continuous case

@@
 way to build a generative model capable of producing samples from random noise
-$z \sim p_Z$. More specifically, the authors pit a generative model $g: {\cal Z}
+$z \sim p_Z$. More specifically, the authors pit a generative model $g: \mathbb{R}
 \mapsto \mathbb{R}^p$ against an adversary classifier $d : \mathbb{R}^p \mapsto \{
@@
 \end{algorithm}
+
+\glnote{Extension to continuous nuisance parameters: have the adversary $r$
+models the conditional distribution $p(z|f(X;\theta_f))$, e.g. with network
+whose 2 outputs are parameters of a truncated gaussian and optimized such that
+$p(z|f(X;\theta_f)) \approx N(\mu(f(X;\theta_f);\theta_r), \sigma(f(X;\theta_f);
+\theta_r)$. The associated loss can be the likelihood (or?), which is maximal when the
+approximation is perfect and minimal when the truncated gaussian degenerates to
+a uniform (when $\sigma$ is large), in which case we should have the guarantee
+that $f$ is pivotal.}
 

### 2016-08-31 tex/paper.tex
save

@@
 \glnote{Extension to continuous nuisance parameters: have the adversary $r$
-models the conditional distribution $p(z|f(X;\theta_f))$, e.g. with network
-whose 2 outputs are parameters of a truncated gaussian and optimized such that
+models the conditional distribution $p(z|f(X;\theta_f))$, e.g. with a network
+whose 2 outputs are the parameters of a truncated gaussian and optimized such that
 $p(z|f(X;\theta_f)) \approx N(\mu(f(X;\theta_f);\theta_r), \sigma(f(X;\theta_f);

### 2016-09-02 tex/paper.tex
savce

@@
 
-\title{Adversarial Training of Neural Networks\\
-against Systematic Uncertainty}
-
-% Using \And between authors leaves it to LaTeX to determine where to
-% break the lines. Using \AND forces a line break at that point. So,
-% if LaTeX puts 3 of 4 authors names on the first line, and the last
-% on the second line, try using \AND instead of \And before the third
-% author name.
+\title{Adversarial Training against Systematic Uncertainty}
 
@@
   \texttt{g.louppe@nyu.edu} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  %% \texttt{email} \\
+  \And
+  Michael Kagan\\
+  SLAC National Accelerator Laboratory\\
+  \texttt{makagan@slac.stanford.edu}\\
+  \And
+  Kyle Cranmer \\
+  New York University\\
+  \texttt{kyle.cranmer@nyu.edu}
 }
@@
 \begin{abstract}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec in faucibus
+lacus. Curabitur in viverra neque. Sed nec urna a sem blandit viverra vel a
+eros. Nam a ultrices leo, a aliquet diam. Proin cursus scelerisque leo, ut
+ullamcorper ante auctor et. Nulla odio nisl, molestie eget consectetur quis,
+eleifend id ligula. Cras lobortis egestas justo ac sagittis. In at ligula est.
+Nulla vel eros sapien. Sed bibendum metus quis dignissim scelerisque. Nulla arcu
+lectus, ornare nec vehicula quis, lobortis a mi. Cras scelerisque mauris quam,
+nec viverra ex convallis sit amet. Ut tellus turpis, lobortis sit amet finibus
+sit amet, maximus ac est.
+
 \end{abstract}
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \caption{Adversarial training of a classifier $f$ against an adversary $r$.\\
@@
 $p(z|f(X;\theta_f)) \approx N(\mu(f(X;\theta_f);\theta_r), \sigma(f(X;\theta_f);
-\theta_r)$. The associated loss can be the likelihood (or?), which is maximal when the
+\theta_r)$. The associated loss should be the negative log-likelihood (= cross-entropy), which is maximal when the
 approximation is perfect and minimal when the truncated gaussian degenerates to
@@
 that $f$ is pivotal.}
+
+\glnote{This section and theoretical results should be rewritten in such a way
+that categorical and continuous nuisance values are solved under the same
+principled approach.}
 
@@
 
-\glnote{This seems to naturally extend to the case where $Z$ takes continuous
-value, where $L_r$ could be a continuous version of the cross-entropy?}
-
 % ==============================================================================

### 2016-09-02 tex/paper.tex
unify categorical and continuous parameters

@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec in faucibus
-lacus. Curabitur in viverra neque. Sed nec urna a sem blandit viverra vel a
-eros. Nam a ultrices leo, a aliquet diam. Proin cursus scelerisque leo, ut
-ullamcorper ante auctor et. Nulla odio nisl, molestie eget consectetur quis,
-eleifend id ligula. Cras lobortis egestas justo ac sagittis. In at ligula est.
-Nulla vel eros sapien. Sed bibendum metus quis dignissim scelerisque. Nulla arcu
-lectus, ornare nec vehicula quis, lobortis a mi. Cras scelerisque mauris quam,
-nec viverra ex convallis sit amet. Ut tellus turpis, lobortis sit amet finibus
-sit amet, maximus ac est.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit...
 
@@
 predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. In
-particular, we pit $f$ against an adversary classifier $r(\cdot ;
-\theta_r) : \mathbb{R} \mapsto {\cal Z}$ of parameters $\theta_r$ and
-associated loss ${\cal L}_r(\theta_f, \theta_r)$.
-Assuming that ${\cal Z}$ defines a finite family of nuisance values $z_l$ (for $l=1, \dots, |{\cal Z}|$),
-this classifier takes as input realizations of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$ parameters,
-and produces as output probability estimates $r(f(X; \theta_f); \theta_r)_l = \hat{p}(z_l| f(X; \theta_f)) $
-that $f(X; \theta_f)$ is generated from the nuisance value $z_l$. If $p(f(X; \theta_f)|z)$ varies with $z$,
-then the corresponding correlation can be captured by $r$. By contrast, if
-$p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then $r$
-should perform poorly and be close to random guessing. Training $f$ such that it additionally minimizes the performance of $r$
-therefore acts as a regularization towards Eqn.~\ref{eqn:criterion}.
+particular, we pit $f$ against an adversary model $r := p_{\theta_r}(z |
+f(X;\theta_f))$ of parameters $\theta_r$ and associated loss ${\cal
+L}_r(\theta_f, \theta_r)$. This model takes  as input realizations of $f(X;
+\theta_f)$, for the current value $\theta_f$ of $f$ parameters, and produces as
+output probability estimates $p_{\theta_r}(z | f(X;\theta_f))$ that $f(X;
+\theta_f)$ is generated from the nuisance value $z$. Intuitively, if $p(f(X; \theta_f)|z)$
+varies with $z$, then the corresponding correlation can be captured by $r$. By
+contrast, if $p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then
+$r$ should perform poorly and be close to random guessing. Training $f$ such
+that it additionally minimizes the performance of $r$ therefore acts as a
+regularization towards Eqn.~\ref{eqn:criterion}.
 
@@
 \end{align}
-The adversarial training procedure to obtain $(\smash{\hat\theta_f},
-\smash{\hat\theta_r})$ is formally presented in
-Algorithm~\ref{alg:adversarial-training} in the case of $f$ being a classifier
-and of the cross-entropy loss for both ${\cal L}_f$ and ${\cal L}_r$. The algorithm
-consists in using stochastic gradient descent alternatively to optimize
-Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.
+Without loss of generality, the adversarial training procedure to obtain
+$(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ is formally presented in
+Algorithm~\ref{alg:adversarial-training} in the case of a binary classifier $f :
+\mathbb{R}^p \mapsto [0,1]$ modeling $p(Y=1|X)$. For reasons further explained
+in Section~\ref{sec:theory}, ${\cal L}_f$ (resp. ${\cal L}_r$) is set to the
+expected value of the
+negative log-likelihood of $Y|X$ under $f$ (resp. of $Z|f(X;\theta_f)$ under
+$r$). The optimization algorithm consists in using stochastic gradient descent
+alternatively for solving Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.
 
@@
             \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
-            $$\nabla_{\theta_r} \sum_{m=1}^M \left[\sum_{z_l \in {\cal Z}} 1(z_m = z_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+            $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|f(x_m;\theta_f))  ;$$}
         \EndFor
@@
         \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
-        $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\sum_{y_c \in {\cal Y}} 1(y_m = y_c) \log f(x_m;\theta_f)_c  +\sum_{z_l \in {\cal Z}} 1(z_m = z_l)\log r(f(x_m;\theta_f);\theta_r)_l \right];$$}
+        $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|f(x_m;\theta_f))  \right],$$
+        \indent where $p_{\theta_f}(y_m|x_m)$ denotes $1(y_m=0)(1-f(x_m;\theta_f)) + 1(y_m=1)f(x_m;\theta_f))$;}
     \EndFor
@@
 
-\glnote{Extension to continuous nuisance parameters: have the adversary $r$
-models the conditional distribution $p(z|f(X;\theta_f))$, e.g. with a network
-whose 2 outputs are the parameters of a truncated gaussian and optimized such that
-$p(z|f(X;\theta_f)) \approx N(\mu(f(X;\theta_f);\theta_r), \sigma(f(X;\theta_f);
-\theta_r)$. The associated loss should be the negative log-likelihood (= cross-entropy), which is maximal when the
-approximation is perfect and minimal when the truncated gaussian degenerates to
-a uniform (when $\sigma$ is large), in which case we should have the guarantee
-that $f$ is pivotal.}
-
-\glnote{This section and theoretical results should be rewritten in such a way
-that categorical and continuous nuisance values are solved under the same
-principled approach.}
-
 % ==============================================================================
@@
 \section{Theoretical results}
+\label{sec:theory}
 
 In this section, we show that in the setting of
-Algorithm~\ref{alg:adversarial-training}, the procedure converges to a
-classifier $f$ which is a pivotal quantity in the sense of
-Eqn.~\ref{eqn:criterion}. Results below are derived in a non-parametric setting,
-by assuming that both $f$ and $r$ have enough capacity. To simplify the
-presentation, we also assume the uniform prior $p(z) =
-\frac{1}{|{\cal Z}|}$ for all $z \in {\cal Z}$, e.g. by having the same
-number of training samples for each modality $z$ of the nuisance
-parameter.
+Algorithm~\ref{alg:adversarial-training} where ${\cal L}_r$ is set to expected
+value of the negative log-likelihood of $Z|f(X;\theta_f)$ under $r$, the
+procedure converges to a classifier $f$ which is a pivotal quantity in the sense
+of Eqn.~\ref{eqn:criterion}. Propositions below are derived in a non-parametric
+setting, by assuming that both $f$ and $r$ have enough capacity. Results hold
+for a nuisance parameter $Z$ taking either categorical values or continuous values within a
+bounded support.  By abuse of notation, $H(p_Z)$ denotes the differential
+entropy in this latter case. Finally, by construction, we assume the uniform
+prior $p(z)$ over the nuisance values $z \in {\cal Z}$ of the parameter.
 
@@
 Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
-\theta_r)$. If $r(f(X;\theta_f) ; \hat\theta_r )_l = \frac{1}{|{\cal Z}|}$
-for all $z_l$, then $f$ is a pivotal quantity.
+\theta_r)$. If $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
+for all $z \in {\cal Z}$, then $f$ is a pivotal quantity.
 \end{proposition}
@@
 \begin{proof}
-Let us first recall that the cross-entropy for distributions $p$ and $q$ is
-minimized when $p=q$. For ${\cal L}_r$ defined as the cross-entropy between the
-true conditional distribution of the nuisance $p_{Z|f(X;\theta_f)}$ and
-the approximate conditional distribution of the nuisance
-$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\smash{\hat\theta}_r =
+The optimal parameters $\smash{\hat\theta}_r =
 \arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
-\theta_r)$ are therefore such that $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$.
-
-In other words, for all $z_l \in {\cal Z}$, we have $r(f(X;\theta_f);\smash{\hat\theta}_r)_l = p(z_l|f(X;\theta_f))$.
-By assumption, $r(f(X;\theta_f);\smash{\hat\theta}_r)_l = \frac{1}{|{\cal Z}|}$,
-and therefore $p(z_l|f(X;\theta_f)) = \frac{1}{|{\cal Z}|}$.
+\theta_r)$ are such that $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
+By assumption, $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$,
+and therefore $p(z|f(X;\theta_f)) = p(z)$.
 Using the Bayes' rule, we write
 \begin{align*}
-    p(f(X;\theta_f)|z_l) &= \frac{ p(z_l|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z_l)} \\
-                         &= \frac{ \tfrac{1}{|{\cal Z}|} p(f(X;\theta_f)) } { \tfrac{1}{|{\cal Z}|} } \\
-                         &= p(f(X;\theta_f)),
+    p(f(X;\theta_f)|z) &= \frac{ p(z|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z)} = p(f(X;\theta_f)),
 \end{align*}
-which holds for all $z_l \in {\cal Z}$ and implies that $f$ is a pivotal quantity.
+which holds for all $z \in {\cal Z}$ and implies that $f$ is a pivotal quantity.
 \end{proof}
@@
 for Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar} such that
-$E(\hat\theta_f, \hat\theta_r) = H(p_{Y|X}) - \log |{\cal Z}|$, then
+$E(\hat\theta_f, \hat\theta_r) = H(p_{Y|X}) - H(p_Z)$, then
 $f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier  and a pivotal quantity.
@@
 \max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
-\theta_r)$, in which case $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$ and $L_r$ reduces to the entropy
+\theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$ and $L_r$ reduces to the entropy
 $H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
@@
 H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
-\log |{\cal Z}| \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
+H(p_Z) \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
 holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
@@
     \item $\smash{\hat\theta}_f$ corresponds to
-    the parameters of an optimal classifier, in which case the log-loss $L_f$
+    the parameters of an optimal classifier, in which case the expected negative log-likelihood $L_f$ of $Y|X$
     reduces to its minimum value $H(p_{Y|X})$,
@@
    outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
-   $p(z_l|f(X;\smash{\hat\theta}_f)) = \frac{1}{|{\cal Z}|}$ for all $z_l \in {\cal
-   Z}$ and $H(p_{Z|f(X;\smash{\hat\theta}_f)}) =
-   -\sum_{z_l \in {\cal Z}} p(z_l|f(X;\smash{\hat\theta}_f)) \log  p(z_l|f(X;\smash{\hat\theta}_f)) =  -\sum_{z_l \in {\cal Z}} \frac{1}{|{\cal Z}|} \log \frac{1}{|{\cal Z}|} = \log |{\cal Z}|$.
+   $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal
+   Z}$ since we assume a uniform prior by construction. Note that in the continuous case, this bound is also realized
+   by the uniform distribution when the support of $Z$ is bounded.
 \end{itemize}
-Accordingly, the second condition implies that $r(f(X;\smash{\hat\theta}_f) ;
-\smash{\hat\theta}_r )_l = \frac{1}{|{\cal Z}|}$ and therefore that at this
+Accordingly, the second condition implies that $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$ and therefore that at this
 point, because of Proposition~\ref{prop:1}, the optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.

### 2016-09-02 tex/paper.tex
save

@@
 In this section, we show that in the setting of
-Algorithm~\ref{alg:adversarial-training} where ${\cal L}_r$ is set to expected
-value of the negative log-likelihood of $Z|f(X;\theta_f)$ under $r$, the
-procedure converges to a classifier $f$ which is a pivotal quantity in the sense
-of Eqn.~\ref{eqn:criterion}. Propositions below are derived in a non-parametric
-setting, by assuming that both $f$ and $r$ have enough capacity. Results hold
-for a nuisance parameter $Z$ taking either categorical values or continuous values within a
-bounded support.  By abuse of notation, $H(p_Z)$ denotes the differential
-entropy in this latter case. Finally, by construction, we assume the uniform
-prior $p(z)$ over the nuisance values $z \in {\cal Z}$ of the parameter.
+Algorithm~\ref{alg:adversarial-training} where ${\cal L}_f$ and ${\cal L}_r$ are
+respectively set to expected value of the negative log-likelihood of $Y|X$ under
+$f$ and of $Z|f(X;\theta_f)$ under $r$, the procedure converges to a classifier
+$f$ which is a pivotal quantity in the sense of Eqn.~\ref{eqn:criterion}.
+Propositions below are derived in a non-parametric setting, by assuming that
+both $f$ and $r$ have enough capacity. Results hold for a nuisance parameter $Z$
+taking either categorical values or continuous values within a bounded support.
+By abuse of notation, $H(p_Z)$ denotes the differential entropy in this latter
+case. Finally, by construction, we assume the uniform prior $p(z)$ over the
+values $z \in {\cal Z}$ of the nuisance parameter.
 
@@
    $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal
-   Z}$ since we assume a uniform prior by construction. Note that in the continuous case, this bound is also realized
-   by the uniform distribution when the support of $Z$ is bounded.
+   Z}$ since we assume a uniform prior by construction. Note that in the continuous case, the supremum of the differential entropy
+   over continuous distributions of bounded support is also realized by the uniform distribution over that support.
 \end{itemize}
@@
 \section{Experiments}
+
+\glnote{Define the architecture of $r$ when $Z$ is categorical ($r$ is a
+standard classifier) or continuous ($r$ is e.g a 2-output NN such that the
+distribution of $Z|f(X)$ is modeled by a truncated gaussian of known support).}
 

### 2016-09-02 tex/paper.tex
save

@@
    Z}$ since we assume a uniform prior by construction. Note that in the continuous case, the supremum of the differential entropy
-   over continuous distributions of bounded support is also realized by the uniform distribution over that support.
+   over continuous distributions of the same bounded support is also realized by the uniform distribution over that support.
 \end{itemize}

### 2016-09-06 tex/paper.tex
save

--- 2016-09-02 tex/paper.tex
@@
 (e.g., a neural network-based classifier if ${\cal Y}$ is a finite set of
-classes) and minimizing  a loss ${\cal L}_f(\theta_f)$. In addition, we require
+classes) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g., the cross-entropy). In addition, we require
 that $f(X_z ; \theta_f)$ should be robust to the value $z$ of the nuisance parameter  --
@@
 Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
-Eqn.~\ref{eqn:criterion-measure}. The converse is however in general not true, since the
+Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true, since the
 sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = y \}$ and $\{
@@
 By abuse of notation, $H(p_Z)$ denotes the differential entropy in this latter
-case. Finally, by construction, we assume the uniform prior $p(z)$ over the
+case. Finally, by construction, we require the uniform prior $p(z)$ over the
 values $z \in {\cal Z}$ of the nuisance parameter.
@@
    $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal
-   Z}$ since we assume a uniform prior by construction. Note that in the continuous case, the supremum of the differential entropy
-   over continuous distributions of bounded support is also realized by the uniform distribution over that support.
+   Z}$ since we require a uniform prior by construction. Note that in the continuous case, the supremum of the differential entropy
+   over continuous distributions on the same bounded support is also realized by the uniform distribution over that support.
 \end{itemize}

### 2016-09-06 tex/paper.tex
typo

@@
 The optimal parameters $\smash{\hat\theta}_r =
-\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
+\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
 \theta_r)$ are such that $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
@@
 For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
-\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
-\theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$ and $L_r$ reduces to the entropy
+\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
+\theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the entropy
 $H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
-value function $E$ can therefore be rewritten as $$E'(\theta_f) = L_f(\theta_f) -
+value function $E$ can therefore be rewritten as $$E'(\theta_f) = {\cal L}_f(\theta_f) -
 H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
-H(p_Z) \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
+H(p_Z) \leq {\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
 holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
@@
     \item $\smash{\hat\theta}_f$ corresponds to
-    the parameters of an optimal classifier, in which case the expected negative log-likelihood $L_f$ of $Y|X$
+    the parameters of an optimal classifier, in which case the expected negative log-likelihood ${\cal L}_f$ of $Y|X$
     reduces to its minimum value $H(p_{Y|X})$,
@@
 We should discuss in which circumstances. In such case, the pivotal quantity
-constraint can however be enforced by outweighting the $L_r$ term, resulting in a
+constraint can however be enforced by outweighting the ${\cal L}_r$ term, resulting in a
 trade-off between classifier optimality and pivotality.}

### 2016-09-06 tex/paper.tex
save

@@
 
-\begin{proposition}
-    \glnote{It remains to prove that the procedure of Algorithm 1
-    converges towards that saddle point. The proof should be similar to the proof of convergence in the GAN paper.}
-\end{proposition}
-
-\glnote{We should further discuss that in practice, the equality in Prop 2 may never hold.
-We should discuss in which circumstances. In such case, the pivotal quantity
-constraint can however be enforced by outweighting the ${\cal L}_r$ term, resulting in a
-trade-off between classifier optimality and pivotality.}
+Proposition~\ref{prop:2} suggests that if at each step of
+Algorithm~\ref{alg:adversarial-training} the adversary $r$ is allowed to reach
+its optimum given $f$ (e.g., by setting $K$ sufficiently high) and if $f$ is
+updated to improve ${\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$, then $f$
+should converge to a classifier which is both optimal and pivotal, provided such
+a classifier exists. On many problems of interest though, such a classifier may
+not exist, in which case the lower bound $H(p_{Y|X}) - H(p_Z) < {\cal
+L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal
+classifier or a pivotal quantity, but not both simultaneously. Accordingly the value
+function $E$ can be rewritten as
+\begin{equation}
+    E_\lambda(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r),
+\end{equation}
+where $\lambda \geq 0$ is a hyper-parameter controlling the trade-off between the performance of $f$
+and its independence with respect to the nuisance parameter. Setting $\lambda$ to a large
+value will preferrably enforces $f$ to be pivotal while setting $\lambda$ to a value close to $0$
+will rather constraint $f$ to be optimal.
 

### 2016-09-06 tex/paper.tex
save

@@
 a classifier exists. On many problems of interest though, such a classifier may
-not exist, in which case the lower bound $H(p_{Y|X}) - H(p_Z) < {\cal
+not exist because the nuisance parameter directly affects the optimal decision
+boundary, in which cases the lower bound $H(p_{Y|X}) - H(p_Z) < {\cal
 L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal
-classifier or a pivotal quantity, but not both simultaneously. Accordingly the value
-function $E$ can be rewritten as
+classifier or a pivotal quantity, but not both simultaneously. Accordingly the
+value function $E$ can be rewritten as
 \begin{equation}
@@
 and its independence with respect to the nuisance parameter. Setting $\lambda$ to a large
-value will preferrably enforces $f$ to be pivotal while setting $\lambda$ to a value close to $0$
+value will preferrably enforces $f$ to be pivotal while setting $\lambda$ close to $0$
 will rather constraint $f$ to be optimal.
+
+\glnote{We should make it clearer what {\it classification optimality} means: it
+is with respect to $Y|X=x$  when  $x$ results from $z\sim p_Z, x \sim X_z$.
+This is often strictly different from optimality with respect to $Y|X=x,Z=z$ for the true value $z$.}
 

### 2016-09-06 tex/paper.tex
save

@@
 $f$ which is a pivotal quantity in the sense of Eqn.~\ref{eqn:criterion}.
-Propositions below are derived in a non-parametric setting, by assuming that
-both $f$ and $r$ have enough capacity. Results hold for a nuisance parameter $Z$
-taking either categorical values or continuous values within a bounded support.
-By abuse of notation, $H(p_Z)$ denotes the differential entropy in this latter
-case. Finally, by construction, we require the uniform prior $p(z)$ over the
-values $z \in {\cal Z}$ of the nuisance parameter.
+
+In this setting, the nuisance parameter $Z$ is considered as a random variable,
+for which we require the uniform prior $p(z)$ (for $z$ in ${\cal Z}$).  Results
+hold for a nuisance parameter $Z$ taking either categorical values or continuous
+values within a bounded support. By abuse of notation, $H(p_Z)$ denotes the
+differential entropy in this latter case. Finally, propositions below are derived in a
+non-parametric setting, by assuming that both $f$ and $r$ have enough capacity.
 
@@
 
-\glnote{We should make it clearer what {\it classification optimality} means: it
-is with respect to $Y|X=x$  when  $x$ results from $z\sim p_Z, x \sim X_z$.
-This is often strictly different from optimality with respect to $Y|X=x,Z=z$ for the true value $z$.}
-
 % ==============================================================================

### 2016-09-06 tex/paper.tex
save

@@
 In this setting, the nuisance parameter $Z$ is considered as a random variable,
-for which we require the uniform prior $p(z)$ (for $z$ in ${\cal Z}$).  Results
+for which we require the uniform prior $p(z)$ (for $z$ in ${\cal Z}$).
+Importantly, classification of $Y$ with respect to $X$ is therefore considered
+in the context where $Z$ is marginalized out, which means that the classifier
+minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not necessarily
+with $Y|X,Z$. Results
 hold for a nuisance parameter $Z$ taking either categorical values or continuous
@@
 a classifier exists. On many problems of interest though, such a classifier may
-not exist because the nuisance parameter directly affects the optimal decision
+not exist because the nuisance parameter directly shapes the decision
 boundary, in which cases the lower bound $H(p_{Y|X}) - H(p_Z) < {\cal
@@
 
+\glnote{Alternative to learning a parameterized classifier for LLR.}
+
 \glnote{Similar to domain adaptation, but with infinitely many domains,
@@
 
-\subsubsection*{Acknowledgments}
+\section*{Acknowledgments}
 

### 2016-09-07 tex/paper.tex
added toy example without continuous case

@@
 \usepackage[final]{nips_2016}
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
+\usepackage[utf8]{inputenc}
+\usepackage{hyperref}
+\usepackage{booktabs}
+\usepackage{amsfonts}
 \usepackage{amssymb}
@@
 
-In this setting, the nuisance parameter $Z$ is considered as a random variable,
-for which we require the uniform prior $p(z)$ (for $z$ in ${\cal Z}$).
+In this setting, the nuisance parameter $Z$ is considered as a random variable of bounded support,
+for which we require the uniform prior $p(z)$ (for $z \in {\cal Z}$).
 Importantly, classification of $Y$ with respect to $X$ is therefore considered
@@
 minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not necessarily
-with $Y|X,Z$. Results
-hold for a nuisance parameter $Z$ taking either categorical values or continuous
-values within a bounded support. By abuse of notation, $H(p_Z)$ denotes the
-differential entropy in this latter case. Finally, propositions below are derived in a
-non-parametric setting, by assuming that both $f$ and $r$ have enough capacity.
+with $Y|X,Z$. Results hold for a nuisance parameter $Z$ taking either
+categorical values or continuous values within a bounded support. By abuse of
+notation, $H(p_Z)$ denotes the differential entropy in this latter case.
+Finally, propositions below are derived in a non-parametric setting, by assuming
+that both $f$ and $r$ have enough capacity.
 
@@
 L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal
-classifier or a pivotal quantity, but not both simultaneously. Accordingly the
-value function $E$ can be rewritten as
+classifier or a pivotal quantity, but not both simultaneously. In this situation, it is natural to rewrite the
+value function $E$  as
 \begin{equation}

### 2016-09-08 tex/paper.tex
switch to nicer APS template

@@
-\documentclass{article}
+\documentclass[twocolumn,preprintnumbers,superscriptaddress,nofootinbib,aps,floatfix,letterpaper]{revtex4-1}
 
-\PassOptionsToPackage{numbers, compress}{natbib}
+\usepackage[utf8]{inputenc}
 
-\usepackage[final]{nips_2016}
-\usepackage[utf8]{inputenc}
-\usepackage{hyperref}
-\usepackage{booktabs}
 \usepackage{amsfonts}
@@
 \usepackage{amsthm}
-\usepackage{caption}
-\usepackage{subcaption}
+
+\usepackage{bm}
+\usepackage{cancel}
+\usepackage{bbold}
+\usepackage{slashed}
 \usepackage{graphicx}
 \usepackage{color}
+\usepackage{hyperref}
 \usepackage{algorithm}
@@
 \newcommand{\glnote}[1]{\textcolor{red}{[GL: #1]}}
-
-\title{Adversarial Training against Systematic Uncertainty}
-
-\author{
-  Gilles Louppe \\
-  New York University\\
-  \texttt{g.louppe@nyu.edu} \\
-  \And
-  Michael Kagan\\
-  SLAC National Accelerator Laboratory\\
-  \texttt{makagan@slac.stanford.edu}\\
-  \And
-  Kyle Cranmer \\
-  New York University\\
-  \texttt{kyle.cranmer@nyu.edu}
-}
 
@@
 
-\maketitle
+% ==============================================================================
 
-% ==============================================================================
+\title{\Large{Adversarial Training against Systematic Uncertainty}}
+\vspace{1cm}
+\author{\small{\bf Gilles Louppe}\thanks{\texttt{g.louppe@nyu.edu}}}
+\affiliation{New York University}
+\author{\small{\bf Michael Kagan}\thanks{\texttt{makagan@slac.stanford.edu}}}
+\affiliation{SLAC National Accelerator Laboratory}
+\author{\small{\bf Kyle Cranmer}\thanks{\texttt{kyle.cranmer@nyu.edu}}}
+\affiliation{New York University}
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit...
+Lorem ipsum dolor sit amet, consectetur adipiscing elit. Etiam commodo, enim
+vitae facilisis pretium, ligula justo aliquet lectus, non interdum erat lorem et
+nisi. Donec pharetra lectus in magna pellentesque vehicula. Praesent dapibus
+lorem sed enim lacinia, a vestibulum sapien mattis. Sed ornare mollis aliquet.
+Nulla tempor lacinia tortor, in rhoncus augue porta nec. Morbi sed convallis
+nibh, eu hendrerit turpis. Curabitur sit amet rhoncus purus. Curabitur eget
+magna lorem. Phasellus pretium nisi quis est tincidunt, faucibus vulputate augue
+viverra. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin a urna a
+ex egestas pulvinar. Curabitur facilisis a dui nec fringilla. Duis bibendum
+fermentum quam. Donec eget urna eget mi placerat malesuada ut et lacus. Nulla ac
+risus scelerisque, congue eros vitae, dignissim risus. Fusce purus ante,
+vulputate sed magna et, varius imperdiet leo.
 
 \end{abstract}
+
+\maketitle
 
@@
 Let consider the multivariate random variables $X_z: \Omega \mapsto
-\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence on a
-nuisance parameter $Z$ whose values $z \in {\cal Z}$  define a
-parameterized family of its systematic uncertainties. That is, $X_z$ and
-$Y$ induce together a joint probability distribution $p(X,Y|z)$, where the
-conditional on $z$ denotes $X_z$. For training, let further assume a
-finite set $\{ x_i, y_i, z_i \}_{i=1}^N$ of realizations
-$X_{z_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
-values $z_i$ of the nuisance parameter. Our goal is to learn a function
-$f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal Y}$ of parameters $\theta_f$
-(e.g., a neural network-based classifier if ${\cal Y}$ is a finite set of
-classes) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g., the cross-entropy). In addition, we require
-that $f(X_z ; \theta_f)$ should be robust to the value $z$ of the nuisance parameter  --
-which remains unknown at test time. More specifically, we aim at building $f$
-such that in the ideal case
+\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
+on a nuisance parameter $Z$ whose values $z \in {\cal Z}$  define a
+parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
+induce together a joint probability distribution $p(X,Y|z)$, where the
+conditional on $z$ denotes $X_z$. For training, let further assume a finite set $\{
+x_i, y_i, z_i \}_{i=1}^N$ of realizations $X_{z_i}(\omega_i), Y(\omega_i)$, for
+$\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
+is to learn a function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal Y}$ of
+parameters $\theta_f$ (e.g., a neural network-based classifier if ${\cal Y}$ is
+a finite set of classes) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
+the cross-entropy). In addition, we require that $f(X_z ; \theta_f)$ should be
+robust to the value $z$ of the nuisance parameter  -- which remains unknown at
+test time. More specifically, we aim at building $f$ such that in the ideal case
 \begin{equation}\label{eqn:criterion-true}
@@
 solve the closely related problem of finding a predictive function $f$ such that
-\begin{equation}\label{eqn:criterion-measure}
-    P(\{ \omega | f(X_{z}(\omega) ; \theta_f) = y \} ) = P( \{ \omega' | f(X_{z^\prime}(\omega') ; \theta_f) = y\}) \text{ for all $y \in {\cal Y}$}.
+\begin{align}\label{eqn:criterion-measure}
+    &P(\{ \omega | f(X_{z}(\omega) ; \theta_f) = y \} ) \nonumber \\
+    &\indent = P( \{ \omega' | f(X_{z^\prime}(\omega') ; \theta_f) = y\})
+\end{align}
+for all $y \in {\cal Y}$. In words, we are looking for a predictive function $f$
+which is a pivotal quantity \citep{degroot1986probability} with respect to the
+nuisance parameter. That is, such that  the distribution of $f(X_z; \theta_f)$
+is invariant with respect to the value $z$ of the nuisance. Note that a function
+$f$ for which Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
+Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true,
+since the sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = y \}$ and $\{
+\omega' | f(X_{z^\prime}(\omega'); \theta_f) = y \}$ do not need to be the same
+for the equality to hold. In order to simplify notations, and as only
+Eqn.~\ref{eqn:criterion-measure} is of direct interest in this work, we denote
+from here on the pivotal quantity criterion as
+\begin{equation}\label{eqn:criterion}
+    p(f(X ; \theta_f) | z ) = p(f(X ; \theta_f) | z^\prime )
 \end{equation}
-In words, we are looking for a predictive function $f$ which is a pivotal
-quantity \citep{degroot1986probability} with respect to the nuisance parameter.
-That is, such that  the distribution of $f(X_z; \theta_f)$ is invariant
-with respect to the value $z$ of the nuisance. Note that a function $f$ for which
-Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
-Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true, since the
-sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = y \}$ and $\{
-\omega' | f(X_{z^\prime}(\omega'); \theta_f) = y \}$ do not need to be the same
-for the equality to hold.
-In order to simplify notations,
-and as only Eqn.~\ref{eqn:criterion-measure} is
-of direct interest in this work, we denote from here on
-the pivotal quantity criterion as
-\begin{equation}\label{eqn:criterion}
-    p(f(X ; \theta_f) | z ) = p(f(X ; \theta_f) | z^\prime ) \text{ for all $z,z^\prime \in  {\cal Z}$.}
-\end{equation}
+for all $z,z^\prime \in  {\cal Z}$.
 
@@
 
-Adversarial training was first proposed by \cite{goodfellow2014generative} as a
+Adversarial training was first proposed by \citep{goodfellow2014generative} as a
 way to build a generative model capable of producing samples from random noise
-$z \sim p_Z$. More specifically, the authors pit a generative model $g: \mathbb{R}
-\mapsto \mathbb{R}^p$ against an adversary classifier $d : \mathbb{R}^p \mapsto \{
-0, 1\}$ whose antagonistic objective is to recognize real data $X$ from generated data $g(Z)$. Both
-models $g$ and $d$ are trained simultaneously, in such a way that $g$ learns to
-produce samples that are difficult to identify by $d$, while $d$ incrementally
-adapts to changes in $g$. At the equilibrium, $g$ models a distribution whose
-samples can be identified by $d$ only by chance. That is, assuming enough
-capacity in $d$ and  $g$, the distribution $p_{g(Z)}$ eventually converges
-towards the real distribution $p_X$.
+$z \sim p_Z$. More specifically, the authors pit a generative model $g:
+\mathbb{R} \mapsto \mathbb{R}^p$ against an adversary classifier $d :
+\mathbb{R}^p \mapsto \{ 0, 1\}$ whose antagonistic objective is to recognize
+real data $X$ from generated data $g(Z)$. Both models $g$ and $d$ are trained
+simultaneously, in such a way that $g$ learns to produce samples that are
+difficult to identify by $d$, while $d$ incrementally adapts to changes in $g$.
+At the equilibrium, $g$ models a distribution whose samples can be identified by
+$d$ only by chance. That is, assuming enough capacity in $d$ and  $g$, the
+distribution $p_{g(Z)}$ eventually converges towards the real distribution
+$p_X$.
 
@@
 output probability estimates $p_{\theta_r}(z | f(X;\theta_f))$ that $f(X;
-\theta_f)$ is generated from the nuisance value $z$. Intuitively, if $p(f(X; \theta_f)|z)$
-varies with $z$, then the corresponding correlation can be captured by $r$. By
-contrast, if $p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then
-$r$ should perform poorly and be close to random guessing. Training $f$ such
-that it additionally minimizes the performance of $r$ therefore acts as a
-regularization towards Eqn.~\ref{eqn:criterion}.
+\theta_f)$ is generated from the nuisance value $z$. Intuitively, if $p(f(X;
+\theta_f)|z)$ varies with $z$, then the corresponding correlation can be
+captured by $r$. By contrast, if $p(f(X; \theta_f)|z)$ is invariant with $z$, as
+we require, then $r$ should perform poorly and be close to random guessing.
+Training $f$ such that it additionally minimizes the performance of $r$
+therefore acts as a regularization towards Eqn.~\ref{eqn:criterion}.
 
@@
 
-\begin{algorithm}[t]
-\caption{Adversarial training of a classifier $f$ against an adversary $r$.\\
-{\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$\\
-{\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$\\
-{\it Hyper-parameters:} Number $T$ of training iterations, Number $K$ of gradient steps to update $r$.}
+\begin{figure*}
+    \begin{minipage}{\linewidth}
+    \begin{algorithm}[H]
+    \caption{Adversarial training of a classifier $f$ against an adversary $r$.}
+
+    \begin{flushleft}
+        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;\\
+        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$;\\
+        {\it Hyper-parameters:} Number $T$ of training iterations,
+                                Number $K$ of gradient steps to update $r$.
+    \end{flushleft}
+
     \label{alg:adversarial-training}
@@
     \end{algorithm}
+    \end{minipage}
+\end{figure*}
 
@@
 
-In this setting, the nuisance parameter $Z$ is considered as a random variable of bounded support,
-for which we require the uniform prior $p(z)$ (for $z \in {\cal Z}$).
-Importantly, classification of $Y$ with respect to $X$ is therefore considered
-in the context where $Z$ is marginalized out, which means that the classifier
-minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not necessarily
-with $Y|X,Z$. Results hold for a nuisance parameter $Z$ taking either
-categorical values or continuous values within a bounded support. By abuse of
-notation, $H(p_Z)$ denotes the differential entropy in this latter case.
-Finally, propositions below are derived in a non-parametric setting, by assuming
-that both $f$ and $r$ have enough capacity.
+In this setting, the nuisance parameter $Z$ is considered as a random variable
+of bounded support, for which we require the uniform prior $p(z)$ (for $z \in
+{\cal Z}$). Importantly, classification of $Y$ with respect to $X$ is therefore
+considered in the context where $Z$ is marginalized out, which means that the
+classifier minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not
+necessarily with $Y|X,Z$. Results hold for a nuisance parameter $Z$ taking
+either categorical values or continuous values within a bounded support. By
+abuse of notation, $H(p_Z)$ denotes the differential entropy in this latter
+case. Finally, propositions below are derived in a non-parametric setting, by
+assuming that both $f$ and $r$ have enough capacity.
 
 \begin{proposition}\label{prop:1}
-Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
-\theta_r)$. If $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
+Let $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
+\theta_r)$ for fixed $\theta_f$. If $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
 for all $z \in {\cal Z}$, then $f$ is a pivotal quantity.
@@
 \begin{proof}
-The optimal parameters $\smash{\hat\theta}_r =
+The optimal parameters $$\smash{\hat\theta}_r =
 \arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
-\theta_r)$ are such that $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
+\theta_r)$$ are such that $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
 By assumption, $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$,
@@
 \begin{align*}
-    p(f(X;\theta_f)|z) &= \frac{ p(z|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z)} = p(f(X;\theta_f)),
+    p(f(X;\theta_f)|z) &= \frac{ p(z|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z)}\\
+                       &= p(f(X;\theta_f)),
 \end{align*}
@@
 $E(\hat\theta_f, \hat\theta_r) = H(p_{Y|X}) - H(p_Z)$, then
-$f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier  and a pivotal quantity.
+$f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier and a pivotal
+quantity.
 \end{proposition}
@@
 \begin{proof}
+
 For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
-\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
-\theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the entropy
+\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal
+L}_r(\theta_f, \theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) =
+p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the entropy
 $H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
-value function $E$ can therefore be rewritten as $$E'(\theta_f) = {\cal L}_f(\theta_f) -
-H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
-H(p_Z) \leq {\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
-holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
+value function $E$ can therefore be rewritten as $$E'(\theta_f) = {\cal
+L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower
+bound $$H(p_{Y|X}) - H(p_Z) \leq {\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$$
+where the equality holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f}
+E'(\theta_f)$ only when
 \begin{itemize}
     \item $\smash{\hat\theta}_f$ corresponds to
-    the parameters of an optimal classifier, in which case the expected negative log-likelihood ${\cal L}_f$ of $Y|X$
-    reduces to its minimum value $H(p_{Y|X})$,
+    the parameters of an optimal classifier, in which case the expected
+    negative log-likelihood ${\cal L}_f$ of $Y|X$ reduces to its minimum
+    value $H(p_{Y|X})$,
     \item all
@@
    $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal
-   Z}$ since we require a uniform prior by construction. Note that in the continuous case, the supremum of the differential entropy
-   over continuous distributions on the same bounded support is also realized by the uniform distribution over that support.
+   Z}$ since we require a uniform prior by construction. Note that in the
+   continuous case, the supremum of the differential entropy
+   over continuous distributions on the same bounded support is also realized
+   by the uniform distribution over that support.
 \end{itemize}
-Accordingly, the second condition implies that $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$ and therefore that at this
-point, because of Proposition~\ref{prop:1}, the optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
+Accordingly, the second condition implies that $p_{\theta_r}(z|f(X;\theta_f)) =
+p(z)$ and therefore that at this point, because of Proposition~\ref{prop:1}, the
+optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
+
 \end{proof}
@@
 a classifier exists. On many problems of interest though, such a classifier may
-not exist because the nuisance parameter directly shapes the decision
-boundary, in which cases the lower bound $H(p_{Y|X}) - H(p_Z) < {\cal
-L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal
-classifier or a pivotal quantity, but not both simultaneously. In this situation, it is natural to rewrite the
-value function $E$  as
+not exist because the nuisance parameter directly shapes the decision boundary,
+in which cases the lower bound $H(p_{Y|X}) - H(p_Z) < {\cal L}_f(\theta_f) -
+H(p_{Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal classifier or a
+pivotal quantity, but not both simultaneously. In this situation, it is natural
+to rewrite the value function $E$  as
 \begin{equation}
@@
 \end{equation}
-where $\lambda \geq 0$ is a hyper-parameter controlling the trade-off between the performance of $f$
-and its independence with respect to the nuisance parameter. Setting $\lambda$ to a large
-value will preferrably enforces $f$ to be pivotal while setting $\lambda$ close to $0$
-will rather constraint $f$ to be optimal.
+where $\lambda \geq 0$ is a hyper-parameter controlling the trade-off between
+the performance of $f$ and its independence with respect to the nuisance
+parameter. Setting $\lambda$ to a large value will preferrably enforces $f$ to
+be pivotal while setting $\lambda$ close to $0$ will rather constraint $f$ to be
+optimal.
 
@@
 \glnote{Alternative to learning a parameterized classifier for LLR.}
-
 \glnote{Similar to domain adaptation, but with infinitely many domains,
 as parameterized by $Z$, also related to transfer learning.}
-
 \glnote{Other applications: removing implicit bias in data (e.g. gender bias).}
@@
 
-\section*{Acknowledgments}
+\begin{acknowledgments}
+The authors would like to thank...
+\end{acknowledgments}
 
-\bibliographystyle{ieeetr}
+% ==============================================================================
+
 \bibliography{bibliography.bib}

### 2016-09-11 tex/paper.tex
save
related work

@@
-\documentclass[twocolumn,preprintnumbers,superscriptaddress,nofootinbib,aps,floatfix,letterpaper]{revtex4-1}
+\documentclass[twocolumn,superscriptaddress,aps]{revtex4-1}
 
@@
 \glnote{Distinction between statistical and systematic uncertainty.}
-\glnote{Define nuisance parameters.}
+\glnote{Outline taxonomy of uncertainties.}
+\glnote{Define nuisance parameters. See refs in Noel's papers.}
 \glnote{We want to build an accurate classifier whose output remains invariant with
@@
 \glnote{Motivate the criterion (which may not be obvious for the ML crowd). See pivotal quantity motivation.}
+\glnote{Discuss what kind of uncertainties the proposed approach is good for.}
 
@@
 
+\glnote{Add summary figure.}
+
 % ==============================================================================
@@
 
+\subsection{Toy example}
+
 \glnote{Define the architecture of $r$ when $Z$ is categorical ($r$ is a
 standard classifier) or continuous ($r$ is e.g a 2-output NN such that the
-distribution of $Z|f(X)$ is modeled by a truncated gaussian of known support).}
+distribution of $Z|f(X)$ is modeled by a truncated gaussian of known support). Cite Nix.}
+
+\subsection{Physics example}
 
@@
 
-\glnote{Alternative to learning a parameterized classifier for LLR.}
-\glnote{Similar to domain adaptation, but with infinitely many domains,
-as parameterized by $Z$, also related to transfer learning.}
-\glnote{Other applications: removing implicit bias in data (e.g. gender bias).}
+To account for systematic uncertainties, experimentalists in high energy physics
+typically take as fixed a classifier $f$ built from training data for a nominal
+value $z_0$ of the nuisance parameter, and then propagate uncertainty
+\glnote{add ref} by estimating $p(f(x)|z)$ with a parameterized calibration
+procedure. Clearly, this classifier is however not optimal for $z \neq z_0$. In
+this setting, parameterized
+classifiers~\citep{cranmer2015approximating,Baldi:2016fzo} directly take
+(nuisance) parameters as additional input variables, hence ultimately providing
+the most statistically powerful approach for incorporating the effect of
+systematics on the underlying classification task.  As argued in
+\citep{Neal:2007zz}, such classifiers can however not be used on real data since
+the correct value $z$ of the nuisance often remains unknown. This is typically
+not an issue in the context of parameter
+inference~\citep{cranmer2015approximating}, where nuisance parameters are
+marginalized out, but otherwise often limits the range of their applications. In
+practice, parameterized classifiers  are also often computationally expensive to
+build and evaluate. In particular, calibrating their decision fucntion, i.e.
+approximating $p(f(x,z)|z)$ as a continuous function of $z$, remains an open
+challenge. By contrast, constraining $f$ to be pivotal yields a classifier which
+may not be optimal, as discussed in Section~\ref{sec:theory}, but that can be
+used in a wider range of applications, since knowing the correct value $z$ of
+the nuisance is not necessary, and for which calibration needs to be carried out
+only once, since  the dependence on the nuisance is now built in the classifier.
+\glnote{Shall we also discuss the relation with \citep{Neal:2007zz} where point
+estimates of the nuisance are used as inputs to $f4$?}
+
+In machine learning, the proposed method can be related to the problem of domain
+adaptation
+\citep{blitzer2006domain,pan2011domain,gopalan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,ganin2014unsupervised},
+where the goal is often stated as trying to learn a domain-invariant
+representation of the data. Similarly, our method also relates to the problem of
+enforcing fairness in classification \citep{zemel2013learning,EdwardsS15}. For
+both families of methods, the problem can equivalently be stated as learning a
+classifier which is a pivotal quantity with respect to either the domain or some
+chosen attribute of the data, such as gender, color or age. In this context,
+\citep{EdwardsS15} is certainly among the closest to our work, in which fairness
+is enforced through an adversarial minimax setup composed of a classifier and an
+adversary discriminator. In particular, our work can be regarded as a
+generalization that also supports the continuous case, which can be viewed as
+handling infinitely many domains, provided they can be continuously
+parameterized, or as enforcing fairness over continuous attributes.
 
@@
 \begin{acknowledgments}
-The authors would like to thank...
 \end{acknowledgments}

### 2016-09-11 tex/paper.tex
save

@@
 challenge. By contrast, constraining $f$ to be pivotal yields a classifier which
-may not be optimal, as discussed in Section~\ref{sec:theory}, but that can be
+may not be optimal with respect to $Y|X,Z$, as discussed in Section~\ref{sec:theory}, but that can be
 used in a wider range of applications, since knowing the correct value $z$ of
@@
 \glnote{Shall we also discuss the relation with \citep{Neal:2007zz} where point
-estimates of the nuisance are used as inputs to $f4$?}
+estimates of the nuisance are used as inputs to $f$?}
 
-In machine learning, the proposed method can be related to the problem of domain
+In machine learning, learning a pivotal quantity can be related to the problem of domain
 adaptation
@@
 chosen attribute of the data, such as gender, color or age. In this context,
-\citep{EdwardsS15} is certainly among the closest to our work, in which fairness
-is enforced through an adversarial minimax setup composed of a classifier and an
-adversary discriminator. In particular, our work can be regarded as a
-generalization that also supports the continuous case, which can be viewed as
-handling infinitely many domains, provided they can be continuously
-parameterized, or as enforcing fairness over continuous attributes.
+\citep{ganin2014unsupervised,EdwardsS15} are certainly among the closest to our
+work, in which domain invariance and fairness are enforced through an adversarial
+minimax setup composed of a classifier and an adversary discriminator. In
+particular, our work can be regarded as a generalization that also supports the
+continuous case, which can be viewed as handling infinitely many domains,
+provided they can be continuously parameterized, or as enforcing fairness over
+continuous attributes.
+
+\glnote{Check related work of cited references to see if important related work are missing.}
 

### 2016-09-11 tex/paper.tex
save

@@
 \theta_f)$, for the current value $\theta_f$ of $f$ parameters, and produces as
-output probability estimates $p_{\theta_r}(z | f(X;\theta_f))$ that $f(X;
-\theta_f)$ is generated from the nuisance value $z$. Intuitively, if $p(f(X;
+output a function $p_{\theta_r}(z | f(X;\theta_f))$ modeling the conditional probability density that $f(X;
+\theta_f)$ is generated from the nuisance value $z$.
+\glnote{Give examples of $r$ for categorical and continuous nuisances.}
+Intuitively, if $p(f(X;
 \theta_f)|z)$ varies with $z$, then the corresponding correlation can be
@@
 the performance of $f$ and its independence with respect to the nuisance
-parameter. Setting $\lambda$ to a large value will preferrably enforces $f$ to
+parameter. Setting $\lambda$ to a large value will preferably enforces $f$ to
 be pivotal while setting $\lambda$ close to $0$ will rather constraint $f$ to be
@@
 standard classifier) or continuous ($r$ is e.g a 2-output NN such that the
-distribution of $Z|f(X)$ is modeled by a truncated gaussian of known support). Cite Nix.}
+distribution of $Z|f(X)$ is modeled by a truncated Gaussian of known support). Cite Nix.}
 
@@
 practice, parameterized classifiers  are also often computationally expensive to
-build and evaluate. In particular, calibrating their decision fucntion, i.e.
+build and evaluate. In particular, calibrating their decision function, i.e.
 approximating $p(f(x,z)|z)$ as a continuous function of $z$, remains an open

### 2016-09-12 tex/paper.tex
save

@@
 \theta_f)$, for the current value $\theta_f$ of $f$ parameters, and produces as
-output a function $p_{\theta_r}(z | f(X;\theta_f))$ modeling the conditional probability density that $f(X;
-\theta_f)$ is generated from the nuisance value $z$.
-\glnote{Give examples of $r$ for categorical and continuous nuisances.}
-Intuitively, if $p(f(X;
-\theta_f)|z)$ varies with $z$, then the corresponding correlation can be
-captured by $r$. By contrast, if $p(f(X; \theta_f)|z)$ is invariant with $z$, as
-we require, then $r$ should perform poorly and be close to random guessing.
-Training $f$ such that it additionally minimizes the performance of $r$
-therefore acts as a regularization towards Eqn.~\ref{eqn:criterion}.
+output a function $p_{\theta_r}(z | f(X;\theta_f))$ modeling the posterior
+probability density that $z$ parameterizes the sample $X$ observed through
+$f(\cdot;\theta_f)$. If $Z$ takes discrete values, then $p_{\theta_r}$ can be
+represented as a probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal
+Z|}}$ whose output $j$ (for $j=1, \dots, |{\cal Z}|$) is the estimated
+probability mass $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly, if $Z$ takes
+continuous values and if we assume some parameteric distribution for
+$Z|f(X;\theta_f)$ (e.g., a gaussian), then $p_{\theta_r}$ can be represented as
+network whose output $j$ is the estimated value of the corresponding parameter
+of that distribution (e.g., its mean and variance), and such that
+$p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$
+under that distribution.  Intuitively, if $p(f(X; \theta_f)|z)$ varies with $z$,
+then the corresponding correlation can be captured by $r$. By contrast, if
+$p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then $r$ should
+perform poorly and be close to random guessing. Training $f$ such that it
+additionally minimizes the performance of $r$ therefore acts as a regularization
+towards Eqn.~\ref{eqn:criterion}.
 
@@
 
+\glnote{Discuss that we could equivalently require an intermediate representation of the data (as learned internally by $f$) to be pivotal, instead of requiring $f(X)$ to be pivotal.}
 \glnote{Add summary figure.}
@@
 marginalized out, but otherwise often limits the range of their applications. In
-practice, parameterized classifiers  are also often computationally expensive to
-build and evaluate. In particular, calibrating their decision function, i.e.
+practice, parameterized classifiers  are also computationally expensive to build
+and evaluate. In particular, calibrating their decision function, i.e.
 approximating $p(f(x,z)|z)$ as a continuous function of $z$, remains an open
 challenge. By contrast, constraining $f$ to be pivotal yields a classifier which
-may not be optimal with respect to $Y|X,Z$, as discussed in Section~\ref{sec:theory}, but that can be
-used in a wider range of applications, since knowing the correct value $z$ of
-the nuisance is not necessary, and for which calibration needs to be carried out
-only once, since  the dependence on the nuisance is now built in the classifier.
-\glnote{Shall we also discuss the relation with \citep{Neal:2007zz} where point
-estimates of the nuisance are used as inputs to $f$?}
+may not be optimal with respect to $Y|X,Z$, as discussed in
+Section~\ref{sec:theory}, but that can otherwise be used in a wider range of
+applications, since knowing the correct value $z$ of the nuisance is not
+necessary. Similarly, calibration needs to be carried out only once, since  the
+dependence on the nuisance is now built-in. \glnote{Shall we also discuss the
+relation with \citep{Neal:2007zz} where point estimates of the nuisance are used
+as inputs to $f$?}
 
-In machine learning, learning a pivotal quantity can be related to the problem of domain
-adaptation
+In machine learning, learning a pivotal quantity can be related to the problem
+of domain adaptation
 \citep{blitzer2006domain,pan2011domain,gopalan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,ganin2014unsupervised},
 where the goal is often stated as trying to learn a domain-invariant
-representation of the data. Similarly, our method also relates to the problem of
-enforcing fairness in classification \citep{zemel2013learning,EdwardsS15}. For
-both families of methods, the problem can equivalently be stated as learning a
-classifier which is a pivotal quantity with respect to either the domain or some
-chosen attribute of the data, such as gender, color or age. In this context,
+representation of the data. Likewise, our method also relates to the problem of
+enforcing fairness in classification \citep{zemel2013learning,EdwardsS15}, which
+is stated as learning a classifier that is independent of some chosen attribute
+such as gender, color or age. For both families of methods, the problem can
+equivalently be stated as learning a classifier which is a pivotal quantity with
+respect to either the domain or the selected feature. In this context,
 \citep{ganin2014unsupervised,EdwardsS15} are certainly among the closest to our
-work, in which domain invariance and fairness are enforced through an adversarial
-minimax setup composed of a classifier and an adversary discriminator. In
-particular, our work can be regarded as a generalization that also supports the
-continuous case, which can be viewed as handling infinitely many domains,
-provided they can be continuously parameterized, or as enforcing fairness over
-continuous attributes.
+work, in which domain invariance and fairness are enforced through an
+adversarial minimax setup composed of a classifier and an adversary
+discriminator. Following this line of work, our method can be regarded as a
+generalization that also supports the continuous case, which can be viewed as
+handling infinitely many domains, provided they can be continuously
+parameterized, or as enforcing fairness over continuous attributes.
 

### 2016-09-12 tex/paper.tex
save

@@
 network whose output $j$ is the estimated value of the corresponding parameter
-of that distribution (e.g., its mean and variance), and such that
+of that distribution (e.g., its mean and variance). As in
+\citep{nix1994estimating}, the estimated probability density
 $p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$

### 2016-09-12 tex/paper.tex
add figure

@@
 \usepackage{algpseudocode}
+\usepackage{tikz}
 
@@
 
+\begin{figure*}
+    \usetikzlibrary{arrows}
+    \def\layersep{1cm}
+
+    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.8]
+    \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!30, line width=0.3mm, fill=white]
+
+    % Classifier f
+    \node at (2,0) {Classifier $f$};
+    \draw (-1,-0.5) rectangle (4,-5.5);
+
+    \path[->, shorten >= 0pt] (-2,-3) edge (-1,-3);
+    \node[left] at (-2,-3) {$X$};
+
+    \path[-o, shorten >= 0pt] (1.5,-6.5) edge (1.5,-5.5);
+    \node[below] at (1.5,-6.5) {$\theta_f$};
+
+    \path[->, shorten >= 0pt] (3.5,-3) edge (6.5,-3);
+    \node[above] at (5.25,-3) {$f(X;\theta_f)$};
+
+    \foreach \name / \y in {1,...,3}
+        \node[neuron] (f-I-\name) at (-0.5,-1-\y) {};
+
+    \foreach \name / \y in {1,...,5}
+        \node[neuron] (f-H1-\name) at (-0.5cm+\layersep,-\y cm) {};
+    \foreach \name / \y in {1,...,5}
+        \node[neuron] (f-H2-\name) at (-0.5cm+3*\layersep,-\y cm) {};
+
+    \node[neuron] (f-O) at (-0.5cm+4*\layersep,-3cm) {};
+
+    \foreach \source in {1,...,3}
+        \foreach \dest in {1,...,5}
+            \path[black!20] (f-I-\source) edge (f-H1-\dest);
+
+    \foreach \source in {1,...,5}
+        \path[black!30] (f-H2-\source) edge (f-O);
+
+    \node[black!20] at (1.5,-3) {...};
+
+    % Adversary r
+    \node at (9,0) {Adversary $r$};
+    \draw (6.5,-0.5) rectangle (11.5,-5.5);
+
+    \node[above] at (12.75,-2.5) {$\mu(f(X;\theta_f))$};
+    \path[-o, shorten >= 0pt] (11,-2.5) edge (14,-2.5);
+    \node[above] at (12.75,-3.5) {$\sigma(f(X;\theta_f))$};
+    \path[-o, shorten >= 0pt] (11,-3.5) edge (14,-3.5);
+
+    \path[-o, shorten >= 0pt] (9,-6.5) edge (9,-5.5);
+    \node[below] at (9,-6.5) {$\theta_r$};
+
+    \foreach \name / \y in {1,...,1}
+        \node[neuron] (r-I-\name) at (7,-2-\y) {};
+
+    \foreach \name / \y in {1,...,5}
+        \node[neuron] (r-H1-\name) at (7cm+\layersep,-\y cm) {};
+    \foreach \name / \y in {1,...,5}
+        \node[neuron] (r-H2-\name) at (7cm+3*\layersep,-\y cm) {};
+
+    \node[neuron] (r-O1) at (7cm+4*\layersep,-2.5cm) {};
+    \node[neuron] (r-O2) at (7cm+4*\layersep,-3.5cm) {};
+
+    \foreach \source in {1,...,1}
+        \foreach \dest in {1,...,5}
+            \path[black!20] (r-I-\source) edge (r-H1-\dest);
+
+    \foreach \source in {1,...,5}
+        \path[black!30] (r-H2-\source) edge (r-O1);
+    \foreach \source in {1,...,5}
+        \path[black!30] (r-H2-\source) edge (r-O2);
+
+    \node[black!20] at (9,-3) {...};
+
+    \draw (14,-1.5) rectangle (17,-4.5);
+    \path[->, shorten >= 0pt] (15.5,-0.5) edge (15.5,-1.5);
+    \node[above] at (15.5,-0.5) {$z$};
+    \path[->, shorten >= 0pt] (15.5,-4.5) edge (15.5,-5.5);
+    \node[below] at (15.5,-5.5) {$p_{\theta_r}(z|f(X;\theta_f))$};
+    \node at (15.5,-3) {Truncated ${\cal N}$};
+
+    \end{tikzpicture}
+
+    \caption{Architecture for the adversarial training of a binary classifier $f$ against a continuous parameter $Z$.}
+    \label{fig:architecture}
+\end{figure*}
+
 In this work, we repurpose adversarial training as a means to constraint the
-predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. In
-particular, we pit $f$ against an adversary model $r := p_{\theta_r}(z |
-f(X;\theta_f))$ of parameters $\theta_r$ and associated loss ${\cal
-L}_r(\theta_f, \theta_r)$. This model takes  as input realizations of $f(X;
-\theta_f)$, for the current value $\theta_f$ of $f$ parameters, and produces as
-output a function $p_{\theta_r}(z | f(X;\theta_f))$ modeling the posterior
-probability density that $z$ parameterizes the sample $X$ observed through
-$f(\cdot;\theta_f)$. If $Z$ takes discrete values, then $p_{\theta_r}$ can be
-represented as a probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal
-Z|}}$ whose output $j$ (for $j=1, \dots, |{\cal Z}|$) is the estimated
-probability mass $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly, if $Z$ takes
-continuous values and if we assume some parameteric distribution for
-$Z|f(X;\theta_f)$ (e.g., a gaussian), then $p_{\theta_r}$ can be represented as
-network whose output $j$ is the estimated value of the corresponding parameter
-of that distribution (e.g., its mean and variance). As in
+predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. As
+illustrated in Figure~\ref{fig:architecture}, we pit $f$ against an adversary
+model $r := p_{\theta_r}(z | f(X;\theta_f))$ of parameters $\theta_r$ and
+associated loss ${\cal L}_r(\theta_f, \theta_r)$. This model takes  as input
+realizations of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$
+parameters, and produces as output a function $p_{\theta_r}(z | f(X;\theta_f))$
+modeling the posterior probability density that $z$ parameterizes the sample $X$
+observed through $f(\cdot;\theta_f)$. If $Z$ takes discrete values, then
+$p_{\theta_r}$ can be represented as a probabilistic classifier $\mathbb{R}
+\mapsto \mathbb{R}^{|{\cal Z|}}$ whose output $j$ (for $j=1, \dots, |{\cal Z}|$)
+is the estimated probability mass $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly,
+if $Z$ takes continuous values and if we assume some parameteric distribution
+for $Z|f(X;\theta_f)$ (e.g., a truncated gaussian over a bounded support), then $p_{\theta_r}$ can be represented
+as network whose output $j$ is the estimated value of the corresponding
+parameter of that distribution (e.g., its mean and variance). As in
 \citep{nix1994estimating}, the estimated probability density
@@
 \glnote{Discuss that we could equivalently require an intermediate representation of the data (as learned internally by $f$) to be pivotal, instead of requiring $f(X)$ to be pivotal.}
-\glnote{Add summary figure.}
 

### 2016-09-12 tex/paper.tex
save

@@
 
-    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.8]
+    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.85]
     \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!30, line width=0.3mm, fill=white]
@@
 
-    \node[above] at (12.75,-2.5) {$\mu(f(X;\theta_f))$};
+    \node[above] at (12.75,-2.5) {$\mu_{\theta_r}(f(X;\theta_f))$};
     \path[-o, shorten >= 0pt] (11,-2.5) edge (14,-2.5);
-    \node[above] at (12.75,-3.5) {$\sigma(f(X;\theta_f))$};
+    \node[above] at (12.75,-3.5) {$\sigma_{\theta_r}(f(X;\theta_f))$};
     \path[-o, shorten >= 0pt] (11,-3.5) edge (14,-3.5);
@@
 
-    \caption{Architecture for the adversarial training of a binary classifier $f$ against a continuous parameter $Z$.}
+    \caption{Architecture for the adversarial training of a binary classifier $f$ against a continuous parameter $Z$. \glnote{Shall we also illustrate the losses?}}
     \label{fig:architecture}

### 2016-09-12 tex/paper.tex
save

@@
 viverra. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin a urna a
-ex egestas pulvinar. Curabitur facilisis a dui nec fringilla. Duis bibendum
-fermentum quam. Donec eget urna eget mi placerat malesuada ut et lacus. Nulla ac
-risus scelerisque, congue eros vitae, dignissim risus. Fusce purus ante,
-vulputate sed magna et, varius imperdiet leo.
+ex egestas pulvinar.
 
@@
 
+    \path[dashed,-] (5.25,-3) edge (5.25,-6.5);
+    \node[below] at (5.25,-6.5) {${\cal L}_f(\theta_f)$};
+
     \foreach \name / \y in {1,...,3}
@@
 
+    \draw[dashed,-] (15.5,-5) -| (12.75,-6.5);
+    \node[below] at (12.75,-6.5) {${\cal L}_r(\theta_f, \theta_r)$};
+
     \end{tikzpicture}
 
-    \caption{Architecture for the adversarial training of a binary classifier $f$ against a continuous parameter $Z$. \glnote{Shall we also illustrate the losses?}}
+    \caption{Architecture for the adversarial training of a binary classifier
+    $f$ against a continuous parameter $Z$, assuming that $Z|f(X;\theta_f)$
+    follows a truncated gaussian distribution.}
     \label{fig:architecture}
@@
 modeling the posterior probability density that $z$ parameterizes the sample $X$
-observed through $f(\cdot;\theta_f)$. If $Z$ takes discrete values, then
-$p_{\theta_r}$ can be represented as a probabilistic classifier $\mathbb{R}
-\mapsto \mathbb{R}^{|{\cal Z|}}$ whose output $j$ (for $j=1, \dots, |{\cal Z}|$)
-is the estimated probability mass $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly,
-if $Z$ takes continuous values and if we assume some parameteric distribution
-for $Z|f(X;\theta_f)$ (e.g., a truncated gaussian over a bounded support), then $p_{\theta_r}$ can be represented
-as network whose output $j$ is the estimated value of the corresponding
-parameter of that distribution (e.g., its mean and variance). As in
-\citep{nix1994estimating}, the estimated probability density
-$p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$
-under that distribution.  Intuitively, if $p(f(X; \theta_f)|z)$ varies with $z$,
+observed through $f(\cdot;\theta_f)$.
+Intuitively, if $p(f(X; \theta_f)|z)$ varies with $z$,
 then the corresponding correlation can be captured by $r$. By contrast, if
@@
 towards Eqn.~\ref{eqn:criterion}.
+
+If $Z$ takes discrete values, then $p_{\theta_r}$ can be represented e.g. as a
+probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal Z|}}$ whose
+output $j$ (for $j=1, \dots, |{\cal Z}|$) is the estimated probability mass
+$p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly, if $Z$ takes continuous values and
+if we assume some parameteric distribution for $Z|f(X;\theta_f)$ (e.g., a
+truncated gaussian over a bounded support), then $p_{\theta_r}$ can be
+represented e.g. as network whose output $j$ is the estimated value of the
+corresponding parameter of that distribution (e.g., its mean and variance). As
+in \citep{nix1994estimating}, the estimated probability density
+$p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$.
+As further explained in the next section, let us note that the adversary $r$ may
+take any form, i.e. it does need to be a neural network, as long as it exposes a
+differentiable function $p_{\theta_r}(z|f(X;\theta_f))$ of sufficient capacity
+to represent the true distribution within its bounded support.
 
@@
 \mathbb{R}^p \mapsto [0,1]$ modeling $p(Y=1|X)$. For reasons further explained
-in Section~\ref{sec:theory}, ${\cal L}_f$ (resp. ${\cal L}_r$) is set to the
+in Section~\ref{sec:theory}, ${\cal L}_f$ and ${\cal L}_r$  are respectively set to the
 expected value of the
-negative log-likelihood of $Y|X$ under $f$ (resp. of $Z|f(X;\theta_f)$ under
-$r$). The optimization algorithm consists in using stochastic gradient descent
+negative log-likelihood of $Y|X$ under $f$ and of of $Z|f(X;\theta_f)$ under
+$r$:
+\begin{align}
+    {\cal L}_f(\theta_f) &= \mathbb{E}_{Y|X} [ -\log p_{\theta_f} (Y|X) ], \\
+    {\cal L}_r(\theta_f, \theta_r) &= \mathbb{E}_{Z|f(X;\theta_f)} [-\log p_{\theta_r} (Z|f(X;\theta_f))].
+\end{align}
+The optimization algorithm consists in using stochastic gradient descent
 alternatively for solving Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.
@@
 \end{figure*}
-
-\glnote{Discuss that we could equivalently require an intermediate representation of the data (as learned internally by $f$) to be pivotal, instead of requiring $f(X)$ to be pivotal.}
 
@@
 
+Interestingly, let us emphasize that these results hold using only the 1D output
+of $f$ (in the case of binary classification) as input to the adversary. We
+could similarly enforce an intermediate representation of the data to be
+pivotal, e.g. as in \citep{ganin2014unsupervised}, but this is in fact not necessary.
+
 % ==============================================================================

### 2016-09-12 tex/paper.tex
save

@@
 
-    \node[black!20] at (1.5,-3) {...};
+    \node[black!30] at (1.5,-3) {...};
 
     % Adversary r
-    \node at (9,0) {Adversary $r$};
+    \node at (11.75,0) {Adversary $r$};
     \draw (6.5,-0.5) rectangle (11.5,-5.5);
@@
 
-    \node[black!20] at (9,-3) {...};
+    \node[black!30] at (9,-3) {...};
 
@@
     \path[->, shorten >= 0pt] (15.5,-0.5) edge (15.5,-1.5);
-    \node[above] at (15.5,-0.5) {$z$};
+    \node[above] at (15.5,-0.5) {$Z$};
     \path[->, shorten >= 0pt] (15.5,-4.5) edge (15.5,-5.5);
-    \node[below] at (15.5,-5.5) {$p_{\theta_r}(z|f(X;\theta_f))$};
+    \node[below] at (15.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
     \node at (15.5,-3) {Truncated ${\cal N}$};
@@
 expected value of the
-negative log-likelihood of $Y|X$ under $f$ and of of $Z|f(X;\theta_f)$ under
+negative log-likelihood of $Y|X$ under $f$ and of $Z|f(X;\theta_f)$ under
 $r$:
@@
 
-Interestingly, let us emphasize that these results hold using only the 1D output
+Interestingly, let us emphasize that these results hold using only the (1D) output
 of $f$ (in the case of binary classification) as input to the adversary. We
 could similarly enforce an intermediate representation of the data to be
-pivotal, e.g. as in \citep{ganin2014unsupervised}, but this is in fact not necessary.
+pivotal, e.g. as in \citep{ganin2014unsupervised}, but this is in fact not
+necessary. \glnote{This also suggests we could use for $r$ a simpler, maybe
+closed form, representation than a neural network (e.g. a GMM?).}
 

### 2016-09-19 tex/paper.tex
save

@@
 \mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
-on a nuisance parameter $Z$ whose values $z \in {\cal Z}$  define a
+of the functional $X$ on a nuisance parameter $Z$ whose values $z \in {\cal Z}$  define a
 parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
@@
 
-    \node[above] at (12.75,-2.5) {$\mu_{\theta_r}(f(X;\theta_f))$};
-    \path[-o, shorten >= 0pt] (11,-2.5) edge (14,-2.5);
-    \node[above] at (12.75,-3.5) {$\sigma_{\theta_r}(f(X;\theta_f))$};
-    \path[-o, shorten >= 0pt] (11,-3.5) edge (14,-3.5);
+    \node[above] at (12.75,-2) {\small $\theta^1(f(X;\theta_f);\theta_r)$};
+    \path[-o, shorten >= 0pt] (11,-2.0) edge (14,-2.0);
+    \node[above] at (12.75,-3) {$\theta^2(f(X;\theta_f);\theta_r)$};
+    \path[-o, shorten >= 0pt] (11,-3) edge (14,-3);
+    \node[above] at (12.75,-4) {$\dots$};
+    \path[-o, shorten >= 0pt] (11,-4) edge (14,-4);
 
@@
 
-    \node[neuron] (r-O1) at (7cm+4*\layersep,-2.5cm) {};
-    \node[neuron] (r-O2) at (7cm+4*\layersep,-3.5cm) {};
+    \node[neuron] (r-O1) at (7cm+4*\layersep,-2cm) {};
+    \node[neuron] (r-O2) at (7cm+4*\layersep,-3cm) {};
+    \node[neuron] (r-O3) at (7cm+4*\layersep,-4cm) {};
 
@@
         \path[black!30] (r-H2-\source) edge (r-O2);
+    \foreach \source in {1,...,5}
+        \path[black!30] (r-H2-\source) edge (r-O3);
 
@@
     \node[below] at (15.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
-    \node at (15.5,-3) {Truncated ${\cal N}$};
+    \node at (15.5,-3) {${\cal D}(\theta^1, \theta^2, \dots)$};
 
@@
     \caption{Architecture for the adversarial training of a binary classifier
-    $f$ against a continuous parameter $Z$, assuming that $Z|f(X;\theta_f)$
-    follows a truncated gaussian distribution.}
+    $f$ against a continuous parameter $Z$.}
     \label{fig:architecture}

### 2016-09-19 tex/paper.tex
save

@@
 
+\tikzstyle{every node}=[font=\scriptsize]
 \begin{figure*}
@@
 
-    \node[above] at (12.75,-2) {\small $\theta^1(f(X;\theta_f);\theta_r)$};
+    \node[above] at (12.75,-2) {$\theta^1(f(X;\theta_f);\theta_r)$};
     \path[-o, shorten >= 0pt] (11,-2.0) edge (14,-2.0);
@@
 $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly, if $Z$ takes continuous values and
-if we assume some parameteric distribution for $Z|f(X;\theta_f)$ (e.g., a
+if we assume some parameteric distribution ${\cal D}$ for $Z|f(X;\theta_f)$ (e.g., a
 truncated gaussian over a bounded support), then $p_{\theta_r}$ can be
 represented e.g. as network whose output $j$ is the estimated value of the
-corresponding parameter of that distribution (e.g., its mean and variance). As
+corresponding parameter $\theta^j$ of that distribution (e.g., its mean and variance). As
 in \citep{nix1994estimating}, the estimated probability density

### 2016-09-19 tex/paper.tex
save

@@
 Let $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
-\theta_r)$ for fixed $\theta_f$. If $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
+\theta_r)$ for fixed $\theta_f$. If $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z)$
 for all $z \in {\cal Z}$, then $f$ is a pivotal quantity.
@@
 \arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
-\theta_r)$$ are such that $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
-By assumption, $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$,
+\theta_r)$$ are such that $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
+By assumption, $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z)$,
 and therefore $p(z|f(X;\theta_f)) = p(z)$.
@@
 \max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal
-L}_r(\theta_f, \theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) =
+L}_r(\theta_f, \theta_r)$, in which case $p_{\hat\theta_r}(z|f(X;\theta_f)) =
 p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the entropy
@@
 \begin{itemize}
-    \item $\smash{\hat\theta}_f$ corresponds to
-    the parameters of an optimal classifier, in which case the expected
-    negative log-likelihood ${\cal L}_f$ of $Y|X$ reduces to its minimum
-    value $H(p_{Y|X})$,
-    \item all
-   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
-   $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal
-   Z}$ since we require a uniform prior by construction. Note that in the
-   continuous case, the supremum of the differential entropy
-   over continuous distributions on the same bounded support is also realized
-   by the uniform distribution over that support.
+    \item $\smash{\hat\theta}_f$ minimizes the negative log-likelihood ${\cal
+    L}_f$ of $X|Y$, which happens when $\smash{\hat\theta}_f$ are the parameters
+    of an optimal classifier and in which case ${\cal L}_f$ reduces to its
+    minimum value $H(p_{Y|X})$,
+
+    \item $\smash{\hat\theta}_f$ maximimizes the entropy
+    $H(p_{Z|f(X;\theta_f)})$, which happens when all outcomes of
+    $Z|f(X;\smash{\hat\theta}_f)$ are equally likely and in which case
+    $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal Z}$ since we
+    require a uniform prior by construction. Note that in the continuous case,
+    the supremum of the differential entropy over continuous distributions on
+    the same bounded support is also realized by the uniform distribution over
+    that support.
 \end{itemize}
-Accordingly, the second condition implies that $p_{\theta_r}(z|f(X;\theta_f)) =
-p(z)$ and therefore that at this point, because of Proposition~\ref{prop:1}, the
-optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
+When the lower bound is active, we have $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
+because of the second condition, and because of Proposition~\ref{prop:1}, the
+optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is therefore also a pivotal
+quantity.
 
@@
 pivotal, e.g. as in \citep{ganin2014unsupervised}, but this is in fact not
-necessary. \glnote{This also suggests we could use for $r$ a simpler, maybe
-closed form, representation than a neural network (e.g. a GMM?).}
+necessary.
 

### 2016-09-19 tex/paper.tex
save

@@
     \node[below] at (15.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
-    \node at (15.5,-3) {${\cal D}(\theta^1, \theta^2, \dots)$};
+    \node at (15.5,-3) {${\cal P}(\theta^1, \theta^2, \dots)$};
 
@@
 $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly, if $Z$ takes continuous values and
-if we assume some parameteric distribution ${\cal D}$ for $Z|f(X;\theta_f)$ (e.g., a
+if we assume some parameteric distribution ${\cal P}$ for $Z|f(X;\theta_f)$ (e.g., a
 truncated gaussian over a bounded support), then $p_{\theta_r}$ can be

### 2016-09-19 tex/paper.tex
save

@@
 When the lower bound is active, we have $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
-because of the second condition, and because of Proposition~\ref{prop:1}, the
-optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is therefore also a pivotal
-quantity.
+because of the second condition, which induces that the
+optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal
+quantity as a consequence of Proposition~\ref{prop:1}.
 

### 2016-09-20 tex/paper.tex
Intro from Michael (#5)
* update intro

* adding stuff to intro

@@
 \section{Introduction}
+
+The discovery of new particles or physical phenomena in high energy
+physics experiments, like those at the Large Hadron Collider, 
+requires the observation of statistically significant deviations from
+the predictions of the Standard Model, the current model
+of the known fundamental particles and their interactions.  Typically,
+discovery requires satisfying the \textit{5-sigma rule}, 
+whereby the deviation must be at least five standard deviations from
+predictions, i.e. having a p-value of $p \lesssim 3 \times 10^{-7}$.
+The evaluation of the statistical significance must not only account
+for statistical uncertainties, but also systematic 
+uncertainties that represent our incomplete knowledge of the theory of
+particle interactions or of the properties of the experimental
+detection apparatus. Systematic uncertainties can alter the expected
+rate of signal and background classes, e.g. their priors, as well as
+the distributions of features. As such, the signal class true-positive
+rate and the background class false-positive rate of a classifier can
+change with variations in the systematic uncertainty thereby
+decreasing the sensitivity of an analysis.
+
+In this paper, we present a new strategy for training a classifier,
+based on adversarial training~\citep{goodfellow2014generative}, 
+which aims to build a classifier whose decision function is invariant
+under variations due to systematic uncertainties.  In this way,
+uncertainties on predictions from simulations will have a reduced
+impact when computing the significance of potential deviations after
+application of the discriminant model and thereby increase
+the sensitivity of the analysis to signals from potential new physics sources.
+
+Given training data $x \in X$ and associated labels $y \in \{s,
+b_{i=1...k} \}$ where $s$ is the signal class and $b_{i=1...k}$ are
+the background classes, then the probability distribution of the data
+can be written
+\[
+p(x) = \pi_s p(x|y=s) + \sum_{i=1}^{k} \pi_{b_i} p(x | y=b_i)
+\]
+where $\pi_{y}$ represents the prior for the signal or a given
+background, and $p(x|y)$ is the conditional distribution of the data
+for the signal or a given background. Systematic uncertainties can
+affect this distribution in the following ways:
+\begin{itemize}
+\item Uncertainties on the priors $\pi_{y}$.  These arise from
+  uncertainties on the rates we expect a given process to occur as
+  predicted from theoretical calculations or from alternative
+  measurements of the rate of a given class.
+
+\item Uncertainties in the simulators of the physical interactions,
+  giving rise to uncertainties on the distributions $p(x|y)$.  These
+  uncertainty are often determined by comparing the predictions of
+  different simulations using differing models of the same physical
+  process.
+
+\item Uncertainties on the detection apparatus and it's effect on the
+  measurement of particle properties, giving rise to uncertainties on
+  the distributions $p(x|y)$.  For instance this includes
+  uncertainties on our knowledge of the mean and variance of the
+  measured energy distribution of a particle.
+
+\end{itemize}
+In this paper, we will not address uncertainties of the first type
+mentioned above (uncertainties on the priors).  Rather, we will focus
+on mitigating the effect of systematic uncertainties that alter the
+conditional distributions $p(x|y)$.
+
+Systematic uncertainties are parameterized by nuisance parameters in
+the statistical analysis of data.  Nuisance parameter often, though
+not always, have a suitable prior coming from alternative measurements
+in data.  For instance, The average measured energy of an electron in
+simulation may be compared with those in data using a relatively pure
+sample of $Z$ boson decays to electrons.  The average measured
+electron energy can not be measured without uncertainty, and the
+measured uncertainty can be used in other analyses to estimate the
+uncertainty deriving from our imperfect knowledge of the electron's
+energy.  A nuisance parameter is assigned to the electron energy
+uncertainty, which controls the variation of the electron energy in
+the statistical analysis of data.  Nuisance parameters are typically,
+though not always, constrained by a suitable prior, e.g. a Gaussian
+distribution with a standard deviation equal to the measured uncertainty.
 

### 2016-09-21 tex/paper.tex
add inline comments in intro + small changes

@@
 
-The discovery of new particles or physical phenomena in high energy
-physics experiments, like those at the Large Hadron Collider, 
-requires the observation of statistically significant deviations from
-the predictions of the Standard Model, the current model
-of the known fundamental particles and their interactions.  Typically,
-discovery requires satisfying the \textit{5-sigma rule}, 
-whereby the deviation must be at least five standard deviations from
-predictions, i.e. having a p-value of $p \lesssim 3 \times 10^{-7}$.
-The evaluation of the statistical significance must not only account
-for statistical uncertainties, but also systematic 
-uncertainties that represent our incomplete knowledge of the theory of
-particle interactions or of the properties of the experimental
-detection apparatus. Systematic uncertainties can alter the expected
-rate of signal and background classes, e.g. their priors, as well as
-the distributions of features. As such, the signal class true-positive
-rate and the background class false-positive rate of a classifier can
-change with variations in the systematic uncertainty thereby
-decreasing the sensitivity of an analysis.
+The discovery of new particles or physical phenomena in high energy physics
+experiments, like those at the Large Hadron Collider, requires the observation
+of statistically significant deviations from the predictions of the Standard
+Model, the current model of the known fundamental particles and their
+interactions.  Typically, discovery requires satisfying the \textit{5-sigma
+rule}, whereby the deviation must be at least five standard deviations from
+predictions, i.e. having a p-value of $p \lesssim 3 \times 10^{-7}$. The
+evaluation of the statistical significance must not only account for statistical
+uncertainties, but also systematic uncertainties that represent our incomplete
+knowledge of the theory of particle interactions or of the properties of the
+experimental detection apparatus. Systematic uncertainties can alter the
+expected rate of signal and background classes, e.g. their priors, as well as
+the distributions of features. As such, the signal class true-positive rate and
+the background class false-positive rate of a classifier can change with
+variations in the systematic uncertainty thereby decreasing the sensitivity of
+an analysis. In this paper, we present a new strategy for training a classifier,
+based on adversarial training~\citep{goodfellow2014generative}, which aims to
+build a classifier whose decision function is invariant under variations due to
+systematic uncertainties.  In this way, uncertainties on predictions from
+simulations will have a reduced impact when computing the significance of
+potential deviations after application of the discriminant model and thereby
+increase the sensitivity of the analysis to signals from potential new physics
+sources. \glnote{Shall we be more explicit somewhere and put a formula
+underlining the effect of having a pivotal classifier on the resulting
+sensitivity? (versus a classifier whose decision function depends on the
+nuisance.)}
 
-In this paper, we present a new strategy for training a classifier,
-based on adversarial training~\citep{goodfellow2014generative}, 
-which aims to build a classifier whose decision function is invariant
-under variations due to systematic uncertainties.  In this way,
-uncertainties on predictions from simulations will have a reduced
-impact when computing the significance of potential deviations after
-application of the discriminant model and thereby increase
-the sensitivity of the analysis to signals from potential new physics sources.
-
-Given training data $x \in X$ and associated labels $y \in \{s,
+Given data $X$ and associated labels $Y$
+taking values $y \in {\cal Y} = \{s,
 b_{i=1...k} \}$ where $s$ is the signal class and $b_{i=1...k}$ are
-the background classes, then the probability distribution of the data
-can be written
-\[
-p(x) = \pi_s p(x|y=s) + \sum_{i=1}^{k} \pi_{b_i} p(x | y=b_i)
-\]
+the background classes, then the probability density of the data
+can be written as a mixture
+\begin{equation}
+p(X) = \pi_s p(X|Y=s) + \sum_{i=1}^{k} \pi_{b_i} p(X | Y=b_i),
+\end{equation}
 where $\pi_{y}$ represents the prior for the signal or a given
-background, and $p(x|y)$ is the conditional distribution of the data
+background, and $p(X|Y)$ is the conditional distribution of the data
 for the signal or a given background. Systematic uncertainties can
-affect this distribution in the following ways:
+affect this distribution in the following ways \glnote{add references}:
 \begin{itemize}
@@
 \item Uncertainties in the simulators of the physical interactions,
-  giving rise to uncertainties on the distributions $p(x|y)$.  These
-  uncertainty are often determined by comparing the predictions of
+  giving rise to uncertainties on the distributions $p(X|Y)$.  These
+  uncertainties are often determined by comparing the predictions of
   different simulations using differing models of the same physical
@@
 
-\item Uncertainties on the detection apparatus and it's effect on the
+\item Uncertainties on the detection apparatus and its effect on the
   measurement of particle properties, giving rise to uncertainties on
-  the distributions $p(x|y)$.  For instance this includes
+  the distributions $p(X|Y)$.  For instance, this includes
   uncertainties on our knowledge of the mean and variance of the
   measured energy distribution of a particle.
-
 \end{itemize}
@@
 on mitigating the effect of systematic uncertainties that alter the
-conditional distributions $p(x|y)$.
+conditional distributions $p(X|Y)$.
 
-Systematic uncertainties are parameterized by nuisance parameters in
-the statistical analysis of data.  Nuisance parameter often, though
-not always, have a suitable prior coming from alternative measurements
-in data.  For instance, The average measured energy of an electron in
-simulation may be compared with those in data using a relatively pure
-sample of $Z$ boson decays to electrons.  The average measured
-electron energy can not be measured without uncertainty, and the
-measured uncertainty can be used in other analyses to estimate the
-uncertainty deriving from our imperfect knowledge of the electron's
-energy.  A nuisance parameter is assigned to the electron energy
-uncertainty, which controls the variation of the electron energy in
-the statistical analysis of data.  Nuisance parameters are typically,
-though not always, constrained by a suitable prior, e.g. a Gaussian
-distribution with a standard deviation equal to the measured uncertainty.
+Systematic uncertainties are parameterized by nuisance parameters in the
+statistical analysis of data \glnote{add references}.  Nuisance parameter often,
+though not always, have a suitable prior coming from alternative measurements in
+data.  For instance, The average measured energy of an electron in simulation
+may be compared with those in data using a relatively pure sample of $Z$ boson
+decays to electrons.  The average measured electron energy can not be measured
+without uncertainty, and the measured uncertainty can be used in other analyses
+to estimate the uncertainty deriving from our imperfect knowledge of the
+electron's energy.  A nuisance parameter is assigned to the electron energy
+uncertainty, which controls the variation of the electron energy in the
+statistical analysis of data.  Nuisance parameters are typically, though not
+always, constrained by a suitable prior, e.g. a Gaussian distribution with a
+standard deviation equal to the measured uncertainty. \glnote{... and
+accordingly, we want to build a classifier independent on the nuisance
+parameter, as a way to be independent of the corresponding systematics.}
 
-\glnote{Distinction between statistical and systematic uncertainty.}
-\glnote{Outline taxonomy of uncertainties.}
-\glnote{Define nuisance parameters. See refs in Noel's papers.}
-\glnote{We want to build an accurate classifier whose output remains invariant with
-respect to systematic uncertainties.}
-% https://www.slac.stanford.edu/econf/C030908/papers/TUAT004.pdf
-\glnote{Motivate the criterion (which may not be obvious for the ML crowd). See pivotal quantity motivation.}
-\glnote{Discuss what kind of uncertainties the proposed approach is good for.}
+\glnote{Add paper outline.}
 

### 2016-09-22 tex/paper.tex
save
losses are in expectation over X

@@
 \begin{align}
-    {\cal L}_f(\theta_f) &= \mathbb{E}_{Y|X} [ -\log p_{\theta_f} (Y|X) ], \\
-    {\cal L}_r(\theta_f, \theta_r) &= \mathbb{E}_{Z|f(X;\theta_f)} [-\log p_{\theta_r} (Z|f(X;\theta_f))].
+    {\cal L}_f(\theta_f) &= \mathbb{E}_{X} \mathbb{E}_{Y|X} [ -\log p_{\theta_f} (Y|X) ], \\
+    {\cal L}_r(\theta_f, \theta_r) &= \mathbb{E}_{X}  \mathbb{E}_{Z|f(X;\theta_f)} [-\log p_{\theta_r} (Z|f(X;\theta_f))].
 \end{align}
@@
 \theta_r)$ for fixed $\theta_f$. If $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z)$
-for all $z \in {\cal Z}$, then $f$ is a pivotal quantity.
+for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$, then $f$ is a pivotal quantity.
 \end{proposition}
@@
 \end{align*}
-which holds for all $z \in {\cal Z}$ and implies that $f$ is a pivotal quantity.
+which holds for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$ and implies that $f$ is a pivotal quantity.
 \end{proof}
@@
 for Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar} such that
-$E(\hat\theta_f, \hat\theta_r) = H(p_{Y|X}) - H(p_Z)$, then
+$E(\hat\theta_f, \hat\theta_r) = \mathbb{E}_{X} [H(p_{Y|X}) - H(p_Z)]$, then
 $f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier and a pivotal
@@
 L}_r(\theta_f, \theta_r)$, in which case $p_{\hat\theta_r}(z|f(X;\theta_f)) =
-p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the entropy
-$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
+p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the expected entropy
+$\mathbb{E}_{X} [ H(p_{Z|f(X;\theta_f)}) ]$ of the conditional distribution of the nuisance. The
 value function $E$ can therefore be rewritten as $$E'(\theta_f) = {\cal
-L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower
-bound $$H(p_{Y|X}) - H(p_Z) \leq {\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$$
+L}_f(\theta_f) - \mathbb{E}_{X} [ H(p_{Z|f(X;\theta_f)}) ].$$  In particular, we have the lower
+bound $$\mathbb{E}_{X} [ H(p_{Y|X}) - H(p_Z)] \leq {\cal L}_f(\theta_f) - \mathbb{E}_{X} [H(p_{Z|f(X;\theta_f)})]$$
 where the equality holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f}
@@
     of an optimal classifier and in which case ${\cal L}_f$ reduces to its
-    minimum value $H(p_{Y|X})$,
+    minimum value $\mathbb{E}_{X} [H(p_{Y|X})]$,
 
-    \item $\smash{\hat\theta}_f$ maximimizes the entropy
-    $H(p_{Z|f(X;\theta_f)})$, which happens when all outcomes of
-    $Z|f(X;\smash{\hat\theta}_f)$ are equally likely and in which case
-    $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal Z}$ since we
+    \item $\smash{\hat\theta}_f$ maximimizes the expected entropy
+    $\mathbb{E}_{X} [H(p_{Z|f(X;\theta_f)})]$, which happens when all outcomes of
+    $Z|f(X;\smash{\hat\theta}_f)$ are equally likely. In other words,
+    $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$ since we
     require a uniform prior by construction. Note that in the continuous case,
@@
 its optimum given $f$ (e.g., by setting $K$ sufficiently high) and if $f$ is
-updated to improve ${\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$, then $f$
+updated to improve ${\cal L}_f(\theta_f) - \mathbb{E}_X [ H(p_{Z|f(X;\theta_f)}) ]$, then $f$
 should converge to a classifier which is both optimal and pivotal, provided such
@@
 not exist because the nuisance parameter directly shapes the decision boundary,
-in which cases the lower bound $H(p_{Y|X}) - H(p_Z) < {\cal L}_f(\theta_f) -
-H(p_{Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal classifier or a
+in which cases the lower bound $\mathbb{E}_{X} [ H(p_{Y|X}) - H(p_Z)] < {\cal L}_f(\theta_f) - \mathbb{E}_{X} [H(p_{Z|f(X;\theta_f)})]$ is strict: $f$ can either be an optimal classifier or a
 pivotal quantity, but not both simultaneously. In this situation, it is natural

### 2016-09-23 tex/paper.tex
save

@@
 evaluation of the statistical significance must not only account for statistical
-uncertainties, but also systematic uncertainties that represent our incomplete
-knowledge of the theory of particle interactions or of the properties of the
-experimental detection apparatus. Systematic uncertainties can alter the
-expected rate of signal and background classes, e.g. their priors, as well as
-the distributions of features. As such, the signal class true-positive rate and
-the background class false-positive rate of a classifier can change with
-variations in the systematic uncertainty thereby decreasing the sensitivity of
-an analysis. In this paper, we present a new strategy for training a classifier,
-based on adversarial training~\citep{goodfellow2014generative}, which aims to
-build a classifier whose decision function is invariant under variations due to
-systematic uncertainties.  In this way, uncertainties on predictions from
-simulations will have a reduced impact when computing the significance of
-potential deviations after application of the discriminant model and thereby
-increase the sensitivity of the analysis to signals from potential new physics
-sources. \glnote{Shall we be more explicit somewhere and put a formula
-underlining the effect of having a pivotal classifier on the resulting
-sensitivity? (versus a classifier whose decision function depends on the
-nuisance.)}
+uncertainties \glnote{Shouldnt we define what are statistical uncertainties, so
+that systematic uncertainties are in constrast better understood?}, but also
+systematic uncertainties that represent our incomplete knowledge of the theory
+of particle interactions or of the properties of the experimental detection
+apparatus. Systematic uncertainties can alter the expected rate of signal and
+background classes, e.g. their priors, as well as the distributions of features.
+As such, the signal class true-positive rate and the background class
+false-positive rate of a classifier can change with variations in the systematic
+uncertainty thereby decreasing the sensitivity of an analysis. In this paper, we
+present a new strategy for training a classifier, based on adversarial
+training~\citep{goodfellow2014generative}, which aims to build a classifier
+whose decision function is invariant under variations due to systematic
+uncertainties.  In this way, uncertainties on predictions from simulations will
+have a reduced impact when computing the significance of potential deviations
+after application of the discriminant model and thereby increase the sensitivity
+of the analysis to signals from potential new physics sources. \glnote{Shall we
+be more explicit somewhere and put a formula underlining the effect of having a
+pivotal classifier on the resulting sensitivity? (versus a classifier whose
+decision function depends on the nuisance.)}
 
@@
 \end{equation}
-for all $z,z^\prime \in  {\cal Z}$.
+for all $z,z^\prime \in  {\cal Z}$ and all values of $f(X ; \theta_f)$.
 
@@
 \begin{proof}
-The optimal parameters $$\smash{\hat\theta}_r =
+Since we assume enough capacity for $r$, the optimal parameters $$\smash{\hat\theta}_r =
 \arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
@@
 
+As an illustrating example, let us consider...
+
 \glnote{Define the architecture of $r$ when $Z$ is categorical ($r$ is a

### 2016-09-24 tex/paper.tex
save

@@
 $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly, if $Z$ takes continuous values and
-if we assume some parameteric distribution ${\cal P}$ for $Z|f(X;\theta_f)$ (e.g., a
-truncated gaussian over a bounded support), then $p_{\theta_r}$ can be
+if we assume some parameteric distribution ${\cal P}$ for $Z|f(X;\theta_f)$
+(e.g., a mixture of gaussians over a bounded support, as modeled with a mixture
+density network~\citep{bishop1994mixture}), then $p_{\theta_r}$ can be
 represented e.g. as network whose output $j$ is the estimated value of the
-corresponding parameter $\theta^j$ of that distribution (e.g., its mean and variance). As
-in \citep{nix1994estimating}, the estimated probability density
+corresponding parameter $\theta^j$ of that distribution (e.g., the mean,
+variance and mixing coefficients of its components). As in
+\citep{nix1994estimating,bishop1994mixture}, the estimated probability density
 $p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$.

### 2016-09-27 tex/paper.tex
bibliography style
save

@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Etiam commodo, enim
-vitae facilisis pretium, ligula justo aliquet lectus, non interdum erat lorem et
-nisi. Donec pharetra lectus in magna pellentesque vehicula. Praesent dapibus
-lorem sed enim lacinia, a vestibulum sapien mattis. Sed ornare mollis aliquet.
-Nulla tempor lacinia tortor, in rhoncus augue porta nec. Morbi sed convallis
-nibh, eu hendrerit turpis. Curabitur sit amet rhoncus purus. Curabitur eget
-magna lorem. Phasellus pretium nisi quis est tincidunt, faucibus vulputate augue
-viverra. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin a urna a
-ex egestas pulvinar.
+Nice and cozy abstract goes here...
 
@@
 evaluation of the statistical significance must not only account for statistical
-uncertainties \glnote{Shouldnt we define what are statistical uncertainties, so
-that systematic uncertainties are in constrast better understood?}, but also
+uncertainties due to random fluctuations, but also for
 systematic uncertainties that represent our incomplete knowledge of the theory
@@
 present a new strategy for training a classifier, based on adversarial
-training~\citep{goodfellow2014generative}, which aims to build a classifier
+training~\citep{goodfellow2014generative}, which aims at building a classifier
 whose decision function is invariant under variations due to systematic
@@
 b_{i=1...k} \}$ where $s$ is the signal class and $b_{i=1...k}$ are
-the background classes, then the probability density of the data
+the background classes, the probability density of the data
 can be written as a mixture
@@
 though not always, have a suitable prior coming from alternative measurements in
-data.  For instance, The average measured energy of an electron in simulation
+data.  For instance, the average measured energy of an electron in simulation
 may be compared with those in data using a relatively pure sample of $Z$ boson
@@
     \caption{Architecture for the adversarial training of a binary classifier
-    $f$ against a continuous parameter $Z$.}
+    $f$ against a nuisance parameter $Z$. The adversary $r$ models the
+    distribution $p(Z|f(X;\theta_f))$ of the nuisance as observed only through the output $f(X;\theta_f)$ of the classifier. By
+    maximizing the antagonistic objective ${\cal L}_r(\theta_f, \theta_r)$ (as part of minimizing ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$), the classifier
+    $f$ forces $p(Z|f(X;\theta_f)$ to become uniform, which happens
+    when $f(X;\theta_f)$ is  independent of the nuisance parameter $Z$ and therefore pivotal.}
+
     \label{fig:architecture}
@@
 \subsection{Toy example}
+\label{sec:toy}
 
 As an illustrating example, let us consider...
-
-\glnote{Define the architecture of $r$ when $Z$ is categorical ($r$ is a
-standard classifier) or continuous ($r$ is e.g a 2-output NN such that the
-distribution of $Z|f(X)$ is modeled by a truncated Gaussian of known support). Cite Nix.}
 
@@
 \glnote{add ref} by estimating $p(f(x)|z)$ with a parameterized calibration
-procedure. Clearly, this classifier is however not optimal for $z \neq z_0$. In
-this setting, parameterized
+procedure. Clearly, this classifier is however not optimal for $z \neq z_0$.
+To overcome this issue, the classifier $f$ is sometimes \glnote{add refs} built instead on a mixture
+of training data generated from several nominal values $z_0, z_1, \dots$ of the nuisance.
+While this certainly improves with respect to classification performance,
+there is however no guarantee that the resulting classifier is pivotal, as shown
+previously in Section~\ref{sec:toy}.
+As an alternative, parameterized
 classifiers~\citep{cranmer2015approximating,Baldi:2016fzo} directly take
@@
 necessary. Similarly, calibration needs to be carried out only once, since  the
-dependence on the nuisance is now built-in. \glnote{Shall we also discuss the
-relation with \citep{Neal:2007zz} where point estimates of the nuisance are used
-as inputs to $f$?}
+dependence on the nuisance is now built-in.
 
@@
 
+\bibliographystyle{acm}
 \bibliography{bibliography.bib}

### 2016-09-27 tex/paper.tex
toy example (work in progress)

@@
 
-As an illustrating example, let us consider...
+\begin{figure*}
+    \label{fig:toy}
+    \begin{center}
+        \includegraphics[width=0.48\textwidth]{figures/f-plain.pdf}
+        \includegraphics[width=0.48\textwidth]{figures/surface-plain.pdf}\\
+        \vspace{-0.8cm}
+        \includegraphics[width=0.48\textwidth]{figures/f-adversary.pdf}
+        \includegraphics[width=0.48\textwidth]{figures/surface-adversary.pdf}
+    \end{center}
+    \vspace{-0.5cm}
+    \caption{Toy example.
+    (Top left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$
+       when $f$ is built without adversarial training. The resulting densities
+       are clearly dependent on $Z$, indicating that $f$ is not pivotal.
+    (Top right) The corresponding decision surface, highlighting
+       the fact that samples are easier to classify for values of $Z$ closer to $5$,
+       hence explaining the dependency.
+    (Bottom left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$ when $f$ is
+       built with adversarial training, as outlined in Section~\ref{sec:method}.
+       The resulting densities are now almost identical to each other, indicating only a
+       small dependency on $Z$.
+    (Bottom right) The corresponding decision surface, illustrating how adversarial
+       training bends the decision function to erase the dependency on $Z$.
+    }
+\end{figure*}
+
+\begin{figure}
+    \label{fig:toy-training}
+    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
+    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$ and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.}
+\end{figure}
+
+As a guiding toy example, let us consider the binary classification problem of 2D
+data drawn from multivariate gaussians with equal priors, such that
+\begin{align}
+    x &\sim {\cal N}(\mu=(0,0), \sigma=2 \times \mathbb{1}) \text{ when } Y=0, \\
+    x &\sim {\cal N}(\mu=(1,-1+z), \sigma=2 \times \mathbb{1}) \text{ when } Y=1.
+\end{align}
+The nuisance parameter $Z$ represents in this case our
+uncertainty about the exact location of the mean of the second gaussian. Our goal is to
+build a classifier $f(\cdot;\theta_f)$ for predicting $Y$ given $X$, but such that
+the probability distribution of $f(X;\theta_f)$ is invariant with respect to the
+nuisance parameter $Z$.
+
+Assuming a uniform prior over the interval ${\cal Z} = [0;5]$, we start by
+generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we train
+a classifier $f$ minimizing $L_f(\theta_f)$ without any adversary.
+\glnote{to be continued}
 

### 2016-09-27 tex/paper.tex
save

@@
     \end{center}
-    \vspace{-0.5cm}
+    \vspace{-1cm}
     \caption{Toy example.
@@
        built with adversarial training, as outlined in Section~\ref{sec:method}.
-       The resulting densities are now almost identical to each other, indicating only a
+       The resulting densities are now closer to each other, indicating only a
        small dependency on $Z$.
@@
     \includegraphics[width=0.48\textwidth]{figures/training.pdf}
+    \vspace{-1cm}
     \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$ and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.}
@@
 \end{align}
-The nuisance parameter $Z$ represents in this case our
+The continuous nuisance parameter $Z$ represents in this case our
 uncertainty about the exact location of the mean of the second gaussian. Our goal is to
@@
 Assuming a uniform prior over the interval ${\cal Z} = [0;5]$, we start by
-generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we train
-a classifier $f$ minimizing $L_f(\theta_f)$ without any adversary.
+generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build
+a classifier $f$ minimizing $L_f(\theta_f)$ without adversarial training.
 \glnote{to be continued}

### 2016-09-27 tex/paper.tex
save

@@
 
-    \item $\smash{\hat\theta}_f$ maximimizes the expected entropy
+    \item $\smash{\hat\theta}_f$ maximizes the expected entropy
     $\mathbb{E}_{X} [H(p_{Z|f(X;\theta_f)})]$, which happens when all outcomes of
@@
 \begin{figure*}
-    \label{fig:toy}
     \begin{center}
@@
     \caption{Toy example.
-    (Top left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$
+    (Top left) Conditional probability densities of the decision scores for $Z=0$, $2.5$ and $5$
        when $f$ is built without adversarial training. The resulting densities
@@
        hence explaining the dependency.
-    (Bottom left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$ when $f$ is
+    (Bottom left) Conditional probability densities of the decision scores for $Z=0$, $2.5$ and $5$ when $f$ is
        built with adversarial training, as outlined in Section~\ref{sec:method}.
@@
     }
+    \label{fig:toy}
 \end{figure*}
@@
 \begin{figure}
-    \label{fig:toy-training}
     \includegraphics[width=0.48\textwidth]{figures/training.pdf}
     \vspace{-1cm}
-    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$ and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.}
+    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$
+             and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.
+             Adversarial training was performed for $1000$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
+    \label{fig:toy-training}
 \end{figure}
@@
 Assuming a uniform prior over the interval ${\cal Z} = [0;5]$, we start by
-generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build
-a classifier $f$ minimizing $L_f(\theta_f)$ without adversarial training.
-\glnote{to be continued}
+generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
+neural network classifier $f$ minimizing $L_f(\theta_f)$ without adversarial
+training. As shown in the first row of Figure~\ref{fig:toy}, the resulting
+classifier is not pivotal, as the conditional probability densities of its
+decision scores $f(X;\theta_f)$ show large discrepancies between values $z$ of
+the nuisance. While not shown here, a classifier trained only from data
+generated at the nominal value $Z=0$ would also not be pivotal.
+
+By contrast, when using Algorithm~\ref{alg:adversarial-training} for training
+$f$ jointly with a mixture density network $r$ modeling the conditional
+distribution of the nuisance as observed through $f$, the resulting classifier
+becomes effectively less dependent on $Z$, as shown in the second row of
+Figure~\ref{fig:toy}. In particular, the conditional probability densities of
+the decision scores $f(X;\theta_f)$ are now closer to each other, indicating only a
+small  dependency on the nuisance, as theoretically expected.
+\glnote{Discuss Figure~\ref{fig:toy-training}}
 

### 2016-09-28 tex/paper.tex
better figures

--- 2016-09-27 tex/paper.tex
@@
     \caption{Toy example.
-    (Top left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$
+    (Upper left) Conditional probability densities of the decision scores for $Z=0, 1, \dots, 4$
        when $f$ is built without adversarial training. The resulting densities
        are clearly dependent on $Z$, indicating that $f$ is not pivotal.
-    (Top right) The corresponding decision surface, highlighting
-       the fact that samples are easier to classify for values of $Z$ closer to $5$,
-       hence explaining the dependency.
-    (Bottom left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$ when $f$ is
+    (Upper right) The corresponding decision surface, highlighting
+       the fact that samples are easier to classify for values of $Z$ closer to $4$,
+       hence partially explaining the dependency.
+    (Lower left) Conditional probability densities of the decision scores for $Z=0, 1, \dots, 4$ when $f$ is
        built with adversarial training, as outlined in Section~\ref{sec:method}.
-       The resulting densities are now closer to each other, indicating only a
+       The resulting densities are now almost identical to each other, indicating only a
        small dependency on $Z$.
-    (Bottom right) The corresponding decision surface, illustrating how adversarial
+    (Lower right) The corresponding decision surface, illustrating how adversarial
        training bends the decision function to erase the dependency on $Z$.
     }
+    \label{fig:toy}
 \end{figure*}
@@
 \begin{figure}
-    \label{fig:toy-training}
     \includegraphics[width=0.48\textwidth]{figures/training.pdf}
     \vspace{-1cm}
-    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$ and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.}
+    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$
+             and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.
+             Adversarial training was performed for $2000$ iterations, mini-batches of size $M=128$, $K=1$ and $\lambda=100$.}
+    \label{fig:toy-training}
 \end{figure}
 
-As a guiding toy example, let us consider the binary classification problem of 2D
+As a guiding toy example, let us consider the binary classification of 2D
 data drawn from multivariate gaussians with equal priors, such that
@@
 
-Assuming a uniform prior over the interval ${\cal Z} = [0;5]$, we start by
-generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build
-a classifier $f$ minimizing $L_f(\theta_f)$ without adversarial training.
-\glnote{to be continued}
+Assuming a uniform prior over the interval ${\cal Z} = [0;4]$, we start by
+generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
+neural network classifier $f$ minimizing $L_f(\theta_f)$ without adversarial
+training. As shown in the upper plots of Figure~\ref{fig:toy}, the resulting
+classifier is not pivotal, as the conditional probability densities of its
+decision scores $f(X;\theta_f)$ show large discrepancies between values $z$ of
+the nuisance. While not shown here, a classifier trained only from data
+generated at the nominal value $Z=0$ would also not be pivotal.
+
+By contrast, when using Algorithm~\ref{alg:adversarial-training} for training
+$f$ jointly with a mixture density network $r$ modeling the conditional
+distribution of the nuisance as observed through $f$, the resulting classifier
+becomes effectively almost independent on $Z$, as shown in the lower plots of
+Figure~\ref{fig:toy}. In particular, the conditional probability densities of
+the decision scores $f(X;\theta_f)$ are now very similar to each other, indicating only a
+small  dependency on the nuisance, as theoretically expected.
+\glnote{Discuss Figure~\ref{fig:toy-training}}
 

### 2016-09-28 tex/paper.tex
save

@@
     \vspace{-1cm}
-    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$
-             and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.
+    \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
+             and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
              Adversarial training was performed for $2000$ iterations, mini-batches of size $M=128$, $K=1$ and $\lambda=100$.}
@@
 generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
-neural network classifier $f$ minimizing $L_f(\theta_f)$ without adversarial
+neural network classifier $f$ minimizing ${\cal L}_f(\theta_f)$ without adversarial
 training. As shown in the upper plots of Figure~\ref{fig:toy}, the resulting
@@
 Figure~\ref{fig:toy}. In particular, the conditional probability densities of
-the decision scores $f(X;\theta_f)$ are now very similar to each other, indicating only a
-small  dependency on the nuisance, as theoretically expected.
-\glnote{Discuss Figure~\ref{fig:toy-training}}
+the decision scores $f(X;\theta_f)$ are now very similar to each other,
+indicating only a small  dependency on the nuisance, as theoretically expected.
+The dynamics of adversarial training is illustrated in
+Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$, ${\cal L}_r$ and
+${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each iteration of
+Algorithm~\ref{alg:adversarial-training}. In the first $200$ iterations, we
+observe that $f$ is worse than random because it rather tries to maximizes the
+adversarial loss ${\cal L}_r$ by producing decision scores $f(X;\theta_f)$ that
+are very unlikely under the current model $p_{\theta_r}$, at the expense of
+making bad predictions with respect to the classification problem. As learning
+goes and $p_{\theta_r}$ becomes more accurate, minimizing $E$ requires making
+predictions that are more accurate, hence decreasing ${\cal L}_f$, or that are
+less dependent on $Z$, hence shaping $p_{\theta_r}$ so that it becomes more
+uniform. Indeed, ${\cal L}_f$ eventually start decreasing, while remaining lower
+bounded by $\min_{\theta_f} {\cal L}_f(\theta_f)$ as approximated by the dashed
+line in the first plot. Similarly,  ${\cal L}_r$ increasingly tends towards the
+differential entropy $H(p_Z) = \log(4)$ of the uniform distribution over ${\cal
+Z}$, as shown by the dashed line in the second plot.
 

### 2016-09-28 tex/paper.tex
save

@@
              and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Adversarial training was performed for $2000$ iterations, mini-batches of size $M=128$, $K=1$ and $\lambda=100$.}
+             Adversarial training was performed for $1000$ iterations, mini-batches of size $M=128$, $K=1$ and $\lambda=10$.}
     \label{fig:toy-training}
@@
 ${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each iteration of
-Algorithm~\ref{alg:adversarial-training}. In the first $200$ iterations, we
-observe that $f$ is worse than random because it rather tries to maximizes the
+Algorithm~\ref{alg:adversarial-training}. In the first iterations, we observe
+that $f$ is worse than random because it rather tries to maximizes the
 adversarial loss ${\cal L}_r$ by producing decision scores $f(X;\theta_f)$ that
@@
 differential entropy $H(p_Z) = \log(4)$ of the uniform distribution over ${\cal
-Z}$, as shown by the dashed line in the second plot.
+Z}$, as shown by the dashed line in the second plot. Finally, let us note that
+the ideal situation of a classifier that is both optimal and pivotal appears to
+be unreachable for this problem, as shown in third plot by the almost constant
+offset between ${\cal L}_f - \lambda {\cal L}_r$ and the dashed line
+approximating $\mathbb{E}_{X} [ H(p_{Y|X}) - \lambda H(p_Z)]$.
 

### 2016-09-28 tex/paper.tex
save

@@
 
+\glnote{Discuss practical considerations regarding the stability of training. A
+better alternative seems to update $r$ until it is good
+enough, then update $f$, and update $r$ again as soon as it becomes too bad.
+This is important e.g. for MDN to prevent weights to explode if the network is
+trained too much, and in which case gradients become useless.} \glnote{Warn against
+setting $\lambda$ too high, since it could simply force $f$ to produce unlikely
+and very bad decisions.}
+
 \subsection{Physics example}

### 2016-09-28 tex/paper.tex
better toy

@@
     \caption{Toy example.
-    (Upper left) Conditional probability densities of the decision scores for $Z=0, 1, \dots, 4$
+    (Upper left) Conditional probability densities of the decision scores for $Z=0, 1, 2$
        when $f$ is built without adversarial training. The resulting densities
@@
     (Upper right) The corresponding decision surface, highlighting
-       the fact that samples are easier to classify for values of $Z$ closer to $4$,
+       the fact that samples are easier to classify for values of $Z$ closer to $2$,
        hence partially explaining the dependency.
-    (Lower left) Conditional probability densities of the decision scores for $Z=0, 1, \dots, 4$ when $f$ is
+    (Lower left) Conditional probability densities of the decision scores for $Z=0, 1, 2$ when $f$ is
        built with adversarial training, as outlined in Section~\ref{sec:method}.
-       The resulting densities are now almost identical to each other, indicating only a
+       The resulting densities are now much closer to each other, indicating only a
        small dependency on $Z$.
@@
              and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Adversarial training was performed for $1000$ iterations, mini-batches of size $M=128$, $K=1$ and $\lambda=10$.}
+             Adversarial training was performed for $300$ iterations, mini-batches of size $M=128$, $K=10$ and $\lambda=10$.}
     \label{fig:toy-training}
@@
 \begin{align}
-    x &\sim {\cal N}(\mu=(0,0), \sigma=2 \times \mathbb{1}) \text{ when } Y=0, \\
-    x &\sim {\cal N}(\mu=(1,-1+z), \sigma=2 \times \mathbb{1}) \text{ when } Y=1.
+    x &\sim {\cal N}(\mu=(0,0), \Sigma= \mathbb{1}) \text{ when } Y=0, \\
+    x &\sim {\cal N}(\mu=(1,z), \Sigma= \mathbb{1}) \text{ when } Y=1.
 \end{align}
@@
 
-Assuming a uniform prior over the interval ${\cal Z} = [0;4]$, we start by
+Assuming a uniform prior over the interval ${\cal Z} = [0;2]$, we start by
 generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
@@
 
-By contrast, when using Algorithm~\ref{alg:adversarial-training} for training
-$f$ jointly with a mixture density network $r$ modeling the conditional
-distribution of the nuisance as observed through $f$, the resulting classifier
-becomes effectively almost independent on $Z$, as shown in the lower plots of
-Figure~\ref{fig:toy}. In particular, the conditional probability densities of
+By contrast, when running Algorithm~\ref{alg:adversarial-training} initialized
+from the classifier obtained previously, in combination with a mixture density
+network $r$ (modeling $Y|f(X;\theta_f)$ as a truncated mixture of four gaussian
+components), adversarial training effectively reshapes the decision function
+$f$, so it that becomes almost independent on $Z$, as shown in the lower plots
+of Figure~\ref{fig:toy}. In particular, the conditional probability densities of
 the decision scores $f(X;\theta_f)$ are now very similar to each other,
@@
 Algorithm~\ref{alg:adversarial-training}. In the first iterations, we observe
-that $f$ is worse than random because it rather tries to maximizes the
-adversarial loss ${\cal L}_r$ by producing decision scores $f(X;\theta_f)$ that
-are very unlikely under the current model $p_{\theta_r}$, at the expense of
-making bad predictions with respect to the classification problem. As learning
-goes and $p_{\theta_r}$ becomes more accurate, minimizing $E$ requires making
-predictions that are more accurate, hence decreasing ${\cal L}_f$, or that are
-less dependent on $Z$, hence shaping $p_{\theta_r}$ so that it becomes more
-uniform. Indeed, ${\cal L}_f$ eventually start decreasing, while remaining lower
-bounded by $\min_{\theta_f} {\cal L}_f(\theta_f)$ as approximated by the dashed
-line in the first plot. Similarly,  ${\cal L}_r$ increasingly tends towards the
-differential entropy $H(p_Z) = \log(4)$ of the uniform distribution over ${\cal
-Z}$, as shown by the dashed line in the second plot. Finally, let us note that
-the ideal situation of a classifier that is both optimal and pivotal appears to
-be unreachable for this problem, as shown in third plot by the almost constant
-offset between ${\cal L}_f - \lambda {\cal L}_r$ and the dashed line
-approximating $\mathbb{E}_{X} [ H(p_{Y|X}) - \lambda H(p_Z)]$.
+that the global objective ${\cal L}_f - \lambda {\cal L}_r$ is minimized by
+making the classifier less optimal, hence the corresponding increase of ${\cal
+L}_f$, but which results in a classifier that is more pivotal, hence the
+corresponding increase of ${\cal L}_r$ and the total net benefit. As learning
+goes, minimizing $E$ requires making predictions that are more accurate, hence
+decreasing ${\cal L}_f$, or that are less dependent on $Z$, hence shaping
+$p_{\theta_r}$ so that it becomes more uniform. Indeed, ${\cal L}_f$ eventually
+start to slighlty decrease, while remaining lower bounded by $\min_{\theta_f}
+{\cal L}_f(\theta_f)$ as approximated by the dashed line in the first plot.
+Similarly,  ${\cal L}_r$ keeps tending towards the differential entropy $H(p_Z) =
+\log(2)$ of the uniform distribution over ${\cal Z}$, as shown by the dashed
+line in the second plot. Finally, let us note that the ideal situation of a
+classifier that is both optimal and pivotal appears to be unreachable for this
+problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
+{\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H(p_{Y|X}) -
+\lambda H(p_Z)]$.
 

### 2016-09-28 tex/paper.tex
save

@@
 
-By contrast, when running Algorithm~\ref{alg:adversarial-training} initialized
-from the classifier obtained previously, in combination with a mixture density
-network $r$ (modeling $Y|f(X;\theta_f)$ as a truncated mixture of four gaussian
+By contrast, when running Algorithm~\ref{alg:adversarial-training} as initialized
+with the classifier obtained previously, and in combination with a mixture density
+network $r$ (modeling $Z|f(X;\theta_f)$ as a truncated mixture of four gaussian
 components), adversarial training effectively reshapes the decision function
-$f$, so it that becomes almost independent on $Z$, as shown in the lower plots
+$f$ so it that becomes almost independent on $Z$, as shown in the lower plots
 of Figure~\ref{fig:toy}. In particular, the conditional probability densities of
@@
 
-\glnote{Discuss practical considerations regarding the stability of training. A
-better alternative seems to update $r$ until it is good
-enough, then update $f$, and update $r$ again as soon as it becomes too bad.
-This is important e.g. for MDN to prevent weights to explode if the network is
-trained too much, and in which case gradients become useless.} \glnote{Warn against
+\glnote{Discuss practical considerations regarding the stability of training.}
+\glnote{Warn against
 setting $\lambda$ too high, since it could simply force $f$ to produce unlikely

### 2016-09-29 tex/paper.tex
save

@@
        built with adversarial training, as outlined in Section~\ref{sec:method}.
-       The resulting densities are now much closer to each other, indicating only a
+       The resulting densities are now almost identical to each other, indicating only a
        small dependency on $Z$.
@@
              and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Adversarial training was performed for $300$ iterations, mini-batches of size $M=128$, $K=10$ and $\lambda=10$.}
+             Adversarial training was performed for $75$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
     \label{fig:toy-training}
@@
 
-By contrast, when running Algorithm~\ref{alg:adversarial-training} as initialized
-with the classifier obtained previously, and in combination with a mixture density
-network $r$ (modeling $Z|f(X;\theta_f)$ as a truncated mixture of four gaussian
-components), adversarial training effectively reshapes the decision function
-$f$ so it that becomes almost independent on $Z$, as shown in the lower plots
-of Figure~\ref{fig:toy}. In particular, the conditional probability densities of
-the decision scores $f(X;\theta_f)$ are now very similar to each other,
-indicating only a small  dependency on the nuisance, as theoretically expected.
-The dynamics of adversarial training is illustrated in
+Let us now consider the adversarial training of $f$ against an adversary $r$
+implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a
+truncated mixture of four gaussians over ${\cal Z}$. When running
+Algorithm~\ref{alg:adversarial-training} as initialized with the classifier $f$
+obtained previously, adversarial training effectively reshapes the decision
+function so it that becomes almost independent on the nuisance parameter, as
+shown in the lower plots of Figure~\ref{fig:toy}. In particular, the conditional
+probability densities of the decision scores $f(X;\theta_f)$ are now very
+similar to each other, indicating only a small  dependency on the nuisance, as
+theoretically expected. The dynamics of adversarial training is illustrated in
 Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$, ${\cal L}_r$ and
@@
 that the global objective ${\cal L}_f - \lambda {\cal L}_r$ is minimized by
-making the classifier less optimal, hence the corresponding increase of ${\cal
+making the classifier less accurate, hence the corresponding increase of ${\cal
 L}_f$, but which results in a classifier that is more pivotal, hence the
 corresponding increase of ${\cal L}_r$ and the total net benefit. As learning
-goes, minimizing $E$ requires making predictions that are more accurate, hence
-decreasing ${\cal L}_f$, or that are less dependent on $Z$, hence shaping
+goes, minimizing $E$ then requires making predictions that are more accurate,
+hence decreasing ${\cal L}_f$, or that are less dependent on $Z$, hence shaping
 $p_{\theta_r}$ so that it becomes more uniform. Indeed, ${\cal L}_f$ eventually
-start to slighlty decrease, while remaining lower bounded by $\min_{\theta_f}
+start to slightly decrease, while remaining lower bounded by $\min_{\theta_f}
 {\cal L}_f(\theta_f)$ as approximated by the dashed line in the first plot.

### 2016-09-29 tex/paper.tex
save

@@
     (Lower right) The corresponding decision surface, illustrating how adversarial
-       training bends the decision function to erase the dependency on $Z$.
+       training bends the decision function vertically to erase the dependency on $Z$.
     }
@@
              and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Adversarial training was performed for $75$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
+             Adversarial training was performed for $100$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
     \label{fig:toy-training}

### 2016-09-29 tex/paper.tex
save

@@
              and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Adversarial training was performed for $100$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
+             Adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
     \label{fig:toy-training}
@@
 \lambda H(p_Z)]$.
-
-\glnote{Discuss practical considerations regarding the stability of training.}
-\glnote{Warn against
-setting $\lambda$ too high, since it could simply force $f$ to produce unlikely
-and very bad decisions.}
 

### 2016-09-29 tex/paper.tex
first draft of abstract (#6)
* first draft of abstract

* now updating pdf after merge

@@
 \begin{abstract}
-
-Nice and cozy abstract goes here...
+Systematic uncertainties in high energy physics, representing
+incomplete knowledge in the simulation of physical processes or
+detection appartus, may cause variations in the class conditional
+distributions of features used for training a classifier.  As such,
+the efficiacy of a classifier may change under the variations caused
+by systematic uncertainties.  In this paper, we propose an adversarial
+training procedure to produce a pivotal
+classifier which is invariant to variations of nuisance parameters
+which parameterize the effects of systematic uncertainties.
+In the machine learning context, this problem is closely related to those of domain
+adaptation and enforcing fairness in classification.  The techniques
+described here can handle the case of continuous nuisance paramater
+variations that represents domain adaptation over infinitely many domains.
 

### 2016-09-29 tex/paper.tex
save

--- 2016-09-29 tex/paper.tex
@@
 
-Nice and cozy abstract goes here...
+Systematic uncertainties in high energy physics, representing our incomplete
+knowledge in the simulation of physical processes or in the detection apparatus,
+may cause variations in the conditional distributions of features used for
+training a classifier.  As such, the efficacy of a classifier may change under
+the variations caused by systematic uncertainties.  In this paper, we repurpose
+adversarial training as a means to produce a pivotal classifier which is
+invariant to variations of the nuisance parameters which parameterize the
+effects of systematic uncertainties. In the machine learning context, this
+problem is closely related to those of domain adaptation and enforcing fairness
+in classification. Following that line of work, the method proposed in this work
+can be regarded as generalization that also supports the continuous case,
+which can be viewed as handling infinitely many domains or as enforcing fairness
+over continuous attributes.
 
@@
 after application of the discriminant model and thereby increase the sensitivity
-of the analysis to signals from potential new physics sources. \glnote{Shall we
-be more explicit somewhere and put a formula underlining the effect of having a
-pivotal classifier on the resulting sensitivity? (versus a classifier whose
-decision function depends on the nuisance.)}
+of the analysis to signals from potential new physics sources.
 

### 2016-09-29 tex/paper.tex
save

@@
 adversarial training as a means to produce a pivotal classifier which is
-invariant to variations of the nuisance parameters which parameterize the
+invariant to variations of the nuisance parameters describing the
 effects of systematic uncertainties. In the machine learning context, this
@@
     \caption{Toy example.
-    (Upper left) Conditional probability densities of the decision scores for $Z=0, 1, 2$
+    (Upper left) Conditional probability densities of the decision scores at $Z=0, 1, 2$
        when $f$ is built without adversarial training. The resulting densities
-       are clearly dependent on $Z$, indicating that $f$ is not pivotal.
+       are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
     (Upper right) The corresponding decision surface, highlighting
@@
        hence partially explaining the dependency.
-    (Lower left) Conditional probability densities of the decision scores for $Z=0, 1, 2$ when $f$ is
+    (Lower left) Conditional probability densities of the decision scores at $Z=0, 1, 2$ when $f$ is
        built with adversarial training, as outlined in Section~\ref{sec:method}.

### 2016-09-29 tex/paper.tex
added a few references

@@
 for the signal or a given background. Systematic uncertainties can
-affect this distribution in the following ways \glnote{add references}:
+affect this distribution in the following ways \citep{sinervo2003definition,bohm2010introduction}:
 \begin{itemize}
@@
 Systematic uncertainties are parameterized by nuisance parameters in the
-statistical analysis of data \glnote{add references}.  Nuisance parameter often,
+statistical analysis of data (see e.g., \citep{bohm2010introduction,cowan2011asymptotic}).  Nuisance parameter often,
 though not always, have a suitable prior coming from alternative measurements in

### 2016-09-29 tex/paper.tex
save

@@
 Systematic uncertainties are parameterized by nuisance parameters in the
-statistical analysis of data (see e.g., \citep{bohm2010introduction,cowan2011asymptotic}).  Nuisance parameter often,
+statistical analysis of data (see e.g.,
+\citep{bohm2010introduction,cowan2011asymptotic}).  Nuisance parameters often,
 though not always, have a suitable prior coming from alternative measurements in
@@
 always, constrained by a suitable prior, e.g. a Gaussian distribution with a
-standard deviation equal to the measured uncertainty. \glnote{... and
-accordingly, we want to build a classifier independent on the nuisance
-parameter, as a way to be independent of the corresponding systematics.}
+standard deviation equal to the measured uncertainty. \glnote{This paragraph
+could be improved, it is not very clear what we want to say...}
 
@@
 value $z_0$ of the nuisance parameter, and then propagate uncertainty
-\glnote{add ref} by estimating $p(f(x)|z)$ with a parameterized calibration
+ by estimating $p(f(x)|z)$ with a parameterized calibration
 procedure. Clearly, this classifier is however not optimal for $z \neq z_0$.
-To overcome this issue, the classifier $f$ is sometimes \glnote{add refs} built instead on a mixture
+To overcome this issue, the classifier $f$ is sometimes built instead on a mixture
 of training data generated from several nominal values $z_0, z_1, \dots$ of the nuisance.

### 2016-09-29 tex/paper.tex
save

@@
 
-Systematic uncertainties in high energy physics, representing our incomplete
-knowledge in the simulation of physical processes or in the detection apparatus,
-may cause variations in the conditional distributions of features used for
-training a classifier.  As such, the efficacy of a classifier may change under
-the variations caused by systematic uncertainties.  In this paper, we repurpose
-adversarial training as a means to produce a pivotal classifier which is
-invariant to variations of the nuisance parameters describing the
-effects of systematic uncertainties. In the machine learning context, this
-problem is closely related to those of domain adaptation and enforcing fairness
-in classification. Following that line of work, the method proposed in this work
-can be regarded as generalization that also supports the continuous case,
+Systematic uncertainties in high energy physics represent our incomplete
+knowledge in the simulation of physical processes or in the properties of the
+detection apparatus. In effect, systematic uncertainties may cause variations in
+the conditional distributions of the features used for training a predictive
+model like a classifier, thereby altering its decisions and resulting efficacy.
+In this paper, we repurpose adversarial training as a means to produce a pivotal
+classifier which is invariant to variations of the nuisance parameters
+describing the effects of systematic uncertainties. In the machine learning
+context, this problem is closely related to those of domain adaptation and
+enforcing fairness in classification. Following that line of work, the proposed
+method can be regarded as generalization that also supports the continuous case,
 which can be viewed as handling infinitely many domains or as enforcing fairness
-over continuous attributes.
+over continuous features.
 
@@
 on mitigating the effect of systematic uncertainties that alter the
-conditional distributions $p(X|Y)$.
+conditional distributions $p(X|Y)$. \glnote{Remove this comment?}
 

### 2016-09-29 tex/paper.tex
save

@@
 
-Systematic uncertainties in high energy physics represent our incomplete
+In high energy physics, systematic uncertainties represent our incomplete
 knowledge in the simulation of physical processes or in the properties of the
-detection apparatus. In effect, systematic uncertainties may cause variations in
-the conditional distributions of the features used for training a predictive
-model like a classifier, thereby altering its decisions and resulting efficacy.
-In this paper, we repurpose adversarial training as a means to produce a pivotal
-classifier which is invariant to variations of the nuisance parameters
-describing the effects of systematic uncertainties. In the machine learning
-context, this problem is closely related to those of domain adaptation and
-enforcing fairness in classification. Following that line of work, the proposed
-method can be regarded as generalization that also supports the continuous case,
-which can be viewed as handling infinitely many domains or as enforcing fairness
-over continuous features.
+detection apparatus. In effect, these uncertainties typically cause variations
+in the conditional distributions of the data, thereby altering the decisions
+and resulting efficacy of a classifier trained from it. In this work, we
+repurpose adversarial training as a means to learn a pivotal classifier which is
+invariant to variations of the nuisance parameters describing the effects of
+systematic uncertainties. In machine learning, this problem is closely related
+to those of domain adaptation and enforcing fairness in classification.
+Following that line of work, the proposed method can be regarded as
+generalization that also supports the continuous case, which can be viewed as
+handling infinitely many domains or as enforcing fairness over continuous
+features.
 

### 2016-09-29 tex/paper.tex
save

@@
 detection apparatus. In effect, these uncertainties typically cause variations
-in the conditional distributions of the data, thereby altering the decisions
-and resulting efficacy of a classifier trained from it. In this work, we
+in the conditional distributions of the data, thereby altering the decisions and
+resulting efficacy of a classifier trained from it. To accomodate for this, we
 repurpose adversarial training as a means to learn a pivotal classifier which is
 invariant to variations of the nuisance parameters describing the effects of
-systematic uncertainties. In machine learning, this problem is closely related
-to those of domain adaptation and enforcing fairness in classification.
-Following that line of work, the proposed method can be regarded as
-generalization that also supports the continuous case, which can be viewed as
-handling infinitely many domains or as enforcing fairness over continuous
-features.
+systematic uncertainties. In particular, we show and derive theoretical
+conditions under which classification optimality and invariance with respect to
+systematics can be achieved, which we confirm experimentally on a couple of
+illustrative examples. In machine learning, this problem is closely related to
+those of domain adaptation and enforcing fairness in classification. Following
+that line of work, the proposed method can be regarded as generalization that
+also supports the continuous case, which can be viewed as handling infinitely
+many domains or as enforcing fairness over continuous features.
 

### 2016-09-29 tex/paper.tex
save

@@
 In high energy physics, systematic uncertainties represent our incomplete
-knowledge in the simulation of physical processes or in the properties of the
-detection apparatus. In effect, these uncertainties typically cause variations
-in the conditional distributions of the data, thereby altering the decisions and
-resulting efficacy of a classifier trained from it. To accomodate for this, we
-repurpose adversarial training as a means to learn a pivotal classifier which is
-invariant to variations of the nuisance parameters describing the effects of
-systematic uncertainties. In particular, we show and derive theoretical
-conditions under which classification optimality and invariance with respect to
-systematics can be achieved, which we confirm experimentally on a couple of
-illustrative examples. In machine learning, this problem is closely related to
-those of domain adaptation and enforcing fairness in classification. Following
-that line of work, the proposed method can be regarded as generalization that
-also supports the continuous case, which can be viewed as handling infinitely
-many domains or as enforcing fairness over continuous features.
+knowledge in the theory of physical processes or in the properties of the
+experimental detection apparatus. In effect, these uncertainties typically cause
+variations in the conditional distributions of the data, thereby altering the
+decisions and resulting efficacy of a classifier trained from it. To contain
+this issue, we propose to repurpose adversarial training as a means to learn a
+pivotal classifier which is invariant to variations of the nuisance parameters
+describing the effects of systematic uncertainties. In particular, we show and
+derive theoretical conditions under which classification optimality and
+invariance with respect to systematics can be achieved, which we confirm
+experimentally on a couple of illustrative examples. In machine learning, this
+problem is closely related to those of domain adaptation and enforcing fairness
+in classification. Following that line of work, the proposed method can be
+regarded as generalization that also supports the continuous case, which can be
+viewed as handling infinitely many domains or as enforcing fairness over
+continuous features.
 

### 2016-10-05 tex/paper.tex
\theta^i -> \gamma_i

@@
 
-    \node[above] at (12.75,-2) {$\theta^1(f(X;\theta_f);\theta_r)$};
+    \node[above] at (12.75,-2) {$\gamma_1(f(X;\theta_f);\theta_r)$};
     \path[-o, shorten >= 0pt] (11,-2.0) edge (14,-2.0);
-    \node[above] at (12.75,-3) {$\theta^2(f(X;\theta_f);\theta_r)$};
+    \node[above] at (12.75,-3) {$\gamma_2(f(X;\theta_f);\theta_r)$};
     \path[-o, shorten >= 0pt] (11,-3) edge (14,-3);
@@
     \node[below] at (15.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
-    \node at (15.5,-3) {${\cal P}(\theta^1, \theta^2, \dots)$};
+    \node at (15.5,-3) {${\cal P}(\gamma_1, \gamma_2, \dots)$};
 
@@
 represented e.g. as network whose output $j$ is the estimated value of the
-corresponding parameter $\theta^j$ of that distribution (e.g., the mean,
+corresponding parameter $\gamma_j$ of that distribution (e.g., the mean,
 variance and mixing coefficients of its components). As in

### 2016-10-09 tex/paper.tex
save
toy: break symmetry, hence avoiding Z=0 to be pivotal
add details about network architectures

@@
              and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
+             Adversarial training was performed for $100$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=100$.}
     \label{fig:toy-training}
@@
 \begin{align}
-    x &\sim {\cal N}(\mu=(0,0), \Sigma= \mathbb{1}) \text{ when } Y=0, \\
-    x &\sim {\cal N}(\mu=(1,z), \Sigma= \mathbb{1}) \text{ when } Y=1.
+    x &\sim {\cal N}((0,0), \begin{bmatrix}
+                              1 & -0.5 \\
+                              -0.5 & 1
+                            \end{bmatrix}) \text{ when } Y=0, \\
+    x &\sim {\cal N}((1,Z),  \begin{bmatrix}
+                              1 & 0 \\
+                              0 & 1
+                             \end{bmatrix}) \text{ when } Y=1.
 \end{align}
@@
 generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
-neural network classifier $f$ minimizing ${\cal L}_f(\theta_f)$ without adversarial
-training. As shown in the upper plots of Figure~\ref{fig:toy}, the resulting
-classifier is not pivotal, as the conditional probability densities of its
-decision scores $f(X;\theta_f)$ show large discrepancies between values $z$ of
-the nuisance. While not shown here, a classifier trained only from data
-generated at the nominal value $Z=0$ would also not be pivotal.
+neural network classifier $f$ minimizing ${\cal L}_f(\theta_f)$ without
+adversarial training. The network architecture comprises 2 dense hidden layers of 20
+nodes with ReLU activations, followed  by a dense output layer with a single
+node with a sigmoid activation. As shown in the upper plots of
+Figure~\ref{fig:toy}, the resulting classifier is not pivotal, as the
+conditional probability densities of its decision scores $f(X;\theta_f)$ show
+large discrepancies between values $z$ of the nuisance. While not shown here, a
+classifier trained only from data generated at the nominal value $Z=0$ would
+also not be pivotal.
 
@@
 implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a
-truncated mixture of four gaussians over ${\cal Z}$. When running
+truncated mixture of four gaussians over ${\cal Z}$. As for $f$, the network
+architecture comprises 2 dense hidden layers of 20 nodes with ReLU activations,
+but is followed by an output layer of 12 nodes corresponding to the means,
+standard deviations and mixture coefficients of the four gaussians. Output nodes
+for the mean values come with tanh activations rescaled over ${\cal Z}$, nodes
+for the standard deviations with exponential activations to ensure positivity,
+while nodes for the mixture coefficients implement the softmax function to
+ensure positivity and normalization. When running
 Algorithm~\ref{alg:adversarial-training} as initialized with the classifier $f$

### 2016-10-20 tex/paper.tex
save
add reference

@@
 representation of the data. Likewise, our method also relates to the problem of
-enforcing fairness in classification \citep{zemel2013learning,EdwardsS15}, which
+enforcing fairness in classification \citep{zemel2013learning,EdwardsS15,zafar2015fairness}, which
 is stated as learning a classifier that is independent of some chosen attribute
@@
 parameterized, or as enforcing fairness over continuous attributes.
-
-\glnote{Check related work of cited references to see if important related work are missing.}
 

### 2016-10-20 tex/paper.tex
update toy example

@@
     \caption{Toy example.
-    (Upper left) Conditional probability densities of the decision scores at $Z=0, 1, 2$
+    (Upper left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
        when $f$ is built without adversarial training. The resulting densities
@@
     (Upper right) The corresponding decision surface, highlighting
-       the fact that samples are easier to classify for values of $Z$ closer to $2$,
+       the fact that samples are easier to classify for values of $Z$ above to $\sigma$,
        hence partially explaining the dependency.
-    (Lower left) Conditional probability densities of the decision scores at $Z=0, 1, 2$ when $f$ is
+    (Lower left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
        built with adversarial training, as outlined in Section~\ref{sec:method}.
@@
              and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Adversarial training was performed for $100$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=100$.}
+             Adversarial training was performed for $100$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
     \label{fig:toy-training}
@@
                             \end{bmatrix}) \text{ when } Y=0, \\
-    x &\sim {\cal N}((1,Z),  \begin{bmatrix}
+    x &\sim {\cal N}((1,1+Z),  \begin{bmatrix}
                               1 & 0 \\
@@
 
-Assuming a uniform prior over the interval ${\cal Z} = [0;2]$, we start by
+Assuming a gaussian prior $z \sim {\cal N}(0,1)$ of unit variance, we start by
 generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
@@
 implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a
-truncated mixture of four gaussians over ${\cal Z}$. As for $f$, the network
+ mixture of five gaussians. As for $f$, the network
 architecture comprises 2 dense hidden layers of 20 nodes with ReLU activations,
-but is followed by an output layer of 12 nodes corresponding to the means,
-standard deviations and mixture coefficients of the four gaussians. Output nodes
-for the mean values come with tanh activations rescaled over ${\cal Z}$, nodes
+but is followed by an output layer of 15 nodes corresponding to the means,
+standard deviations and mixture coefficients of the five gaussians. Output nodes
+for the mean values come with linear activations, nodes
 for the standard deviations with exponential activations to ensure positivity,
@@
 Similarly,  ${\cal L}_r$ keeps tending towards the differential entropy $H(p_Z) =
-\log(2)$ of the uniform distribution over ${\cal Z}$, as shown by the dashed
+\log(\sigma \sqrt{2 \pi e}) = 1.419$ of the gaussian distribution with unit variance, as shown by the dashed
 line in the second plot. Finally, let us note that the ideal situation of a

### 2016-10-20 tex/paper.tex
simpler notation for entropy

@@
 either categorical values or continuous values within a bounded support. By
-abuse of notation, $H(p_Z)$ denotes the differential entropy in this latter
+abuse of notation, $H(Z)$ denotes the differential entropy in this latter
 case. Finally, propositions below are derived in a non-parametric setting, by
@@
 for Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar} such that
-$E(\hat\theta_f, \hat\theta_r) = \mathbb{E}_{X} [H(p_{Y|X}) - H(p_Z)]$, then
+$E(\hat\theta_f, \hat\theta_r) = \mathbb{E}_{X} [H({Y|X}) - H(Z)]$, then
 $f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier and a pivotal
@@
 p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the expected entropy
-$\mathbb{E}_{X} [ H(p_{Z|f(X;\theta_f)}) ]$ of the conditional distribution of the nuisance. The
+$\mathbb{E}_{X} [ H({Z|f(X;\theta_f)}) ]$ of the conditional distribution of the nuisance. The
 value function $E$ can therefore be rewritten as $$E'(\theta_f) = {\cal
-L}_f(\theta_f) - \mathbb{E}_{X} [ H(p_{Z|f(X;\theta_f)}) ].$$  In particular, we have the lower
-bound $$\mathbb{E}_{X} [ H(p_{Y|X}) - H(p_Z)] \leq {\cal L}_f(\theta_f) - \mathbb{E}_{X} [H(p_{Z|f(X;\theta_f)})]$$
+L}_f(\theta_f) - \mathbb{E}_{X} [ H({Z|f(X;\theta_f)}) ].$$  In particular, we have the lower
+bound $$\mathbb{E}_{X} [ H({Y|X}) - H(Z)] \leq {\cal L}_f(\theta_f) - \mathbb{E}_{X} [H({Z|f(X;\theta_f)})]$$
 where the equality holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f}
@@
     of an optimal classifier and in which case ${\cal L}_f$ reduces to its
-    minimum value $\mathbb{E}_{X} [H(p_{Y|X})]$,
+    minimum value $\mathbb{E}_{X} [H({Y|X})]$,
 
     \item $\smash{\hat\theta}_f$ maximizes the expected entropy
-    $\mathbb{E}_{X} [H(p_{Z|f(X;\theta_f)})]$, which happens when all outcomes of
+    $\mathbb{E}_{X} [H({Z|f(X;\theta_f)})]$, which happens when all outcomes of
     $Z|f(X;\smash{\hat\theta}_f)$ are equally likely. In other words,
@@
 its optimum given $f$ (e.g., by setting $K$ sufficiently high) and if $f$ is
-updated to improve ${\cal L}_f(\theta_f) - \mathbb{E}_X [ H(p_{Z|f(X;\theta_f)}) ]$, then $f$
+updated to improve ${\cal L}_f(\theta_f) - \mathbb{E}_X [ H({Z|f(X;\theta_f)}) ]$, then $f$
 should converge to a classifier which is both optimal and pivotal, provided such
@@
 not exist because the nuisance parameter directly shapes the decision boundary,
-in which cases the lower bound $\mathbb{E}_{X} [ H(p_{Y|X}) - H(p_Z)] < {\cal L}_f(\theta_f) - \mathbb{E}_{X} [H(p_{Z|f(X;\theta_f)})]$ is strict: $f$ can either be an optimal classifier or a
+in which cases the lower bound $\mathbb{E}_{X} [ H({Y|X}) - H(Z)] < {\cal L}_f(\theta_f) - \mathbb{E}_{X} [H({Z|f(X;\theta_f)})]$ is strict: $f$ can either be an optimal classifier or a
 pivotal quantity, but not both simultaneously. In this situation, it is natural
@@
 {\cal L}_f(\theta_f)$ as approximated by the dashed line in the first plot.
-Similarly,  ${\cal L}_r$ keeps tending towards the differential entropy $H(p_Z) =
+Similarly,  ${\cal L}_r$ keeps tending towards the differential entropy $H(Z) =
 \log(\sigma \sqrt{2 \pi e}) = 1.419$ of the gaussian distribution with unit variance, as shown by the dashed
@@
 problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
-{\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H(p_{Y|X}) -
-\lambda H(p_Z)]$.
+{\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H({Y|X}) -
+\lambda H(Z)]$.
 

### 2016-10-20 tex/paper.tex
update toy

@@
 if we assume some parameteric distribution ${\cal P}$ for $Z|f(X;\theta_f)$
-(e.g., a mixture of gaussians over a bounded support, as modeled with a mixture
+(e.g., a mixture of gaussians, as modeled with a mixture
 density network~\citep{bishop1994mixture}), then $p_{\theta_r}$ can be
@@
                               -0.5 & 1
-                            \end{bmatrix}) \text{ when } Y=0, \\
+                            \end{bmatrix}) &\text{ when } Y=0, \\
     x &\sim {\cal N}((1,1+Z),  \begin{bmatrix}
@@
                               0 & 1
-                             \end{bmatrix}) \text{ when } Y=1.
+                             \end{bmatrix}) &\text{ when } Y=1.
 \end{align}
@@
 Let us now consider the adversarial training of $f$ against an adversary $r$
-implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a
- mixture of five gaussians. As for $f$, the network
-architecture comprises 2 dense hidden layers of 20 nodes with ReLU activations,
-but is followed by an output layer of 15 nodes corresponding to the means,
-standard deviations and mixture coefficients of the five gaussians. Output nodes
-for the mean values come with linear activations, nodes
-for the standard deviations with exponential activations to ensure positivity,
-while nodes for the mixture coefficients implement the softmax function to
-ensure positivity and normalization. When running
-Algorithm~\ref{alg:adversarial-training} as initialized with the classifier $f$
-obtained previously, adversarial training effectively reshapes the decision
-function so it that becomes almost independent on the nuisance parameter, as
-shown in the lower plots of Figure~\ref{fig:toy}. In particular, the conditional
-probability densities of the decision scores $f(X;\theta_f)$ are now very
-similar to each other, indicating only a small  dependency on the nuisance, as
-theoretically expected. The dynamics of adversarial training is illustrated in
-Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$, ${\cal L}_r$ and
-${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each iteration of
-Algorithm~\ref{alg:adversarial-training}. In the first iterations, we observe
-that the global objective ${\cal L}_f - \lambda {\cal L}_r$ is minimized by
-making the classifier less accurate, hence the corresponding increase of ${\cal
-L}_f$, but which results in a classifier that is more pivotal, hence the
-corresponding increase of ${\cal L}_r$ and the total net benefit. As learning
-goes, minimizing $E$ then requires making predictions that are more accurate,
-hence decreasing ${\cal L}_f$, or that are less dependent on $Z$, hence shaping
-$p_{\theta_r}$ so that it becomes more uniform. Indeed, ${\cal L}_f$ eventually
-start to slightly decrease, while remaining lower bounded by $\min_{\theta_f}
-{\cal L}_f(\theta_f)$ as approximated by the dashed line in the first plot.
-Similarly,  ${\cal L}_r$ keeps tending towards the differential entropy $H(Z) =
-\log(\sigma \sqrt{2 \pi e}) = 1.419$ of the gaussian distribution with unit variance, as shown by the dashed
-line in the second plot. Finally, let us note that the ideal situation of a
-classifier that is both optimal and pivotal appears to be unreachable for this
-problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
-{\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H({Y|X}) -
-\lambda H(Z)]$.
+implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a mixture
+of five gaussians. As for $f$, the network architecture comprises 2 dense hidden
+layers of 20 nodes with ReLU activations, but is followed by an output layer of
+15 nodes corresponding to the means, standard deviations and mixture
+coefficients of the five gaussians. Output nodes for the mean values come with
+linear activations, nodes for the standard deviations with exponential
+activations to ensure positivity, while nodes for the mixture coefficients
+implement the softmax function to ensure positivity and normalization. When
+running Algorithm~\ref{alg:adversarial-training} as initialized with the
+classifier $f$ obtained previously, adversarial training effectively reshapes
+the decision function so it that becomes almost independent on the nuisance
+parameter, as shown in the lower plots of Figure~\ref{fig:toy}. In particular,
+the conditional probability densities of the decision scores $f(X;\theta_f)$ are
+now very similar to each other, indicating only a small  dependency on the
+nuisance, as theoretically expected. The dynamics of adversarial training is
+illustrated in Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$,
+${\cal L}_r$ and ${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each
+iteration of Algorithm~\ref{alg:adversarial-training}. In the first iterations,
+we observe that the global objective ${\cal L}_f - \lambda {\cal L}_r$ is
+minimized by making the classifier less accurate, hence the corresponding
+increase of ${\cal L}_f$, but which results in a classifier that is more
+pivotal, hence the corresponding increase of ${\cal L}_r$ and the total net
+benefit. As learning goes, minimizing $E$ then requires making predictions that
+are more accurate, hence decreasing ${\cal L}_f$, or that are less dependent on
+$Z$, hence shaping $p_{\theta_r}$ towards the prior $p(z)$. Indeed, ${\cal L}_f$
+eventually starts to slightly decrease, while remaining lower bounded by
+$\min_{\theta_f} {\cal L}_f(\theta_f)$ as approximated by the dashed line in the
+first plot. Similarly,  ${\cal L}_r$ tends towards the differential entropy
+$H(Z)$ of the prior (where $H(Z) = \log(\sigma \sqrt{2 \pi e}) = 1.419$ in the
+case of a gaussian with unit variance), as shown by the dashed line in the
+second plot.
+% Finally, let us note that the ideal situation of a
+% classifier that is both optimal and pivotal appears to be unreachable for this
+% problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
+% {\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H({Y|X}) -
+% \lambda H(Z)]$.
 

### 2016-10-21 tex/paper.tex
update method and theory sections

@@
 parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
-induce together a joint probability distribution $p(X,Y|z)$, where the
-conditional on $z$ denotes $X_z$. For training, let further assume a finite set $\{
+induce together a joint probability distribution $p(X,Y|Z=z)$, where the
+conditional denotes $X_z$. For training, let further assume a finite set $\{
 x_i, y_i, z_i \}_{i=1}^N$ of realizations $X_{z_i}(\omega_i), Y(\omega_i)$, for
 $\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
-is to learn a function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal Y}$ of
-parameters $\theta_f$ (e.g., a neural network-based classifier if ${\cal Y}$ is
-a finite set of classes) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
+is to learn a score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S} \subseteq {\cal Y}$ of
+parameters $\theta_f$ (e.g., a neural network-based probabilistic classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
 the cross-entropy). In addition, we require that $f(X_z ; \theta_f)$ should be
@@
 
-Since we do not have training tuples $(X_{z}(\omega),
+Since in general we do not have training tuples $(X_{z}(\omega),
 X_{z^\prime}(\omega))$ (for the same unknown $\omega$), we propose instead to
@@
 \begin{align}\label{eqn:criterion-measure}
-    &P(\{ \omega | f(X_{z}(\omega) ; \theta_f) = y \} ) \nonumber \\
-    &\indent = P( \{ \omega' | f(X_{z^\prime}(\omega') ; \theta_f) = y\})
+    &P(\{ \omega | f(X_{z}(\omega) ; \theta_f) = s \} ) \nonumber \\
+    &\indent = P( \{ \omega' | f(X_{z^\prime}(\omega') ; \theta_f) = s\})
 \end{align}
-for all $y \in {\cal Y}$. In words, we are looking for a predictive function $f$
+for all $s \in {\cal S}$. In words, we are looking for a predictive function $f$
 which is a pivotal quantity \citep{degroot1986probability} with respect to the
@@
 Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true,
-since the sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = y \}$ and $\{
-\omega' | f(X_{z^\prime}(\omega'); \theta_f) = y \}$ do not need to be the same
+since the sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = s \}$ and $\{
+\omega' | f(X_{z^\prime}(\omega'); \theta_f) = s \}$ do not need to be the same
 for the equality to hold. In order to simplify notations, and as only
@@
 \begin{equation}\label{eqn:criterion}
-    p(f(X ; \theta_f) | z ) = p(f(X ; \theta_f) | z^\prime )
+    p(f(X ; \theta_f) = s | z ) = p(f(X ; \theta_f) = s | z^\prime )
 \end{equation}
-for all $z,z^\prime \in  {\cal Z}$ and all values of $f(X ; \theta_f)$.
+for all $z,z^\prime \in  {\cal Z}$ and all values $s$ of $f(X ; \theta_f)$.
 
@@
 way to build a generative model capable of producing samples from random noise
-$z \sim p_Z$. More specifically, the authors pit a generative model $g:
+$z$. More specifically, the authors pit a generative model $g:
 \mathbb{R} \mapsto \mathbb{R}^p$ against an adversary classifier $d :
-\mathbb{R}^p \mapsto \{ 0, 1\}$ whose antagonistic objective is to recognize
+\mathbb{R}^p \mapsto [0, 1]$ whose antagonistic objective is to recognize
 real data $X$ from generated data $g(Z)$. Both models $g$ and $d$ are trained
@@
 $d$ only by chance. That is, assuming enough capacity in $d$ and  $g$, the
-distribution $p_{g(Z)}$ eventually converges towards the real distribution
-$p_X$.
+distribution of $g(Z)$ eventually converges towards the real distribution
+of $X$.
 
@@
     $f$ against a nuisance parameter $Z$. The adversary $r$ models the
-    distribution $p(Z|f(X;\theta_f))$ of the nuisance as observed only through the output $f(X;\theta_f)$ of the classifier. By
+    distribution $p(z|f(X;\theta_f)=s)$ of the nuisance as observed only through the output $f(X;\theta_f)$ of the classifier. By
     maximizing the antagonistic objective ${\cal L}_r(\theta_f, \theta_r)$ (as part of minimizing ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$), the classifier
-    $f$ forces $p(Z|f(X;\theta_f)$ to become uniform, which happens
+    $f$ forces $p(z|f(X;\theta_f)=s)$ towards the prior $p(z)$, which happens
     when $f(X;\theta_f)$ is  independent of the nuisance parameter $Z$ and therefore pivotal.}
@@
 illustrated in Figure~\ref{fig:architecture}, we pit $f$ against an adversary
-model $r := p_{\theta_r}(z | f(X;\theta_f))$ of parameters $\theta_r$ and
+model $r := p_{\theta_r}(z | f(X;\theta_f)=s)$ of parameters $\theta_r$ and
 associated loss ${\cal L}_r(\theta_f, \theta_r)$. This model takes  as input
-realizations of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$
-parameters, and produces as output a function $p_{\theta_r}(z | f(X;\theta_f))$
-modeling the posterior probability density that $z$ parameterizes the sample $X$
-observed through $f(\cdot;\theta_f)$.
-Intuitively, if $p(f(X; \theta_f)|z)$ varies with $z$,
+realizations $s$ of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$
+parameters, and produces as output a function $p_{\theta_r}(z | f(X;\theta_f)=s)$
+modeling the posterior probability density that $z$ parameterizes the sample
+observed as $s$.
+Intuitively, if $p(f(X; \theta_f)=s|z)$ varies with $z$,
 then the corresponding correlation can be captured by $r$. By contrast, if
-$p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then $r$ should
+$p(f(X; \theta_f)=s|z)$ is invariant with $z$, as we require, then $r$ should
 perform poorly and be close to random guessing. Training $f$ such that it
@@
 output $j$ (for $j=1, \dots, |{\cal Z}|$) is the estimated probability mass
-$p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly, if $Z$ takes continuous values and
-if we assume some parameteric distribution ${\cal P}$ for $Z|f(X;\theta_f)$
+$p_{\theta_r}(z_j|f(X;\theta_f)=s)$. Similarly, if $Z$ takes continuous values and
+if we assume some parameteric distribution ${\cal P}$ for $Z|f(X;\theta_f)=s$
 (e.g., a mixture of gaussians, as modeled with a mixture
@@
 \citep{nix1994estimating,bishop1994mixture}, the estimated probability density
-$p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$.
+$p_{\theta_r}(z|f(X;\theta_f)=s)$ can then be evaluated for any $z \in {\cal Z}$ and any score $s \in {\cal S}$.
 As further explained in the next section, let us note that the adversary $r$ may
 take any form, i.e. it does need to be a neural network, as long as it exposes a
-differentiable function $p_{\theta_r}(z|f(X;\theta_f))$ of sufficient capacity
-to represent the true distribution within its bounded support.
+differentiable function $p_{\theta_r}(z|f(X;\theta_f)=s)$ of sufficient capacity
+to represent the true distribution.
 
@@
 \begin{align}
-    {\cal L}_f(\theta_f) &= \mathbb{E}_{X} \mathbb{E}_{Y|X} [ -\log p_{\theta_f} (Y|X) ], \\
-    {\cal L}_r(\theta_f, \theta_r) &= \mathbb{E}_{X}  \mathbb{E}_{Z|f(X;\theta_f)} [-\log p_{\theta_r} (Z|f(X;\theta_f))].
+    {\cal L}_f(\theta_f) &= \mathbb{E}_{x \sim X}  \mathbb{E}_{y \sim Y|x} [ -\log p_{\theta_f} (y|x) ], \\
+    {\cal L}_r(\theta_f, \theta_r) &= \mathbb{E}_{s \sim f(X;\theta_f)}  \mathbb{E}_{z \sim Z|s} [-\log p_{\theta_r} (z|s)].
 \end{align}
@@
             \For{$k=1$ to $K$} \Comment{Update $r$}
-                \State{Sample minibatch $\{x_m, z_m \}_{m=1}^M$} of size $M$;
+                \State{Sample minibatch $\{x_m, z_m, s_m = f(x_m;\theta_f) \}_{m=1}^M$} of size $M$;
                 \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
-                $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|f(x_m;\theta_f))  ;$$}
+                $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|s_m)  ;$$}
             \EndFor
-            \State{Sample minibatch $\{x_m, y_m, z_m \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
+            \State{Sample minibatch $\{x_m, y_m, z_m, s_m = f(x_m;\theta_f)  \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
             \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
-            $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|f(x_m;\theta_f))  \right],$$
-            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $1(y_m=0)(1-f(x_m;\theta_f)) + 1(y_m=1)f(x_m;\theta_f))$;}
+            $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|s_m)  \right],$$
+            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $1(y_m=0)(1-s_m) + 1(y_m=1)s_m$;}
         \EndFor
@@
 In this setting, the nuisance parameter $Z$ is considered as a random variable
-of bounded support, for which we require the uniform prior $p(z)$ (for $z \in
+of prior $p(z)$ (for $z \in
 {\cal Z}$). Importantly, classification of $Y$ with respect to $X$ is therefore
@@
 abuse of notation, $H(Z)$ denotes the differential entropy in this latter
-case. Finally, propositions below are derived in a non-parametric setting, by
+case. Finally, the  proposition below is derived in a non-parametric setting, by
 assuming that both $f$ and $r$ have enough capacity.
 
-\begin{proposition}\label{prop:1}
-Let $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
-\theta_r)$ for fixed $\theta_f$. If $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z)$
-for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$, then $f$ is a pivotal quantity.
-\end{proposition}
-
-\begin{proof}
-Since we assume enough capacity for $r$, the optimal parameters $$\smash{\hat\theta}_r =
-\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
-\theta_r)$$ are such that $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
-By assumption, $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z)$,
-and therefore $p(z|f(X;\theta_f)) = p(z)$.
-Using the Bayes' rule, we write
-\begin{align*}
-    p(f(X;\theta_f)|z) &= \frac{ p(z|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z)}\\
-                       &= p(f(X;\theta_f)),
-\end{align*}
-which holds for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$ and implies that $f$ is a pivotal quantity.
-\end{proof}
+% \begin{proposition}\label{prop:1}
+% Let $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
+% \theta_r)$ for fixed $\theta_f$. If $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z)$
+% for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$, then $f$ is a pivotal quantity.
+% \end{proposition}
+%
+% \begin{proof}
+% Since we assume enough capacity for $r$, the optimal parameters $$\smash{\hat\theta}_r =
+% \arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
+% \theta_r)$$ are such that $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
+% By assumption, $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z)$,
+% and therefore $p(z|f(X;\theta_f)) = p(z)$.
+% Using the Bayes' rule, we write
+% \begin{align*}
+%     p(f(X;\theta_f)|z) &= \frac{ p(z|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z)}\\
+%                        &= p(f(X;\theta_f)),
+% \end{align*}
+% which holds for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$ and implies that $f$ is a pivotal quantity.
+% \end{proof}
 
@@
 for Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar} such that
-$E(\hat\theta_f, \hat\theta_r) = \mathbb{E}_{X} [H({Y|X}) - H(Z)]$, then
+$E(\hat\theta_f, \hat\theta_r) = H({Y|X}) - H(Z)$, then
 $f(\cdot;\smash{\hat\theta}_f)$ is both an optimal classifier and a pivotal
@@
 
-For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
+For fixed $\theta_f$, the adversary $r$ is optimal at $$\hat\theta_r = \arg
 \max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal
-L}_r(\theta_f, \theta_r)$, in which case $p_{\hat\theta_r}(z|f(X;\theta_f)) =
+L}_r(\theta_f, \theta_r),$$ in which case $p_{\hat\theta_r}(z|f(X;\theta_f)) =
 p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the expected entropy
-$\mathbb{E}_{X} [ H({Z|f(X;\theta_f)}) ]$ of the conditional distribution of the nuisance. The
-value function $E$ can therefore be rewritten as $$E'(\theta_f) = {\cal
-L}_f(\theta_f) - \mathbb{E}_{X} [ H({Z|f(X;\theta_f)}) ].$$  In particular, we have the lower
-bound $$\mathbb{E}_{X} [ H({Y|X}) - H(Z)] \leq {\cal L}_f(\theta_f) - \mathbb{E}_{X} [H({Z|f(X;\theta_f)})]$$
+$\mathbb{E}_{s \sim f(X;\theta_f)} [ H({Z|f(X;\theta_f)=s}) ]$ of the conditional distribution of the nuisance.
+This expectation is nothing else than the conditional entropy of the random variables
+$Z$ and $f(X;\theta_f)$ and can be written as $H(Z|f(X;\theta_f))$.
+Accordingly, the
+value function $E$ can be restated as $$E'(\theta_f) = {\cal
+L}_f(\theta_f) -  H({Z|f(X;\theta_f)}).$$  In particular, we have the lower
+bound $$H({Y|X}) - H(Z) \leq {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})$$
 where the equality holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f}
@@
     of an optimal classifier and in which case ${\cal L}_f$ reduces to its
-    minimum value $\mathbb{E}_{X} [H({Y|X})]$,
+    minimum value $H({Y|X})$,
 
-    \item $\smash{\hat\theta}_f$ maximizes the expected entropy
-    $\mathbb{E}_{X} [H({Z|f(X;\theta_f)})]$, which happens when all outcomes of
-    $Z|f(X;\smash{\hat\theta}_f)$ are equally likely. In other words,
-    $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$ since we
-    require a uniform prior by construction. Note that in the continuous case,
-    the supremum of the differential entropy over continuous distributions on
-    the same bounded support is also realized by the uniform distribution over
-    that support.
+    \item $\smash{\hat\theta}_f$ maximizes the conditional entropy
+    $H({Z|f(X;\theta_f)})$ since $H(Z|f(X;\theta)) \leq H(Z)$.
 \end{itemize}
-When the lower bound is active, we have $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
-because of the second condition, which induces that the
+When the lower bound is active, we have $H(Z|f(X;\theta_f)) = H(Z)$
+because of the second condition, which happens exactly when $Z$ and $f(X;\theta_f)$
+are independent variables. In other words,  the
 optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal
-quantity as a consequence of Proposition~\ref{prop:1}.
+quantity.
 
@@
 its optimum given $f$ (e.g., by setting $K$ sufficiently high) and if $f$ is
-updated to improve ${\cal L}_f(\theta_f) - \mathbb{E}_X [ H({Z|f(X;\theta_f)}) ]$, then $f$
+updated to improve ${\cal L}_f(\theta_f) -  H({Z|f(X;\theta_f)})$, then $f$
 should converge to a classifier which is both optimal and pivotal, provided such
@@
 not exist because the nuisance parameter directly shapes the decision boundary,
-in which cases the lower bound $\mathbb{E}_{X} [ H({Y|X}) - H(Z)] < {\cal L}_f(\theta_f) - \mathbb{E}_{X} [H({Z|f(X;\theta_f)})]$ is strict: $f$ can either be an optimal classifier or a
+in which cases the lower bound $ H({Y|X}) - H(Z) < {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal classifier or a
 pivotal quantity, but not both simultaneously. In this situation, it is natural
@@
 
-Interestingly, let us emphasize that these results hold using only the (1D) output
-of $f$ (in the case of binary classification) as input to the adversary. We
+Interestingly, let us emphasize that these results hold using only the (1D) output $s$
+of $f(\cdot;\theta_f)$ (in the case of binary classification) as input to the adversary. We
 could similarly enforce an intermediate representation of the data to be

### 2016-10-21 tex/paper.tex
save

@@
 In this setting, the nuisance parameter $Z$ is considered as a random variable
-of prior $p(z)$ (for $z \in
-{\cal Z}$). Importantly, classification of $Y$ with respect to $X$ is therefore
+of prior $p(z)$ (for $z \in {\cal Z}$), and our goal is to find a function
+$f(\cdot;\theta_f)$ such that $f(X;\theta_f)$ and $Z$ are independent random
+variables.   Importantly, classification of $Y$ with respect to $X$ is therefore
 considered in the context where $Z$ is marginalized out, which means that the
@@
 either categorical values or continuous values within a bounded support. By
-abuse of notation, $H(Z)$ denotes the differential entropy in this latter
-case. Finally, the  proposition below is derived in a non-parametric setting, by
+abuse of notation, $H(Z)$ denotes the differential entropy in this latter case.
+Finally, the  proposition below is derived in a non-parametric setting, by
 assuming that both $f$ and $r$ have enough capacity.
@@
 \max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal
-L}_r(\theta_f, \theta_r),$$ in which case $p_{\hat\theta_r}(z|f(X;\theta_f)) =
-p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the expected entropy
+L}_r(\theta_f, \theta_r),$$ in which case $p_{\hat\theta_r}(z|f(X;\theta_f)=s) =
+p(z|f(X;\theta_f)=s)$ for all $z$ and all $s$, and ${\cal L}_r$ reduces to the expected entropy
 $\mathbb{E}_{s \sim f(X;\theta_f)} [ H({Z|f(X;\theta_f)=s}) ]$ of the conditional distribution of the nuisance.

### 2016-10-21 tex/paper.tex
save

@@
 Accordingly, the
-value function $E$ can be restated as $$E'(\theta_f) = {\cal
+value function $E$ can be restated as a function depending on $\theta_f$ only: $$E'(\theta_f) = {\cal
 L}_f(\theta_f) -  H({Z|f(X;\theta_f)}).$$  In particular, we have the lower
@@
     \item $\smash{\hat\theta}_f$ maximizes the conditional entropy
-    $H({Z|f(X;\theta_f)})$ since $H(Z|f(X;\theta)) \leq H(Z)$.
+    $H({Z|f(X;\theta_f)})$ since $H(Z|f(X;\theta)) \leq H(Z)$. Note that this
+    latter inequality holds for both the discrete and the differential definitions of entropy.
 \end{itemize}
@@
 second plot.
-% Finally, let us note that the ideal situation of a
-% classifier that is both optimal and pivotal appears to be unreachable for this
-% problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
-% {\cal L}_r$ and the dashed line approximating $\mathbb{E}_{X} [ H({Y|X}) -
-% \lambda H(Z)]$.
+Finally, let us note that the ideal situation of a
+classifier that is both optimal and pivotal appears to be unreachable for this
+problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
+{\cal L}_r$ and the dashed line approximating $H({Y|X}) -
+\lambda H(Z)$.
 

### 2016-10-22 tex/paper.tex
save

@@
 $f(\cdot;\theta_f)$ such that $f(X;\theta_f)$ and $Z$ are independent random
-variables.   Importantly, classification of $Y$ with respect to $X$ is therefore
+variables.   Importantly, classification of $Y$ with respect to $X$ is
 considered in the context where $Z$ is marginalized out, which means that the
@@
 necessarily with $Y|X,Z$. Results hold for a nuisance parameter $Z$ taking
-either categorical values or continuous values within a bounded support. By
+either categorical or continuous values. By
 abuse of notation, $H(Z)$ denotes the differential entropy in this latter case.
@@
 assuming that both $f$ and $r$ have enough capacity.
-
-% \begin{proposition}\label{prop:1}
-% Let $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
-% \theta_r)$ for fixed $\theta_f$. If $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z)$
-% for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$, then $f$ is a pivotal quantity.
-% \end{proposition}
-%
-% \begin{proof}
-% Since we assume enough capacity for $r$, the optimal parameters $$\smash{\hat\theta}_r =
-% \arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
-% \theta_r)$$ are such that $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
-% By assumption, $p_{\hat\theta_r}(z|f(X;\theta_f)) = p(z)$,
-% and therefore $p(z|f(X;\theta_f)) = p(z)$.
-% Using the Bayes' rule, we write
-% \begin{align*}
-%     p(f(X;\theta_f)|z) &= \frac{ p(z|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z)}\\
-%                        &= p(f(X;\theta_f)),
-% \end{align*}
-% which holds for all $z \in {\cal Z}$ and all values of $f(X;\theta_f)$ and implies that $f$ is a pivotal quantity.
-% \end{proof}
 
@@
 should converge to a classifier which is both optimal and pivotal, provided such
-a classifier exists. On many problems of interest though, such a classifier may
+a classifier exists. A formal proof of convergence of the alternating
+stochastic gradient descent procedure of
+Algorithm~\ref{alg:adversarial-training} remains however to be proven.
+
+On many problems of interest, let us also note that such a classifier may
 not exist because the nuisance parameter directly shapes the decision boundary,
@@
 
-Interestingly, let us emphasize that these results hold using only the (1D) output $s$
+Interestingly, let us finally emphasize that these results hold using only the (1D) output $s$
 of $f(\cdot;\theta_f)$ (in the case of binary classification) as input to the adversary. We
@@
              and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=100$ and $\lambda=50$.}
+             Adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=500$ and $\lambda=50$.}
     \label{fig:toy-training}
@@
 the conditional probability densities of the decision scores $f(X;\theta_f)$ are
-now very similar to each other, indicating only a small  dependency on the
+now very similar to each other, indicating only a small residual  dependency on the
 nuisance, as theoretically expected. The dynamics of adversarial training is
@@
 benefit. As learning goes, minimizing $E$ then requires making predictions that
-are more accurate, hence decreasing ${\cal L}_f$, or that are less dependent on
+are more accurate, hence decreasing ${\cal L}_f$, or that are even less dependent on
 $Z$, hence shaping $p_{\theta_r}$ towards the prior $p(z)$. Indeed, ${\cal L}_f$
@@
 classifier that is both optimal and pivotal appears to be unreachable for this
-problem, as shown in third plot by the  offset between ${\cal L}_f - \lambda
+problem, as shown in the third plot by the  offset between ${\cal L}_f - \lambda
 {\cal L}_r$ and the dashed line approximating $H({Y|X}) -

### 2016-10-24 tex/paper.tex
physics example (wip)

@@
 
-\subsection{Physics example}
+\subsection{High energy physics example}
+
+In collisions at the Large Hadron Collider, massive particles can be produced at
+such high energies that their decay products are collimated and the resulting
+jets overlap. Still, deducing whether an observed jet is either due  to a single
+hadronic particle or to the decay of a massive object into multiple hadronic
+particles  is critical for understanding the nature of the particles produced in
+the collision \citep{butterworth2008jet,guest2016jet}. More importantly, this
+classification problem is often made difficult by the presence of simulatenous
+interactions, known as pileup, which contributes significant energy depositions
+unrelated to the particles of interest.
+
+\begin{figure}
+    \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
+    \vspace{-1cm}
+    \caption{(to write)}
+    \label{fig:physics-ams}
+\end{figure}
 

### 2016-10-24 tex/paper.tex
physics example

@@
 predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. As
-illustrated in Figure~\ref{fig:architecture}, we pit $f$ against an adversary
+illustrated in Fig.~\ref{fig:architecture}, we pit $f$ against an adversary
 model $r := p_{\theta_r}(z | f(X;\theta_f)=s)$ of parameters $\theta_r$ and
@@
 to rewrite the value function $E$  as
-\begin{equation}
+\begin{equation}\label{eqn:vf-lambda}
     E_\lambda(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r),
@@
 node with a sigmoid activation. As shown in the upper plots of
-Figure~\ref{fig:toy}, the resulting classifier is not pivotal, as the
+Fig.~\ref{fig:toy}, the resulting classifier is not pivotal, as the
 conditional probability densities of its decision scores $f(X;\theta_f)$ show
@@
 the decision function so it that becomes almost independent on the nuisance
-parameter, as shown in the lower plots of Figure~\ref{fig:toy}. In particular,
+parameter, as shown in the lower plots of Fig.~\ref{fig:toy}. In particular,
 the conditional probability densities of the decision scores $f(X;\theta_f)$ are
@@
 nuisance, as theoretically expected. The dynamics of adversarial training is
-illustrated in Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$,
+illustrated in Fig.~\ref{fig:toy-training}, where the losses ${\cal L}_f$,
 ${\cal L}_r$ and ${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each
@@
 the collision \citep{butterworth2008jet,guest2016jet}. More importantly, this
-classification problem is often made difficult by the presence of simulatenous
+classification problem is often made difficult by the presence of simultaneous
 interactions, known as pileup, which contributes significant energy depositions
 unrelated to the particles of interest.
+
+In this setting, we reuse the classification problem of \citep{guest2016jet}
+between single jets produced in quark or gluon fragmentation and two overlapping
+jets produced when a high-velocity $W$ boson decays to a collimated pair of
+quarks. As nuisance parameter, we consider events without pileup ($Z=0$) and
+events with pileup ($Z=1$), for which an average number of $\langle \mu \rangle =
+50$ unrelated interactions are overlaid. Our goal is to build an accurate
+classifier, for which we also want to minimize the effects due to the
+uncertainties on the nuisance. More specifically, we choose to recast the
+classification problem as a hypothesis test between signal+background (jets
+originating from $W$ bosons or from single quarks and gluons) and background
+only (jets originating from single quarks and gluons only) and tune the decision
+threshold of our classifier by maximizing its approximate median significance
+(AMS), when uncertainties in the background are taken into account (see Eqn.~19
+of \citep{adam2014higgs}). Our motivation is that reducing the effects of
+uncertainties by requiring independence of $Z$ with the classifier output should
+allow for a larger maximum significance.
+
+To minimize the effects of the nuisance $Z$ in the background events, we train a
+classifier using  Algorithm~\ref{alg:adversarial-training} but consider the
+adversarial term ${\cal L}_r$ conditioned on $Y=0$ only. Both $f$ and $r$ are
+neural networks with 3 hidden layers of 64 nodes and ReLU activations, each
+terminated by a single final output node with a sigmoid activation. Training is
+performed on a subset of 150000 samples while AMS is evaluated on an independent
+test set of 5000000 samples. Results reported below are averages over 5 runs.
+
+As shown in  Fig.~\ref{fig:physics-ams}, trading classification accuracy for
+independence with respect to pileup results in a positive total net benefit in
+terms of statistical significance. Without adversarial training (at
+$\lambda=0$), the maximum significance peaks at $7$. By contrast, as the
+independence constraint is made stronger (for $\lambda > 0$) the AMS peak moves
+higher, with a maximum value around  $7.8$ for $\lambda=10$. In other words,
+sacrificing accuracy to reduce  the  effects of the systematics in the
+background jets does indeed allow one to improve the overall significance of the
+results. Setting $\lambda$ too high however (e.g. $\lambda=500$) results in a
+decrease of the maximum significance, by focusing $f$ too much on independence,
+at the expense of too large a drop in accuracy.
 
@@
     \vspace{-1cm}
-    \caption{(to write)}
+    \caption{Approximate median signifance as a function of the decision threshold
+             on the output of $f$.  As shown at $\lambda=10$, trading
+             classification accuracy for independence with respect to pileup
+             results in a positive total net benefit in terms of statistical significance.}
     \label{fig:physics-ams}

### 2016-10-24 tex/paper.tex
typos

@@
 threshold of our classifier by maximizing its approximate median significance
-(AMS), when uncertainties in the background are taken into account (see Eqn.~19
+(AMS), when uncertainties in the background are taken into account (see Eqn.~20
 of \citep{adam2014higgs}). Our motivation is that reducing the effects of
@@
     \vspace{-1cm}
-    \caption{Approximate median signifance as a function of the decision threshold
+    \caption{Approximate median significance as a function of the decision threshold
              on the output of $f$.  As shown at $\lambda=10$, trading

### 2016-10-24 tex/paper.tex
save

@@
 
-To minimize the effects of the nuisance $Z$ in the background events, we train a
+\begin{figure}
+    \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
+    \vspace{-1cm}
+    \caption{Approximate median significance as a function of the decision threshold
+             on the output of $f$.  As shown at $\lambda=10$, trading
+             classification accuracy for independence with respect to pileup
+             results in a positive total net benefit in terms of statistical significance.}
+    \label{fig:physics-ams}
+\end{figure}
+
+To minimize the effects of $Z$ in the background events, we train a
 classifier using  Algorithm~\ref{alg:adversarial-training} but consider the
@@
 at the expense of too large a drop in accuracy.
-
-\begin{figure}
-    \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
-    \vspace{-1cm}
-    \caption{Approximate median significance as a function of the decision threshold
-             on the output of $f$.  As shown at $\lambda=10$, trading
-             classification accuracy for independence with respect to pileup
-             results in a positive total net benefit in terms of statistical significance.}
-    \label{fig:physics-ams}
-\end{figure}
 

### 2016-10-24 tex/paper.tex
save

@@
 
+\begin{figure}
+    \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
+    \vspace{-1cm}
+    \caption{Approximate median significance as a function of the decision threshold
+             on the output of $f$.  As shown at $\lambda=10$, trading
+             classification accuracy for independence with respect to pileup
+             results in a positive total net benefit in terms of statistical significance.}
+    \label{fig:physics-ams}
+\end{figure}
+
 In collisions at the Large Hadron Collider, massive particles can be produced at
@@
 
-\begin{figure}
-    \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
-    \vspace{-1cm}
-    \caption{Approximate median significance as a function of the decision threshold
-             on the output of $f$.  As shown at $\lambda=10$, trading
-             classification accuracy for independence with respect to pileup
-             results in a positive total net benefit in terms of statistical significance.}
-    \label{fig:physics-ams}
-\end{figure}
-
-To minimize the effects of $Z$ in the background events, we train a
-classifier using  Algorithm~\ref{alg:adversarial-training} but consider the
-adversarial term ${\cal L}_r$ conditioned on $Y=0$ only. Both $f$ and $r$ are
-neural networks with 3 hidden layers of 64 nodes and ReLU activations, each
-terminated by a single final output node with a sigmoid activation. Training is
-performed on a subset of 150000 samples while AMS is evaluated on an independent
-test set of 5000000 samples. Results reported below are averages over 5 runs.
+To minimize the effects of $Z$ in the background events, we train a classifier
+using  Algorithm~\ref{alg:adversarial-training} but consider the adversarial
+term ${\cal L}_r$ conditioned on $Y=0$ only. Both $f$ and $r$ are neural
+networks with 3 hidden layers of 64 nodes and ReLU activations, each terminated
+by a single final output node with a sigmoid activation. Experiments are performed
+on the high-level features data described in \citep{guest2016jet}, on a subset
+of 150000 samples for training while AMS is evaluated on an independent test set of 5000000
+samples. Results reported below are averages over 5 runs.
 

### 2016-10-24 tex/paper.tex
save

@@
 classifier minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not
-necessarily with $Y|X,Z$. Results hold for a nuisance parameter $Z$ taking
-either categorical or continuous values. By
-abuse of notation, $H(Z)$ denotes the differential entropy in this latter case.
-Finally, the  proposition below is derived in a non-parametric setting, by
-assuming that both $f$ and $r$ have enough capacity.
+necessarily with $Y|X,Z$ (unless $Z$ is made explicit and is included among the
+input variables in $X$). Results hold for a nuisance parameter $Z$ taking either
+categorical or continuous values. By abuse of notation, $H(Z)$ denotes the
+differential entropy in this latter case. Finally, the  proposition below is
+derived in a non-parametric setting, by assuming that both $f$ and $r$ have
+enough capacity.
 
@@
 particles  is critical for understanding the nature of the particles produced in
-the collision \citep{butterworth2008jet,guest2016jet}. More importantly, this
+the collision \citep{butterworth2008jet,guest2016jet}. Critically, this
 classification problem is often made difficult by the presence of simultaneous
@@
 of \citep{adam2014higgs}). Our motivation is that reducing the effects of
-uncertainties by requiring independence of $Z$ with the classifier output should
+uncertainties by requiring independence of $Z$ with the classifier output $f(X;\theta_f)$ should
 allow for a larger maximum significance.
@@
 
-As shown in  Fig.~\ref{fig:physics-ams}, trading classification accuracy for
-independence with respect to pileup results in a positive total net benefit in
-terms of statistical significance. Without adversarial training (at
+As  Fig.~\ref{fig:physics-ams} illustrates, without adversarial training (at
 $\lambda=0$), the maximum significance peaks at $7$. By contrast, as the
@@
 higher, with a maximum value around  $7.8$ for $\lambda=10$. In other words,
-sacrificing accuracy to reduce  the  effects of the systematics in the
-background jets does indeed allow one to improve the overall significance of the
-results. Setting $\lambda$ too high however (e.g. $\lambda=500$) results in a
-decrease of the maximum significance, by focusing $f$ too much on independence,
-at the expense of too large a drop in accuracy.
+trading classification accuracy for independence with respect to pileup results
+in a positive total net benefit in terms of statistical significance. Setting
+$\lambda$ too high however (e.g. $\lambda=500$) results in a decrease of the
+maximum significance, by focusing the capacity of $f$ too strongly on
+independence, at the expense of classification accuracy. As demonstrated in this
+example, controlling the classification versus pivot trade-off through $\lambda$
+therefore gives us a principled and effective approach for maximizing
+significance by desensitizing the classifier output $f(X;\theta_f)$ in the most
+beneficial way.
 

### 2016-10-25 tex/paper.tex
save

@@
 \begin{itemize}
-    \item $\smash{\hat\theta}_f$ minimizes the negative log-likelihood ${\cal
-    L}_f$ of $X|Y$, which happens when $\smash{\hat\theta}_f$ are the parameters
+    \item $\smash{\hat\theta}_f$ minimizes the negative log-likelihood of $Y|X$ under $f$,
+    which happens when $\smash{\hat\theta}_f$ are the parameters
     of an optimal classifier and in which case ${\cal L}_f$ reduces to its

### 2016-10-27 tex/paper.tex
add ref

@@
 representation of the data. Likewise, our method also relates to the problem of
-enforcing fairness in classification \citep{zemel2013learning,EdwardsS15,zafar2015fairness}, which
+enforcing fairness in classification \citep{zemel2013learning,feldman2015certifying,EdwardsS15,zafar2015fairness}, which
 is stated as learning a classifier that is independent of some chosen attribute

### 2016-10-27 tex/paper.tex
save

@@
 \mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
-of the functional $X$ on a nuisance parameter $Z$ whose values $z \in {\cal Z}$  define a
+of the functional $X$ on a nuisance parameter $Z$, whose values $z \in {\cal Z}$  define a
 parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
@@
 $\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
-is to learn a score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S} \subseteq {\cal Y}$ of
+is to learn a score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of
 parameters $\theta_f$ (e.g., a neural network-based probabilistic classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
@@
 \end{align}
-for all $s \in {\cal S}$. In words, we are looking for a predictive function $f$
+for all values $s \in {\cal S}$ taken by the score function $f$. In words, we are looking for a predictive function $f$
 which is a pivotal quantity \citep{degroot1986probability} with respect to the
@@
 \end{equation}
-for all $z,z^\prime \in  {\cal Z}$ and all values $s$ of $f(X ; \theta_f)$.
+for all $z,z^\prime \in  {\cal Z}$ and all values $s \in {\cal S}$ of $f(X ; \theta_f)$.
 
@@
 
-Adversarial training was first proposed by \citep{goodfellow2014generative} as a
+Joint training of adversarial networks was first proposed by \citep{goodfellow2014generative} as a
 way to build a generative model capable of producing samples from random noise
@@
 
-In this work, we repurpose adversarial training as a means to constraint the
+In this work, we repurpose adversarial networks as a means to constraint the
 predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. As
@@
 where the equality holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f}
-E'(\theta_f)$ only when
+E'(\theta_f)$  when:
 \begin{itemize}
@@
     which happens when $\smash{\hat\theta}_f$ are the parameters
-    of an optimal classifier and in which case ${\cal L}_f$ reduces to its
-    minimum value $H({Y|X})$,
+    of an optimal classifier. In this case, ${\cal L}_f$ reduces to its
+    minimum value $H({Y|X})$.
 
@@
 
-On many problems of interest, let us also note that such a classifier may
-not exist because the nuisance parameter directly shapes the decision boundary,
-in which cases the lower bound $ H({Y|X}) - H(Z) < {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal classifier or a
+On many practical problems, let us note that such a classifier may
+not exist because the nuisance parameter directly shapes the decision boundary.
+In this case, the lower bound $ H({Y|X}) - H(Z) < {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal classifier or a
 pivotal quantity, but not both simultaneously. In this situation, it is natural

### 2016-10-27 tex/paper.tex
save

@@
 
-Assuming a gaussian prior $z \sim {\cal N}(0,1)$ of unit variance, we start by
-generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
+Assuming a gaussian prior $z \sim {\cal N}(0,1)$, we start by
+generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we train a
 neural network classifier $f$ minimizing ${\cal L}_f(\theta_f)$ without
-adversarial training. The network architecture comprises 2 dense hidden layers of 20
+considering its adversary $r$. The network architecture comprises 2 dense hidden layers of 20
 nodes with ReLU activations, followed  by a dense output layer with a single
@@
 
-Let us now consider the adversarial training of $f$ against an adversary $r$
+Let us now consider the joint training of $f$ against an adversary $r$
 implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a mixture
@@
 coefficients of the five gaussians. Output nodes for the mean values come with
-linear activations, nodes for the standard deviations with exponential
-activations to ensure positivity, while nodes for the mixture coefficients
+linear activations, output nodes for the standard deviations with exponential
+activations to ensure positivity, while output nodes for the mixture coefficients
 implement the softmax function to ensure positivity and normalization. When

### 2016-10-27 tex/paper.tex
fix ref

@@
 particles  is critical for understanding the nature of the particles produced in
-the collision \citep{butterworth2008jet,guest2016jet}. Critically, this
+the collision \citep{butterworth2008jet,baldi2016jet}. Critically, this
 classification problem is often made difficult by the presence of simultaneous
@@
 
-In this setting, we reuse the classification problem of \citep{guest2016jet}
+In this setting, we reuse the classification problem of \citep{baldi2016jet}
 between single jets produced in quark or gluon fragmentation and two overlapping
@@
 by a single final output node with a sigmoid activation. Experiments are performed
-on the high-level features data described in \citep{guest2016jet}, on a subset
+on the high-level features data described in \citep{baldi2016jet}, on a subset
 of 150000 samples for training while AMS is evaluated on an independent test set of 5000000

### 2016-10-27 tex/paper.tex
edit physics toy

@@
 
-In collisions at the Large Hadron Collider, massive particles can be produced at
-such high energies that their decay products are collimated and the resulting
-jets overlap. Still, deducing whether an observed jet is either due  to a single
-hadronic particle or to the decay of a massive object into multiple hadronic
-particles  is critical for understanding the nature of the particles produced in
-the collision \citep{butterworth2008jet,baldi2016jet}. Critically, this
-classification problem is often made difficult by the presence of simultaneous
-interactions, known as pileup, which contributes significant energy depositions
-unrelated to the particles of interest.
+Jets, or collimated sprays of particles produced
+by high energy quarks and gluons, are produced ubiqitiously at high
+energy colliders like the Large Hadron Collider \citep{LHCMachine}.  In a wide
+array of new physcis models, new highly massive particles can decay
+though known heavy particles in Standard Model (SM), like the $W$, $Z$, Higgs bosons, or top
+quarks which subsequently decay to mutiple quarks.  When these these
+heavy SM particles have significant energy, the Lorentz boost causes
+their quark decay products to merge into a single jet which has a rich
+internal substructure (see e.g.  \citep{Altheimer:2012mn,Altheimer:2013yza} and references
+therin).  Distinguishing these \textit{boosted jets} produced by heavy SM particles
+from vanilla quarks and  gluons is a fundamental challenge in
+searching for signs of new high mass particles.
 
-In this setting, we reuse the classification problem of \citep{baldi2016jet}
-between single jets produced in quark or gluon fragmentation and two overlapping
-jets produced when a high-velocity $W$ boson decays to a collimated pair of
-quarks. As nuisance parameter, we consider events without pileup ($Z=0$) and
+Challenging in it's own right, this classification problem is
+made all the more difficult by the presence of pileup, or multiple simultaneous
+proton-proton interactions occuring at the same time as the
+primary interaction.  These pileup interactions produce additional
+particles that can contributie singificant energies to jets unrelated to
+the underlying discrimnating information. Thus pileup acts as a source of noise in
+the classification problem.
+
+% In collisions at the Large Hadron Collider, massive particles can be produced at
+% such high energies that their decay products are collimated and the resulting
+% jets overlap. Still, deducing whether an observed jet is either due  to a single
+% hadronic particle or to the decay of a massive object into multiple hadronic
+% particles  is critical for understanding the nature of the particles produced in
+% the collision \citep{butterworth2008jet,guest2016jet}. Critically, this
+% classification problem is often made difficult by the presence of simultaneous
+% interactions, known as pileup, which contributes significant energy depositions
+% unrelated to the particles of interest.
+
+The classification challenge used here is common in jet substructure
+studies (see \citep{Khachatryan:2014vla,ATL-PHYS-PUB-2015-033,wbosonATLAS} and
+references therein): we aim to discriminate between jets produced by quark or gluon fragmentation (the
+background) and  boosted $W$ boson jets containing a collimated pair of
+quarks (the signal).  We reuse the datasets used in reference
+\citep{guest2016jet}.  Briefly summarizing, jets are built with the
+anti-$k_T$ algoritm \citep{Cacciari:2008gp} 
+with radius parameter $R=1.2$, and trimmed \citep{Krohn:2009th} with subjets built with
+the $k_T$ algorithm and parameter $f_{cut}=0.2$. The features used
+in the classifiers are: trimmed jet invariant mass, N-subjettiness
+$\tau_{21}^{\beta=1}$ \citep{nsub,Thaler2012}, and the energy correlation
+functions \citep{Larkoski2013} $C_{2}^{\beta=1}$, $C_{2}^{\beta=2}$,
+$D_{2}^{\beta=1}$, $D_{2}^{\beta=2}$.
+
+As nuisance parameter, we consider events without pileup ($Z=0$) and
 events with pileup ($Z=1$), for which an average number of $\langle \mu \rangle =
-50$ unrelated interactions are overlaid. Our goal is to build an accurate
+50$ unrelated additional pileup interactions are overlaid. Our goal is to build an accurate
 classifier, for which we also want to minimize the effects due to the
@@
 classification problem as a hypothesis test between signal+background (jets
-originating from $W$ bosons or from single quarks and gluons) and background
-only (jets originating from single quarks and gluons only) and tune the decision
+originating from $W$ bosons or from quarks and gluons) and background
+only (jets originating from quarks and gluons only) and tune the decision
 threshold of our classifier by maximizing its approximate median significance
@@
 by a single final output node with a sigmoid activation. Experiments are performed
-on the high-level features data described in \citep{baldi2016jet}, on a subset
-of 150000 samples for training while AMS is evaluated on an independent test set of 5000000
-samples. Results reported below are averages over 5 runs.
+on a subset of 150000 samples for training while AMS is evaluated on an independent test set of 5000000
+samples. Both training and testing samples are weighted such
+that the signal-to-background ratio is 1/10, allowing us to probe the
+efficacy of the method proposed here in a background dominated
+environment. Results reported below are averages over 5 runs.
 

### 2016-10-27 tex/paper.tex
fix typos

@@
 
-\begin{figure}
-    \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
-    \vspace{-1cm}
-    \caption{Approximate median significance as a function of the decision threshold
-             on the output of $f$.  As shown at $\lambda=10$, trading
-             classification accuracy for independence with respect to pileup
-             results in a positive total net benefit in terms of statistical significance.}
-    \label{fig:physics-ams}
-\end{figure}
-
-Jets, or collimated sprays of particles produced
-by high energy quarks and gluons, are produced ubiqitiously at high
-energy colliders like the Large Hadron Collider \citep{LHCMachine}.  In a wide
-array of new physcis models, new highly massive particles can decay
-though known heavy particles in Standard Model (SM), like the $W$, $Z$, Higgs bosons, or top
-quarks which subsequently decay to mutiple quarks.  When these these
-heavy SM particles have significant energy, the Lorentz boost causes
-their quark decay products to merge into a single jet which has a rich
-internal substructure (see e.g.  \citep{Altheimer:2012mn,Altheimer:2013yza} and references
-therin).  Distinguishing these \textit{boosted jets} produced by heavy SM particles
-from vanilla quarks and  gluons is a fundamental challenge in
-searching for signs of new high mass particles.
-
-Challenging in it's own right, this classification problem is
-made all the more difficult by the presence of pileup, or multiple simultaneous
-proton-proton interactions occuring at the same time as the
-primary interaction.  These pileup interactions produce additional
-particles that can contributie singificant energies to jets unrelated to
-the underlying discrimnating information. Thus pileup acts as a source of noise in
-the classification problem.
-
-% In collisions at the Large Hadron Collider, massive particles can be produced at
-% such high energies that their decay products are collimated and the resulting
-% jets overlap. Still, deducing whether an observed jet is either due  to a single
-% hadronic particle or to the decay of a massive object into multiple hadronic
-% particles  is critical for understanding the nature of the particles produced in
-% the collision \citep{butterworth2008jet,guest2016jet}. Critically, this
-% classification problem is often made difficult by the presence of simultaneous
-% interactions, known as pileup, which contributes significant energy depositions
-% unrelated to the particles of interest.
+Jets, or collimated sprays of particles produced by high energy quarks and
+gluons, are produced ubiquitously at high energy colliders like the Large Hadron
+Collider \citep{LHCMachine}.  In a wide array of new physics models, new highly
+massive particles can decay though known heavy particles in Standard Model (SM),
+like the $W$, $Z$, Higgs bosons, or top quarks which subsequently decay to
+multiple quarks.  When these these heavy SM particles have significant energy,
+the Lorentz boost causes their quark decay products to merge into a single jet
+which has a rich internal substructure (see e.g.
+\citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
+Distinguishing these \textit{boosted jets} produced by heavy SM particles from
+vanilla quarks and  gluons is a fundamental challenge in searching for signs of
+new high mass particles. Challenging in it's own right, this classification
+problem is made all the more difficult by the presence of pileup, or multiple
+simultaneous proton-proton interactions occurring at the same time as the primary
+interaction.  These pileup interactions produce additional particles that can
+contribute significant energies to jets unrelated to the underlying
+discriminating information. Thus pileup acts as a source of noise in the
+classification problem.
 
@@
 \citep{guest2016jet}.  Briefly summarizing, jets are built with the
-anti-$k_T$ algoritm \citep{Cacciari:2008gp} 
+anti-$k_T$ algorithm \citep{Cacciari:2008gp}
 with radius parameter $R=1.2$, and trimmed \citep{Krohn:2009th} with subjets built with
@@
 environment. Results reported below are averages over 5 runs.
+
+\begin{figure}
+    \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
+    \vspace{-1cm}
+    \caption{Approximate median significance as a function of the decision threshold
+             on the output of $f$.  As shown at $\lambda=10$, trading
+             classification accuracy for independence with respect to pileup
+             results in a positive total net benefit in terms of statistical significance.}
+    \label{fig:physics-ams}
+\end{figure}
 

### 2016-10-27 tex/paper.tex
save

@@
 Jets, or collimated sprays of particles produced by high energy quarks and
-gluons, are produced ubiquitously at high energy colliders like the Large Hadron
-Collider \citep{LHCMachine}.  In a wide array of new physics models, new highly
-massive particles can decay though known heavy particles in Standard Model (SM),
+gluons, are produced ubiquitously at high energy colliders like the LHC \citep{LHCMachine}.
+In a wide array of new physics models, new highly
+massive particles can decay through known heavy particles in the Standard Model (SM),
 like the $W$, $Z$, Higgs bosons, or top quarks which subsequently decay to
-multiple quarks.  When these these heavy SM particles have significant energy,
+multiple quarks.  When these heavy SM particles have significant energy,
 the Lorentz boost causes their quark decay products to merge into a single jet
@@
 \citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
-Distinguishing these \textit{boosted jets} produced by heavy SM particles from
+Distinguishing these boosted jets produced by heavy SM particles from
 vanilla quarks and  gluons is a fundamental challenge in searching for signs of
-new high mass particles. Challenging in it's own right, this classification
+new high mass particles. Challenging in its own right, this classification
 problem is made all the more difficult by the presence of pileup, or multiple

### 2016-10-30 tex/paper.tex
No content change, just encoding
Some issue with encoding. No content changes.
http://stackoverflow.com/questions/21732903/git-sees-entire-file-as-one-
line-due-to-mac-line-endings
new abstract and intro, not complete

@@
 
-In high energy physics, systematic uncertainties represent our incomplete
-knowledge in the theory of physical processes or in the properties of the
-experimental detection apparatus. In effect, these uncertainties typically cause
-variations in the conditional distributions of the data, thereby altering the
-decisions and resulting efficacy of a classifier trained from it. To contain
-this issue, we propose to repurpose adversarial training as a means to learn a
-pivotal classifier which is invariant to variations of the nuisance parameters
-describing the effects of systematic uncertainties. In particular, we show and
-derive theoretical conditions under which classification optimality and
-invariance with respect to systematics can be achieved, which we confirm
-experimentally on a couple of illustrative examples. In machine learning, this
-problem is closely related to those of domain adaptation and enforcing fairness
-in classification. Following that line of work, the proposed method can be
-regarded as generalization that also supports the continuous case, which can be
-viewed as handling infinitely many domains or as enforcing fairness over
-continuous features.
+Many inference problems involve data generation processes that are not uniquely specified or are uncertain in some way. 
+In a scientific context, the presence of several plausible data generation processes is often
+associated to the presence of systematic uncertainties. 
+Robust inference is possible if it is based on a pivot -- a quantity whose
+distribution is invariant to the unknown value of the nuisance parameters that parametrize a family of data generation processes. 
+We introduce an adversarial training procedure to enforce the pivotal property on a predictive model, which 
+allows one to tune the tradeoff between power and robustness. 
+We demonstrate the effectiveness of this approach with a toy and real world example.
 
+%Many inference problems involve data generation processes that are not uniquely specified or uncertain in some way. 
+%In a scientific context, the presence of several plausible data generation processes is often
+%associated to the presence of systematic uncertainty. 
+%We consider the situation of a 
+%family of data generation processes parametrized by nuisance parameters, which can
+%be categorical or continuous. Robust inference is possible if it is based on a pivot -- a quantity whose
+%distribution is invariant to the unknown value of these nuisance parameters. 
+%We introduce an adversarial training procedure to enforce the pivotal property and
+%allow one to tune the tradeoff between power and robustness.
+
+%Our goal is to have an inferential procedure that is robust to the uncertainties in the data generation process. 
 \end{abstract}
@@
 \maketitle
-
 % ==============================================================================
@@
 
-The discovery of new particles or physical phenomena in high energy physics
-experiments, like those at the Large Hadron Collider, requires the observation
-of statistically significant deviations from the predictions of the Standard
-Model, the current model of the known fundamental particles and their
-interactions.  Typically, discovery requires satisfying the \textit{5-sigma
-rule}, whereby the deviation must be at least five standard deviations from
-predictions, i.e. having a p-value of $p \lesssim 3 \times 10^{-7}$. The
-evaluation of the statistical significance must not only account for statistical
-uncertainties due to random fluctuations, but also for
-systematic uncertainties that represent our incomplete knowledge of the theory
-of particle interactions or of the properties of the experimental detection
-apparatus. Systematic uncertainties can alter the expected rate of signal and
-background classes, e.g. their priors, as well as the distributions of features.
-As such, the signal class true-positive rate and the background class
-false-positive rate of a classifier can change with variations in the systematic
-uncertainty thereby decreasing the sensitivity of an analysis. In this paper, we
-present a new strategy for training a classifier, based on adversarial
-training~\citep{goodfellow2014generative}, which aims at building a classifier
-whose decision function is invariant under variations due to systematic
-uncertainties.  In this way, uncertainties on predictions from simulations will
-have a reduced impact when computing the significance of potential deviations
-after application of the discriminant model and thereby increase the sensitivity
-of the analysis to signals from potential new physics sources.
-
-Given data $X$ and associated labels $Y$
-taking values $y \in {\cal Y} = \{s,
-b_{i=1...k} \}$ where $s$ is the signal class and $b_{i=1...k}$ are
-the background classes, the probability density of the data
-can be written as a mixture
-\begin{equation}
-p(X) = \pi_s p(X|Y=s) + \sum_{i=1}^{k} \pi_{b_i} p(X | Y=b_i),
-\end{equation}
-where $\pi_{y}$ represents the prior for the signal or a given
-background, and $p(X|Y)$ is the conditional distribution of the data
-for the signal or a given background. Systematic uncertainties can
-affect this distribution in the following ways \citep{sinervo2003definition,bohm2010introduction}:
-\begin{itemize}
-\item Uncertainties on the priors $\pi_{y}$.  These arise from
-  uncertainties on the rates we expect a given process to occur as
-  predicted from theoretical calculations or from alternative
-  measurements of the rate of a given class.
-
-\item Uncertainties in the simulators of the physical interactions,
-  giving rise to uncertainties on the distributions $p(X|Y)$.  These
-  uncertainties are often determined by comparing the predictions of
-  different simulations using differing models of the same physical
-  process.
-
-\item Uncertainties on the detection apparatus and its effect on the
-  measurement of particle properties, giving rise to uncertainties on
-  the distributions $p(X|Y)$.  For instance, this includes
-  uncertainties on our knowledge of the mean and variance of the
-  measured energy distribution of a particle.
-\end{itemize}
-In this paper, we will not address uncertainties of the first type
-mentioned above (uncertainties on the priors).  Rather, we will focus
-on mitigating the effect of systematic uncertainties that alter the
-conditional distributions $p(X|Y)$. \glnote{Remove this comment?}
-
-Systematic uncertainties are parameterized by nuisance parameters in the
-statistical analysis of data (see e.g.,
-\citep{bohm2010introduction,cowan2011asymptotic}).  Nuisance parameters often,
-though not always, have a suitable prior coming from alternative measurements in
-data.  For instance, the average measured energy of an electron in simulation
-may be compared with those in data using a relatively pure sample of $Z$ boson
-decays to electrons.  The average measured electron energy can not be measured
-without uncertainty, and the measured uncertainty can be used in other analyses
-to estimate the uncertainty deriving from our imperfect knowledge of the
-electron's energy.  A nuisance parameter is assigned to the electron energy
-uncertainty, which controls the variation of the electron energy in the
-statistical analysis of data.  Nuisance parameters are typically, though not
-always, constrained by a suitable prior, e.g. a Gaussian distribution with a
-standard deviation equal to the measured uncertainty. \glnote{This paragraph
-could be improved, it is not very clear what we want to say...}
-
-\glnote{Add paper outline.}
+\kcnote{ need some text like abstract here before diving into formal statement.}
 
@@
 \label{sec:problem}
+%
+%Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
+%sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
+%Let consider the multivariate random variables $X_z: \Omega \mapsto
+%\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
+%of the functional $X$ on a nuisance parameter $Z$, whose values $z \in {\cal Z}$  define a
+%parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
+%induce together a joint probability distribution $p(X,Y|Z=z)$, where the
+%conditional denotes $X_z$. For training, let further assume a finite set $\{
+%x_i, y_i, z_i \}_{i=1}^N$ of realizations $X_{z_i}(\omega_i), Y(\omega_i)$, for
+%$\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
+%is to learn a score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of
+%parameters $\theta_f$ (e.g., a neural network-based probabilistic classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
+%the cross-entropy). In addition, we require that $f(X_z ; \theta_f)$ should be
+%robust to the value $z$ of the nuisance parameter  -- which remains unknown at
+%test time. More specifically, we aim at building $f$ such that in the ideal case
+%\begin{equation}\label{eqn:criterion-true}
+%f(X_{z}(\omega) ; \theta_f) = f(X_{z^\prime}(\omega) ; \theta_f)
+%\end{equation} for all
+%samples $\omega \in \Omega$ and all $z, z^\prime$ pairs of values of the
+%nuisance parameter.
+%
+%Since in general we do not have training tuples $(X_{z}(\omega),
+%X_{z^\prime}(\omega))$ (for the same unknown $\omega$),
 
-Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
-sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
-Let consider the multivariate random variables $X_z: \Omega \mapsto
-\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
-of the functional $X$ on a nuisance parameter $Z$, whose values $z \in {\cal Z}$  define a
-parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
-induce together a joint probability distribution $p(X,Y|Z=z)$, where the
-conditional denotes $X_z$. For training, let further assume a finite set $\{
-x_i, y_i, z_i \}_{i=1}^N$ of realizations $X_{z_i}(\omega_i), Y(\omega_i)$, for
-$\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
+Our goal
 is to learn a score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of
 parameters $\theta_f$ (e.g., a neural network-based probabilistic classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
-the cross-entropy). In addition, we require that $f(X_z ; \theta_f)$ should be
+the cross-entropy). In addition, we require that the distribution of $f(X_z ; \theta_f)$ should be
 robust to the value $z$ of the nuisance parameter  -- which remains unknown at
-test time. More specifically, we aim at building $f$ such that in the ideal case
-\begin{equation}\label{eqn:criterion-true}
-f(X_{z}(\omega) ; \theta_f) = f(X_{z^\prime}(\omega) ; \theta_f)
-\end{equation} for all
-samples $\omega \in \Omega$ and all $z, z^\prime$ pairs of values of the
-nuisance parameter.
+test time.
 
-Since in general we do not have training tuples $(X_{z}(\omega),
-X_{z^\prime}(\omega))$ (for the same unknown $\omega$), we propose instead to
-solve the closely related problem of finding a predictive function $f$ such that
-\begin{align}\label{eqn:criterion-measure}
-    &P(\{ \omega | f(X_{z}(\omega) ; \theta_f) = s \} ) \nonumber \\
-    &\indent = P( \{ \omega' | f(X_{z^\prime}(\omega') ; \theta_f) = s\})
-\end{align}
-for all values $s \in {\cal S}$ taken by the score function $f$. In words, we are looking for a predictive function $f$
-which is a pivotal quantity \citep{degroot1986probability} with respect to the
-nuisance parameter. That is, such that  the distribution of $f(X_z; \theta_f)$
-is invariant with respect to the value $z$ of the nuisance. Note that a function
-$f$ for which Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
-Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true,
-since the sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = s \}$ and $\{
-\omega' | f(X_{z^\prime}(\omega'); \theta_f) = s \}$ do not need to be the same
-for the equality to hold. In order to simplify notations, and as only
-Eqn.~\ref{eqn:criterion-measure} is of direct interest in this work, we denote
-from here on the pivotal quantity criterion as
+We propose  to
+solve the  problem of finding a predictive function $f$ such that
 \begin{equation}\label{eqn:criterion}
@@
 for all $z,z^\prime \in  {\cal Z}$ and all values $s \in {\cal S}$ of $f(X ; \theta_f)$.
+In words, we are looking for a predictive function $f$
+which is a pivotal quantity \citep{degroot1986probability} with respect to the
+nuisance parameter. That is, such that  the distribution of $f(X_z; \theta_f)$
+is invariant with respect to the value $z$ of the nuisance.
+
+%Note that a function
+%$f$ for which Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
+%Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true,
+%since the sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = s \}$ and $\{
+%\omega' | f(X_{z^\prime}(\omega'); \theta_f) = s \}$ do not need to be the same
+%for the equality to hold. In order to simplify notations, and as only
+%Eqn.~\ref{eqn:criterion-measure} is of direct interest in this work, we denote
+%from here on the pivotal quantity criterion as
+
+\glnote{Add paper outline.}
 

### 2016-10-30 tex/paper.tex
one more step of cleanup
updated intro

@@
 \newcommand{\glnote}[1]{\textcolor{red}{[GL: #1]}}
+\newcommand{\kcnote}[1]{\textcolor{red}{[KC: #1]}}
 
@@
 
-\title{\Large{Adversarial Training against Systematic Uncertainty}}
+\title{\Large{Learning to Pivot with Adversarial Networks}}
 \vspace{1cm}
@@
 
+Machine learning techniques have been used to enhance a number of scientific disciplines, 
+and they have the potential to transform even more of the scientific process.
+One of the challenges of applying machine learning techniques to scientific 
+problems is the need to incorporate systematic uncertainties. The presence of systematic
+uncertainties affects both the robustness of inference and the metrics used to evaluate the 
+performance of a particular data analysis strategy.
+
+We will focus on supervised learning techniques where the presence of systematic 
+uncertainty can be associated to the fact that the data generation process is not uniquely 
+specified. In other words, the lack of systematic uncertainties corresponds to the case 
+that the process that generates training data is unique, fully specified, 
+and an accurate representative of the real world data.
+
+A common situation when systematic uncertainty is present is when the training data are not representative of
+the real data that a predictive model will be applied to in practice. Several techniques for domain adaptation have
+been developed to create predictive models that are more robust to this type of uncertainty.
+
+Another common situation is that there are several plausible data generation processes. We consider the situation of a 
+family of data generation processes parametrized by nuisance parameters, which can be categorical or continuous.
+In particular, we consider the probability model $p(X,Y,Z)$, where $X$ are the data, $Y$ are the target labels, and $Z$ are the nuisance parameters.
+We will consider the problem of learning a predictive model for $Y$ conditional on the observed values of $X$ that is robust to uncertainty in the unknown value of $Z$.
+
+\kcnote{Fill in outline}
+
+In Sec.~\ref{sec:problem}  Sec.~\ref{sec:method}  Sec.~\ref{sec:theory}  Sec.~\ref{sec:toy}  Sec.~\ref{sec:hep} 
+ Sec.~\ref{sec:related}  Sec.~\ref{sec:conclusions} .
+
 % ==============================================================================
@@
 
-Our goal
-is to learn a score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of
+We begin with a family of data generation processes $p(X,Y,Z)$, where $X$ are the data, $Y$ are the target labels, and $Z$ are the nuisance parameters that can be continuous or categorical.
+We can use this model to generate training data for $X$ conditional on both $Y$ and $Z$.
+Our goal is to learn a predictive model $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of
 parameters $\theta_f$ (e.g., a neural network-based probabilistic classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
-the cross-entropy). In addition, we require that the distribution of $f(X_z ; \theta_f)$ should be
+the cross-entropy).
+
+In addition, we would like that inference based on $f(X ; \theta_f)$ should be
 robust to the value $z$ of the nuisance parameter  -- which remains unknown at
-test time.
-
-We propose  to
-solve the  problem of finding a predictive function $f$ such that
+test time. A formal way of enforcing robustness is to require that the distribution of 
+$f(X ; \theta_f)$ conditional on $Z$ (and possibly $Y$) be invariant to the value $Z$
+ of the nuisance parameter. Thus, we wish find a predictive function $f$ such that
 \begin{equation}\label{eqn:criterion}
@@
 
+\kcnote{adjust text for conditional on Y as well}
+
 %Note that a function
@@
 %from here on the pivotal quantity criterion as
-
-\glnote{Add paper outline.}
 
@@
 \subsection{High energy physics example}
+\label{sec:hep}
 
@@
 \section{Related work}
+\label{sec:related}
 
@@
 \section{Conclusions}
+\label{sec:conclusions}
 

### 2016-10-30 tex/paper.tex
small things

@@
 
-We begin with a family of data generation processes $p(X,Y,Z)$, where $X$ are the data, $Y$ are the target labels, and $Z$ are the nuisance parameters that can be continuous or categorical.
-We can use this model to generate training data for $X$ conditional on both $Y$ and $Z$.
-Our goal is to learn a predictive model $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of
-parameters $\theta_f$ (e.g., a neural network-based probabilistic classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
-the cross-entropy).
+We begin with a family of data generation processes $p(X,Y,Z)$, where $X$ are
+the data, $Y$ are the target labels, and $Z$ are the nuisance parameters that
+can be continuous or categorical. We can use this model to generate training
+data $\{x_i, y_i, z_i\}_{i=1}^N$ for $X$ conditional on both $Y$ and $Z$. Our goal is to learn a predictive
+score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of parameters
+$\theta_f$ (e.g., a neural network-based probabilistic classifier) and
+minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g., the cross-entropy).
 
 In addition, we would like that inference based on $f(X ; \theta_f)$ should be
-robust to the value $z$ of the nuisance parameter  -- which remains unknown at
+robust to the value $z \in {\cal Z}$ of the nuisance parameter $Z$  -- which remains unknown at
 test time. A formal way of enforcing robustness is to require that the distribution of
-$f(X ; \theta_f)$ conditional on $Z$ (and possibly $Y$) be invariant to the value $Z$
- of the nuisance parameter. Thus, we wish find a predictive function $f$ such that
+$f(X ; \theta_f)$ conditional on $Z$ (and possibly $Y$) be invariant to the value
+ of the nuisance parameter $Z$. Thus, we wish find a predictive function $f$ such that
 \begin{equation}\label{eqn:criterion}

### 2016-10-31 tex/paper.tex
move fig1
abstract + intro

@@
 
-Many inference problems involve data generation processes that are not uniquely specified or are uncertain in some way.
-In a scientific context, the presence of several plausible data generation processes is often
-associated to the presence of systematic uncertainties.
-Robust inference is possible if it is based on a pivot -- a quantity whose
-distribution is invariant to the unknown value of the nuisance parameters that parametrize a family of data generation processes.
-We introduce an adversarial training procedure to enforce the pivotal property on a predictive model, which
-allows one to tune the tradeoff between power and robustness.
-We demonstrate the effectiveness of this approach with a toy and real world example.
+Many inference problems involve data generation processes that are not uniquely
+specified or are uncertain in some way. In a scientific context, the presence of
+several plausible data generation processes is often associated to the presence
+of systematic uncertainties. Robust inference is possible if it is based on a
+pivot -- a quantity whose distribution is invariant to the unknown value of the
+nuisance parameters that parametrize the family of data generation processes. In
+this work, we introduce an adversarial training procedure to enforce the pivotal
+property on a predictive model, which allows one to tune the tradeoff between
+power and robustness. We derive theoretical results proving that the proposed
+procedure converges towards a predictive model that is both optimal and independent on
+the nuisance parameters, and demonstrate the effectiveness of this approach with a
+toy and real world examples.
 
@@
 
-Machine learning techniques have been used to enhance a number of scientific disciplines,
-and they have the potential to transform even more of the scientific process.
-One of the challenges of applying machine learning techniques to scientific
-problems is the need to incorporate systematic uncertainties. The presence of systematic
-uncertainties affects both the robustness of inference and the metrics used to evaluate the
-performance of a particular data analysis strategy.
+Machine learning techniques have been used to enhance a number of scientific
+disciplines, and they have the potential to transform even more of the
+scientific process. One of the challenges of applying machine learning
+techniques to scientific problems is the need to incorporate systematic
+uncertainties. The presence of systematic uncertainties affects both the
+robustness of inference and the metrics used to evaluate the performance of a
+particular data analysis strategy.
 
-We will focus on supervised learning techniques where the presence of systematic
-uncertainty can be associated to the fact that the data generation process is not uniquely
-specified. In other words, the lack of systematic uncertainties corresponds to the case
-that the process that generates training data is unique, fully specified,
-and an accurate representative of the real world data.
+In this work, we focus on supervised learning techniques where the presence of
+systematic uncertainty can be associated to the fact that the data generation
+process is not uniquely specified. In other words, the lack of systematic
+uncertainties corresponds to the (rare) case that the process that generates
+training data is unique, fully specified, and an accurate representative of the
+real world data.
+A common situation when systematic uncertainty is present is when the training
+data are not representative of the real data that a predictive model will be
+applied to in practice. Several techniques for domain adaptation have been
+developed to create predictive models that are more robust to this type of
+uncertainty.
+Another common situation is that there are several plausible data generation
+processes, specificied as a family of data generation processes
+parametrized by nuisance parameters, which can be categorical or continuous.
 
-A common situation when systematic uncertainty is present is when the training data are not representative of
-the real data that a predictive model will be applied to in practice. Several techniques for domain adaptation have
-been developed to create predictive models that are more robust to this type of uncertainty.
+Assuming a probability model $p(X,Y,Z)$, where $X$ are the
+data, $Y$ are the target labels, and $Z$ are the nuisance parameters,
+we consider the problem of learning a predictive model for $Y$ conditional on the
+observed values of $X$ that is robust to uncertainty in the unknown value of
+$Z$. More specifically, we introduce a learning procedure for enforcing the
+pivotal property on a predictive model by jointly training adversarial networks.
+We derive theoretical results proving that the proposed procedure converges
+towards a predictive model that is both optimal and independent on the nuisance
+parameters, and demonstrate the effectiveness of this approach with a toy and
+real world examples.
 
-Another common situation is that there are several plausible data generation processes. We consider the situation of a
-family of data generation processes parametrized by nuisance parameters, which can be categorical or continuous.
-In particular, we consider the probability model $p(X,Y,Z)$, where $X$ are the data, $Y$ are the target labels, and $Z$ are the nuisance parameters.
-We will consider the problem of learning a predictive model for $Y$ conditional on the observed values of $X$ that is robust to uncertainty in the unknown value of $Z$.
-
-\kcnote{Fill in outline}
-
-In Sec.~\ref{sec:problem}  Sec.~\ref{sec:method}  Sec.~\ref{sec:theory}  Sec.~\ref{sec:toy}  Sec.~\ref{sec:hep}
- Sec.~\ref{sec:related}  Sec.~\ref{sec:conclusions} .
+The remainder of the paper is structured as follows. In Sec.~\ref{sec:problem},
+we first formally define the problem of learning a pivotal predictive model. In
+sections \ref{sec:method} and \ref{sec:theory}, we describe how adversarial
+networks can be trained to obtain predictive models that are independent on
+nuisance parameters and show that the proposed procedure indeed converges
+towards a model that is both optimal and robust. The  effectiveness of the
+approach is then demonstrated in sections \ref{sec:toy} and \ref{sec:hep} with a
+toy and real world examples. In Sec.~\ref{sec:related}, we relate our work
+to close contributions from statistical inference, domain adaptation
+and fairness in classification. Finally, Sec.~\ref{sec:conclusions} gathers
+our concluding remarks.
 
@@
 \label{sec:method}
-
-Joint training of adversarial networks was first proposed by \citep{goodfellow2014generative} as a
-way to build a generative model capable of producing samples from random noise
-$z$. More specifically, the authors pit a generative model $g:
-\mathbb{R} \mapsto \mathbb{R}^p$ against an adversary classifier $d :
-\mathbb{R}^p \mapsto [0, 1]$ whose antagonistic objective is to recognize
-real data $X$ from generated data $g(Z)$. Both models $g$ and $d$ are trained
-simultaneously, in such a way that $g$ learns to produce samples that are
-difficult to identify by $d$, while $d$ incrementally adapts to changes in $g$.
-At the equilibrium, $g$ models a distribution whose samples can be identified by
-$d$ only by chance. That is, assuming enough capacity in $d$ and  $g$, the
-distribution of $g(Z)$ eventually converges towards the real distribution
-of $X$.
 
@@
 \end{figure*}
+
+Joint training of adversarial networks was first proposed by \citep{goodfellow2014generative} as a
+way to build a generative model capable of producing samples from random noise
+$z$. More specifically, the authors pit a generative model $g:
+\mathbb{R} \mapsto \mathbb{R}^p$ against an adversary classifier $d :
+\mathbb{R}^p \mapsto [0, 1]$ whose antagonistic objective is to recognize
+real data $X$ from generated data $g(Z)$. Both models $g$ and $d$ are trained
+simultaneously, in such a way that $g$ learns to produce samples that are
+difficult to identify by $d$, while $d$ incrementally adapts to changes in $g$.
+At the equilibrium, $g$ models a distribution whose samples can be identified by
+$d$ only by chance. That is, assuming enough capacity in $d$ and  $g$, the
+distribution of $g(Z)$ eventually converges towards the real distribution
+of $X$.
 
@@
 using  Algorithm~\ref{alg:adversarial-training} but consider the adversarial
-term ${\cal L}_r$ conditioned on $Y=0$ only. Both $f$ and $r$ are neural
-networks with 3 hidden layers of 64 nodes and ReLU activations, each terminated
-by a single final output node with a sigmoid activation. Experiments are performed
-on a subset of 150000 samples for training while AMS is evaluated on an independent test set of 5000000
-samples. Both training and testing samples are weighted such
-that the signal-to-background ratio is 1/10, allowing us to probe the
-efficacy of the method proposed here in a background dominated
-environment. Results reported below are averages over 5 runs.
+term ${\cal L}_r$ conditioned on $Y=0$ only, as outlined in
+Sec.~\ref{sec:problem}. Both $f$ and $r$ are neural networks with 3 hidden
+layers of 64 nodes and ReLU activations, each terminated by a single final
+output node with a sigmoid activation. Experiments are performed on a subset of
+150000 samples for training while AMS is evaluated on an independent test set of
+5000000 samples. Both training and testing samples are weighted such that the
+signal-to-background ratio is 1/10, allowing us to probe the efficacy of the
+method proposed here in a background dominated environment. Results reported
+below are averages over 5 runs.
 
@@
 As  Fig.~\ref{fig:physics-ams} illustrates, without adversarial training (at
-$\lambda=0$), the maximum significance peaks at $7$. By contrast, as the
-independence constraint is made stronger (for $\lambda > 0$) the AMS peak moves
-higher, with a maximum value around  $7.8$ for $\lambda=10$. In other words,
-trading classification accuracy for independence with respect to pileup results
-in a positive total net benefit in terms of statistical significance. Setting
-$\lambda$ too high however (e.g. $\lambda=500$) results in a decrease of the
-maximum significance, by focusing the capacity of $f$ too strongly on
-independence, at the expense of classification accuracy. As demonstrated in this
-example, controlling the classification versus pivot trade-off through $\lambda$
-therefore gives us a principled and effective approach for maximizing
-significance by desensitizing the classifier output $f(X;\theta_f)$ in the most
-beneficial way.
+$\lambda=0|Z=0$ when building a classifier at the nominal value $Z=0$ only, or
+at $\lambda=0$ when building a classifier on data sampled from $p(X,Y,Z)$), the
+maximum significance peaks at $7$. By contrast, as the independence constraint
+is made stronger (for $\lambda > 0$) the AMS peak moves higher, with a maximum
+value around  $7.8$ for $\lambda=10$. In other words, trading classification
+accuracy for independence with respect to pileup results in a positive total net
+benefit in terms of statistical significance. Setting $\lambda$ too high however
+(e.g. $\lambda=500$) results in a decrease of the maximum significance, by
+focusing the capacity of $f$ too strongly on independence, at the expense of
+classification accuracy. As demonstrated in this example, controlling the
+classification versus pivot trade-off through $\lambda$ therefore gives us a
+principled and effective approach for maximizing significance by desensitizing
+the classifier output $f(X;\theta_f)$ in the most beneficial way.
 

### 2016-10-31 tex/paper.tex
save

@@
 
-\kcnote{ need some text like abstract here before diving into formal statement.}
+% \kcnote{ need some text like abstract here before diving into formal statement.}
 
@@
 real world data.
-A common situation when systematic uncertainty is present is when the training
+By contrast, a common situation when systematic uncertainty is present is when the training
 data are not representative of the real data that a predictive model will be
@@
 
-Assuming a probability model $p(X,Y,Z)$, where $X$ are the
-data, $Y$ are the target labels, and $Z$ are the nuisance parameters,
-we consider the problem of learning a predictive model for $Y$ conditional on the
-observed values of $X$ that is robust to uncertainty in the unknown value of
-$Z$. More specifically, we introduce a learning procedure for enforcing the
-pivotal property on a predictive model by jointly training adversarial networks.
-We derive theoretical results proving that the proposed procedure converges
-towards a predictive model that is both optimal and independent on the nuisance
-parameters, and demonstrate the effectiveness of this approach with a toy and
-real world examples.
+Assuming a probability model $p(X,Y,Z)$, where $X$ are the data, $Y$ are the
+target labels, and $Z$ are the nuisance parameters, we consider the problem of
+learning a predictive model for $Y$ conditional on the observed values of $X$
+that is robust to uncertainty in the unknown value of $Z$. More specifically, we
+introduce a learning procedure for enforcing the pivotal property on a
+predictive model by jointly training two neural networks in an adversarial
+fashion. We derive theoretical results proving that the proposed procedure
+converges towards a predictive model that is both optimal and independent on the
+nuisance parameters (if that models exists) or for which one can tune the
+trade-off between power and robustness. Finally, we demonstrate the
+effectiveness of this approach with a toy and real world examples.
 
@@
 networks can be trained to obtain predictive models that are independent on
-nuisance parameters and show that the proposed procedure indeed converges
+nuisance parameters and show that the proposed procedure converges
 towards a model that is both optimal and robust. The  effectiveness of the
 approach is then demonstrated in sections \ref{sec:toy} and \ref{sec:hep} with a
-toy and real world examples. In Sec.~\ref{sec:related}, we relate our work
-to close contributions from statistical inference, domain adaptation
-and fairness in classification. Finally, Sec.~\ref{sec:conclusions} gathers
-our concluding remarks.
+toy and real world examples. In Sec.~\ref{sec:related}, we relate our work to
+close contributions from statistical inference, domain adaptation and fairness
+in classification. Finally, we gather our concluding remarks in
+Sec.~\ref{sec:conclusions}.
 
 % ==============================================================================
-
-\section{Problem statement}
-\label{sec:problem}
-%
-%Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
-%sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
-%Let consider the multivariate random variables $X_z: \Omega \mapsto
-%\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
-%of the functional $X$ on a nuisance parameter $Z$, whose values $z \in {\cal Z}$  define a
-%parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
-%induce together a joint probability distribution $p(X,Y|Z=z)$, where the
-%conditional denotes $X_z$. For training, let further assume a finite set $\{
-%x_i, y_i, z_i \}_{i=1}^N$ of realizations $X_{z_i}(\omega_i), Y(\omega_i)$, for
-%$\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
-%is to learn a score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of
-%parameters $\theta_f$ (e.g., a neural network-based probabilistic classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
-%the cross-entropy). In addition, we require that $f(X_z ; \theta_f)$ should be
-%robust to the value $z$ of the nuisance parameter  -- which remains unknown at
-%test time. More specifically, we aim at building $f$ such that in the ideal case
-%\begin{equation}\label{eqn:criterion-true}
-%f(X_{z}(\omega) ; \theta_f) = f(X_{z^\prime}(\omega) ; \theta_f)
-%\end{equation} for all
-%samples $\omega \in \Omega$ and all $z, z^\prime$ pairs of values of the
-%nuisance parameter.
-%
-%Since in general we do not have training tuples $(X_{z}(\omega),
-%X_{z^\prime}(\omega))$ (for the same unknown $\omega$),
-
-We begin with a family of data generation processes $p(X,Y,Z)$, where $X$ are
-the data, $Y$ are the target labels, and $Z$ are the nuisance parameters that
-can be continuous or categorical. We can use this model to generate training
-data $\{x_i, y_i, z_i\}_{i=1}^N$ for $X$ conditional on both $Y$ and $Z$. Our goal is to learn a predictive
-score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of parameters
-$\theta_f$ (e.g., a neural network-based probabilistic classifier) and
-minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g., the cross-entropy).
-
-In addition, we would like that inference based on $f(X ; \theta_f)$ should be
-robust to the value $z \in {\cal Z}$ of the nuisance parameter $Z$  -- which remains unknown at
-test time. A formal way of enforcing robustness is to require that the distribution of
-$f(X ; \theta_f)$ conditional on $Z$ (and possibly $Y$) be invariant to the value
- of the nuisance parameter $Z$. Thus, we wish find a predictive function $f$ such that
-\begin{equation}\label{eqn:criterion}
-    p(f(X ; \theta_f) = s | z ) = p(f(X ; \theta_f) = s | z^\prime )
-\end{equation}
-for all $z,z^\prime \in  {\cal Z}$ and all values $s \in {\cal S}$ of $f(X ; \theta_f)$.
-In words, we are looking for a predictive function $f$
-which is a pivotal quantity \citep{degroot1986probability} with respect to the
-nuisance parameter. That is, such that  the distribution of $f(X_z; \theta_f)$
-is invariant with respect to the value $z$ of the nuisance.
-
-\kcnote{adjust text for conditional on Y as well}
-
-%Note that a function
-%$f$ for which Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
-%Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true,
-%since the sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = s \}$ and $\{
-%\omega' | f(X_{z^\prime}(\omega'); \theta_f) = s \}$ do not need to be the same
-%for the equality to hold. In order to simplify notations, and as only
-%Eqn.~\ref{eqn:criterion-measure} is of direct interest in this work, we denote
-%from here on the pivotal quantity criterion as
-
-% ==============================================================================
-
-\section{Method}
-\label{sec:method}
 
@@
 
+\section{Problem statement}
+\label{sec:problem}
+%
+%Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
+%sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
+%Let consider the multivariate random variables $X_z: \Omega \mapsto
+%\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
+%of the functional $X$ on a nuisance parameter $Z$, whose values $z \in {\cal Z}$  define a
+%parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
+%induce together a joint probability distribution $p(X,Y|Z=z)$, where the
+%conditional denotes $X_z$. For training, let further assume a finite set $\{
+%x_i, y_i, z_i \}_{i=1}^N$ of realizations $X_{z_i}(\omega_i), Y(\omega_i)$, for
+%$\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
+%is to learn a score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of
+%parameters $\theta_f$ (e.g., a neural network-based probabilistic classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
+%the cross-entropy). In addition, we require that $f(X_z ; \theta_f)$ should be
+%robust to the value $z$ of the nuisance parameter  -- which remains unknown at
+%test time. More specifically, we aim at building $f$ such that in the ideal case
+%\begin{equation}\label{eqn:criterion-true}
+%f(X_{z}(\omega) ; \theta_f) = f(X_{z^\prime}(\omega) ; \theta_f)
+%\end{equation} for all
+%samples $\omega \in \Omega$ and all $z, z^\prime$ pairs of values of the
+%nuisance parameter.
+%
+%Since in general we do not have training tuples $(X_{z}(\omega),
+%X_{z^\prime}(\omega))$ (for the same unknown $\omega$),
+
+We begin with a family of data generation processes $p(X,Y,Z)$, where $X$ are
+the data, $Y$ are the target labels, and $Z$ are the nuisance parameters that
+can be continuous or categorical. We assume training
+data $\{x_i, y_i, z_i\}_{i=1}^N$ for $X$ conditional on both $Y$ and $Z$. Our goal is to learn a predictive
+score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of parameters
+$\theta_f$ (e.g., a neural network-based probabilistic classifier) and
+minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g., the cross-entropy).
+
+In addition, we would like that inference based on $f(X ; \theta_f)$ should be
+robust to the value $z \in {\cal Z}$ of the nuisance parameter $Z$  -- which remains unknown at
+test time. A formal way of enforcing robustness is to require that the distribution of
+$f(X ; \theta_f)$ conditional on $Z$ (and possibly $Y$) be invariant to the value
+ of the nuisance parameter $Z$. Thus, we wish find a predictive function $f$ such that
+\begin{equation}\label{eqn:criterion}
+    p(f(X ; \theta_f) = s | z ) = p(f(X ; \theta_f) = s | z^\prime )
+\end{equation}
+for all $z,z^\prime \in  {\cal Z}$ and all values $s \in {\cal S}$ of $f(X ; \theta_f)$.
+In words, we are looking for a predictive function $f$
+which is a pivotal quantity \citep{degroot1986probability} with respect to the
+nuisance parameter. That is, such that  $f(X; \theta_f)$ and $Z$ are independent random variables.
+
+As stated in Eqn.~\ref{eqn:criterion}, the pivotal quantity criterion is
+considered in the case where $Y$ is marginalized out. In some situations
+however (see e.g., Sec.~\ref{sec:hep}), class conditional independence of $f(X;
+\theta_f)$ on the nuisance $Z$ is preferred, which can then be stated as
+requiring
+\begin{equation}\label{eqn:criterion-class}
+    p(f(X ; \theta_f) = s | z, y ) = p(f(X ; \theta_f) = s | z^\prime, y )
+\end{equation}
+for one or several specified values $y \in {\cal Y}$ of $Y$.
+
+%Note that a function
+%$f$ for which Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
+%Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true,
+%since the sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = s \}$ and $\{
+%\omega' | f(X_{z^\prime}(\omega'); \theta_f) = s \}$ do not need to be the same
+%for the equality to hold. In order to simplify notations, and as only
+%Eqn.~\ref{eqn:criterion-measure} is of direct interest in this work, we denote
+%from here on the pivotal quantity criterion as
+
+% ==============================================================================
+
+\section{Method}
+\label{sec:method}
+
+\begin{figure*}
+    \begin{minipage}{\linewidth}
+    \begin{algorithm}[H]
+    \caption{Adversarial training of a classifier $f$ against an adversary $r$.}
+
+    \begin{flushleft}
+        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;\\
+        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$;\\
+        {\it Hyper-parameters:} Number $T$ of training iterations,
+                                Number $K$ of gradient steps to update $r$.
+    \end{flushleft}
+
+    \label{alg:adversarial-training}
+    \begin{algorithmic}[1]
+        \For{$t=1$ to $T$}
+            \For{$k=1$ to $K$} \Comment{Update $r$}
+                \State{Sample minibatch $\{x_m, z_m, s_m = f(x_m;\theta_f) \}_{m=1}^M$} of size $M$;
+                \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
+                $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|s_m)  ;$$}
+            \EndFor
+            \State{Sample minibatch $\{x_m, y_m, z_m, s_m = f(x_m;\theta_f)  \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
+            \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
+            $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|s_m)  \right],$$
+            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $1(y_m=0)(1-s_m) + 1(y_m=1)s_m$;}
+        \EndFor
+    \end{algorithmic}
+    \end{algorithm}
+    \end{minipage}
+\end{figure*}
+
 Joint training of adversarial networks was first proposed by \citep{goodfellow2014generative} as a
@@
 alternatively for solving Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.
-
-\begin{figure*}
-    \begin{minipage}{\linewidth}
-    \begin{algorithm}[H]
-    \caption{Adversarial training of a classifier $f$ against an adversary $r$.}
-
-    \begin{flushleft}
-        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;\\
-        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$;\\
-        {\it Hyper-parameters:} Number $T$ of training iterations,
-                                Number $K$ of gradient steps to update $r$.
-    \end{flushleft}
-
-    \label{alg:adversarial-training}
-    \begin{algorithmic}[1]
-        \For{$t=1$ to $T$}
-            \For{$k=1$ to $K$} \Comment{Update $r$}
-                \State{Sample minibatch $\{x_m, z_m, s_m = f(x_m;\theta_f) \}_{m=1}^M$} of size $M$;
-                \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
-                $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|s_m)  ;$$}
-            \EndFor
-            \State{Sample minibatch $\{x_m, y_m, z_m, s_m = f(x_m;\theta_f)  \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
-            \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
-            $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|s_m)  \right],$$
-            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $1(y_m=0)(1-s_m) + 1(y_m=1)s_m$;}
-        \EndFor
-    \end{algorithmic}
-    \end{algorithm}
-    \end{minipage}
-\end{figure*}
+Finally, in the case of a class conditional pivot, the settings are the
+same, except that the adversarial term ${\cal L}_r(\theta_f, \theta_r)$ is restricted to $Y=y$.
 
@@
 its optimum given $f$ (e.g., by setting $K$ sufficiently high) and if $f$ is
-updated to improve ${\cal L}_f(\theta_f) -  H({Z|f(X;\theta_f)})$, then $f$
-should converge to a classifier which is both optimal and pivotal, provided such
-a classifier exists. A formal proof of convergence of the alternating
-stochastic gradient descent procedure of
-Algorithm~\ref{alg:adversarial-training} remains however to be proven.
+updated to improve ${\cal L}_f(\theta_f) -  H({Z|f(X;\theta_f)})$ with
+sufficiently small steps, then $f$ should converge to a classifier which is both
+optimal and pivotal, provided such a classifier exists. A formal proof of
+convergence of the alternating stochastic gradient descent procedure of
+Algorithm~\ref{alg:adversarial-training} in the case where a finite number $K$
+of steps is taken for $r$ remains however to be proven.
 

### 2016-10-31 tex/paper.tex
save

@@
 property on a predictive model, which allows one to tune the tradeoff between
-power and robustness. We derive theoretical results proving that the proposed
-procedure converges towards a predictive model that is both optimal and independent on
-the nuisance parameters, and demonstrate the effectiveness of this approach with a
-toy and real world examples.
+power and robustness. Under the assumption of existence, we derive theoretical
+results proving that the proposed procedure converges towards a predictive model
+that is both optimal and independent on the nuisance parameters, and demonstrate
+the effectiveness of this approach with a toy and real world examples.
 
@@
 In this setting, the nuisance parameter $Z$ is considered as a random variable
-of prior $p(z)$ (for $z \in {\cal Z}$), and our goal is to find a function
+of prior $p(Z)$, and our goal is to find a function
 $f(\cdot;\theta_f)$ such that $f(X;\theta_f)$ and $Z$ are independent random
@@
 
-On many practical problems, let us note that such a classifier may
-not exist because the nuisance parameter directly shapes the decision boundary.
-In this case, the lower bound $ H({Y|X}) - H(Z) < {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal classifier or a
+On many practical problems, the assumption of existence of such a classifier may
+not hold because the nuisance parameter directly shapes the decision boundary.
+In this case, the lower bound $$ H({Y|X}) - H(Z) < {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})$$ is strict: $f$ can either be an optimal classifier or a
 pivotal quantity, but not both simultaneously. In this situation, it is natural
@@
 of domain adaptation
-\citep{blitzer2006domain,pan2011domain,gopalan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,ganin2014unsupervised},
+\citep{blitzer2006domain,pan2011domain,gopalan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,ajakan2014domain,ganin2014unsupervised},
 where the goal is often stated as trying to learn a domain-invariant
@@
 adversarial minimax setup composed of a classifier and an adversary
-discriminator. Following this line of work, our method can be regarded as a
+discriminator. Following this line of work and without making assumptions on the prior $p(Z)$, our method can be regarded as a
 generalization that also supports the continuous case, which can be viewed as

### 2016-10-31 tex/paper.tex
acks
save
resize fig 2

@@
 that is both optimal and independent on the nuisance parameters, and demonstrate
-the effectiveness of this approach with a toy and real world examples.
+the effectiveness of this approach with a toy and physics examples.
 
@@
 trade-off between power and robustness. Finally, we demonstrate the
-effectiveness of this approach with a toy and real world examples.
+effectiveness of this approach with a toy and physics examples.
 
@@
 towards a model that is both optimal and robust. The  effectiveness of the
-approach is then demonstrated in sections \ref{sec:toy} and \ref{sec:hep} with a
-toy and real world examples. In Sec.~\ref{sec:related}, we relate our work to
+approach is then demonstrated empirically in sections \ref{sec:toy} and \ref{sec:hep}.
+In Sec.~\ref{sec:related}, we relate our work to
 close contributions from statistical inference, domain adaptation and fairness
@@
     \begin{center}
-        \includegraphics[width=0.48\textwidth]{figures/f-plain.pdf}
-        \includegraphics[width=0.48\textwidth]{figures/surface-plain.pdf}\\
-        \vspace{-0.8cm}
-        \includegraphics[width=0.48\textwidth]{figures/f-adversary.pdf}
-        \includegraphics[width=0.48\textwidth]{figures/surface-adversary.pdf}
+        \includegraphics[width=0.245\textwidth]{figures/f-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{figures/surface-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{figures/f-adversary.pdf}
+        \includegraphics[width=0.245\textwidth]{figures/surface-adversary.pdf}
     \end{center}
@@
     \caption{Toy example.
-    (Upper left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
+    (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
        when $f$ is built without adversarial training. The resulting densities
        are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
-    (Upper right) The corresponding decision surface, highlighting
+    (Middle left) The corresponding decision surface, highlighting
        the fact that samples are easier to classify for values of $Z$ above to $\sigma$,
        hence partially explaining the dependency.
-    (Lower left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
+    (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
        built with adversarial training, as outlined in Section~\ref{sec:method}.
@@
        small dependency on $Z$.
-    (Lower right) The corresponding decision surface, illustrating how adversarial
+    (Right) The corresponding decision surface, illustrating how adversarial
        training bends the decision function vertically to erase the dependency on $Z$.
@@
 quarks (the signal).  We reuse the datasets used in reference
-\citep{guest2016jet}.  Briefly summarizing, jets are built with the
+\citep{baldi2016jet}.  Briefly summarizing, jets are built with the
 anti-$k_T$ algorithm \citep{Cacciari:2008gp}
@@
     \vspace{-1cm}
-    \caption{Approximate median significance as a function of the decision threshold
+    \caption{Physics example. Approximate median significance as a function of the decision threshold
              on the output of $f$.  As shown at $\lambda=10$, trading
@@
 \begin{acknowledgments}
+    KC and GL are both supported through NSF ACI-1450310, additionally KC is supported
+    through PHY-1505463 and PHY-1205376.
 \end{acknowledgments}

### 2016-10-31 tex/paper.tex
conclusions (work in progress)

@@
 pivot -- a quantity whose distribution is invariant to the unknown value of the
+(categorical or continuous)
 nuisance parameters that parametrize the family of data generation processes. In
-this work, we introduce an adversarial training procedure to enforce the pivotal
+this work, we introduce a flexible adversarial training procedure to enforce the pivotal
 property on a predictive model, which allows one to tune the tradeoff between
@@
     \end{center}
-    \vspace{-1cm}
+    \vspace{-0.5cm}
     \caption{Toy example.
@@
 
+In this work, we proposed a flexible learning procedure for building a
+predictive model that is independent on continuous or categorical nuisance
+parameters by jointly training two neural networks in an adversarial fashion.
+From a theoretical perspective, we showed that the approach leads to a
+predictive model that is both optimal and pivotal (if that models exists) or for
+which one can tune the trade-off between power and robustness. From an empirical
+point of view, we confirmed the effectiveness of our method on a toy and physics
+example.
+
+\glnote{Add future works.}
+\glnote{Mention limitations of the current algorithm.}
+
 % ==============================================================================

### 2016-10-31 tex/paper.tex
save

--- 2016-10-31 tex/paper.tex
@@
 pivot -- a quantity whose distribution is invariant to the unknown value of the
-nuisance parameters that parametrize the family of data generation processes. In
-this work, we introduce an adversarial training procedure to enforce the pivotal
-property on a predictive model, which allows one to tune the tradeoff between
-power and robustness. Under the assumption of existence, we derive theoretical
-results proving that the proposed procedure converges towards a predictive model
+(categorical or continuous)
+nuisance parameters that parametrizes the family of data generation processes. In
+this work, we introduce a flexible adversarial training procedure to enforce the pivotal
+property on a predictive model, which allows one to tune the trade-off between
+power and robustness. We derive theoretical
+results proving that the proposed procedure tends towards a minimax solution corresponding to a predictive model
 that is both optimal and independent on the nuisance parameters, and demonstrate
-the effectiveness of this approach with a toy and physics examples.
+its effectiveness  with a toy and physics examples.
 
@@
 Another common situation is that there are several plausible data generation
-processes, specificied as a family of data generation processes
+processes, specified as a family of data generation processes
 parametrized by nuisance parameters, which can be categorical or continuous.
@@
 $p_{\theta_r}(z_j|f(X;\theta_f)=s)$. Similarly, if $Z$ takes continuous values and
-if we assume some parameteric distribution ${\cal P}$ for $Z|f(X;\theta_f)=s$
+if we assume some parametric distribution ${\cal P}$ for $Z|f(X;\theta_f)=s$
 (e.g., a mixture of gaussians, as modeled with a mixture
@@
 \end{equation}
-that we optimize by finding the saddle point $(\smash{\hat\theta_f}, \smash{\hat\theta_r})$ such that
-\begin{align}
-    \smash{\hat\theta_f} &= \arg \min_{\theta_f} E(\theta_f, \smash{\hat\theta_r}), \label{eqn:min_thetaf} \\
-    \smash{\hat\theta_r} &= \arg \max_{\theta_r} E(\smash{\hat\theta_f}, \theta_r) \label{eqn:max_thetar}.
-\end{align}
+that we optimize by finding the minimax solution
+\begin{equation}\label{eqn:min_thetaf}
+    \smash{\hat\theta_f, \hat\theta_f} = \arg \min_{\theta_f} \max_{\theta_r} E(\theta_f, \theta_r).
+\end{equation}
 Without loss of generality, the adversarial training procedure to obtain
@@
 The optimization algorithm consists in using stochastic gradient descent
-alternatively for solving Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar}.
+alternatively for solving Eqn.~\ref{eqn:min_thetaf}.
 Finally, in the case of a class conditional pivot, the settings are the
@@
 \begin{proposition}\label{prop:2}
-If there exists a saddle point $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$
-for Eqn.~\ref{eqn:min_thetaf} and \ref{eqn:max_thetar} such that
+If there exists a minimax solution $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$
+for Eqn.~\ref{eqn:min_thetaf} such that
 $E(\hat\theta_f, \hat\theta_r) = H({Y|X}) - H(Z)$, then
@@
 value $z_0$ of the nuisance parameter, and then propagate uncertainty
- by estimating $p(f(x)|z)$ with a parameterized calibration
+ by estimating $p(f(x)|z)$ with a parametrized calibration
 procedure. Clearly, this classifier is however not optimal for $z \neq z_0$.
@@
 previously in Section~\ref{sec:toy}.
-As an alternative, parameterized
+As an alternative, parametrized
 classifiers~\citep{cranmer2015approximating,Baldi:2016fzo} directly take
@@
 marginalized out, but otherwise often limits the range of their applications. In
-practice, parameterized classifiers  are also computationally expensive to build
+practice, parametrized classifiers  are also computationally expensive to build
 and evaluate. In particular, calibrating their decision function, i.e.
@@
 handling infinitely many domains, provided they can be continuously
-parameterized, or as enforcing fairness over continuous attributes.
+parametrized, or as enforcing fairness over continuous attributes.
 
@@
 \label{sec:conclusions}
+
+In this work, we proposed a flexible learning procedure for building a
+predictive model that is independent on continuous or categorical nuisance
+parameters by jointly training two neural networks in an adversarial fashion.
+From a theoretical perspective, we motivated the proposed algorithm by showing
+that the minimax value  of its value function corresponds to a predictive model
+that is both optimal and pivotal (if that models exists) or for which one can
+tune the trade-off between power and robustness. From an empirical point of
+view, we confirmed the effectiveness of our method on a toy and physics examples.
+
+As previously discussed, the proposed solution can be used in any situation
+where the training data may not be representative of the real data the
+predictive model will be applied to in practice. In the scientific context, this
+includes situations where the data generation processes are associated to
+systematic uncertainties (e.g., as in high energy physics), and which would be
+certainly interesting to revisit in light of the proposed solution. More generally,
+the approach also extends to cases where independence of the predictive
+model with respect to random variables is desired, as in domain adaptation
+or fairness in classification.
 

### 2016-10-31 tex/paper.tex
add axis labels

@@
 respectively set to expected value of the negative log-likelihood of $Y|X$ under
-$f$ and of $Z|f(X;\theta_f)$ under $r$, the procedure converges to a classifier
-$f$ which is a pivotal quantity in the sense of Eqn.~\ref{eqn:criterion}.
+$f$ and of $Z|f(X;\theta_f)$ under $r$, the minimax solution of Eqn.~\ref{eqn:min_thetaf}
+corresponds to a classifier
+$f$ which is a pivotal quantity.
 
@@
 nodes with ReLU activations, followed  by a dense output layer with a single
-node with a sigmoid activation. As shown in the upper plots of
+node with a sigmoid activation. As shown in 
 Fig.~\ref{fig:toy}, the resulting classifier is not pivotal, as the
@@
 the decision function so it that becomes almost independent on the nuisance
-parameter, as shown in the lower plots of Fig.~\ref{fig:toy}. In particular,
+parameter, as shown in Fig.~\ref{fig:toy}. In particular,
 the conditional probability densities of the decision scores $f(X;\theta_f)$ are
@@
     \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
-    \vspace{-1cm}
     \caption{Physics example. Approximate median significance as a function of the decision threshold

### 2016-10-31 tex/paper.tex
typo

@@
     (Middle left) The corresponding decision surface, highlighting
-       the fact that samples are easier to classify for values of $Z$ above to $\sigma$,
+       the fact that samples are easier to classify for values of $Z$ above $\sigma$,
        hence partially explaining the dependency.

### 2016-10-31 tex/paper.tex
save

@@
 includes situations where the data generation processes are associated to
-systematic uncertainties (e.g., as in high energy physics), and which would be
-certainly interesting to revisit in light of the proposed solution. More generally,
+systematic uncertainties (e.g., as in high energy physics), and which would certainly be
+ interesting to revisit in light of the proposed solution. More generally,
 the approach also extends to cases where independence of the predictive
 model with respect to random variables is desired, as in domain adaptation
-or fairness in classification.
+or fairness in classification. Again, trying our algorithm on these
+use cases is certaintly worth investigating.
 

### 2016-10-31 tex/paper.tex
Misc changes

@@
 uncertainty.
-Another common situation is that there are several plausible data generation
+A more generic situation is that there are several plausible data generation
 processes, specified as a family of data generation processes
@@
 trade-off between power and robustness. Finally, we demonstrate the
-effectiveness of this approach with a toy and physics examples.
+effectiveness of this approach with a toy example and an example from particle physics.
 
@@
 
-We begin with a family of data generation processes $p(X,Y,Z)$, where $X$ are
-the data, $Y$ are the target labels, and $Z$ are the nuisance parameters that
+We begin with a family of data generation processes $p(X,Y,Z)$, where $X \in \mathcal{X}$ are
+the data, $Y\in \mathcal{Y}$ are the target labels, and $Z\in \mathcal{Z}$ are the nuisance parameters that
 can be continuous or categorical. We assume training
-data $\{x_i, y_i, z_i\}_{i=1}^N$ for $X$ conditional on both $Y$ and $Z$. Our goal is to learn a predictive
-score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of parameters
+data $\{x_i, y_i, z_i\}_{i=1}^N$ for $X$ conditional on both $Y$ and $Z$. 
+Let us assume that prior to incorporating the effect of uncertainty in $Z$, 
+our goal is to learn a 
+regression function $f : \mathcal{X} \mapsto {\cal S}$ of parameters
 $\theta_f$ (e.g., a neural network-based probabilistic classifier) and
@@
 
-In addition, we would like that inference based on $f(X ; \theta_f)$ should be
+We augment our initial objective so that inference based on $f(X ; \theta_f)$ will be
 robust to the value $z \in {\cal Z}$ of the nuisance parameter $Z$  -- which remains unknown at
@@
 which is a pivotal quantity \citep{degroot1986probability} with respect to the
-nuisance parameter. That is, such that  $f(X; \theta_f)$ and $Z$ are independent random variables.
+nuisance parameter. This implies that  $f(X; \theta_f)$ and $Z$ are independent random variables.
 
@@
 \end{equation}
-for one or several specified values $y \in {\cal Y}$ of $Y$.
+for one or several specified values $y \in {\cal Y}$.
 
@@
     \begin{center}
-        \includegraphics[width=0.245\textwidth]{figures/f-plain.pdf}
-        \includegraphics[width=0.245\textwidth]{figures/surface-plain.pdf}
-        \includegraphics[width=0.245\textwidth]{figures/f-adversary.pdf}
-        \includegraphics[width=0.245\textwidth]{figures/surface-adversary.pdf}
+        \includegraphics[width=0.23\textwidth]{figures/f-plain.pdf}
+        \includegraphics[width=0.23\textwidth]{figures/surface-plain.pdf}
+        \includegraphics[width=0.23\textwidth]{figures/f-adversary.pdf}
+        \includegraphics[width=0.23\textwidth]{figures/surface-adversary.pdf}
     \end{center}
@@
 \begin{align}
-    x &\sim {\cal N}((0,0), \begin{bmatrix}
+    x &\sim {\cal N}\left ((0,0), \begin{bmatrix}
                               1 & -0.5 \\
                               -0.5 & 1
-                            \end{bmatrix}) &\text{ when } Y=0, \\
-    x &\sim {\cal N}((1,1+Z),  \begin{bmatrix}
+                            \end{bmatrix}\right) &\text{ when } Y=0, \\
+    x &\sim {\cal N}\left ((1,1+Z),  \begin{bmatrix}
                               1 & 0 \\
                               0 & 1
-                             \end{bmatrix}) &\text{ when } Y=1.
+                             \end{bmatrix}\right) &\text{ when } Y=1.
 \end{align}
@@
 
-To account for systematic uncertainties, experimentalists in high energy physics
-typically take as fixed a classifier $f$ built from training data for a nominal
-value $z_0$ of the nuisance parameter, and then propagate uncertainty
- by estimating $p(f(x)|z)$ with a parametrized calibration
-procedure. Clearly, this classifier is however not optimal for $z \neq z_0$.
-To overcome this issue, the classifier $f$ is sometimes built instead on a mixture
-of training data generated from several nominal values $z_0, z_1, \dots$ of the nuisance.
-While this certainly improves with respect to classification performance,
-there is however no guarantee that the resulting classifier is pivotal, as shown
-previously in Section~\ref{sec:toy}.
-As an alternative, parametrized
-classifiers~\citep{cranmer2015approximating,Baldi:2016fzo} directly take
-(nuisance) parameters as additional input variables, hence ultimately providing
-the most statistically powerful approach for incorporating the effect of
-systematics on the underlying classification task.  As argued in
-\citep{Neal:2007zz}, such classifiers can however not be used on real data since
-the correct value $z$ of the nuisance often remains unknown. This is typically
-not an issue in the context of parameter
-inference~\citep{cranmer2015approximating}, where nuisance parameters are
-marginalized out, but otherwise often limits the range of their applications. In
-practice, parametrized classifiers  are also computationally expensive to build
-and evaluate. In particular, calibrating their decision function, i.e.
-approximating $p(f(x,z)|z)$ as a continuous function of $z$, remains an open
-challenge. By contrast, constraining $f$ to be pivotal yields a classifier which
-may not be optimal with respect to $Y|X,Z$, as discussed in
-Section~\ref{sec:theory}, but that can otherwise be used in a wider range of
-applications, since knowing the correct value $z$ of the nuisance is not
-necessary. Similarly, calibration needs to be carried out only once, since  the
-dependence on the nuisance is now built-in.
-
 In machine learning, learning a pivotal quantity can be related to the problem
@@
 adversarial minimax setup composed of a classifier and an adversary
-discriminator. Following this line of work and without making assumptions on the prior $p(Z)$, our method can be regarded as a
-generalization that also supports the continuous case, which can be viewed as
-handling infinitely many domains, provided they can be continuously
-parametrized, or as enforcing fairness over continuous attributes.
+discriminator. 
+Following this line of work and without making assumptions on the prior $p(Z)$, our method can be regarded as a
+generalization that also supports a continuously parametrized family of domains or as enforcing fairness over continuous attributes.
+
+\bigskip
+
+To account for systematic uncertainties, experimentalists in high energy physics
+typically take as fixed a classifier $f$ built from training data for a nominal
+value $z_0$ of the nuisance parameter, and then propagate uncertainty
+ by estimating $p(f(x)|z)$ with a parametrized calibration
+procedure. Clearly, this classifier is however not optimal for $z \neq z_0$.
+To overcome this issue, the classifier $f$ is sometimes built instead on a mixture
+of training data generated from several plausible values $z_0, z_1, \dots$ of the nuisance parameter.
+While this certainly improves classification performance with respect to the marginal model $p(X,Y)$,
+there is no reason to expect the resulting classifier to be pivotal, as shown
+previously in Section~\ref{sec:toy}.
+As an alternative, parametrized
+classifiers~\citep{cranmer2015approximating,Baldi:2016fzo} directly take
+(nuisance) parameters as additional input variables, hence ultimately providing
+the most statistically powerful approach for incorporating the effect of
+systematics on the underlying classification task.  As argued in
+\citep{Neal:2007zz}, it is not obvious how such classifiers can be used on real data since
+the correct value $z$ of the nuisance often remains unknown. This is
+not an issue in the context of parameter
+inference~\citep{cranmer2015approximating}, where nuisance parameters are
+eliminated via optimization or marginalization.
+%, but otherwise often limits the range of their applications. 
+In practice, parametrized classifiers  are  computationally expensive to build
+and evaluate. In particular, calibrating their decision function, i.e.
+approximating $p(f(x,z)|y,z)$ as a continuous function of $z$, remains an open
+challenge. By contrast, constraining $f$ to be pivotal yields a classifier which
+may not be optimal with respect to $Y|X,Z$, as discussed in
+Section~\ref{sec:theory}, but that can otherwise be directly used in a wider range of
+applications, since the dependence on the nuisance parameter $z$ has already been eliminated. 
+Moreover, if $f$ is sufficiently close to being pivotal, then calibration only needs to be carried out once.
+
+\bigskip
+
+%Following this line of work and without making assumptions on the prior $p(Z)$, our method can be regarded as a
+%generalization that also supports the continuous case, which can be viewed as
+%handling infinitely many domains, provided they can be continuously
+%parametrized, or as enforcing fairness over continuous attributes.
 
@@
 In this work, we proposed a flexible learning procedure for building a
-predictive model that is independent on continuous or categorical nuisance
+predictive model that is independent of continuous or categorical nuisance
 parameters by jointly training two neural networks in an adversarial fashion.
@@
 tune the trade-off between power and robustness. From an empirical point of
-view, we confirmed the effectiveness of our method on a toy and physics examples.
+view, we confirmed the effectiveness of our method on a toy example and particle physics example.
 

### 2016-10-31 tex/paper.tex
large font for fig2
larger figures
Reworked HEP example
I reworded the HEP example to have a little more motivation at the
beginning and a more developed connection to the hypothesis test,
motivation for AMS, motivation for pivoting on y=0, and optimization of
the hyper parameter lambda.

@@
     \begin{center}
-        \includegraphics[width=0.23\textwidth]{figures/f-plain.pdf}
-        \includegraphics[width=0.23\textwidth]{figures/surface-plain.pdf}
-        \includegraphics[width=0.23\textwidth]{figures/f-adversary.pdf}
-        \includegraphics[width=0.23\textwidth]{figures/surface-adversary.pdf}
+        \includegraphics[width=0.245\textwidth]{figures/f-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{figures/surface-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{figures/f-adversary.pdf}
+        \includegraphics[width=0.245\textwidth]{figures/surface-adversary.pdf}
     \end{center}
@@
 
-Jets, or collimated sprays of particles produced by high energy quarks and
-gluons, are produced ubiquitously at high energy colliders like the LHC \citep{LHCMachine}.
-In a wide array of new physics models, new highly
-massive particles can decay through known heavy particles in the Standard Model (SM),
-like the $W$, $Z$, Higgs bosons, or top quarks which subsequently decay to
-multiple quarks.  When these heavy SM particles have significant energy,
-the Lorentz boost causes their quark decay products to merge into a single jet
-which has a rich internal substructure (see e.g.
+%Experiments at high energy colliders like the LHC \citep{LHCMachine}
+%are searching for evidence of new particles beyond the ones
+%described by the so-called Standard Model (SM) of particle physics.
+%A wide array of extended theories predict the existence of new massive
+%particles that would decay to known heavy particles in the Standard Model,
+%like the $W$, $Z$, Higgs boson, or top quark, which subsequently decay to
+%multiple quarks.
+%Finally, those quarks would produce collimated sprays of particles known as \textit{jets}.
+%However, jets are produced ubiquitously at high energy colliders through
+%more mundane processes in the SM, which leads to a challenging classification problem
+%that is beset with a number of sources of systematic uncertainty.
+
+Experiments at high energy colliders like the LHC \citep{LHCMachine}
+are searching for evidence of new particles beyond the ones
+described by the well established Standard Model (SM) of particle physics.
+%Searches for these new particles in the LHC data involve a challenging
+%classification problem that is beset with a number of sources of systematic uncertainty.
+%
+A wide array of extended theories predict the existence of new massive
+particles that would decay to known particles in the SM such as the $W$ boson.
+The $W$ boson is unstable and can decay to
+two quarks, each of which produce collimated sprays of particles known as \textit{jets}
+(collectively denoted $W\to jj$).
+If the exotic particle is heavy, then the $W$ boson will be moving very fast,
+and  relativistic effects will cause the two jets from its decay to merge into a single `fat jet'.
+These fat jets have a rich internal substructure (see e.g.
 \citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
-Distinguishing these boosted jets produced by heavy SM particles from
-vanilla quarks and  gluons is a fundamental challenge in searching for signs of
-new high mass particles. Challenging in its own right, this classification
-problem is made all the more difficult by the presence of pileup, or multiple
-simultaneous proton-proton interactions occurring at the same time as the primary
-interaction.  These pileup interactions produce additional particles that can
-contribute significant energies to jets unrelated to the underlying
-discriminating information. Thus pileup acts as a source of noise in the
-classification problem.
+However, jets are produced ubiquitously at high energy colliders through
+more mundane processes in the SM, which leads to a challenging classification problem
+that is beset with a number of sources of systematic uncertainty.
+%Distinguishing these ($y=1$) fat jets produced by heavy SM particles from
+%the copious background of normal ($y=0$) jets is a fundamental challenge in searching
+%for signs of new high mass particles.
+
+%A wide array of extended theories predict the existence of new massive
+%particles that would decay to known heavy particles in the Standard Model
+%(e.g. the $W$ and $Z$ bosons), which subsequently decay to
+%multiple quarks.
+%Those quarks subsequently produce collimated sprays of particles known as \textit{jets} (denoted $j$).
+%However, jets are produced ubiquitously at high energy colliders through
+%more mundane processes in the SM, which leads to a challenging classification problem
+%that is beset with a number of sources of systematic uncertainty.
+
+%Consider the binary classification problem where $y=0$ corresponds to data produced from the
+%ubiquitous SM process $pp \to jj$ and $y=1$ corresponds to data produced from an exotic
+%process $pp \to W' \to W(\to jj) Z(\to jj)$ predicted by an extended theory.
+%When the mass of the $W'$ is much larger than the mass of the $W$ and $Z$ bosons,
+%then the $W$ and $Z$ particles are moving very fast and the relativistic Lorentz boost
+%causes the two jets from their decay to merge into a single `fat jet',
+
+%Consider the search for an exotic particle whose decay leads to a $W$ boson
+%that subsequently decays to two jets via $W\to jj$.  If the $W$ boson is moving
+%very fast, then the relativistic Lorentz boost
+%causes the two jets from its decay to merge into a single `fat jet',
+%which has a rich internal substructure (see e.g.
+%\citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
+%%We can cast this as a binary classification problem where $y=0$
+%%corresponds to data produced from the ubiquitous SM process
+%%that produce jets and $y=1$  corresponds to data produced from an exotic
+%%process that gives rise to fat jets from $W\to jj$.
+
+%binary classification problem where $y=0$ corresponds to data produced from the
+%ubiquitous SM process $pp \to jj$ and $y=1$ corresponds to data produced from an exotic
+%process $pp \to W' \to W(\to jj) Z(\to jj)$ predicted by an extended theory.
+%When the mass of the $W'$ is much larger than the mass of the $W$ and $Z$ bosons,
+%then the $W$ and $Z$ particles are moving very fast and the relativistic Lorentz boost
+%causes the two jets from their decay to merge into a single `fat jet',
+%which has a rich internal substructure (see e.g.
+%\citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
+
+%Distinguishing these ($y=1$) fat jets produced by heavy SM particles from
+%the copious background of normal ($y=0$) jets is a fundamental challenge in searching
+%for signs of new high mass particles.
 
@@
 studies (see \citep{Khachatryan:2014vla,ATL-PHYS-PUB-2015-033,wbosonATLAS} and
-references therein): we aim to discriminate between jets produced by quark or gluon fragmentation (the
-background) and  boosted $W$ boson jets containing a collimated pair of
-quarks (the signal).  We reuse the datasets used in reference
-\citep{baldi2016jet}.  Briefly summarizing, jets are built with the
+references therein): we aim to discriminate between  the normal jets produced copiously
+at the LHC ($y=0$) and  the fat jets coming from an exotic process ($y=1$).
+We reuse the datasets used in reference
+\citep{baldi2016jet}.  For completeness, the jets are built with the
 anti-$k_T$ algorithm \citep{Cacciari:2008gp}
@@
 
-As nuisance parameter, we consider events without pileup ($Z=0$) and
-events with pileup ($Z=1$), for which an average number of $\langle \mu \rangle =
-50$ unrelated additional pileup interactions are overlaid. Our goal is to build an accurate
-classifier, for which we also want to minimize the effects due to the
-uncertainties on the nuisance. More specifically, we choose to recast the
-classification problem as a hypothesis test between signal+background (jets
-originating from $W$ bosons or from quarks and gluons) and background
-only (jets originating from quarks and gluons only) and tune the decision
-threshold of our classifier by maximizing its approximate median significance
-(AMS), when uncertainties in the background are taken into account (see Eqn.~20
-of \citep{adam2014higgs}). Our motivation is that reducing the effects of
-uncertainties by requiring independence of $Z$ with the classifier output $f(X;\theta_f)$ should
-allow for a larger maximum significance.
+Challenging in its own right, this classification
+problem is made all the more difficult by the presence of pileup, or multiple
+ proton-proton interactions occurring simultaneously with the primary
+interaction.  These pileup interactions produce additional particles that can
+contribute significant energies to jets unrelated to the underlying
+discriminating information. The number of pileup interactions can vary with
+the running conditions of the collider, and we want the classifier to be robust
+to these conditions. Taking some liberty, we consider an extreme case with
+a categorical nuisance parameter, where $Z=0$ corresponds to events without pileup
+and $Z=1$ corresponds to events with pileup, for which there are an average  of
+50 independent pileup interactions overlaid.
 
-To minimize the effects of $Z$ in the background events, we train a classifier
+We do not expect that we will be able to find a function $f$ that simultaneously
+minimizes the classification loss $L_f$ and is pivotal. Thus, we need to optimize
+the hyperparameter $\lambda$ of Eqtn.~\ref{eqn:vf-lambda} with respect to
+a higher-level objective. In this case, the natural higher-level context is a
+hypothesis test of a null hypothesis with no $Y=1$ events against an
+alternate hypothesis that is a mixture of $Y=0$ and $Y=1$ events.
+In the absence of systematic uncertainties optimizing $L_f$ simultaneously
+optimizes the power of a classical hypothesis test in the Neyman-Pearson sense.
+When we include systematic uncertainties we will need to balance the
+classification performance against the robustness to uncertainty in $Z$.
+Since we are still performing a hypothesis test against the null, we only
+wish to impose the pivotal property on $Y=0$ events. Furthermore,
+we use the Approximate Median Significance (AMS), that was used
+in the Higgs boson machine learning challenge that is a natural generalization
+of the power of a hypothesis test when systematic uncertainties are taken into account
+(see Eqn.~20 of \citep{adam2014higgs}).
+
+%Our goal is to build an accurate classifier, for which we also want to minimize the effects due to the
+%uncertainties on the nuisance. More specifically, we choose to recast the
+%classification problem as a hypothesis test between signal+background (jets
+%originating from $W$ bosons or from quarks and gluons) and background
+%only (jets originating from quarks and gluons only) and tune the decision
+%threshold of our classifier by maximizing its approximate median significance
+%(AMS), when uncertainties in the background are taken into account (see Eqn.~20
+%of \citep{adam2014higgs}). Our motivation is that reducing the effects of
+%uncertainties by requiring independence of $Z$ with the classifier output $f(X;\theta_f)$ should
+%allow for a larger maximum significance.
+
+For several values of $\lambda$, we train a classifier
 using  Algorithm~\ref{alg:adversarial-training} but consider the adversarial
@@
 at $\lambda=0$ when building a classifier on data sampled from $p(X,Y,Z)$), the
-maximum significance peaks at $7$. By contrast, as the independence constraint
+AMS peaks at $7$. By contrast, as the pivotal constraint
 is made stronger (for $\lambda > 0$) the AMS peak moves higher, with a maximum
 value around  $7.8$ for $\lambda=10$. In other words, trading classification
-accuracy for independence with respect to pileup results in a positive total net
-benefit in terms of statistical significance. Setting $\lambda$ too high however
-(e.g. $\lambda=500$) results in a decrease of the maximum significance, by
-focusing the capacity of $f$ too strongly on independence, at the expense of
-classification accuracy. As demonstrated in this example, controlling the
-classification versus pivot trade-off through $\lambda$ therefore gives us a
-principled and effective approach for maximizing significance by desensitizing
-the classifier output $f(X;\theta_f)$ in the most beneficial way.
+accuracy for robustness to pileup results in a net
+benefit in terms of the power of the resulting hypothesis test. Setting $\lambda$ too high however
+(e.g. $\lambda=500$) results in a decrease of the maximum AMS, by
+focusing the capacity of $f$ too strongly on independence with $Z$, at the expense of
+classification accuracy. As demonstrated in this example, optimizing $\lambda$
+gives us a principled and effective approach to control the tradeoff between
+classification accuracy and robustness that ultimately maximizes the
+power of the enveloping hypothesis test.
+%significance by desensitizing
+%the classifier output $f(X;\theta_f)$ in the most beneficial way.
 
@@
 generalization that also supports a continuously parametrized family of domains or as enforcing fairness over continuous attributes.
-
-\bigskip
 
@@
 Section~\ref{sec:theory}, but that can otherwise be directly used in a wider range of
-applications, since the dependence on the nuisance parameter $z$ has already been eliminated. 
+applications, since the dependence on the nuisance parameter $Z$ has already been eliminated.
 Moreover, if $f$ is sufficiently close to being pivotal, then calibration only needs to be carried out once.
-
-\bigskip
 
@@
 
+\bigskip
+
 % ==============================================================================
@@
 \begin{acknowledgments}
+
     KC and GL are both supported through NSF ACI-1450310, additionally KC is supported
     through PHY-1505463 and PHY-1205376.
+
 \end{acknowledgments}
+
+\vfill
 

### 2016-10-31 tex/paper.tex
save

@@
 references therein): we aim to discriminate between  the normal jets produced copiously
-at the LHC ($y=0$) and  the fat jets coming from an exotic process ($y=1$).
+at the LHC ($Y=0$) and  the fat jets coming from an exotic process ($Y=1$).
 We reuse the datasets used in reference

### 2016-10-31 tex/paper.tex
save
some trimming for better layout

@@
 minimizes the classification loss $L_f$ and is pivotal. Thus, we need to optimize
-the hyperparameter $\lambda$ of Eqtn.~\ref{eqn:vf-lambda} with respect to
+the hyperparameter $\lambda$ of Eqn.~\ref{eqn:vf-lambda} with respect to
 a higher-level objective. In this case, the natural higher-level context is a
@@
 alternate hypothesis that is a mixture of $Y=0$ and $Y=1$ events.
-In the absence of systematic uncertainties optimizing $L_f$ simultaneously
+In the absence of systematic uncertainties, optimizing $L_f$ simultaneously
 optimizes the power of a classical hypothesis test in the Neyman-Pearson sense.
-When we include systematic uncertainties we will need to balance the
+When we include systematic uncertainties we need to balance the
 classification performance against the robustness to uncertainty in $Z$.
@@
 gives us a principled and effective approach to control the tradeoff between
-classification accuracy and robustness that ultimately maximizes the
+accuracy and robustness that ultimately maximizes the
 power of the enveloping hypothesis test.
@@
 
+\vfill
+
 \end{document}

### 2016-10-31 tex/paper.tex
trimming

@@
 test time. A formal way of enforcing robustness is to require that the distribution of
-$f(X ; \theta_f)$ conditional on $Z$ (and possibly $Y$) be invariant to the value
- of the nuisance parameter $Z$. Thus, we wish find a predictive function $f$ such that
+$f(X ; \theta_f)$ conditional on $Z$ (and possibly $Y$) be invariant with
+ the nuisance parameter $Z$. Thus, we wish find a predictive function $f$ such that
 \begin{equation}\label{eqn:criterion}

### 2016-11-01 tex/paper.tex
small changes

@@
 (categorical or continuous)
-nuisance parameters that parametrizes the family of data generation processes. In
+nuisance parameters that parametrizes this family of data generation processes. In
 this work, we introduce a flexible adversarial training procedure to enforce the pivotal
-property on a predictive model, which allows one to tune the trade-off between
-power and robustness. We derive theoretical
+property on a predictive model (e.g., a classifier), which allows one to tune the trade-off between
+accuracy and robustness. We derive theoretical
 results proving that the proposed procedure tends towards a minimax solution corresponding to a predictive model
@@
     \item $\smash{\hat\theta}_f$ maximizes the conditional entropy
-    $H({Z|f(X;\theta_f)})$ since $H(Z|f(X;\theta)) \leq H(Z)$. Note that this
+    $H({Z|f(X;\theta_f)})$, since $H(Z|f(X;\theta)) \leq H(Z)$ from the properties of entropy. Note that this
     latter inequality holds for both the discrete and the differential definitions of entropy.
@@
 sufficiently small steps, then $f$ should converge to a classifier which is both
-optimal and pivotal, provided such a classifier exists. A formal proof of
-convergence of the alternating stochastic gradient descent procedure of
+optimal and pivotal, provided such a classifier exists. Let us note however
+that despite the former theoretical characterization of the minimax solution
+of Eqn.~\ref{eqn:min_thetaf}, formal guarantees of
+convergence towards that solution by
 Algorithm~\ref{alg:adversarial-training} in the case where a finite number $K$
-of steps is taken for $r$ remains however to be proven.
+of steps is taken for $r$ remains to be proven.
 
@@
 generalization that also supports a continuously parametrized family of domains or as enforcing fairness over continuous attributes.
+In that sense, our work is also original in the theoretical characterization
+of the minimax solution of Eqn.~\ref{eqn:min_thetaf}.
 
@@
 
-As previously discussed, the proposed solution can be used in any situation
+Despite the theoretical characterization of the minimax solution of Eqn.~\ref{eqn:min_thetaf}
+that motivates this work,
+Algorithm~\ref{alg:adversarial-training} remains a heuristic in the sense
+that we do not provide formal guarantees of convergence of the procedure towards
+that solution. Empirical evidence gathered here and in similar minimax-based
+related works (see e.g., \citep{goodfellow2014generative,EdwardsS15}) suggests
+that the alternating optimization algorithm works sufficiently well in practice,
+but further research studying its stability and convergence would certainly
+be beneficial for a large body of works.
+
+In terms of applications, the proposed solution can be used in any situation
 where the training data may not be representative of the real data the
@@
 includes situations where the data generation processes are associated to
-systematic uncertainties (e.g., as in high energy physics), and which would certainly be
- interesting to revisit in light of the proposed solution. More generally,
-the approach also extends to cases where independence of the predictive
-model with respect to random variables is desired, as in domain adaptation
-or fairness in classification. Again, trying our algorithm on these
-use cases is certaintly worth investigating.
+systematic uncertainties (e.g., as in high energy physics), and which would be
+worth revisiting in light of the proposed solution. More generally, the approach
+also extends to cases where independence of the predictive model with respect to
+random variables is desired, as in domain adaptation or fairness in
+classification. Again, trying our algorithm on these use cases is certaintly
+worth investigating.
 

### 2016-11-01 tex/paper.tex
typos

@@
 minimizes the classification loss $L_f$ and is pivotal. Thus, we need to optimize
-the hyperparameter $\lambda$ of Eqn.~\ref{eqn:vf-lambda} with respect to
+the hyper-parameter $\lambda$ of Eqn.~\ref{eqn:vf-lambda} with respect to
 a higher-level objective. In this case, the natural higher-level context is a
@@
 classification accuracy. As demonstrated in this example, optimizing $\lambda$
-gives us a principled and effective approach to control the tradeoff between
+gives us a principled and effective approach to control the trade-off between
 accuracy and robustness that ultimately maximizes the
@@
 random variables is desired, as in domain adaptation or fairness in
-classification. Again, trying our algorithm on these use cases is certaintly
+classification. Again, trying our algorithm on these use cases is certainly
 worth investigating.

### 2016-11-01 code/jets.py
code

@@ -1,0 +1,110 @@
+import numpy as np
+np.random.seed = 777
+
+import sys
+
+lam = float(sys.argv[1])
+seed = int(sys.argv[2])
+print(lam, seed)
+
+
+import pickle
+fd = open("jets-pile.pickle", "rb")
+X_train, X_test, y_train, y_test, z_train, z_test = pickle.load(fd)
+X = X_train
+y = y_train
+z = z_train
+fd.close()
+
+mask = (z_train[:, 0] == 1)
+X_train = X_train[mask]
+y_train = y_train[mask]
+z_train = z_train[mask]
+
+
+
+import keras.backend as K
+from keras.layers import Input, Dense
+from keras.models import Model
+
+np.random.seed = seed
+
+inputs = Input(shape=(X.shape[1],))
+Dx = Dense(64, activation="tanh")(inputs)
+Dx = Dense(64, activation="relu")(Dx)
+Dx = Dense(64, activation="relu")(Dx)
+Dx = Dense(1, activation="sigmoid")(Dx)
+D = Model(input=[inputs], output=[Dx])
+
+Rx = D(inputs)
+Rx = Dense(64, activation="relu")(Rx)
+Rx = Dense(64, activation="relu")(Rx)
+Rx = Dense(64, activation="relu")(Rx)
+Rx = Dense(z.shape[1], activation="softmax")(Rx)
+R = Model(input=[inputs], output=[Rx])
+
+
+from keras.optimizers import SGD, Adam
+import keras.backend as K
+
+def make_loss_D(c):
+    def loss_D(y_true, y_pred):
+        return c * K.binary_crossentropy(y_pred, y_true)
+    return loss_D
+
+def make_loss_R(c):
+    def loss_R(z_true, z_pred):
+        return c * K.categorical_crossentropy(z_pred, z_true)
+    return loss_R
+
+opt_D = Adam()
+D.compile(loss=[make_loss_D(c=1.0)], optimizer=opt_D)
+
+opt_DRf = SGD(momentum=0)
+DRf = Model(input=[inputs], output=[D(inputs), R(inputs)])
+DRf.compile(loss=[make_loss_D(c=1.0),
+                  make_loss_R(c=-lam)],   # compare with c=0.0, ie. when no there is no adversary
+            optimizer=opt_DRf)
+
+opt_DfR = SGD(momentum=0)
+DfR = Model(input=[inputs], output=[R(inputs)])
+DfR.compile(loss=[make_loss_R(c=1.0)],
[…]

### 2016-11-01 tex/paper.tex
small things

@@
 (categorical or continuous)
-nuisance parameters that parametrizes this family of data generation processes. In
-this work, we introduce a flexible adversarial training procedure to enforce the pivotal
-property on a predictive model (e.g., a classifier), which allows one to tune the trade-off between
+nuisance parameters that parametrizes this family of generation processes. In
+this work, we introduce a flexible training procedure based on adversarial networks to enforce the pivotal
+property on a predictive model, which allows one to tune the trade-off between
 accuracy and robustness. We derive theoretical
-results proving that the proposed procedure tends towards a minimax solution corresponding to a predictive model
+results showing that the proposed procedure tends towards a minimax solution corresponding to a predictive model
 that is both optimal and independent on the nuisance parameters, and demonstrate

### 2016-11-01 tex/paper.tex
tiny things

@@
 $f(X ; \theta_f)$ conditional on $Z$ (and possibly $Y$) be invariant with
- the nuisance parameter $Z$. Thus, we wish find a predictive function $f$ such that
+ the nuisance parameter $Z$. Thus, we wish to find a function $f$ such that
 \begin{equation}\label{eqn:criterion}
@@
         {\it Hyper-parameters:} Number $T$ of training iterations,
-                                Number $K$ of gradient steps to update $r$.
+                                Number $K$ of gradient steps to update $r$,
+                                Size $M$ of the minibatches.
     \end{flushleft}
@@
 \mathbb{R}^p \mapsto [0,1]$ modeling $p(Y=1|X)$. For reasons further explained
-in Section~\ref{sec:theory}, ${\cal L}_f$ and ${\cal L}_r$  are respectively set to the
+in Sec.~\ref{sec:theory}, ${\cal L}_f$ and ${\cal L}_r$  are respectively set to the
 expected value of the
@@
 updated to improve ${\cal L}_f(\theta_f) -  H({Z|f(X;\theta_f)})$ with
-sufficiently small steps, then $f$ should converge to a classifier which is both
-optimal and pivotal, provided such a classifier exists. Let us note however
+sufficiently small steps, then $f$ should converge to a classifier that is both
+optimal and pivotal, provided such a classifier exists. Let us note
 that despite the former theoretical characterization of the minimax solution
@@
 
-On many practical problems, the assumption of existence of such a classifier may
+On many practical problems, the assumption of existence of an optimal and pivotal classifier may
 not hold because the nuisance parameter directly shapes the decision boundary.
@@
     (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
-       built with adversarial training, as outlined in Section~\ref{sec:method}.
+       built with adversarial training, as outlined in Sec.~\ref{sec:method}.
        The resulting densities are now almost identical to each other, indicating only a
@@
 $Z$, hence shaping $p_{\theta_r}$ towards the prior $p(z)$. Indeed, ${\cal L}_f$
-eventually starts to slightly decrease, while remaining lower bounded by
+eventually starts to  decrease, while remaining lower bounded by
 $\min_{\theta_f} {\cal L}_f(\theta_f)$ as approximated by the dashed line in the
@@
 \citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
-However, jets are produced ubiquitously at high energy colliders through
+However, jets are also produced ubiquitously at high energy colliders through
 more mundane processes in the SM, which leads to a challenging classification problem
@@
 at the LHC ($Y=0$) and  the fat jets coming from an exotic process ($Y=1$).
-We reuse the datasets used in reference
+We reuse the datasets used in
 \citep{baldi2016jet}.  For completeness, the jets are built with the
@@
 $\tau_{21}^{\beta=1}$ \citep{nsub,Thaler2012}, and the energy correlation
-functions \citep{Larkoski2013} $C_{2}^{\beta=1}$, $C_{2}^{\beta=2}$,
-$D_{2}^{\beta=1}$, $D_{2}^{\beta=2}$.
+functions  $C_{2}^{\beta=1}$, $C_{2}^{\beta=2}$,
+$D_{2}^{\beta=1}$, $D_{2}^{\beta=2}$ \citep{Larkoski2013}.
 
@@
 Since we are still performing a hypothesis test against the null, we only
-wish to impose the pivotal property on $Y=0$ events. Furthermore,
+wish to impose the pivotal property on $Y=0$ events. To this end,
 we use the Approximate Median Significance (AMS), that was used
@@
              classification accuracy for independence with respect to pileup
-             results in a positive total net benefit in terms of statistical significance.}
+             results in a net benefit in terms of statistical significance.}
     \label{fig:physics-ams}
@@
 there is no reason to expect the resulting classifier to be pivotal, as shown
-previously in Section~\ref{sec:toy}.
+previously in Sec.~\ref{sec:toy}.
 As an alternative, parametrized
@@
 %, but otherwise often limits the range of their applications.
-In practice, parametrized classifiers  are  computationally expensive to build
+In practice, parametrized classifiers  are also computationally expensive to build
 and evaluate. In particular, calibrating their decision function, i.e.
@@
 may not be optimal with respect to $Y|X,Z$, as discussed in
-Section~\ref{sec:theory}, but that can otherwise be directly used in a wider range of
+Sec.~\ref{sec:theory}, but that can otherwise be directly used in a wider range of
 applications, since the dependence on the nuisance parameter $Z$ has already been eliminated.

### 2016-11-01 tex/paper.tex
minor changes (#14)
- fix some equations
- conditional MLE notation
- acknowledge data
- specify 1000, 10000 sig, bkg events

@@
             $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|s_m)  \right],$$
-            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $1(y_m=0)(1-s_m) + 1(y_m=1)s_m$;}
+            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $\mathbf{1}(y_m=0)(1-s_m) + \mathbf{1}(y_m=1)s_m$;}
         \EndFor
@@
 \begin{equation}\label{eqn:min_thetaf}
-    \smash{\hat\theta_f, \hat\theta_f} = \arg \min_{\theta_f} \max_{\theta_r} E(\theta_f, \theta_r).
+    \smash{\hat\theta_f, \hat\theta_r} = \arg \min_{\theta_f} \max_{\theta_r} E(\theta_f, \theta_r).
 \end{equation}
@@
 
-For fixed $\theta_f$, the adversary $r$ is optimal at $$\hat\theta_r = \arg
+For fixed $\theta_f$, the adversary $r$ is optimal at $$\hat{\hat\theta}_r = \arg
 \max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal
-L}_r(\theta_f, \theta_r),$$ in which case $p_{\hat\theta_r}(z|f(X;\theta_f)=s) =
+L}_r(\theta_f, \theta_r),$$ in which case $p_{\hat{\hat\theta}_r}(z|f(X;\theta_f)=s) =
 p(z|f(X;\theta_f)=s)$ for all $z$ and all $s$, and ${\cal L}_r$ reduces to the expected entropy
@@
 5000000 samples. Both training and testing samples are weighted such that the
-signal-to-background ratio is 1/10, allowing us to probe the efficacy of the
-method proposed here in a background dominated environment. Results reported
-below are averages over 5 runs.
+null hypothesis corresponded to 10,000  of $Y=0$ events and the alternate
+hypothesis included an additional 1,000 $Y=1$ events prior to any thresholding on $f$.
+This allows us to probe the efficacy of the
+method proposed here in a representative background-dominated high energy physics environment. 
+Results reported below are averages over 5 runs.
 
@@
 
+    We would like to thank the authors of Ref.\citep{baldi2016jet} for sharing the data used in their
+    studies.
     KC and GL are both supported through NSF ACI-1450310, additionally KC is supported

### 2016-11-01 tex/paper.tex
add numbers to all equations

@@
 
-For fixed $\theta_f$, the adversary $r$ is optimal at $$\hat{\hat\theta}_r = \arg
-\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal
-L}_r(\theta_f, \theta_r),$$ in which case $p_{\hat{\hat\theta}_r}(z|f(X;\theta_f)=s) =
+For fixed $\theta_f$, the adversary $r$ is optimal at
+\begin{equation}
+    \hat{\hat\theta}_r = \arg \max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal L}_r(\theta_f, \theta_r),
+\end{equation}
+in which case $p_{\hat{\hat\theta}_r}(z|f(X;\theta_f)=s) =
 p(z|f(X;\theta_f)=s)$ for all $z$ and all $s$, and ${\cal L}_r$ reduces to the expected entropy
@@
 Accordingly, the
-value function $E$ can be restated as a function depending on $\theta_f$ only: $$E'(\theta_f) = {\cal
-L}_f(\theta_f) -  H({Z|f(X;\theta_f)}).$$  In particular, we have the lower
-bound $$H({Y|X}) - H(Z) \leq {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})$$
+value function $E$ can be restated as a function depending on $\theta_f$ only:
+\begin{equation}
+    E'(\theta_f) = {\cal L}_f(\theta_f) -  H({Z|f(X;\theta_f)}).
+\end{equation}
+In particular, we have the lower bound
+\begin{equation}
+    H({Y|X}) - H(Z) \leq {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})
+\end{equation}
 where the equality holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f}
@@
 not hold because the nuisance parameter directly shapes the decision boundary.
-In this case, the lower bound $$ H({Y|X}) - H(Z) < {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})$$ is strict: $f$ can either be an optimal classifier or a
+In this case, the lower bound
+\begin{equation}
+    H({Y|X}) - H(Z) < {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})
+\end{equation} is strict: $f$ can either be an optimal classifier or a
 pivotal quantity, but not both simultaneously. In this situation, it is natural
@@
 5000000 samples. Both training and testing samples are weighted such that the
-null hypothesis corresponded to 10,000  of $Y=0$ events and the alternate
-hypothesis included an additional 1,000 $Y=1$ events prior to any thresholding on $f$.
+null hypothesis corresponded to 1000  of $Y=0$ events and the alternate
+hypothesis included an additional 100 $Y=1$ events prior to any thresholding on $f$.
 This allows us to probe the efficacy of the
@@
 
-\vfill
-
 % ==============================================================================
@@
 
-\vfill
-
 \end{document}

### 2016-11-01 tex/paper.tex
mk small changes

--- 2016-11-01 tex/paper.tex
@@
 that is both optimal and independent on the nuisance parameters, and demonstrate
-its effectiveness  with a toy and physics examples.
+its effectiveness  with a toy and a particle physics examples.
 
@@
 that is robust to uncertainty in the unknown value of $Z$. More specifically, we
-introduce a learning procedure for enforcing the pivotal property on a
+introduce a learning procedure for enforcing the pivotal property
+\citep{degroot1986probability}  on a
 predictive model by jointly training two neural networks in an adversarial
@@
 $p_{\theta_r}(z_j|f(X;\theta_f)=s)$. Similarly, if $Z$ takes continuous values and
-if we assume some parametric distribution ${\cal P}$ for $Z|f(X;\theta_f)=s$
+if we assume a flexible enough parametric distribution ${\cal P}$ for $Z|f(X;\theta_f)=s$
 (e.g., a mixture of gaussians, as modeled with a mixture
 density network~\citep{bishop1994mixture}), then $p_{\theta_r}$ can be
-represented e.g. as network whose output $j$ is the estimated value of the
+represented e.g. as a network whose output $j$ is the estimated value of the
 corresponding parameter $\gamma_j$ of that distribution (e.g., the mean,
@@
 implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a mixture
-of five gaussians. As for $f$, the network architecture comprises 2 dense hidden
+of five gaussians. Similar to $f$, the network architecture of $r$ comprises 2 dense hidden
 layers of 20 nodes with ReLU activations, but is followed by an output layer of
@@
 If the exotic particle is heavy, then the $W$ boson will be moving very fast,
-and  relativistic effects will cause the two jets from its decay to merge into a single `fat jet'.
-These fat jets have a rich internal substructure (see e.g.
+and  relativistic effects will cause the two jets from its decay to
+merge into a single `$W$-jet'.
+These  $W$-jets have a rich internal substructure (see e.g.
 \citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
@@
 that is beset with a number of sources of systematic uncertainty.
+
 %Distinguishing these ($y=1$) fat jets produced by heavy SM particles from
@@
 references therein): we aim to discriminate between the normal jets produced copiously
-at the LHC ($Y=0$) and  the fat jets coming from an exotic process ($Y=1$).
+at the LHC ($Y=0$) and the $W$-jets ($Y=1$) potentially coming from an exotic proces.
 We reuse the datasets used in
@@
 tune the trade-off between power and robustness. From an empirical point of
-view, we confirmed the effectiveness of our method on a toy example and particle physics example.
+view, we confirmed the effectiveness of our method on a toy example
+and a particle physics example.
 
@@
     KC and GL are both supported through NSF ACI-1450310, additionally KC is supported
-    through PHY-1505463 and PHY-1205376.
+    through PHY-1505463 and PHY-1205376.  MK is support by the US
+    Department of Energy (DOE) under grant DE-AC02-76SF00515 and by the
+    SLAC Panofsky Fellowship.
 

### 2016-11-01 tex/paper.tex

@@
 
-For fixed $\theta_f$, the adversary $r$ is optimal at $$\hat\theta_r = \arg
-\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal
-L}_r(\theta_f, \theta_r),$$ in which case $p_{\hat\theta_r}(z|f(X;\theta_f)=s) =
+For fixed $\theta_f$, the adversary $r$ is optimal at
+\begin{equation}
+    \hat{\hat\theta}_r = \arg \max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal L}_r(\theta_f, \theta_r),
+\end{equation}
+in which case $p_{\hat{\hat\theta}_r}(z|f(X;\theta_f)=s) =
 p(z|f(X;\theta_f)=s)$ for all $z$ and all $s$, and ${\cal L}_r$ reduces to the expected entropy
@@
 5000000 samples. Both training and testing samples are weighted such that the
-signal-to-background ratio is 1/10, allowing us to probe the efficacy of the
-method proposed here in a background dominated environment. Results reported
-below are averages over 5 runs.
+null hypothesis corresponded to 1000  of $Y=0$ events and the alternate
+hypothesis included an additional 100 $Y=1$ events prior to any thresholding on $f$.
+This allows us to probe the efficacy of the
+method proposed here in a representative background-dominated high energy physics environment.
+Results reported below are averages over 5 runs.
 

### 2016-11-01 tex/paper.tex
typo

@@
 references therein): we aim to discriminate between the normal jets produced copiously
-at the LHC ($Y=0$) and the $W$-jets ($Y=1$) potentially coming from an exotic proces.
+at the LHC ($Y=0$) and the $W$-jets ($Y=1$) potentially coming from an exotic process.
 We reuse the datasets used in
@@
 
-\begin{acknowledgments}
+\section*{Acknowledgments}
 
@@
     KC and GL are both supported through NSF ACI-1450310, additionally KC is supported
-    through PHY-1505463 and PHY-1205376.  MK is support by the US
+    through PHY-1505463 and PHY-1205376.  MK is supported by the US
     Department of Energy (DOE) under grant DE-AC02-76SF00515 and by the
@@
 
-\end{acknowledgments}
+\vfill
 

### 2016-11-01 tex/paper.tex
save

@@
 generalization that also supports a continuously parametrized family of domains or as enforcing fairness over continuous attributes.
-In that sense, our work is also original in the theoretical characterization
-of the minimax solution of Eqn.~\ref{eqn:min_thetaf}.
+In that sense, and to the best of our knowledge, our work is also original in the theoretical characterization
+of the minimax solution of Eqn.~\ref{eqn:min_thetaf}, as well as in how one can trade-off accuracy for robustness through Eqn.~\ref{eqn:vf-lambda}.
 

### 2016-11-01 tex/paper.tex
added remark on non-uniqueness of solutions

@@
 In this case, the lower bound
-\begin{equation}
+\begin{equation}\label{eqn:lower-bound-strict}
     H({Y|X}) - H(Z) < {\cal L}_f(\theta_f) - H({Z|f(X;\theta_f)})
@@
 optimal.
+When the lower bound is strict, let us note however that there may
+exist distinct but equally good solutions $\theta_f,\theta_r$ minimizing Eqn.~\ref{eqn:vf-lambda}.
+In this zero-sum game, an increase in accuracy would exactly be compensated by a decrease in pivotality and vice-versa.
+How to best navigate this Pareto frontier to maximize a higher-level objective remains a question open for future works.
 

### 2016-11-02 tex/paper.tex
small things

@@
              and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=500$ and $\lambda=50$.}
+             Initialized with a pre-trained classifier $f$, adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=500$ and $\lambda=50$.}
     \label{fig:toy-training}
@@
 layers of 64 nodes and ReLU activations, each terminated by a single final
-output node with a sigmoid activation. Experiments are performed on a subset of
+output node with a sigmoid activation. As in the previous example, adversarial training is initialized
+with $f$ pre-trained. Experiments are performed on a subset of
 150000 samples for training while AMS is evaluated on an independent test set of
@@
 generalization that also supports a continuously parametrized family of domains or as enforcing fairness over continuous attributes.
-In that sense, and to the best of our knowledge, our work is also original in the theoretical characterization
-of the minimax solution of Eqn.~\ref{eqn:min_thetaf}, as well as in how one can trade-off accuracy for robustness through Eqn.~\ref{eqn:vf-lambda}.
+% In that sense, and to the best of our knowledge, our work is also original in the theoretical characterization
+% of the minimax solution of Eqn.~\ref{eqn:min_thetaf}, as well as in how one can trade-off accuracy for robustness through Eqn.~\ref{eqn:vf-lambda}.
 

### 2016-11-02 code/jets.py
Code cleanup

@@ -1,14 +1,18 @@
 import numpy as np
-np.random.seed = 777
-
+import pickle
 import sys
 
+import keras.backend as K
+from keras.layers import Input, Dense
+from keras.models import Model
+from keras.optimizers import SGD, Adam
+
+# Command line arguments
 lam = float(sys.argv[1])
 seed = int(sys.argv[2])
-print(lam, seed)
+np.random.seed = seed
 
-
-import pickle
+# Prepare data
 fd = open("jets-pile.pickle", "rb")
 X_train, X_test, y_train, y_test, z_train, z_test = pickle.load(fd)
 X = X_train
@@ -16,19 +20,12 @@ y = y_train
 z = z_train
 fd.close()
 
-mask = (z_train[:, 0] == 1)
-X_train = X_train[mask]
-y_train = y_train[mask]
-z_train = z_train[mask]
+# mask = (z_train[:, 0] == 1)
+# X_train = X_train[mask]
+# y_train = y_train[mask]
+# z_train = z_train[mask]
 
-
-
-import keras.backend as K
-from keras.layers import Input, Dense
-from keras.models import Model
-
-np.random.seed = seed
-
+# Set network architectures
 inputs = Input(shape=(X.shape[1],))
 Dx = Dense(64, activation="tanh")(inputs)
 Dx = Dense(64, activation="relu")(Dx)
@@ -44,18 +41,17 @@ Rx = Dense(z.shape[1], activation="softmax")(Rx)
 R = Model(input=[inputs], output=[Rx])
 
 
-from keras.optimizers import SGD, Adam
-import keras.backend as K
-
 def make_loss_D(c):
     def loss_D(y_true, y_pred):
         return c * K.binary_crossentropy(y_pred, y_true)
     return loss_D
 
+
 def make_loss_R(c):
     def loss_R(z_true, z_pred):
         return c * K.categorical_crossentropy(z_pred, z_true)
     return loss_R
+
 
 opt_D = Adam()
 D.compile(loss=[make_loss_D(c=1.0)], optimizer=opt_D)
@@ -63,7 +59,7 @@ D.compile(loss=[make_loss_D(c=1.0)], optimizer=opt_D)
 opt_DRf = SGD(momentum=0)
 DRf = Model(input=[inputs], output=[D(inputs), R(inputs)])
 DRf.compile(loss=[make_loss_D(c=1.0),
[…]

### 2016-11-02 tex/paper.tex
Tex cleanup

@@
 pivot -- a quantity whose distribution is invariant to the unknown value of the
-(categorical or continuous)
-nuisance parameters that parametrizes this family of generation processes. In
-this work, we introduce a flexible training procedure based on adversarial networks to enforce the pivotal
-property on a predictive model, which allows one to tune the trade-off between
-accuracy and robustness. We derive theoretical
-results showing that the proposed procedure tends towards a minimax solution corresponding to a predictive model
-that is both optimal and independent on the nuisance parameters, and demonstrate
-its effectiveness  with a toy and a particle physics examples.
+(categorical or continuous) nuisance parameters that parametrizes this family of
+generation processes. In this work, we introduce a flexible training procedure
+based on adversarial networks to enforce the pivotal property on a predictive
+model, which allows one to tune the trade-off between accuracy and robustness.
+We derive theoretical results showing that the proposed procedure tends towards
+a minimax solution corresponding to a predictive model that is both optimal and
+independent on the nuisance parameters, and demonstrate its effectiveness  with
+a toy and a particle physics examples.
 
-%Many inference problems involve data generation processes that are not uniquely specified or uncertain in some way.
-%In a scientific context, the presence of several plausible data generation processes is often
-%associated to the presence of systematic uncertainty.
-%We consider the situation of a
-%family of data generation processes parametrized by nuisance parameters, which can
-%be categorical or continuous. Robust inference is possible if it is based on a pivot -- a quantity whose
-%distribution is invariant to the unknown value of these nuisance parameters.
-%We introduce an adversarial training procedure to enforce the pivotal property and
-%allow one to tune the tradeoff between power and robustness.
-
-%Our goal is to have an inferential procedure that is robust to the uncertainties in the data generation process.
 \end{abstract}
@@
 \section{Introduction}
-
-% \kcnote{ need some text like abstract here before diving into formal statement.}
 
@@
 training data is unique, fully specified, and an accurate representative of the
-real world data.
-By contrast, a common situation when systematic uncertainty is present is when the training
-data are not representative of the real data that a predictive model will be
-applied to in practice. Several techniques for domain adaptation have been
-developed to create predictive models that are more robust to this type of
-uncertainty.
-A more generic situation is that there are several plausible data generation
-processes, specified as a family of data generation processes
-parametrized by nuisance parameters, which can be categorical or continuous.
+real world data. By contrast, a common situation when systematic uncertainty is
+present is when the training data are not representative of the real data that a
+predictive model will be applied to in practice. Several techniques for domain
+adaptation have been developed to create predictive models that are more robust
+to this type of uncertainty. A more generic situation is that there are several
+plausible data generation processes, specified as a family of data generation
+processes parametrized by nuisance parameters, which can be categorical or
+continuous.
 
@@
 \label{sec:problem}
-%
-%Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
-%sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
-%Let consider the multivariate random variables $X_z: \Omega \mapsto
-%\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
-%of the functional $X$ on a nuisance parameter $Z$, whose values $z \in {\cal Z}$  define a
-%parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
-%induce together a joint probability distribution $p(X,Y|Z=z)$, where the
-%conditional denotes $X_z$. For training, let further assume a finite set $\{
-%x_i, y_i, z_i \}_{i=1}^N$ of realizations $X_{z_i}(\omega_i), Y(\omega_i)$, for
-%$\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
-%is to learn a score function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal S}$ of
-%parameters $\theta_f$ (e.g., a neural network-based probabilistic classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
-%the cross-entropy). In addition, we require that $f(X_z ; \theta_f)$ should be
-%robust to the value $z$ of the nuisance parameter  -- which remains unknown at
-%test time. More specifically, we aim at building $f$ such that in the ideal case
-%\begin{equation}\label{eqn:criterion-true}
-%f(X_{z}(\omega) ; \theta_f) = f(X_{z^\prime}(\omega) ; \theta_f)
-%\end{equation} for all
-%samples $\omega \in \Omega$ and all $z, z^\prime$ pairs of values of the
-%nuisance parameter.
-%
-%Since in general we do not have training tuples $(X_{z}(\omega),
-%X_{z^\prime}(\omega))$ (for the same unknown $\omega$),
 
-We begin with a family of data generation processes $p(X,Y,Z)$, where $X \in \mathcal{X}$ are
-the data, $Y\in \mathcal{Y}$ are the target labels, and $Z\in \mathcal{Z}$ are the nuisance parameters that
-can be continuous or categorical. We assume training
-data $\{x_i, y_i, z_i\}_{i=1}^N$ for $X$ conditional on both $Y$ and $Z$.
-Let us assume that prior to incorporating the effect of uncertainty in $Z$,
-our goal is to learn a
-regression function $f : \mathcal{X} \mapsto {\cal S}$ of parameters
-$\theta_f$ (e.g., a neural network-based probabilistic classifier) and
-minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g., the cross-entropy).
+We begin with a family of data generation processes $p(X,Y,Z)$, where $X \in
+\mathcal{X}$ are the data, $Y\in \mathcal{Y}$ are the target labels, and $Z\in
+\mathcal{Z}$ are the nuisance parameters that can be continuous or categorical.
+We assume training data $\{x_i, y_i, z_i\}_{i=1}^N$ for $X$ conditional on both
+$Y$ and $Z$. Let us assume that prior to incorporating the effect of uncertainty
+in $Z$, our goal is to learn a regression function $f : \mathcal{X} \mapsto
+{\cal S}$ of parameters $\theta_f$ (e.g., a neural network-based probabilistic
+classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g., the
+cross-entropy).
 
@@
 for one or several specified values $y \in {\cal Y}$.
-
-%Note that a function
-%$f$ for which Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
-%Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true,
-%since the sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = s \}$ and $\{
-%\omega' | f(X_{z^\prime}(\omega'); \theta_f) = s \}$ do not need to be the same
-%for the equality to hold. In order to simplify notations, and as only
-%Eqn.~\ref{eqn:criterion-measure} is of direct interest in this work, we denote
-%from here on the pivotal quantity criterion as
 
@@
 the performance of $f$ and its independence with respect to the nuisance
-parameter. Setting $\lambda$ to a large value will preferably enforces $f$ to
-be pivotal while setting $\lambda$ close to $0$ will rather constraint $f$ to be
-optimal.
-When the lower bound is strict, let us note however that there may
-exist distinct but equally good solutions $\theta_f,\theta_r$ minimizing Eqn.~\ref{eqn:vf-lambda}.
-In this zero-sum game, an increase in accuracy would exactly be compensated by a decrease in pivotality and vice-versa.
-How to best navigate this Pareto frontier to maximize a higher-level objective remains a question open for future works.
+parameter. Setting $\lambda$ to a large value will preferably enforces $f$ to be
+pivotal while setting $\lambda$ close to $0$ will rather constraint $f$ to be
+optimal. When the lower bound is strict, let us note however that there may
+exist distinct but equally good solutions $\theta_f,\theta_r$ minimizing
+Eqn.~\ref{eqn:vf-lambda}. In this zero-sum game, an increase in accuracy would
+exactly be compensated by a decrease in pivotality and vice-versa. How to best
+navigate this Pareto frontier to maximize a higher-level objective remains a
+question open for future works.
 
-Interestingly, let us finally emphasize that these results hold using only the (1D) output $s$
-of $f(\cdot;\theta_f)$ (in the case of binary classification) as input to the adversary. We
-could similarly enforce an intermediate representation of the data to be
-pivotal, e.g. as in \citep{ganin2014unsupervised}, but this is in fact not
-necessary.
+Interestingly, let us finally emphasize that these results hold using only the
+(1D) output $s$ of $f(\cdot;\theta_f)$ (in the case of binary classification) as
+input to the adversary. We could similarly enforce an intermediate
+representation of the data to be pivotal, e.g. as in
+\citep{ganin2014unsupervised}, but this is in fact not necessary.
 
@@
 
-Assuming a gaussian prior $z \sim {\cal N}(0,1)$, we start by
-generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we train a
-neural network classifier $f$ minimizing ${\cal L}_f(\theta_f)$ without
-considering its adversary $r$. The network architecture comprises 2 dense hidden layers of 20
+Assuming a gaussian prior $z \sim {\cal N}(0,1)$, we start by generating
+training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we train a neural
+network classifier $f$ minimizing ${\cal L}_f(\theta_f)$ without considering its
+adversary $r$. The network architecture comprises 2 dense hidden layers of 20
 nodes with ReLU activations, followed  by a dense output layer with a single
-node with a sigmoid activation. As shown in
-Fig.~\ref{fig:toy}, the resulting classifier is not pivotal, as the
-conditional probability densities of its decision scores $f(X;\theta_f)$ show
-large discrepancies between values $z$ of the nuisance. While not shown here, a
-classifier trained only from data generated at the nominal value $Z=0$ would
-also not be pivotal.
+node with a sigmoid activation. As shown in Fig.~\ref{fig:toy}, the resulting
+classifier is not pivotal, as the conditional probability densities of its
+decision scores $f(X;\theta_f)$ show large discrepancies between values $z$ of
+the nuisance. While not shown here, a classifier trained only from data
+generated at the nominal value $Z=0$ would also not be pivotal.
 
@@
 implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a mixture
-of five gaussians. Similar to $f$, the network architecture of $r$ comprises 2 dense hidden
-layers of 20 nodes with ReLU activations, but is followed by an output layer of
-15 nodes corresponding to the means, standard deviations and mixture
-coefficients of the five gaussians. Output nodes for the mean values come with
-linear activations, output nodes for the standard deviations with exponential
-activations to ensure positivity, while output nodes for the mixture coefficients
-implement the softmax function to ensure positivity and normalization. When
-running Algorithm~\ref{alg:adversarial-training} as initialized with the
-classifier $f$ obtained previously, adversarial training effectively reshapes
-the decision function so it that becomes almost independent on the nuisance
-parameter, as shown in Fig.~\ref{fig:toy}. In particular,
-the conditional probability densities of the decision scores $f(X;\theta_f)$ are
-now very similar to each other, indicating only a small residual  dependency on the
+of five gaussians. Similar to $f$, the network architecture of $r$ comprises 2
+dense hidden layers of 20 nodes with ReLU activations, but is followed by an
+output layer of 15 nodes corresponding to the means, standard deviations and
+mixture coefficients of the five gaussians. Output nodes for the mean values
+come with linear activations, output nodes for the standard deviations with
+exponential activations to ensure positivity, while output nodes for the mixture
+coefficients implement the softmax function to ensure positivity and
+normalization. When running Algorithm~\ref{alg:adversarial-training} as
+initialized with the classifier $f$ obtained previously, adversarial training
+effectively reshapes the decision function so it that becomes almost independent
+on the nuisance parameter, as shown in Fig.~\ref{fig:toy}. In particular, the
+conditional probability densities of the decision scores $f(X;\theta_f)$ are now
+very similar to each other, indicating only a small residual  dependency on the
 nuisance, as theoretically expected. The dynamics of adversarial training is
@@
 benefit. As learning goes, minimizing $E$ then requires making predictions that
-are more accurate, hence decreasing ${\cal L}_f$, or that are even less dependent on
-$Z$, hence shaping $p_{\theta_r}$ towards the prior $p(z)$. Indeed, ${\cal L}_f$
-eventually starts to  decrease, while remaining lower bounded by
+are more accurate, hence decreasing ${\cal L}_f$, or that are even less
+dependent on $Z$, hence shaping $p_{\theta_r}$ towards the prior $p(z)$. Indeed,
+${\cal L}_f$ eventually starts to  decrease, while remaining lower bounded by
 $\min_{\theta_f} {\cal L}_f(\theta_f)$ as approximated by the dashed line in the
@@
 case of a gaussian with unit variance), as shown by the dashed line in the
-second plot.
-Finally, let us note that the ideal situation of a
-classifier that is both optimal and pivotal appears to be unreachable for this
-problem, as shown in the third plot by the  offset between ${\cal L}_f - \lambda
-{\cal L}_r$ and the dashed line approximating $H({Y|X}) -
-\lambda H(Z)$.
+second plot. Finally, let us note that the ideal situation of a classifier that
+is both optimal and pivotal appears to be unreachable for this problem, as shown
+in the third plot by the  offset between ${\cal L}_f - \lambda {\cal L}_r$ and
+the dashed line approximating $H({Y|X}) - \lambda H(Z)$.
 
@@
 
-%Experiments at high energy colliders like the LHC \citep{LHCMachine}
-%are searching for evidence of new particles beyond the ones
-%described by the so-called Standard Model (SM) of particle physics.
-%A wide array of extended theories predict the existence of new massive
-%particles that would decay to known heavy particles in the Standard Model,
-%like the $W$, $Z$, Higgs boson, or top quark, which subsequently decay to
-%multiple quarks.
-%Finally, those quarks would produce collimated sprays of particles known as \textit{jets}.
-%However, jets are produced ubiquitously at high energy colliders through
-%more mundane processes in the SM, which leads to a challenging classification problem
-%that is beset with a number of sources of systematic uncertainty.
-
-Experiments at high energy colliders like the LHC \citep{LHCMachine}
-are searching for evidence of new particles beyond the ones
-described by the well established Standard Model (SM) of particle physics.
-%Searches for these new particles in the LHC data involve a challenging
-%classification problem that is beset with a number of sources of systematic uncertainty.
-%
-A wide array of extended theories predict the existence of new massive
-particles that would decay to known particles in the SM such as the $W$ boson.
-The $W$ boson is unstable and can decay to
-two quarks, each of which produce collimated sprays of particles known as \textit{jets}
-(collectively denoted $W\to jj$).
-If the exotic particle is heavy, then the $W$ boson will be moving very fast,
-and  relativistic effects will cause the two jets from its decay to
-merge into a single `$W$-jet'.
-These  $W$-jets have a rich internal substructure (see e.g.
-\citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
-However, jets are also produced ubiquitously at high energy colliders through
-more mundane processes in the SM, which leads to a challenging classification problem
+Experiments at high energy colliders like the LHC \citep{LHCMachine} are
+searching for evidence of new particles beyond the ones described by the well
+established Standard Model (SM) of particle physics. A wide array of extended
+theories predict the existence of new massive particles that would decay to
+known particles in the SM such as the $W$ boson. The $W$ boson is unstable and
+can decay to two quarks, each of which produce collimated sprays of particles
+known as \textit{jets} (collectively denoted $W\to jj$). If the exotic particle
+is heavy, then the $W$ boson will be moving very fast, and  relativistic effects
+will cause the two jets from its decay to merge into a single `$W$-jet'. These
+$W$-jets have a rich internal substructure (see e.g.
+\citep{Altheimer:2012mn,Altheimer:2013yza} and references therein). However,
+jets are also produced ubiquitously at high energy colliders through more
+mundane processes in the SM, which leads to a challenging classification problem
 that is beset with a number of sources of systematic uncertainty.
-
-%Distinguishing these ($y=1$) fat jets produced by heavy SM particles from
-%the copious background of normal ($y=0$) jets is a fundamental challenge in searching
-%for signs of new high mass particles.
-
-%A wide array of extended theories predict the existence of new massive
-%particles that would decay to known heavy particles in the Standard Model
-%(e.g. the $W$ and $Z$ bosons), which subsequently decay to
-%multiple quarks.
-%Those quarks subsequently produce collimated sprays of particles known as \textit{jets} (denoted $j$).
-%However, jets are produced ubiquitously at high energy colliders through
-%more mundane processes in the SM, which leads to a challenging classification problem
-%that is beset with a number of sources of systematic uncertainty.
-
-%Consider the binary classification problem where $y=0$ corresponds to data produced from the
-%ubiquitous SM process $pp \to jj$ and $y=1$ corresponds to data produced from an exotic
-%process $pp \to W' \to W(\to jj) Z(\to jj)$ predicted by an extended theory.
-%When the mass of the $W'$ is much larger than the mass of the $W$ and $Z$ bosons,
-%then the $W$ and $Z$ particles are moving very fast and the relativistic Lorentz boost
-%causes the two jets from their decay to merge into a single `fat jet',
-
-%Consider the search for an exotic particle whose decay leads to a $W$ boson
-%that subsequently decays to two jets via $W\to jj$.  If the $W$ boson is moving
-%very fast, then the relativistic Lorentz boost
-%causes the two jets from its decay to merge into a single `fat jet',
-%which has a rich internal substructure (see e.g.
-%\citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
-%%We can cast this as a binary classification problem where $y=0$
-%%corresponds to data produced from the ubiquitous SM process
-%%that produce jets and $y=1$  corresponds to data produced from an exotic
-%%process that gives rise to fat jets from $W\to jj$.
-
-%binary classification problem where $y=0$ corresponds to data produced from the
-%ubiquitous SM process $pp \to jj$ and $y=1$ corresponds to data produced from an exotic
-%process $pp \to W' \to W(\to jj) Z(\to jj)$ predicted by an extended theory.
-%When the mass of the $W'$ is much larger than the mass of the $W$ and $Z$ bosons,
-%then the $W$ and $Z$ particles are moving very fast and the relativistic Lorentz boost
-%causes the two jets from their decay to merge into a single `fat jet',
-%which has a rich internal substructure (see e.g.
-%\citep{Altheimer:2012mn,Altheimer:2013yza} and references therein).
-
-%Distinguishing these ($y=1$) fat jets produced by heavy SM particles from
-%the copious background of normal ($y=0$) jets is a fundamental challenge in searching
-%for signs of new high mass particles.
 
@@
 
-%Our goal is to build an accurate classifier, for which we also want to minimize the effects due to the
-%uncertainties on the nuisance. More specifically, we choose to recast the
-%classification problem as a hypothesis test between signal+background (jets
-%originating from $W$ bosons or from quarks and gluons) and background
-%only (jets originating from quarks and gluons only) and tune the decision
-%threshold of our classifier by maximizing its approximate median significance
-%(AMS), when uncertainties in the background are taken into account (see Eqn.~20
-%of \citep{adam2014higgs}). Our motivation is that reducing the effects of
-%uncertainties by requiring independence of $Z$ with the classifier output $f(X;\theta_f)$ should
-%allow for a larger maximum significance.
-
 For several values of $\lambda$, we train a classifier
@@
 power of the enveloping hypothesis test.
-%significance by desensitizing
-%the classifier output $f(X;\theta_f)$ in the most beneficial way.
 
@@
 representation of the data. Likewise, our method also relates to the problem of
-enforcing fairness in classification \citep{zemel2013learning,feldman2015certifying,EdwardsS15,zafar2015fairness}, which
-is stated as learning a classifier that is independent of some chosen attribute
-such as gender, color or age. For both families of methods, the problem can
-equivalently be stated as learning a classifier which is a pivotal quantity with
-respect to either the domain or the selected feature. In this context,
-\citep{ganin2014unsupervised,EdwardsS15} are certainly among the closest to our
-work, in which domain invariance and fairness are enforced through an
-adversarial minimax setup composed of a classifier and an adversary
-discriminator.
-Following this line of work and without making assumptions on the prior $p(Z)$, our method can be regarded as a
-generalization that also supports a continuously parametrized family of domains or as enforcing fairness over continuous attributes.
-% In that sense, and to the best of our knowledge, our work is also original in the theoretical characterization
-% of the minimax solution of Eqn.~\ref{eqn:min_thetaf}, as well as in how one can trade-off accuracy for robustness through Eqn.~\ref{eqn:vf-lambda}.
+enforcing fairness in classification
+\citep{zemel2013learning,feldman2015certifying,EdwardsS15,zafar2015fairness},
+which is stated as learning a classifier that is independent of some chosen
+attribute such as gender, color or age. For both families of methods, the
+problem can equivalently be stated as learning a classifier which is a pivotal
+quantity with respect to either the domain or the selected feature. In this
+context, \citep{ganin2014unsupervised,EdwardsS15} are certainly among the
+closest to our work, in which domain invariance and fairness are enforced
+through an adversarial minimax setup composed of a classifier and an adversary
+discriminator. Following this line of work and without making assumptions on the
+prior $p(Z)$, our method can be regarded as a generalization that also supports
+a continuously parametrized family of domains or as enforcing fairness over
+continuous attributes.
 
@@
 typically take as fixed a classifier $f$ built from training data for a nominal
-value $z_0$ of the nuisance parameter, and then propagate uncertainty
- by estimating $p(f(x)|z)$ with a parametrized calibration
-procedure. Clearly, this classifier is however not optimal for $z \neq z_0$.
-To overcome this issue, the classifier $f$ is sometimes built instead on a mixture
-of training data generated from several plausible values $z_0, z_1, \dots$ of the nuisance parameter.
-While this certainly improves classification performance with respect to the marginal model $p(X,Y)$,
-there is no reason to expect the resulting classifier to be pivotal, as shown
-previously in Sec.~\ref{sec:toy}.
-As an alternative, parametrized
+value $z_0$ of the nuisance parameter, and then propagate uncertainty by
+estimating $p(f(x)|z)$ with a parametrized calibration procedure. Clearly, this
+classifier is however not optimal for $z \neq z_0$. To overcome this issue, the
+classifier $f$ is sometimes built instead on a mixture of training data
+generated from several plausible values $z_0, z_1, \dots$ of the nuisance
+parameter. While this certainly improves classification performance with respect
+to the marginal model $p(X,Y)$, there is no reason to expect the resulting
+classifier to be pivotal, as shown previously in Sec.~\ref{sec:toy}. As an
+alternative, parametrized
 classifiers~\citep{cranmer2015approximating,Baldi:2016fzo} directly take
@@
 systematics on the underlying classification task.  As argued in
-\citep{Neal:2007zz}, it is not obvious how such classifiers can be used on real data since
-the correct value $z$ of the nuisance often remains unknown. This is
+\citep{Neal:2007zz}, it is not obvious how such classifiers can be used on real
+data since the correct value $z$ of the nuisance often remains unknown. This is
 not an issue in the context of parameter
 inference~\citep{cranmer2015approximating}, where nuisance parameters are
-eliminated via optimization or marginalization.
-%, but otherwise often limits the range of their applications.
-In practice, parametrized classifiers  are also computationally expensive to build
-and evaluate. In particular, calibrating their decision function, i.e.
-approximating $p(f(x,z)|y,z)$ as a continuous function of $z$, remains an open
-challenge. By contrast, constraining $f$ to be pivotal yields a classifier which
-may not be optimal with respect to $Y|X,Z$, as discussed in
-Sec.~\ref{sec:theory}, but that can otherwise be directly used in a wider range of
-applications, since the dependence on the nuisance parameter $Z$ has already been eliminated.
-Moreover, if $f$ is sufficiently close to being pivotal, then calibration only needs to be carried out once.
-
-%Following this line of work and without making assumptions on the prior $p(Z)$, our method can be regarded as a
-%generalization that also supports the continuous case, which can be viewed as
-%handling infinitely many domains, provided they can be continuously
-%parametrized, or as enforcing fairness over continuous attributes.
+eliminated via optimization or marginalization. In practice, parametrized
+classifiers  are also computationally expensive to build and evaluate. In
+particular, calibrating their decision function, i.e. approximating
+$p(f(x,z)|y,z)$ as a continuous function of $z$, remains an open challenge. By
+contrast, constraining $f$ to be pivotal yields a classifier which may not be
+optimal with respect to $Y|X,Z$, as discussed in Sec.~\ref{sec:theory}, but that
+can otherwise be directly used in a wider range of applications, since the
+dependence on the nuisance parameter $Z$ has already been eliminated. Moreover,
+if $f$ is sufficiently close to being pivotal, then calibration only needs to be
+carried out once.
 
@@
 
-\bigskip
-
 % ==============================================================================

### 2016-11-02 tex/paper.tex
typo

@@
 As further explained in the next section, let us note that the adversary $r$ may
-take any form, i.e. it does need to be a neural network, as long as it exposes a
+take any form, i.e. it does not need to be a neural network, as long as it exposes a
 differentiable function $p_{\theta_r}(z|f(X;\theta_f)=s)$ of sufficient capacity

### 2016-11-02 tex/paper.tex
small things

@@
 adversary $r$. The network architecture comprises 2 dense hidden layers of 20
-nodes with ReLU activations, followed  by a dense output layer with a single
+nodes respectively with tanh and ReLU activations, followed  by a dense output layer with a single
 node with a sigmoid activation. As shown in Fig.~\ref{fig:toy}, the resulting
@@
 implemented as a mixture density network modeling $Z|f(X;\theta_f)$ as a mixture
-of five gaussians. Similar to $f$, the network architecture of $r$ comprises 2
-dense hidden layers of 20 nodes with ReLU activations, but is followed by an
+of five gaussians. The network architecture of $r$ comprises 2
+dense hidden layers of 20 nodes with ReLU activations, followed by an
 output layer of 15 nodes corresponding to the means, standard deviations and
@@
 term ${\cal L}_r$ conditioned on $Y=0$ only, as outlined in
-Sec.~\ref{sec:problem}. Both $f$ and $r$ are neural networks with 3 hidden
-layers of 64 nodes and ReLU activations, each terminated by a single final
-output node with a sigmoid activation. As in the previous example, adversarial training is initialized
+Sec.~\ref{sec:problem}. The architecture of $f$ comprises 3 hidden layers of 64 nodes
+respectively with tanh, ReLU and ReLU activations, and is terminated by a single final
+output node with a sigmoid activation. The architecture of $r$ is the same,
+but uses only ReLU activations in its hidden nodes.
+As in the previous example, adversarial training is initialized
 with $f$ pre-trained. Experiments are performed on a subset of

### 2016-11-02 tex/paper.tex
update abstract

@@
 generation processes. In this work,  we introduce a flexible training procedure
-based on adversarial networks to enforce the pivotal property on a predictive
-model, which allows one to tune the trade-off between accuracy and robustness.
-We derive theoretical results showing that the proposed procedure tends towards
-a minimax solution corresponding to a predictive model that is both optimal and
-independent on the nuisance parameters, and demonstrate its effectiveness  with
-a toy and a particle physics examples.
+based on adversarial networks for enforcing the pivotal property on a predictive
+model. We derive theoretical results proving that it
+converges towards a predictive model that is both optimal and independent on the
+nuisance parameters (if that models exists) or for which one can tune the
+trade-off between power and robustness. Finally, we demonstrate the
+effectiveness of this approach with a toy example and an example from particle
+physics.
 

### 2016-11-02 tex/paper.tex
save

@@
 based on adversarial networks for enforcing the pivotal property on a predictive
-model. We derive theoretical results proving that it
-converges towards a predictive model that is both optimal and independent on the
-nuisance parameters (if that models exists) or for which one can tune the
-trade-off between power and robustness. Finally, we demonstrate the
+model. We derive theoretical results showing that the proposed algorithm tends towards a minimax
+solution corresponding to a predictive model that is both optimal and
+independent on the nuisance parameters (if that models exists) or for which one
+can tune the trade-off between power and robustness. Finally, we demonstrate the
 effectiveness of this approach with a toy example and an example from particle

### 2016-11-02 tex/paper.tex
several edits

--- 2016-11-02 tex/paper.tex
@@
 based on adversarial networks for enforcing the pivotal property on a predictive
-model. We derive theoretical results proving that it
-converges towards a predictive model that is both optimal and independent on the
-nuisance parameters (if that models exists) or for which one can tune the
-trade-off between power and robustness. Finally, we demonstrate the
+model. We derive theoretical results showing that the proposed algorithm tends towards a minimax
+solution corresponding to a predictive model that is both optimal and
+independent of the nuisance parameters (if that models exists) or for which one
+can tune the trade-off between power and robustness. Finally, we demonstrate the
 effectiveness of this approach with a toy example and an example from particle
@@
 fashion. We derive theoretical results proving that the proposed procedure
-converges towards a predictive model that is both optimal and independent on the
+converges towards a predictive model that is both optimal and independent of the
 nuisance parameters (if that models exists) or for which one can tune the
@@
 The remainder of the paper is structured as follows. In Sec.~\ref{sec:problem},
-we first formally define the problem of learning a pivotal predictive model. In
-sections \ref{sec:method} and \ref{sec:theory}, we describe how adversarial
-networks can be trained to obtain predictive models that are independent on
+we first formally define the problem of learning a pivotal, predictive model. In
+Sec. \ref{sec:method} and \ref{sec:theory}, we describe how adversarial
+networks can be trained to obtain predictive models that are independent of
 nuisance parameters and show that the proposed procedure converges
 towards a model that is both optimal and robust. The  effectiveness of the
-approach is then demonstrated empirically in sections \ref{sec:toy} and \ref{sec:hep}.
+approach is then demonstrated empirically in Sec. \ref{sec:toy} and \ref{sec:hep}.
 In Sec.~\ref{sec:related}, we relate our work to
-close contributions from statistical inference, domain adaptation and fairness
+close contributions from statistical inference, domain adaptation, and fairness
 in classification. Finally, we gather our concluding remarks in
@@
     \caption{Architecture for the adversarial training of a binary classifier
-    $f$ against a nuisance parameter $Z$. The adversary $r$ models the
-    distribution $p(z|f(X;\theta_f)=s)$ of the nuisance as observed only through the output $f(X;\theta_f)$ of the classifier. By
+    $f$ against a nuisance parameters $Z$. The adversary $r$ models the
+    distribution $p(z|f(X;\theta_f)=s)$ of the nuisance parameters as observed only through the output $f(X;\theta_f)$ of the classifier. By
     maximizing the antagonistic objective ${\cal L}_r(\theta_f, \theta_r)$ (as part of minimizing ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$), the classifier
@@
 \mathcal{Z}$ are the nuisance parameters that can be continuous or categorical.
-We assume training data $\{x_i, y_i, z_i\}_{i=1}^N$ for $X$ conditional on both
-$Y$ and $Z$. Let us assume that prior to incorporating the effect of uncertainty
+We assume training data $\{x_i, y_i, z_i\}_{i=1}^N$. Let us assume that prior to incorporating the effect of uncertainty
 in $Z$, our goal is to learn a regression function $f : \mathcal{X} \mapsto
 {\cal S}$ of parameters $\theta_f$ (e.g., a neural network-based probabilistic
-classifier) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g., the
+classifier) that minimizes a loss ${\cal L}_f(\theta_f)$ (e.g., the
 cross-entropy).
@@
 which is a pivotal quantity \citep{degroot1986probability} with respect to the
-nuisance parameter. This implies that  $f(X; \theta_f)$ and $Z$ are independent random variables.
+nuisance parameters. This implies that  $f(X; \theta_f)$ and $Z$ are independent random variables.
 
 As stated in Eqn.~\ref{eqn:criterion}, the pivotal quantity criterion is
-considered in the case where $Y$ is marginalized out. In some situations
+imposed with respect to the $p(X|Z)$ with $Y$ is marginalized out. In some situations
 however (see e.g., Sec.~\ref{sec:hep}), class conditional independence of $f(X;
@@
 $z$. More specifically, the authors pit a generative model $g:
-\mathbb{R} \mapsto \mathbb{R}^p$ against an adversary classifier $d :
+\mathbb{R}^n \mapsto \mathbb{R}^p$ against an adversary classifier $d :
 \mathbb{R}^p \mapsto [0, 1]$ whose antagonistic objective is to recognize
@@
 
-In this work, we repurpose adversarial networks as a means to constraint the
+In this work, we repurpose adversarial networks as a means to constrain the
 predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. As
@@
 
-As for generative adversarial networks, we propose to
+As with generative adversarial networks, we propose to
 train $f$ and $r$ simultaneously, which we carry out by considering
@@
 necessarily with $Y|X,Z$ (unless $Z$ is made explicit and is included among the
-input variables in $X$). Results hold for a nuisance parameter $Z$ taking either
+input variables in $X$). Results hold for a nuisance parameters $Z$ taking either
 categorical or continuous values. By abuse of notation, $H(Z)$ denotes the
@@
 p(z|f(X;\theta_f)=s)$ for all $z$ and all $s$, and ${\cal L}_r$ reduces to the expected entropy
-$\mathbb{E}_{s \sim f(X;\theta_f)} [ H({Z|f(X;\theta_f)=s}) ]$ of the conditional distribution of the nuisance.
+$\mathbb{E}_{s \sim f(X;\theta_f)} [ H({Z|f(X;\theta_f)=s}) ]$ of the conditional distribution of the nuisance parameters.
 This expectation is nothing else than the conditional entropy of the random variables
@@
 \end{itemize}
-When the lower bound is active, we have $H(Z|f(X;\theta_f)) = H(Z)$
+By assumption, the lower bound is active, thus we have $H(Z|f(X;\theta_f)) = H(Z)$
 because of the second condition, which happens exactly when $Z$ and $f(X;\theta_f)$
@@
 decision scores $f(X;\theta_f)$ show large discrepancies between values $z$ of
-the nuisance. While not shown here, a classifier trained only from data
+the nuisance parameters. While not shown here, a classifier trained only from data
 generated at the nominal value $Z=0$ would also not be pivotal.
@@
 dependent on $Z$, hence shaping $p_{\theta_r}$ towards the prior $p(z)$. Indeed,
-${\cal L}_f$ eventually starts to  decrease, while remaining lower bounded by
+${\cal L}_f$ eventually starts to  decrease, while remaining bounded from below by
 $\min_{\theta_f} {\cal L}_f(\theta_f)$ as approximated by the dashed line in the
@@
 second plot. Finally, let us note that the ideal situation of a classifier that
-is both optimal and pivotal appears to be unreachable for this problem, as shown
+is both optimal and pivotal is unreachable for this problem, as shown
 in the third plot by the  offset between ${\cal L}_f - \lambda {\cal L}_r$ and
@@
 We do not expect that we will be able to find a function $f$ that simultaneously
-minimizes the classification loss $L_f$ and is pivotal. Thus, we need to optimize
+minimizes the classification loss ${\cal L}_f$ and is pivotal. Thus, we need to optimize
 the hyper-parameter $\lambda$ of Eqn.~\ref{eqn:vf-lambda} with respect to
@@
 alternate hypothesis that is a mixture of $Y=0$ and $Y=1$ events.
-In the absence of systematic uncertainties, optimizing $L_f$ simultaneously
+In the absence of systematic uncertainties, optimizing ${\cal L}_f$ simultaneously
 optimizes the power of a classical hypothesis test in the Neyman-Pearson sense.
@@
 wish to impose the pivotal property on $Y=0$ events. To this end,
-we use the Approximate Median Significance (AMS), that was used
-in the Higgs boson machine learning challenge that is a natural generalization
+we use as a higher level objective the Approximate Median Significance (AMS), which was used
+in the Higgs boson machine learning challenge and is a natural generalization
 of the power of a hypothesis test when systematic uncertainties are taken into account
@@
 
-To account for systematic uncertainties, experimentalists in high energy physics
-typically take as fixed a classifier $f$ built from training data for a nominal
+A common approach to account for systematic uncertainties in 
+a scientific context (e.g. in high energy physics)
+is to take as fixed a classifier $f$ built from training data for a nominal
 value $z_0$ of the nuisance parameter, and then propagate uncertainty by
@@
 that solution. Empirical evidence gathered here and in similar minimax-based
-related works (see e.g., \citep{goodfellow2014generative,EdwardsS15}) suggests
+ works (see e.g., \citep{goodfellow2014generative,EdwardsS15}) suggests
 that the alternating optimization algorithm works sufficiently well in practice,
@@
 where the training data may not be representative of the real data the
-predictive model will be applied to in practice. In the scientific context, this
-includes situations where the data generation processes are associated to
-systematic uncertainties (e.g., as in high energy physics), and which would be
-worth revisiting in light of the proposed solution. More generally, the approach
+predictive model will be applied to in practice. In the scientific context, the 
+presence of systematic uncertainty can be incorporated by considering a family
+of data generation processes, and it would be worth revisiting those scientific problems 
+that utilize machine learning in light of this technique. Moreover, the approach
 also extends to cases where independence of the predictive model with respect to
-random variables is desired, as in domain adaptation or fairness in
+observed random variables is desired, as in domain adaptation or fairness in
 classification. Again, trying our algorithm on these use cases is certainly

### 2016-11-02 tex/paper.tex
changes to section 3

@@
 associated loss ${\cal L}_r(\theta_f, \theta_r)$. This model takes  as input
-realizations $s$ of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$
-parameters, and produces as output a function $p_{\theta_r}(z | f(X;\theta_f)=s)$
-modeling the posterior probability density that $z$ parameterizes the sample
-observed as $s$.
+realizations $f(X; \theta_f)$ and produces as output a function 
+modeling the posterior probability density $p_{\theta_r}(z | f(X;\theta_f)=s)$.
 Intuitively, if $p(f(X; \theta_f)=s|z)$ varies with $z$,
@@
 probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal Z|}}$ whose
-output $j$ (for $j=1, \dots, |{\cal Z}|$) is the estimated probability mass
-$p_{\theta_r}(z_j|f(X;\theta_f)=s)$. Similarly, if $Z$ takes continuous values and
-if we assume a flexible enough parametric distribution ${\cal P}$ for $Z|f(X;\theta_f)=s$
-(e.g., a mixture of gaussians, as modeled with a mixture
-density network~\citep{bishop1994mixture}), then $p_{\theta_r}$ can be
-represented e.g. as a network whose output $j$ is the estimated value of the
+$j^\textrm{th}$ output (for $j=1, \dots, |{\cal Z}|$) is the estimated probability mass
+$p_{\theta_r}(z_j|f(X;\theta_f)=s)$. Similarly, if $Z$ takes continuous values, 
+then we can model the posterior probability density $p(z | f(X;\theta_f)=s)$
+with a sufficiently flexible  parametric distribution $p_{\theta_r}(z | f(X;\theta_f)=s)$. 
+The adversary $r$ may
+take any form, i.e. it does not need to be a neural network, as long as it exposes a
+differentiable function $p_{\theta_r}(z|f(X;\theta_f)=s)$ of sufficient capacity
+to represent the true distribution. Fig.~\ref{fig:architecture} illustrates 
+a concrete example where $p_{\theta_r}(z | f(X;\theta_f)=s)$ is a mixture of 
+gaussians, as modeled with a mixture density network~\citep{bishop1994mixture}).
+The $j^\textrm{th}$ output corresponds to the estimated value of the
 corresponding parameter $\gamma_j$ of that distribution (e.g., the mean,
@@
 $p_{\theta_r}(z|f(X;\theta_f)=s)$ can then be evaluated for any $z \in {\cal Z}$ and any score $s \in {\cal S}$.
-As further explained in the next section, let us note that the adversary $r$ may
-take any form, i.e. it does not need to be a neural network, as long as it exposes a
-differentiable function $p_{\theta_r}(z|f(X;\theta_f)=s)$ of sufficient capacity
-to represent the true distribution.
+
+%If $Z$ takes discrete values, then $p_{\theta_r}$ can be represented e.g. as a
+%probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal Z|}}$ whose
+%output $j$ (for $j=1, \dots, |{\cal Z}|$) is the estimated probability mass
+%$p_{\theta_r}(z_j|f(X;\theta_f)=s)$. Similarly, if $Z$ takes continuous values and
+%if we assume a flexible enough parametric distribution ${\cal P}$ for $Z|f(X;\theta_f)=s$
+%(e.g., a mixture of gaussians, as modeled with a mixture
+%density network~\citep{bishop1994mixture}), then $p_{\theta_r}$ can be
+%represented e.g. as a network whose output $j$ is the estimated value of the
+%corresponding parameter $\gamma_j$ of that distribution (e.g., the mean,
+%variance and mixing coefficients of its components). As in
+%\citep{nix1994estimating,bishop1994mixture}, the estimated probability density
+%$p_{\theta_r}(z|f(X;\theta_f)=s)$ can then be evaluated for any $z \in {\cal Z}$ and any score $s \in {\cal S}$.
+%As further explained in the next section, let us note that the adversary $r$ may
+%take any form, i.e. it does not need to be a neural network, as long as it exposes a
+%differentiable function $p_{\theta_r}(z|f(X;\theta_f)=s)$ of sufficient capacity
+%to represent the true distribution.
 

### 2016-11-02 tex/paper.tex
one more edit to section 3

@@
 then we can model the posterior probability density $p(z | f(X;\theta_f)=s)$
-with a sufficiently flexible  parametric distribution $p_{\theta_r}(z | f(X;\theta_f)=s)$. 
+with a sufficiently flexible  parametric family of distributions $\mathcal{P}(\gamma_1, \gamma_2, \dots)$,
+where the parameters $\gamma_j$ depend on $f(X, \theta_f)$ and $\theta_r$. 
 The adversary $r$ may

### 2016-11-02 tex/paper.tex
last changes?

@@
 \mathcal{Z}$ are the nuisance parameters that can be continuous or categorical.
-We assume training data $\{x_i, y_i, z_i\}_{i=1}^N$. Let us assume that prior to incorporating the effect of uncertainty
+% We assume training data $\{x_i, y_i, z_i\}_{i=1}^N$.
+Let us assume that prior to incorporating the effect of uncertainty
 in $Z$, our goal is to learn a regression function $f : \mathcal{X} \mapsto
@@
 As stated in Eqn.~\ref{eqn:criterion}, the pivotal quantity criterion is
-imposed with respect to the $p(X|Z)$ with $Y$ is marginalized out. In some situations
+imposed with respect to $p(X|Z)$ where $Y$ is marginalized out. In some situations
 however (see e.g., Sec.~\ref{sec:hep}), class conditional independence of $f(X;
@@
 updated to improve ${\cal L}_f(\theta_f) -  H({Z|f(X;\theta_f)})$ with
-sufficiently small steps, then $f$ should converge to a classifier that is both
-optimal and pivotal, provided such a classifier exists. Let us note
-that despite the former theoretical characterization of the minimax solution
-of Eqn.~\ref{eqn:min_thetaf}, formal guarantees of
+sufficiently small steps, then $f$ should eventually
+converge to a classifier that is both
+optimal and pivotal. Provided this classifier exists, the adversarial term ${\cal L}_r$
+can therefore be regarded as a way to select among
+the class of all optimal classifiers a function $f$ that is also pivotal.
+Despite the former theoretical characterization of the minimax solution
+of Eqn.~\ref{eqn:min_thetaf}, let us note that formal guarantees of
 convergence towards that solution by
@@
 problem can equivalently be stated as learning a classifier which is a pivotal
-quantity with respect to either the domain or the selected feature. In this
+quantity with respect to either the domain or the selected feature. As an
+example, unsupervised domain adaptation with labeled data from a source domain
+and unlabeled data from a target domain can be recast as learning a predictive
+model $f$ (i.e., trained to minimize ${\cal L}_f$ evaluated on labeled source
+data only) that is also a pivot with respect to the domain $Z$ (i.e., trained to
+maximize ${\cal L}_r$ evaluated on both source and target data). In this
 context, \citep{ganin2014unsupervised,EdwardsS15} are certainly among the

### 2016-11-02 tex/paper.tex
last last changes

@@
 updated to improve ${\cal L}_f(\theta_f) -  H({Z|f(X;\theta_f)})$ with
-sufficiently small steps, then $f$ should eventually
-converge to a classifier that is both
-optimal and pivotal. Provided this classifier exists, the adversarial term ${\cal L}_r$
-can therefore be regarded as a way to select among
+sufficiently small steps, then $f$ should converge to a classifier that is both
+optimal and pivotal, provided such a classifier exists. Therefore, the adversarial term ${\cal L}_r$
+can be regarded as a way to select among
 the class of all optimal classifiers a function $f$ that is also pivotal.

### 2017-01-24 tex/icml2017/paper.tex
add ICML version

--- 2016-11-02 tex/paper.tex
@@
-\documentclass[twocolumn,superscriptaddress,aps]{revtex4-1}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%% ICML 2016 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage[utf8]{inputenc}
+% Use the following line _only_ if you're still using LaTeX 2.09.
+%\documentstyle[icml2016,epsf,natbib]{article}
+% If you rely on Latex2e packages, like most moden people use this:
+\documentclass{article}
 
@@
 
-\usepackage{bm}
-\usepackage{cancel}
-\usepackage{bbold}
-\usepackage{slashed}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{hyperref}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
+% use Times
+\usepackage{times}
+% For figures
+\usepackage{graphicx} % more modern
+%\usepackage{epsfig} % less modern
+\usepackage{subfigure}
+
 \usepackage{tikz}
-
-\newcommand{\glnote}[1]{\textcolor{red}{[GL: #1]}}
-\newcommand{\kcnote}[1]{\textcolor{red}{[KC: #1]}}
-
 \theoremstyle{plain}
@@
 
+% For citations
+\usepackage{natbib}
+
+% For algorithms
+\usepackage{algorithm}
+\usepackage{algorithmic}
+
+% As of 2011, we use the hyperref package to produce hyperlinks in the
+% resulting PDF.  If this breaks your system, please commend out the
+% following usepackage line and replace \usepackage{icml2016} with
+% \usepackage[nohyperref]{icml2016} above.
+\usepackage{hyperref}
+
+% Packages hyperref and algorithmic misbehave sometimes.  We can fix
+% this with the following command.
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
+
+% Employ the following version of the ``usepackage'' statement for
+% submitting the draft version of the paper for review.  This will set
+% the note in the first column to ``Under review.  Do not distribute.''
+\usepackage{icml2016}
+
+% Employ this version of the ``usepackage'' statement after the paper has
+% been accepted, when creating the final version.  This will set the
+% note in the first column to ``Proceedings of the...''
+%\usepackage[accepted]{icml2016}
+
+% The \icmltitle you define below is probably too long as a header.
+% Therefore, a short form for the running title is supplied here:
+\icmltitlerunning{Learning to Pivot with Adversarial Networks}
+
 \begin{document}
 
-% ==============================================================================
+\twocolumn[
+\icmltitle{Learning to Pivot with Adversarial Networks}
 
-\title{\Large{Learning to Pivot with Adversarial Networks}}
-\vspace{1cm}
-\author{\small{\bf Gilles Louppe}\thanks{\texttt{g.louppe@nyu.edu}}}
-\affiliation{New York University}
-\author{\small{\bf Michael Kagan}\thanks{\texttt{makagan@slac.stanford.edu}}}
-\affiliation{SLAC National Accelerator Laboratory}
-\author{\small{\bf Kyle Cranmer}\thanks{\texttt{kyle.cranmer@nyu.edu}}}
-\affiliation{New York University}
+% It is OKAY to include author information, even for blind
+% submissions: the style file will automatically remove it for you
+% unless you've provided the [accepted] option to the icml2016
+% package.
+\icmlauthor{Gilles Louppe}{g.louppe@nyu.edu}
+\icmladdress{New York University}
+\icmlauthor{Michael Kagan}{makagan@slac.stanford.edu}
+\icmladdress{SLAC National Accelerator Laboratory}
+\icmlauthor{Kyle Cranmer}{kyle.cranmer@nyu.edu}
+\icmladdress{New York University}
+
+% You may provide any keywords that you
+% find helpful for describing your paper; these are used to populate
+% the "keywords" metadata in the PDF but will not be shown in the document
+%\icmlkeywords{boring formatting information, machine learning, ICML}
+
+\vskip 0.3in
+]
 
@@
 
-\maketitle
 % ==============================================================================
@@
 effectiveness of this approach with a toy example and an example from particle physics.
-
-The remainder of the paper is structured as follows. In Sec.~\ref{sec:problem},
-we first formally define the problem of learning a pivotal, predictive model. In
-Sec. \ref{sec:method} and \ref{sec:theory}, we describe how adversarial
-networks can be trained to obtain predictive models that are independent of
-nuisance parameters and show that the proposed procedure converges
-towards a model that is both optimal and robust. The  effectiveness of the
-approach is then demonstrated empirically in Sec. \ref{sec:toy} and \ref{sec:hep}.
-In Sec.~\ref{sec:related}, we relate our work to
-close contributions from statistical inference, domain adaptation, and fairness
-in classification. Finally, we gather our concluding remarks in
-Sec.~\ref{sec:conclusions}.
-
-% ==============================================================================
 
@@
 
+The remainder of the paper is structured as follows. In Sec.~\ref{sec:problem},
+we first formally define the problem of learning a pivotal, predictive model. In
+Sec. \ref{sec:method} and \ref{sec:theory}, we describe how adversarial
+networks can be trained to obtain predictive models that are independent of
+nuisance parameters and show that the proposed procedure converges
+towards a model that is both optimal and robust. The  effectiveness of the
+approach is then demonstrated empirically in Sec. \ref{sec:toy} and \ref{sec:hep}.
+In Sec.~\ref{sec:related}, we relate our work to
+close contributions from statistical inference, domain adaptation, and fairness
+in classification. Finally, we gather our concluding remarks in
+Sec.~\ref{sec:conclusions}.
+
+% ==============================================================================
+
 \section{Problem statement}
@@
 \label{sec:method}
-
-\begin{figure*}
-    \begin{minipage}{\linewidth}
-    \begin{algorithm}[H]
-    \caption{Adversarial training of a classifier $f$ against an adversary $r$.}
-
-    \begin{flushleft}
-        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;\\
-        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$;\\
-        {\it Hyper-parameters:} Number $T$ of training iterations,
-                                Number $K$ of gradient steps to update $r$,
-                                Size $M$ of the minibatches.
-    \end{flushleft}
-
-    \label{alg:adversarial-training}
-    \begin{algorithmic}[1]
-        \For{$t=1$ to $T$}
-            \For{$k=1$ to $K$} \Comment{Update $r$}
-                \State{Sample minibatch $\{x_m, z_m, s_m = f(x_m;\theta_f) \}_{m=1}^M$} of size $M$;
-                \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
-                $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|s_m)  ;$$}
-            \EndFor
-            \State{Sample minibatch $\{x_m, y_m, z_m, s_m = f(x_m;\theta_f)  \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
-            \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
-            $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|s_m)  \right],$$
-            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $\mathbf{1}(y_m=0)(1-s_m) + \mathbf{1}(y_m=1)s_m$;}
-        \EndFor
-    \end{algorithmic}
-    \end{algorithm}
-    \end{minipage}
-\end{figure*}
 
@@
 same, except that the adversarial term ${\cal L}_r(\theta_f, \theta_r)$ is restricted to $Y=y$.
+
+\begin{figure*}
+    \begin{minipage}{\linewidth}
+    \begin{algorithm}[H]
+    \caption{Adversarial training of a classifier $f$ against an adversary $r$.}
+
+    \begin{flushleft}
+        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;\\
+        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$;\\
+        {\it Hyper-parameters:} Number $T$ of training iterations,
+                                Number $K$ of gradient steps to update $r$,
+                                Size $M$ of the minibatches.
+    \end{flushleft}
+
+    \label{alg:adversarial-training}
+    \begin{algorithmic}[1]
+        \FOR{$t=1$ to $T$}
+            \FOR{$k=1$ to $K$} %\COMMENT{Update $r$}
+                \STATE{Sample minibatch $\{x_m, z_m, s_m = f(x_m;\theta_f) \}_{m=1}^M$ of size $M$;}
+                \STATE{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
+                $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|s_m)  ;$$}
+            \ENDFOR
+            \STATE{Sample minibatch $\{x_m, y_m, z_m, s_m = f(x_m;\theta_f)  \}_{m=1}^M$ of size $M$;} %\Comment{Update $f$}
+            \STATE{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
+            $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|s_m)  \right],$$
+            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $\mathbf{1}(y_m=0)(1-s_m) + \mathbf{1}(y_m=1)s_m$;}
+        \ENDFOR
+    \end{algorithmic}
+    \end{algorithm}
+    \end{minipage}
+\end{figure*}
 
@@
     \begin{center}
-        \includegraphics[width=0.245\textwidth]{figures/f-plain.pdf}
-        \includegraphics[width=0.245\textwidth]{figures/surface-plain.pdf}
-        \includegraphics[width=0.245\textwidth]{figures/f-adversary.pdf}
-        \includegraphics[width=0.245\textwidth]{figures/surface-adversary.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/f-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/surface-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/f-adversary.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/surface-adversary.pdf}
     \end{center}
@@
 \end{figure*}
-
-\begin{figure}
-    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
-    \vspace{-1cm}
-    \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
-             and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
-             Initialized with a pre-trained classifier $f$, adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=500$ and $\lambda=50$.}
-    \label{fig:toy-training}
-\end{figure}
 
@@
 the dashed line approximating $H({Y|X}) - \lambda H(Z)$.
+
+\begin{figure}
+    \includegraphics[width=0.48\textwidth]{../figures/training.pdf}
+    \vspace{-1cm}
+    \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
+             and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
+             Initialized with a pre-trained classifier $f$, adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=500$ and $\lambda=50$.}
+    \label{fig:toy-training}
+\end{figure}
 
@@
 \begin{figure}
-    \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
+    \includegraphics[width=0.48\textwidth]{../figures/ams.pdf}
     \caption{Physics example. Approximate median significance as a function of the decision threshold
@@
 
-\section*{Acknowledgments}
+% \section*{Acknowledgements}
+%
+% We would like to thank the authors of \citep{baldi2016jet} for sharing the data used in their
+% studies.
+% KC and GL are both supported through NSF ACI-1450310, additionally KC is supported
+% through PHY-1505463 and PHY-1205376.  MK is supported by the US
+% Department of Energy (DOE) under grant DE-AC02-76SF00515 and by the
+% SLAC Panofsky Fellowship.
 
-    We would like to thank the authors of Ref.\citep{baldi2016jet} for sharing the data used in their
-    studies.
-    KC and GL are both supported through NSF ACI-1450310, additionally KC is supported
-    through PHY-1505463 and PHY-1205376.  MK is supported by the US
-    Department of Energy (DOE) under grant DE-AC02-76SF00515 and by the
-    SLAC Panofsky Fellowship.
+% In the unusual situation where you want a paper to appear in the
+% references without citing it in the main text, use \nocite
 
-% ==============================================================================
-
-\bibliographystyle{acm}
-\bibliography{bibliography.bib}
+\bibliography{bibliography}
+\bibliographystyle{icml2016}
 
 \end{document}
+
+% This document was modified from the file originally made available by
+% Pat Langley and Andrea Danyluk for ICML-2K. This version was
+% created by Lise Getoor and Tobias Scheffer, it was slightly modified
+% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz,
+% slightly modified from the 2009 version by Kiri Wagstaff and
+% Sam Roweis's 2008 version, which is slightly modified from
+% Prasad Tadepalli's 2007 version which is a lightly
+% changed version of the previous year's version by Andrew Moore,
+% which was in turn edited from those of Kristian Kersting and
+% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2017-02-01 tex/icml2017/paper.tex
save

@@
 based on adversarial networks for enforcing the pivotal property on a predictive
-model. We derive theoretical results showing that the proposed algorithm tends towards a minimax
-solution corresponding to a predictive model that is both optimal and
-independent of the nuisance parameters (if that models exists) or for which one
-can tune the trade-off between power and robustness. Finally, we demonstrate the
-effectiveness of this approach with a toy example and an example from particle
-physics.
+model. We derive theoretical results showing that the proposed algorithm tends
+towards a minimax solution corresponding to a predictive model that is both
+optimal and independent of the nuisance parameters (if that model exists) or
+for which one can tune the trade-off between power and robustness. Finally, we
+demonstrate the effectiveness of this approach with a toy example and an example
+from particle physics.
 
@@
     \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.85]
-    \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!30, line width=0.3mm, fill=white]
+    \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!20, line width=0.3mm, fill=white]
 
@@
         \foreach \dest in {1,...,5}
-            \path[black!20] (f-I-\source) edge (f-H1-\dest);
+            \path[black] (f-I-\source) edge (f-H1-\dest);
 
     \foreach \source in {1,...,5}
-        \path[black!30] (f-H2-\source) edge (f-O);
+        \path[black] (f-H2-\source) edge (f-O);
 
-    \node[black!30] at (1.5,-3) {...};
+    \node[black] at (1.5,-3) {...};
 
@@
         \foreach \dest in {1,...,5}
-            \path[black!20] (r-I-\source) edge (r-H1-\dest);
+            \path[black] (r-I-\source) edge (r-H1-\dest);
 
     \foreach \source in {1,...,5}
-        \path[black!30] (r-H2-\source) edge (r-O1);
+        \path[black] (r-H2-\source) edge (r-O1);
     \foreach \source in {1,...,5}
-        \path[black!30] (r-H2-\source) edge (r-O2);
+        \path[black] (r-H2-\source) edge (r-O2);
     \foreach \source in {1,...,5}
-        \path[black!30] (r-H2-\source) edge (r-O3);
+        \path[black] (r-H2-\source) edge (r-O3);
 
-    \node[black!30] at (9,-3) {...};
+    \node[black] at (9,-3) {...};
 

### 2017-02-01 tex/icml2017/paper.tex
save

@@
 % the note in the first column to ``Under review.  Do not distribute.''
-\usepackage{icml2016}
+\usepackage[accepted]{icml2016}
 
@@
 of systematic uncertainties. Robust inference is possible if it is based on a
-pivot -- a quantity whose distribution is invariant to the unknown value of the
+pivot -- a quantity whose distribution does not depend on the unknown value of the
 (categorical or continuous) nuisance parameters that parametrizes this family of
 generation processes. In this work,  we introduce a flexible training procedure
-based on adversarial networks for enforcing the pivotal property on a predictive
-model. We derive theoretical results showing that the proposed algorithm tends
+based on adversarial networks for enforcing the pivotal property, i.e., the stastistical independence of a predictive
+model with respect to nuisance parameters. We derive theoretical results showing that the proposed algorithm tends
 towards a minimax solution corresponding to a predictive model that is both
 optimal and independent of the nuisance parameters (if that model exists) or
-for which one can tune the trade-off between power and robustness. Finally, we
+for which one can tune a trade-off between power and robustness. Finally, we
 demonstrate the effectiveness of this approach with a toy example and an example
@@
 target labels, and $Z$ are the nuisance parameters, we consider the problem of
-learning a predictive model for $Y$ conditional on the observed values of $X$
+learning a predictive model $f(X)$ for $Y$ conditional on the observed values of $X$
 that is robust to uncertainty in the unknown value of $Z$. More specifically, we
-introduce a learning procedure for enforcing the pivotal property
-\citep{degroot1986probability}  on a
-predictive model by jointly training two neural networks in an adversarial
-fashion. We derive theoretical results proving that the proposed procedure
-converges towards a predictive model that is both optimal and independent of the
-nuisance parameters (if that models exists) or for which one can tune the
-trade-off between power and robustness. Finally, we demonstrate the
+introduce a learning procedure based on adversarial networks
+for enforcing that  $f(X)$ is a pivot, i.e., a quantity whose distribution
+does not depend on $Z$.
+We derive theoretical results proving that the proposed procedure
+converges towards a predictive model that is both optimal and statistically independent of the
+nuisance parameters (if that model exists) or for which one can tune a
+trade-off between power and robustness, e.g., as driven by the requirements of a higher level statistical analysis.
+Finally, we demonstrate the
 effectiveness of this approach with a toy example and an example from particle physics.
@@
 In words, we are looking for a predictive function $f$
-which is a pivotal quantity \citep{degroot1986probability} with respect to the
+which is a pivotal quantity (see e.g., \citep{degroot1986probability}) with respect to the
 nuisance parameters. This implies that  $f(X; \theta_f)$ and $Z$ are independent random variables.
@@
 associated loss ${\cal L}_r(\theta_f, \theta_r)$. This model takes  as input
-realizations $f(X; \theta_f)$ and produces as output a function
+realizations $s$ of $f(X; \theta_f)$ and produces as output a function
 modeling the posterior probability density $p_{\theta_r}(z | f(X;\theta_f)=s)$.
@@
 that solution. Empirical evidence gathered here and in similar minimax-based
- works (see e.g., \citep{goodfellow2014generative,EdwardsS15}) suggests
+ works suggests
 that the alternating optimization algorithm works sufficiently well in practice,

### 2017-02-01 tex/icml2017/paper.tex
save

@@
 % the note in the first column to ``Under review.  Do not distribute.''
-\usepackage[accepted]{icml2016}
+\usepackage{icml2016}
 
@@
 generation processes. In this work,  we introduce a flexible training procedure
-based on adversarial networks for enforcing the pivotal property, i.e., the stastistical independence of a predictive
+based on adversarial networks for enforcing the pivotal property, i.e., the statistical independence of a predictive
 model with respect to nuisance parameters. We derive theoretical results showing that the proposed algorithm tends

### 2017-02-01 tex/icml2017/paper.tex
save

@@
 generation processes. In this work,  we introduce a flexible training procedure
-based on adversarial networks for enforcing the pivotal property, i.e., the statistical independence of a predictive
+based on adversarial networks for enforcing the pivotal property, i.e., the independence of a predictive
 model with respect to nuisance parameters. We derive theoretical results showing that the proposed algorithm tends

### 2017-02-02 tex/icml2017/paper.tex
save

@@
 generation processes. In this work,  we introduce a flexible training procedure
-based on adversarial networks for enforcing the pivotal property, i.e., the independence of a predictive
-model with respect to nuisance parameters. We derive theoretical results showing that the proposed algorithm tends
+based on adversarial networks for enforcing the pivotal property on a predictive model, i.e., its independence
+with respect to nuisance parameters. We derive theoretical results showing that the proposed algorithm tends
 towards a minimax solution corresponding to a predictive model that is both
@@
 uncertainties. The presence of systematic uncertainties affects both the
-robustness of inference and the metrics used to evaluate the performance of a
-particular data analysis strategy.
+robustness of inference and the performance metrics used to evaluate a
+particular analysis strategy.
 
@@
 present is when the training data are not representative of the real data that a
-predictive model will be applied to in practice. Several techniques for domain
+predictive model will be applied to in practice.
+Several techniques for domain
 adaptation have been developed to create predictive models that are more robust
-to this type of uncertainty. A more generic situation is that there are several
+to this type of uncertainty.
+A more generic situation is that there are several
 plausible data generation processes, specified as a family of data generation
@@
 continuous.
+In this broader context, statisticians have for long been working on the development of robust inference techniques
+based on the concept of a pivot -- a quantity whose distribution is invariant
+with the nuisance parameters. Pivotal quantities are e.g. fundamental to the construction
+of test statistics and confidence intervals.
 
@@
 nuisance parameters (if that model exists) or for which one can tune a
-trade-off between power and robustness, e.g., as driven by the requirements of a higher level statistical analysis.
+trade-off between power and robustness (e.g., as driven by the requirements of a higher level statistical analysis).
 Finally, we demonstrate the

### 2017-05-15 tex/nips/nips_2017.tex
Use NIPS template

--- 2017-02-02 tex/icml2017/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%% ICML 2016 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass{article}
 
-% Use the following line _only_ if you're still using LaTeX 2.09.
-%\documentstyle[icml2016,epsf,natbib]{article}
-% If you rely on Latex2e packages, like most moden people use this:
-\documentclass{article}
+% if you need to pass options to natbib, use, e.g.:
+% \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading nips_2017
+%
+% to avoid loading the natbib package, add option nonatbib:
+% \usepackage[nonatbib]{nips_2017}
+
+%\usepackage{nips_2017}
+
+% to compile a camera-ready version, add the [final] option, e.g.:
+\usepackage[final]{nips_2017}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
 
@@
 \usepackage{amsthm}
-
-% use Times
-\usepackage{times}
-% For figures
-\usepackage{graphicx} % more modern
-%\usepackage{epsfig} % less modern
+\usepackage{graphicx}
 \usepackage{subfigure}
+\usepackage{natbib}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 
@@
 
-% For citations
-\usepackage{natbib}
+%\newcommand{\theHalgorithm}{\arabic{algorithm}}
 
-% For algorithms
-\usepackage{algorithm}
-\usepackage{algorithmic}
+\title{Learning to Pivot with Adversarial Networks}
 
-% As of 2011, we use the hyperref package to produce hyperlinks in the
-% resulting PDF.  If this breaks your system, please commend out the
-% following usepackage line and replace \usepackage{icml2016} with
-% \usepackage[nohyperref]{icml2016} above.
-\usepackage{hyperref}
+% The \author macro works with any number of authors. There are two
+% commands used to separate the names and addresses of multiple
+% authors: \And and \AND.
+%
+% Using \And between authors leaves it to LaTeX to determine where to
+% break the lines. Using \AND forces a line break at that point. So,
+% if LaTeX puts 3 of 4 authors names on the first line, and the last
+% on the second line, try using \AND instead of \And before the third
+% author name.
 
-% Packages hyperref and algorithmic misbehave sometimes.  We can fix
-% this with the following command.
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-% Employ the following version of the ``usepackage'' statement for
-% submitting the draft version of the paper for review.  This will set
-% the note in the first column to ``Under review.  Do not distribute.''
-\usepackage{icml2016}
-
-% Employ this version of the ``usepackage'' statement after the paper has
-% been accepted, when creating the final version.  This will set the
-% note in the first column to ``Proceedings of the...''
-%\usepackage[accepted]{icml2016}
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Learning to Pivot with Adversarial Networks}
+\author{
+  Gilles Louppe\\
+  New York Univeristy\\
+  \texttt{g.louppe@nyu.edu}\\
+  \And
+  Michael Kagan\\
+  SLAC National Accelerator Laboratory\\
+  \texttt{makagan@slac.stanford.edu}\\
+  \And
+  Kyle Cranmer\\
+  New York Univeristy\\
+  \texttt{kyle.cranmer@nyu.edu}
+}
 
 \begin{document}
+% \nipsfinalcopy is no longer used
 
-\twocolumn[
-\icmltitle{Learning to Pivot with Adversarial Networks}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2016
-% package.
-\icmlauthor{Gilles Louppe}{g.louppe@nyu.edu}
-\icmladdress{New York University}
-\icmlauthor{Michael Kagan}{makagan@slac.stanford.edu}
-\icmladdress{SLAC National Accelerator Laboratory}
-\icmlauthor{Kyle Cranmer}{kyle.cranmer@nyu.edu}
-\icmladdress{New York University}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-%\icmlkeywords{boring formatting information, machine learning, ICML}
-
-\vskip 0.3in
-]
+\maketitle
 
 \begin{abstract}
-
     Many inference problems involve data generation processes that are not uniquely
@@
     from particle physics.
-
 \end{abstract}
@@
 
-    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.85]
+    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.65]
     \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!20, line width=0.3mm, fill=white]
@@
     \begin{flushleft}
-        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;\\
-        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$;\\
-        {\it Hyper-parameters:} Number $T$ of training iterations,
-                                Number $K$ of gradient steps to update $r$,
-                                Size $M$ of the minibatches.
+        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;
+        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$.
+        % {\it Hyper-parameters:} Number $T$ of iterations,
+        %                         Number $K$ of gradient steps to update $r$,
+        %                         Size $M$ of the minibatches.
     \end{flushleft}
@@
 
-\begin{figure*}
-    \begin{center}
-        \includegraphics[width=0.245\textwidth]{../figures/f-plain.pdf}
-        \includegraphics[width=0.245\textwidth]{../figures/surface-plain.pdf}
-        \includegraphics[width=0.245\textwidth]{../figures/f-adversary.pdf}
-        \includegraphics[width=0.245\textwidth]{../figures/surface-adversary.pdf}
-    \end{center}
-    \vspace{-0.5cm}
-    \caption{Toy example.
-    (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
-       when $f$ is built without adversarial training. The resulting densities
-       are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
-    (Middle left) The corresponding decision surface, highlighting
-       the fact that samples are easier to classify for values of $Z$ above $\sigma$,
-       hence partially explaining the dependency.
-    (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
-       built with adversarial training, as outlined in Sec.~\ref{sec:method}.
-       The resulting densities are now almost identical to each other, indicating only a
-       small dependency on $Z$.
-    (Right) The corresponding decision surface, illustrating how adversarial
-       training bends the decision function vertically to erase the dependency on $Z$.
-    }
-    \label{fig:toy}
-\end{figure*}
-
 As a guiding toy example, let us consider the binary classification of 2D
@@
 
+\begin{figure*}
+    \begin{center}
+        \includegraphics[width=0.245\textwidth]{../figures/f-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/surface-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/f-adversary.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/surface-adversary.pdf}
+    \end{center}
+    \caption{Toy example.
+    (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
+       when $f$ is built without adversarial training. The resulting densities
+       are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
+    (Middle left) The corresponding decision surface, highlighting
+       the fact that samples are easier to classify for values of $Z$ above $\sigma$,
+       hence partially explaining the dependency.
+    (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
+       built with adversarial training, as outlined in Sec.~\ref{sec:method}.
+       The resulting densities are now almost identical to each other, indicating only a
+       small dependency on $Z$.
+    (Right) The corresponding decision surface, illustrating how adversarial
+       training bends the decision function vertically to erase the dependency on $Z$.
+    }
+    \label{fig:toy}
+\end{figure*}
+
 \begin{figure}
-    \includegraphics[width=0.48\textwidth]{../figures/training.pdf}
-    \vspace{-1cm}
+\centering
+\begin{minipage}{.46\linewidth}
+    \centering
+        \includegraphics[width=0.95\textwidth]{../figures/training.pdf}
         \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
@@
         \label{fig:toy-training}
+\end{minipage}
+\hspace{.05\linewidth}
+\begin{minipage}{.46\linewidth}
+    \centering
+        \includegraphics[width=0.95\textwidth]{../figures/ams.pdf}
+        \caption{Physics example. Approximate median significance as a function of the decision threshold
+                 on the output of $f$.  As shown at $\lambda=10$, trading
+                 classification accuracy for independence with respect to pileup
+                 results in a net benefit in terms of statistical significance.}
+        \label{fig:physics-ams}
+\end{minipage}
 \end{figure}
@@
 Results reported below are averages over 5 runs.
-
-\begin{figure}
-    \includegraphics[width=0.48\textwidth]{../figures/ams.pdf}
-    \caption{Physics example. Approximate median significance as a function of the decision threshold
-             on the output of $f$.  As shown at $\lambda=10$, trading
-             classification accuracy for independence with respect to pileup
-             results in a net benefit in terms of statistical significance.}
-    \label{fig:physics-ams}
-\end{figure}
 
@@
 
-% \section*{Acknowledgements}
-%
-% We would like to thank the authors of \citep{baldi2016jet} for sharing the data used in their
-% studies.
-% KC and GL are both supported through NSF ACI-1450310, additionally KC is supported
-% through PHY-1505463 and PHY-1205376.  MK is supported by the US
-% Department of Energy (DOE) under grant DE-AC02-76SF00515 and by the
-% SLAC Panofsky Fellowship.
-
-% In the unusual situation where you want a paper to appear in the
-% references without citing it in the main text, use \nocite
-
 \bibliography{bibliography}
-\bibliographystyle{icml2016}
+\bibliographystyle{apalike}
 
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was
-% created by Lise Getoor and Tobias Scheffer, it was slightly modified
-% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz,
-% slightly modified from the 2009 version by Kiri Wagstaff and
-% Sam Roweis's 2008 version, which is slightly modified from
-% Prasad Tadepalli's 2007 version which is a lightly
-% changed version of the previous year's version by Andrew Moore,
-% which was in turn edited from those of Kristian Kersting and
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2017-05-15 tex/nips/nips_2017.tex
save

@@
 \usepackage{subfigure}
-\usepackage{natbib}
 \usepackage{algorithm}
@@
     (categorical or continuous) nuisance parameters that parametrizes this family of
-    generation processes. In this work,  we introduce a flexible training procedure
+    generation processes. In this work,  we introduce a training procedure
     based on adversarial networks for enforcing the pivotal property on a predictive model, i.e., its independence
@@
 based on the concept of a pivot -- a quantity whose distribution is invariant
-with the nuisance parameters. Pivotal quantities are e.g. fundamental to the construction
+with the nuisance parameters (see e.g., \citep{degroot1986probability}). Pivotal quantities are e.g. fundamental to the construction
 of test statistics and confidence intervals.
@@
 
-The remainder of the paper is structured as follows. In Sec.~\ref{sec:problem},
-we first formally define the problem of learning a pivotal, predictive model. In
-Sec. \ref{sec:method} and \ref{sec:theory}, we describe how adversarial
-networks can be trained to obtain predictive models that are independent of
-nuisance parameters and show that the proposed procedure converges
-towards a model that is both optimal and robust. The  effectiveness of the
-approach is then demonstrated empirically in Sec. \ref{sec:toy} and \ref{sec:hep}.
-In Sec.~\ref{sec:related}, we relate our work to
-close contributions from statistical inference, domain adaptation, and fairness
-in classification. Finally, we gather our concluding remarks in
-Sec.~\ref{sec:conclusions}.
+% % this adds nothing
+% The remainder of the paper is structured as follows. In Sec.~\ref{sec:problem},
+% we first formally define the problem of learning a pivotal, predictive model. In
+% Sec. \ref{sec:method} and \ref{sec:theory}, we describe how adversarial
+% networks can be trained to obtain predictive models that are independent of
+% nuisance parameters and show that the proposed procedure converges
+% towards a model that is both optimal and robust. The  effectiveness of the
+% approach is then demonstrated empirically in Sec. \ref{sec:toy} and \ref{sec:hep}.
+% In Sec.~\ref{sec:related}, we relate our work to
+% close contributions from statistical inference, domain adaptation, and fairness
+% in classification. Finally, we gather our concluding remarks in
+% Sec.~\ref{sec:conclusions}.
 
@@
 classifier) that minimizes a loss ${\cal L}_f(\theta_f)$ (e.g., the
-cross-entropy).
+cross-entropy). In classification, values $\smash{s \in {\cal S} = \mathbb{R}^{|{\cal Y}|}}$
+correspond to scores used for mapping hard predictions $y \in {\cal Y}$, while ${\cal S} = {\cal Y}$ for regression.
 
@@
 In words, we are looking for a predictive function $f$
-which is a pivotal quantity (see e.g., \citep{degroot1986probability}) with respect to the
+which is a pivotal quantity with respect to the
 nuisance parameters. This implies that  $f(X; \theta_f)$ and $Z$ are independent random variables.
@@
 
-If $Z$ takes discrete values, then $p_{\theta_r}$ can be represented e.g. as a
+If $Z$ takes discrete values, then $p_{\theta_r}$ can be represented as a
 probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal Z|}}$ whose
@@
 corresponding parameter $\gamma_j$ of that distribution (e.g., the mean,
-variance and mixing coefficients of its components). As in
-\citep{nix1994estimating,bishop1994mixture}, the estimated probability density
+variance and mixing coefficients of its components). The estimated probability density
 $p_{\theta_r}(z|f(X;\theta_f)=s)$ can then be evaluated for any $z \in {\cal Z}$ and any score $s \in {\cal S}$.

### 2017-05-16 tex/nips/nips_2017.tex
minor edits

@@
 \usepackage{algorithmic}
+
+\usepackage{color}
+\newcommand{\glnote}[1]{\textcolor{red}{[GL: #1]}}
 
@@
 
+\glnote{improve abstract}
+
 \section{Introduction}
@@
 that is robust to uncertainty in the unknown value of $Z$. More specifically, we
-introduce a learning procedure based on adversarial networks
+introduce a learning procedure based on adversarial networks~\citep{goodfellow2014generative}
 for enforcing that  $f(X)$ is a pivot, i.e., a quantity whose distribution
@@
 Let us assume that prior to incorporating the effect of uncertainty
-in $Z$, our goal is to learn a regression function $f : \mathcal{X} \mapsto
-{\cal S}$ of parameters $\theta_f$ (e.g., a neural network-based probabilistic
+in $Z$, our goal is to learn a regression function $f : \mathcal{X} \to
+{\cal S}$ with parameters $\theta_f$ (e.g., a neural network-based probabilistic
 classifier) that minimizes a loss ${\cal L}_f(\theta_f)$ (e.g., the
@@
 $z$. More specifically, the authors pit a generative model $g:
-\mathbb{R}^n \mapsto \mathbb{R}^p$ against an adversary classifier $d :
-\mathbb{R}^p \mapsto [0, 1]$ whose antagonistic objective is to recognize
+\mathbb{R}^n \to \mathbb{R}^p$ against an adversarial classifier $d :
+\mathbb{R}^p \to [0, 1]$ whose antagonistic objective is to recognize
 real data $X$ from generated data $g(Z)$. Both models $g$ and $d$ are trained
@@
 predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. As
-illustrated in Fig.~\ref{fig:architecture}, we pit $f$ against an adversary
-model $r := p_{\theta_r}(z | f(X;\theta_f)=s)$ of parameters $\theta_r$ and
+illustrated in Fig.~\ref{fig:architecture}, we pit $f$ against an adversarial
+model $r := p_{\theta_r}(z | f(X;\theta_f)=s)$ with parameters $\theta_r$ and
 associated loss ${\cal L}_r(\theta_f, \theta_r)$. This model takes  as input
@@
 If $Z$ takes discrete values, then $p_{\theta_r}$ can be represented as a
-probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal Z|}}$ whose
+probabilistic classifier $\mathbb{R} \to \mathbb{R}^{|{\cal Z|}}$ whose
 $j^\textrm{th}$ output (for $j=1, \dots, |{\cal Z}|$) is the estimated probability mass
@@
 %If $Z$ takes discrete values, then $p_{\theta_r}$ can be represented e.g. as a
-%probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal Z|}}$ whose
+%probabilistic classifier $\mathbb{R} \to \mathbb{R}^{|{\cal Z|}}$ whose
 %output $j$ (for $j=1, \dots, |{\cal Z}|$) is the estimated probability mass
@@
 Algorithm~\ref{alg:adversarial-training} in the case of a binary classifier $f :
-\mathbb{R}^p \mapsto [0,1]$ modeling $p(Y=1|X)$. For reasons further explained
+\mathbb{R}^p \to [0,1]$ modeling $p(Y=1|X)$. For reasons further explained
 in Sec.~\ref{sec:theory}, ${\cal L}_f$ and ${\cal L}_r$  are respectively set to the
@@
                             \end{bmatrix}\right) &\text{ when } Y=0, \\
-    x &\sim {\cal N}\left ((1,1+Z),  \begin{bmatrix}
+    x|Z=z &\sim {\cal N}\left ((1,1+z),  \begin{bmatrix}
                               1 & 0 \\
@@
 closest to our work, in which domain invariance and fairness are enforced
-through an adversarial minimax setup composed of a classifier and an adversary
+through an adversarial minimax setup composed of a classifier and an adversarial
 discriminator. Following this line of work and without making assumptions on the

### 2017-05-16 tex/nips/nips_2017.tex
trim

@@
 \usepackage{algorithmic}
+
+\usepackage[font=footnotesize]{caption}
 
@@
 \begin{abstract}
-    Many inference problems involve data generation processes that are not uniquely
-    specified or are uncertain in some way. In a scientific context, the presence of
-    several plausible data generation processes is often associated to the presence
-    of systematic uncertainties. Robust inference is possible if it is based on a
+    % Many inference problems involve data generation processes that are not uniquely
+    % specified or are uncertain in some way.
+    In a scientific context, the presence of
+    several plausible data generation processes is often associated to systematic uncertainties. Robust inference is possible if it is based on a
     pivot -- a quantity whose distribution does not depend on the unknown value of the
-    (categorical or continuous) nuisance parameters that parametrizes this family of
+    (typically continuous) nuisance parameters that parametrizes this family of
     generation processes. In this work,  we introduce a training procedure
-    based on adversarial networks for enforcing the pivotal property on a predictive model, i.e., its independence
-    with respect to nuisance parameters. We derive theoretical results showing that the proposed algorithm tends
+    based on adversarial networks for enforcing the pivotal property on a predictive model.
+    We derive theoretical results showing that the procedure tends
     towards a minimax solution corresponding to a predictive model that is both
     optimal and independent of the nuisance parameters (if that model exists) or
-    for which one can tune a trade-off between power and robustness. Finally, we
+    for which one can tune a trade-off between accuracy and robustness. We
     demonstrate the effectiveness of this approach with a toy example and an example
@@
 
-\glnote{improve abstract}
-
 \section{Introduction}
@@
 scientific process. One of the challenges of applying machine learning
-techniques to scientific problems is the need to incorporate systematic
-uncertainties. The presence of systematic uncertainties affects both the
-robustness of inference and the performance metrics used to evaluate a
+to scientific problems is the need to incorporate systematic
+uncertainties, which affect both the
+robustness of inference and the metrics used to evaluate a
 particular analysis strategy.
 
-In this work, we focus on supervised learning techniques where the presence of
+In this work, we focus on supervised learning where the presence of
 systematic uncertainty can be associated to the fact that the data generation
@@
 real world data. By contrast, a common situation when systematic uncertainty is
-present is when the training data are not representative of the real data that a
-predictive model will be applied to in practice.
+present is when the training data are not representative of the real data.
 Several techniques for domain
 adaptation have been developed to create predictive models that are more robust
-to this type of uncertainty.
+to this binary type of uncertainty.
 A more generic situation is that there are several
 plausible data generation processes, specified as a family of data generation
-processes parametrized by nuisance parameters, which can be categorical or
-continuous.
-In this broader context, statisticians have for long been working on the development of robust inference techniques
+processes parametrized by continuous nuisance parameters, as is typically found in scientific domains.
+In this broader context, statisticians have for long been working on robust inference techniques
 based on the concept of a pivot -- a quantity whose distribution is invariant
-with the nuisance parameters (see e.g., \citep{degroot1986probability}). Pivotal quantities are e.g. fundamental to the construction
-of test statistics and confidence intervals.
+with the nuisance parameters (see e.g., \citep{degroot1986probability}).
+% Pivotal quantities are e.g. fundamental to the construction
+% of test statistics and confidence intervals.
 
@@
 learning a predictive model $f(X)$ for $Y$ conditional on the observed values of $X$
-that is robust to uncertainty in the unknown value of $Z$. More specifically, we
+that is robust to uncertainty in the unknown value of $Z$. We
 introduce a learning procedure based on adversarial networks~\citep{goodfellow2014generative}
-for enforcing that  $f(X)$ is a pivot, i.e., a quantity whose distribution
-does not depend on $Z$.
+for enforcing that  $f(X)$ is a pivot with respect to $Z$.
 We derive theoretical results proving that the proposed procedure
@@
 nuisance parameters (if that model exists) or for which one can tune a
-trade-off between power and robustness (e.g., as driven by the requirements of a higher level statistical analysis).
+trade-off between accuracy and robustness (e.g., as driven by a higher level objective).
 Finally, we demonstrate the
@@
     distribution $p(z|f(X;\theta_f)=s)$ of the nuisance parameters as observed only through the output $f(X;\theta_f)$ of the classifier. By
-    maximizing the antagonistic objective ${\cal L}_r(\theta_f, \theta_r)$ (as part of minimizing ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$), the classifier
+    maximizing the antagonistic objective ${\cal L}_r(\theta_f, \theta_r)$, the classifier
     $f$ forces $p(z|f(X;\theta_f)=s)$ towards the prior $p(z)$, which happens
@@
 \label{sec:method}
+
+\begin{figure*}
+    \begin{minipage}{\linewidth}
+    \begin{algorithm}[H]
+    \caption{Adversarial training of a classifier $f$ against an adversary $r$.}
+
+    {\footnotesize
+    \begin{flushleft}
+        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;
+        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$.
+        % {\it Hyper-parameters:} Number $T$ of iterations,
+        %                         Number $K$ of gradient steps to update $r$,
+        %                         Size $M$ of the minibatches.
+    \end{flushleft}
+
+    \label{alg:adversarial-training}
+    \begin{algorithmic}[1]
+        \FOR{$t=1$ to $T$}
+            \FOR{$k=1$ to $K$} %\COMMENT{Update $r$}
+                \STATE{Sample minibatch $\{x_m, z_m, s_m = f(x_m;\theta_f) \}_{m=1}^M$ of size $M$;}
+                \STATE{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
+                $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|s_m)  ;$$}
+            \ENDFOR
+            \STATE{Sample minibatch $\{x_m, y_m, z_m, s_m = f(x_m;\theta_f)  \}_{m=1}^M$ of size $M$;} %\Comment{Update $f$}
+            \STATE{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
+            $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|s_m)  \right],$$
+            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $\mathbf{1}(y_m=0)(1-s_m) + \mathbf{1}(y_m=1)s_m$;}
+        \ENDFOR
+    \end{algorithmic}
+    }
+    \end{algorithm}
+    \end{minipage}
+\end{figure*}
 
@@
 
-\begin{figure*}
-    \begin{minipage}{\linewidth}
-    \begin{algorithm}[H]
-    \caption{Adversarial training of a classifier $f$ against an adversary $r$.}
-
-    \begin{flushleft}
-        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;
-        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$.
-        % {\it Hyper-parameters:} Number $T$ of iterations,
-        %                         Number $K$ of gradient steps to update $r$,
-        %                         Size $M$ of the minibatches.
-    \end{flushleft}
-
-    \label{alg:adversarial-training}
-    \begin{algorithmic}[1]
-        \FOR{$t=1$ to $T$}
-            \FOR{$k=1$ to $K$} %\COMMENT{Update $r$}
-                \STATE{Sample minibatch $\{x_m, z_m, s_m = f(x_m;\theta_f) \}_{m=1}^M$ of size $M$;}
-                \STATE{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
-                $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|s_m)  ;$$}
-            \ENDFOR
-            \STATE{Sample minibatch $\{x_m, y_m, z_m, s_m = f(x_m;\theta_f)  \}_{m=1}^M$ of size $M$;} %\Comment{Update $f$}
-            \STATE{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
-            $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|s_m)  \right],$$
-            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $\mathbf{1}(y_m=0)(1-s_m) + \mathbf{1}(y_m=1)s_m$;}
-        \ENDFOR
-    \end{algorithmic}
-    \end{algorithm}
-    \end{minipage}
-\end{figure*}
-
 % ==============================================================================
@@
 
-On many practical problems, the assumption of existence of an optimal and pivotal classifier may
+In practice, the assumption of existence of an optimal and pivotal classifier may
 not hold because the nuisance parameter directly shapes the decision boundary.
@@
 
-Interestingly, let us finally emphasize that these results hold using only the
-(1D) output $s$ of $f(\cdot;\theta_f)$ (in the case of binary classification) as
+Interestingly, let us finally emphasize that our results hold using only the
+(1D) output $s$ of $f(\cdot;\theta_f)$ as
 input to the adversary. We could similarly enforce an intermediate
 representation of the data to be pivotal, e.g. as in
-\citep{ganin2014unsupervised}, but this is in fact not necessary.
+\citep{ganin2014unsupervised}, but this is not necessary.
 
@@
 \label{sec:toy}
+
+\begin{figure*}
+    \begin{center}
+        \includegraphics[width=0.245\textwidth]{../figures/f-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/surface-plain.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/f-adversary.pdf}
+        \includegraphics[width=0.245\textwidth]{../figures/surface-adversary.pdf}
+    \end{center}
+    \caption{Toy example.
+    (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
+        without adversarial training. The resulting densities
+       are dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
+    (Middle left) The associated decision surface, highlighting
+       the fact that samples are easier to classify for values of $Z$ above $\sigma$,
+       hence explaining the dependency.
+    (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
+       built with adversarial training.
+       The resulting densities are now almost identical to each other, indicating only a
+       small dependency on $Z$.
+    (Right) The associated decision surface, illustrating how adversarial
+       training bends the decision function vertically to erase the dependency on $Z$.
+    }
+    \label{fig:toy}
+\end{figure*}
 
@@
 \end{align}
-The continuous nuisance parameter $Z$ represents in this case our
-uncertainty about the exact location of the mean of the second gaussian. Our goal is to
+The continuous nuisance parameter $Z$ here represents our
+uncertainty about the location of the mean of the second gaussian. Our goal is to
 build a classifier $f(\cdot;\theta_f)$ for predicting $Y$ given $X$, but such that
@@
 
-Assuming a gaussian prior $z \sim {\cal N}(0,1)$, we start by generating
-training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we train a neural
-network classifier $f$ minimizing ${\cal L}_f(\theta_f)$ without considering its
+Assuming a gaussian prior $z \sim {\cal N}(0,1)$, we generate
+data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we train a neural
+network  $f$ minimizing ${\cal L}_f(\theta_f)$ without considering its
 adversary $r$. The network architecture comprises 2 dense hidden layers of 20
@@
 output layer of 15 nodes corresponding to the means, standard deviations and
-mixture coefficients of the five gaussians. Output nodes for the mean values
+mixture coefficients of the gaussians. Output nodes for the mean values
 come with linear activations, output nodes for the standard deviations with
@@
 effectively reshapes the decision function so it that becomes almost independent
-on the nuisance parameter, as shown in Fig.~\ref{fig:toy}. In particular, the
+on the nuisance parameter, as shown in Fig.~\ref{fig:toy}. The
 conditional probability densities of the decision scores $f(X;\theta_f)$ are now
-very similar to each other, indicating only a small residual  dependency on the
+very similar to each other, indicating only a residual  dependency on the
 nuisance, as theoretically expected. The dynamics of adversarial training is
@@
 ${\cal L}_r$ and ${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each
-iteration of Algorithm~\ref{alg:adversarial-training}. In the first iterations,
+iteration. In the first iterations,
 we observe that the global objective ${\cal L}_f - \lambda {\cal L}_r$ is
@@
 increase of ${\cal L}_f$, but which results in a classifier that is more
-pivotal, hence the corresponding increase of ${\cal L}_r$ and the total net
-benefit. As learning goes, minimizing $E$ then requires making predictions that
+pivotal, hence the associated increase of ${\cal L}_r$ and the total net
+benefit. As learning goes, minimizing $E$ requires making predictions that
 are more accurate, hence decreasing ${\cal L}_f$, or that are even less
-dependent on $Z$, hence shaping $p_{\theta_r}$ towards the prior $p(z)$. Indeed,
-${\cal L}_f$ eventually starts to  decrease, while remaining bounded from below by
+dependent on $Z$, hence shaping $p_{\theta_r}$ towards the prior $p(Z)$. Indeed,
+${\cal L}_f$ eventually starts decreasing, while remaining bounded from below by
 $\min_{\theta_f} {\cal L}_f(\theta_f)$ as approximated by the dashed line in the
@@
 $H(Z)$ of the prior (where $H(Z) = \log(\sigma \sqrt{2 \pi e}) = 1.419$ in the
-case of a gaussian with unit variance), as shown by the dashed line in the
+case of a standard normal), as shown by the dashed line in the
 second plot. Finally, let us note that the ideal situation of a classifier that
@@
 
-\begin{figure*}
-    \begin{center}
-        \includegraphics[width=0.245\textwidth]{../figures/f-plain.pdf}
-        \includegraphics[width=0.245\textwidth]{../figures/surface-plain.pdf}
-        \includegraphics[width=0.245\textwidth]{../figures/f-adversary.pdf}
-        \includegraphics[width=0.245\textwidth]{../figures/surface-adversary.pdf}
-    \end{center}
-    \caption{Toy example.
-    (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
-       when $f$ is built without adversarial training. The resulting densities
-       are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
-    (Middle left) The corresponding decision surface, highlighting
-       the fact that samples are easier to classify for values of $Z$ above $\sigma$,
-       hence partially explaining the dependency.
-    (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
-       built with adversarial training, as outlined in Sec.~\ref{sec:method}.
-       The resulting densities are now almost identical to each other, indicating only a
-       small dependency on $Z$.
-    (Right) The corresponding decision surface, illustrating how adversarial
-       training bends the decision function vertically to erase the dependency on $Z$.
-    }
-    \label{fig:toy}
-\end{figure*}
-
 \begin{figure}
@@
     \centering
-        \includegraphics[width=0.95\textwidth]{../figures/training.pdf}
+        \includegraphics[width=\textwidth]{../figures/training.pdf}
         \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
@@
     \centering
-        \includegraphics[width=0.95\textwidth]{../figures/ams.pdf}
+        \includegraphics[width=\textwidth]{../figures/ams.pdf}
         \caption{Physics example. Approximate median significance as a function of the decision threshold
-                 on the output of $f$.  As shown at $\lambda=10$, trading
-                 classification accuracy for independence with respect to pileup
+                 on the output of $f$. At $\lambda=10$, trading
+                 accuracy for independence to pileup
                  results in a net benefit in terms of statistical significance.}
@@
 Experiments at high energy colliders like the LHC \citep{LHCMachine} are
-searching for evidence of new particles beyond the ones described by the well
-established Standard Model (SM) of particle physics. A wide array of extended
+searching for evidence of new particles beyond those described by the
+Standard Model (SM) of particle physics. A wide array of
 theories predict the existence of new massive particles that would decay to
@@
 can decay to two quarks, each of which produce collimated sprays of particles
-known as \textit{jets} (collectively denoted $W\to jj$). If the exotic particle
+known as jets.
+%(collectively denoted $W\to jj$).
+If the exotic particle
 is heavy, then the $W$ boson will be moving very fast, and  relativistic effects
@@
 $W$-jets have a rich internal substructure (see e.g.
-\citep{Altheimer:2012mn,Altheimer:2013yza} and references therein). However,
+\citep{Altheimer:2012mn,Altheimer:2013yza}). However,
 jets are also produced ubiquitously at high energy colliders through more
@@
 that is beset with a number of sources of systematic uncertainty.
-
 The classification challenge used here is common in jet substructure
-studies (see \citep{Khachatryan:2014vla,ATL-PHYS-PUB-2015-033,wbosonATLAS} and
-references therein): we aim to discriminate between the normal jets produced copiously
-at the LHC ($Y=0$) and the $W$-jets ($Y=1$) potentially coming from an exotic process.
+studies (see e.g. \citep{Khachatryan:2014vla,ATL-PHYS-PUB-2015-033,wbosonATLAS}): we aim to distinguish normal jets produced copiously
+at the LHC ($Y=0$) and from $W$-jets ($Y=1$) potentially coming from an exotic process.
 We reuse the datasets used in
-\citep{baldi2016jet}.  For completeness, the jets are built with the
-anti-$k_T$ algorithm \citep{Cacciari:2008gp}
-with radius parameter $R=1.2$, and trimmed \citep{Krohn:2009th} with subjets built with
-the $k_T$ algorithm and parameter $f_{cut}=0.2$. The features used
-in the classifiers are: trimmed jet invariant mass, N-subjettiness
-$\tau_{21}^{\beta=1}$ \citep{nsub,Thaler2012}, and the energy correlation
-functions  $C_{2}^{\beta=1}$, $C_{2}^{\beta=2}$,
-$D_{2}^{\beta=1}$, $D_{2}^{\beta=2}$ \citep{Larkoski2013}.
+\citep{baldi2016jet}.
+% For completeness, the jets are built with the
+% anti-$k_T$ algorithm \citep{Cacciari:2008gp}
+% with radius parameter $R=1.2$, and trimmed \citep{Krohn:2009th} with subjets built with
+% the $k_T$ algorithm and parameter $f_{cut}=0.2$. The features used
+% in the classifiers are: trimmed jet invariant mass, N-subjettiness
+% $\tau_{21}^{\beta=1}$ \citep{nsub,Thaler2012}, and the energy correlation
+% functions  $C_{2}^{\beta=1}$, $C_{2}^{\beta=2}$,
+% $D_{2}^{\beta=1}$, $D_{2}^{\beta=2}$ \citep{Larkoski2013}.
 
@@
 wish to impose the pivotal property on $Y=0$ events. To this end,
-we use as a higher level objective the Approximate Median Significance (AMS), which was used
-in the Higgs boson machine learning challenge and is a natural generalization
+we use as a higher level objective the Approximate Median Significance (AMS), which is a natural generalization
 of the power of a hypothesis test when systematic uncertainties are taken into account
@@
 is made stronger (for $\lambda > 0$) the AMS peak moves higher, with a maximum
-value around  $7.8$ for $\lambda=10$. In other words, trading classification
-accuracy for robustness to pileup results in a net
-benefit in terms of the power of the resulting hypothesis test. Setting $\lambda$ too high however
+value around  $7.8$ for $\lambda=10$. Trading classification
+accuracy for robustness to pileup thereby results in a net
+benefit in terms of the power of the hypothesis test. Setting $\lambda$ too high however
 (e.g. $\lambda=500$) results in a decrease of the maximum AMS, by
 focusing the capacity of $f$ too strongly on independence with $Z$, at the expense of
-classification accuracy. As demonstrated in this example, optimizing $\lambda$
-gives us a principled and effective approach to control the trade-off between
+accuracy. In effect, optimizing $\lambda$
+yields a principled and effective approach to control the trade-off between
 accuracy and robustness that ultimately maximizes the
@@
 
-In machine learning, learning a pivotal quantity can be related to the problem
+Learning a pivotal quantity can be related to the problem
 of domain adaptation
@@
 enforcing fairness in classification
-\citep{zemel2013learning,feldman2015certifying,EdwardsS15,zafar2015fairness},
+\citep{kamishima2012fairness,zemel2013learning,feldman2015certifying,EdwardsS15,zafar2015fairness,louizos2015variational},
 which is stated as learning a classifier that is independent of some chosen
@@
 through an adversarial minimax setup composed of a classifier and an adversarial
-discriminator. Following this line of work and without making assumptions on the
-prior $p(Z)$, our method can be regarded as a generalization that also supports
+discriminator. Following this line of work, our method can be regarded as a unified generalization that also supports
 a continuously parametrized family of domains or as enforcing fairness over
 continuous attributes.
+
+While our contribution may first appear as a minor extension, most previous
+works are based on the strong and limiting assumption that $Z$ is a binary
+random variable (e.g., $Z=0$ for the source domain, and
+$Z=1$ for the discrete domain). For
+instance, the motivation behind \citep{pan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,zemel2013learning,ganin2014unsupervised,ajakan2014domain,EdwardsS15,louizos2015variational}
+takes root in the minimization of some form of divergence between the two distributions of
+$f(X)|Z=0$ and $f(X)|Z=1$. For this reason, and in contrast with our procedure,
+these works cannot intrinsically be generalized to non-binary or
+continuous nuisance parameters, both from a practical and theoretical point of view.
+Of notable interest however, \cite{kamishima2012fairness} define
+a prejudice regularization term based on empirical estimates of $p(f(X)|Z)$.
+While this approach is in principle sufficient for handling non-binary
+nuisance parameters $Z$, it requires accurate
+empirical estimates of $p(f(X)|Z=z)$ for all values $z$, which quickly becomes
+impractical as the cardinality of $Z$ increases. By contrast, our approach models the conditional
+dependence through an adversarial network, which allows for generalization without
+necessarily requiring a growing number of training examples.
 
@@
 
-Despite the theoretical characterization of the minimax solution of Eqn.~\ref{eqn:min_thetaf}
-that motivates this work,
-Algorithm~\ref{alg:adversarial-training} remains a heuristic in the sense
-that we do not provide formal guarantees of convergence of the procedure towards
-that solution. Empirical evidence gathered here and in similar minimax-based
- works suggests
-that the alternating optimization algorithm works sufficiently well in practice,
-but further research studying its stability and convergence would certainly
-be beneficial for a large body of works.
+% Despite the theoretical characterization of the minimax solution of Eqn.~\ref{eqn:min_thetaf}
+% that motivates this work,
+% Algorithm~\ref{alg:adversarial-training} remains a heuristic in the sense
+% that we do not provide formal guarantees of convergence of the procedure towards
+% that solution. Empirical evidence gathered here and in similar minimax-based
+%  works suggests
+% that the alternating optimization algorithm works sufficiently well in practice,
+% but further research studying its stability and convergence would certainly
+% be beneficial for a large body of works.
 

### 2017-05-16 tex/nips/nips_2017.tex
save

@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-\usepackage[final]{nips_2017}
+\usepackage{nips_2017}
 
@@
 
-In this work, we focus on supervised learning where the presence of
+In this work, we focus on supervised learning
 systematic uncertainty can be associated to the fact that the data generation
@@
 Several techniques for domain
-adaptation have been developed to create predictive models that are more robust
+adaptation have been developed to create models that are more robust
 to this binary type of uncertainty.
 A more generic situation is that there are several
-plausible data generation processes, specified as a family of data generation
-processes parametrized by continuous nuisance parameters, as is typically found in scientific domains.
+plausible data generation processes, specified as a family
+parametrized by continuous nuisance parameters, as is typically found in scientific domains.
 In this broader context, statisticians have for long been working on robust inference techniques
@@
 that is robust to uncertainty in the unknown value of $Z$. We
-introduce a learning procedure based on adversarial networks~\citep{goodfellow2014generative}
+introduce a flexible learning procedure based on adversarial networks~\citep{goodfellow2014generative}
 for enforcing that  $f(X)$ is a pivot with respect to $Z$.
-We derive theoretical results proving that the proposed procedure
-converges towards a predictive model that is both optimal and statistically independent of the
+We derive theoretical results proving that the procedure
+converges towards a model that is both optimal and statistically independent of the
 nuisance parameters (if that model exists) or for which one can tune a
 trade-off between accuracy and robustness (e.g., as driven by a higher level objective).
+In particular, and to the best of our knowledge, our contribution is the first
+working solution for imposing statistical independence on the predictive model, regardless of the type of
+the nuisance parameter (discrete or continuous) or of its prior.
 Finally, we demonstrate the
-effectiveness of this approach with a toy example and an example from particle physics.
+effectiveness of the approach with a toy example and an example from particle physics.
 
@@
 continuous nuisance parameters, both from a practical and theoretical point of view.
-Of notable interest however, \cite{kamishima2012fairness} define
+Of notable interest however, \cite{kamishima2012fairness} enforces fairness through
 a prejudice regularization term based on empirical estimates of $p(f(X)|Z)$.

### 2017-05-17 tex/nips/nips_2017.tex
save

@@
 In particular, and to the best of our knowledge, our contribution is the first
-working solution for imposing statistical independence on the predictive model, regardless of the type of
+solution for imposing pivotal constraints on a predictive model,
+working regardless of the type of
 the nuisance parameter (discrete or continuous) or of its prior.
@@
 random variable (e.g., $Z=0$ for the source domain, and
-$Z=1$ for the discrete domain). For
-instance, the motivation behind \citep{pan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,zemel2013learning,ganin2014unsupervised,ajakan2014domain,EdwardsS15,louizos2015variational}
-takes root in the minimization of some form of divergence between the two distributions of
+$Z=1$ for the target domain).
+In particular, \citep{pan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,zemel2013learning,ganin2014unsupervised,ajakan2014domain,EdwardsS15,louizos2015variational}
+are all based on the minimization of some form of divergence between the two distributions of
 $f(X)|Z=0$ and $f(X)|Z=1$. For this reason, and in contrast with our procedure,

### 2017-05-18 tex/nips/nips_2017.tex
bring in Shimmin into experiments
Had to chop some sentences to make it fit.

@@
   Gilles Louppe\\
-  New York Univeristy\\
+  New York University\\
   \texttt{g.louppe@nyu.edu}\\
@@
   Kyle Cranmer\\
-  New York Univeristy\\
+  New York University\\
   \texttt{kyle.cranmer@nyu.edu}
@@
     pivot -- a quantity whose distribution does not depend on the unknown value of the
-    (typically continuous) nuisance parameters that parametrizes this family of
+    (typically continuous) nuisance parameters that parametrize this family of
     generation processes. In this work,  we introduce a training procedure
@@
 quantity.
-
 \end{proof}
@@
 
-\subsection{Toy example}
+\subsection{A toy example with a continous nuisance parameter}
 \label{sec:toy}
@@
     \centering
-        \includegraphics[width=\textwidth]{../figures/training.pdf}
+        \includegraphics[width=\textwidth]{../figures/training.pdf}\vspace{-1em}
         \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
@@
     \centering
-        \includegraphics[width=\textwidth]{../figures/ams.pdf}
+        \includegraphics[width=\textwidth]{../figures/ams.pdf}\vspace{-1em}
         \caption{Physics example. Approximate median significance as a function of the decision threshold
@@
 
-\subsection{High energy physics example}
+\subsection{High energy physics examples}
 \label{sec:hep}
 
-Experiments at high energy colliders like the LHC \citep{LHCMachine} are
+\paragraph{Binary Case} Experiments at high energy colliders like the LHC \citep{LHCMachine} are
 searching for evidence of new particles beyond those described by the
@@
 of the power of a hypothesis test when systematic uncertainties are taken into account
-(see Eqn.~20 of \citep{adam2014higgs}).
+(see Eqn.~20 of \cite{adam2014higgs}).
 
@@
 
+\paragraph{Continous Case} 
+%Our method also applies to cases where one desires the predictive model to be independent of 
+%some other observed random variables, as in domain adaptation or fairness in
+%classification. Importantly, our approach is able to enforce fairness over
+%continuous attributes.
+Recently, an independent group has used our approach to learn jet classifiers 
+that are independent of the jet mass, which is a continuous attribute~\citep{Shimmin:2017mfk}. 
+The results of their studies show that the adversarial training strategy works very well for 
+real-world problems with continuous attributes, thus enhancing the sensitivity of searches for 
+new physics at the LHC.
+
 % ==============================================================================
@@
 
-While our contribution may first appear as a minor extension, most previous
-works are based on the strong and limiting assumption that $Z$ is a binary
+%While our contribution may first appear as a minor extension, most
+Most related
+work is based on the strong and limiting assumption that $Z$ is a binary
 random variable (e.g., $Z=0$ for the source domain, and
@@
 are all based on the minimization of some form of divergence between the two distributions of
-$f(X)|Z=0$ and $f(X)|Z=1$. For this reason, and in contrast with our procedure,
-these works cannot intrinsically be generalized to non-binary or
+$f(X)|Z=0$ and $f(X)|Z=1$. For this reason, 
+%and in contrast with our procedure,
+these works cannot directly be generalized to non-binary or
 continuous nuisance parameters, both from a practical and theoretical point of view.
-Of notable interest however, \cite{kamishima2012fairness} enforces fairness through
+Notably, \cite{kamishima2012fairness} enforces fairness through
 a prejudice regularization term based on empirical estimates of $p(f(X)|Z)$.
@@
 the most statistically powerful approach for incorporating the effect of
-systematics on the underlying classification task.  As argued in
-\citep{Neal:2007zz}, it is not obvious how such classifiers can be used on real
-data since the correct value $z$ of the nuisance often remains unknown. This is
-not an issue in the context of parameter
-inference~\citep{cranmer2015approximating}, where nuisance parameters are
-eliminated via optimization or marginalization. In practice, parametrized
+systematics on the underlying classification task.  
+%As argued in
+%\citep{Neal:2007zz}, it is not obvious how such classifiers can be used on real
+%data since the correct value $z$ of the nuisance often remains unknown. This is
+%not an issue in the context of parameter
+%inference~\citep{cranmer2015approximating}, where nuisance parameters are
+%eliminated via optimization or marginalization. 
+In practice, parametrized
 classifiers  are also computationally expensive to build and evaluate. In
@@
 $p(f(x,z)|y,z)$ as a continuous function of $z$, remains an open challenge. By
-contrast, constraining $f$ to be pivotal yields a classifier which may not be
-optimal with respect to $Y|X,Z$, as discussed in Sec.~\ref{sec:theory}, but that
-can otherwise be directly used in a wider range of applications, since the
-dependence on the nuisance parameter $Z$ has already been eliminated. Moreover,
-if $f$ is sufficiently close to being pivotal, then calibration only needs to be
-carried out once.
+contrast, constraining $f$ to be pivotal yields a classifier 
+%which may not be
+%optimal with respect to $Y|X,Z$, as discussed in Sec.~\ref{sec:theory}, but 
+that can 
+%otherwise 
+be directly used in a wider range of applications, since the
+dependence on the nuisance parameter $Z$ has already been eliminated. 
+%Moreover, if $f$ is sufficiently close to being pivotal, then calibration only needs to be
+%carried out once.
 
@@
 observed random variables is desired, as in domain adaptation or fairness in
-classification. Again, trying our algorithm on these use cases is certainly
-worth investigating.
+classification (e.g. \citep{Shimmin:2017mfk}).
+% not sure this is a great sentence to end with, sounds like a sales pitch
+% Again, trying our algorithm on these use cases is certainly
+%worth investigating.
 

### 2017-05-18 tex/nips/nips_2017.tex
new abstract

@@
 \begin{abstract}
-    % Many inference problems involve data generation processes that are not uniquely
-    % specified or are uncertain in some way.
-    In a scientific context, the presence of
-    several plausible data generation processes is often associated to systematic uncertainties. Robust inference is possible if it is based on a
+Several techniques for domain adaptation have been proposed to account for differences in the
+distribution of the data used for training and testing. The majority of this work focuses on
+a binary domain label. Similar problems occur in a scientific context where there may be a continuous family of
+plausible data generation processes associated to the presence of systematic uncertainties. 
+Robust inference is possible if it is based on a
     pivot -- a quantity whose distribution does not depend on the unknown value of the
-    (typically continuous) nuisance parameters that parametrize this family of
-    generation processes. In this work,  we introduce a training procedure
-    based on adversarial networks for enforcing the pivotal property on a predictive model.
-    We derive theoretical results showing that the procedure tends
-    towards a minimax solution corresponding to a predictive model that is both
-    optimal and independent of the nuisance parameters (if that model exists) or
-    for which one can tune a trade-off between accuracy and robustness. We
-    demonstrate the effectiveness of this approach with a toy example and an example
+    nuisance parameters that parametrize this family of
+    data generation processes. In this work,  we introduce and derive theoretical results
+     for ta training procedure
+    based on adversarial networks for enforcing the pivotal property (or, equivalently, fairness with respect to continuous attributes) on a predictive model.
+    The method includes a hyperparameter, which one can tune a trade-off between accuracy and robustness. We
+    demonstrate the effectiveness of this approach with a toy example and an examples
     from particle physics.
 \end{abstract}
+
+%\begin{abstract}
+%    % Many inference problems involve data generation processes that are not uniquely
+%    % specified or are uncertain in some way.
+%    In a scientific context, the presence of
+%    several plausible data generation processes is often associated to systematic uncertainties. Robust inference is possible if it is based on a
+%    pivot -- a quantity whose distribution does not depend on the unknown value of the
+%    (typically continuous) nuisance parameters that parametrize this family of
+%    generation processes. In this work,  we introduce a training procedure
+%    based on adversarial networks for enforcing the pivotal property on a predictive model.
+%    We derive theoretical results showing that the procedure tends
+%    towards a minimax solution corresponding to a predictive model that is both
+%    optimal and independent of the nuisance parameters (if that model exists) or
+%    for which one can tune a trade-off between accuracy and robustness. We
+%    demonstrate the effectiveness of this approach with a toy example and an example
+%    from particle physics.
+%\end{abstract}
 

### 2017-05-18 tex/nips/nips_2017.tex
fix typo

@@
 
-In this work, we focus on supervised learning
-systematic uncertainty can be associated to the fact that the data generation
-process is not uniquely specified. In other words, the lack of systematic
+In this work, we focus on supervised learning techniques where 
+systematic uncertainties can be associated to a data generation
+process that is not uniquely specified. In other words, the lack of systematic
 uncertainties corresponds to the (rare) case that the process that generates

### 2017-05-18 tex/nips/nips_2017.tex
another grammatical fix
typos

@@
 \begin{abstract}
-Several techniques for domain adaptation have been proposed to account for differences in the
-distribution of the data used for training and testing. The majority of this work focuses on
-a binary domain label. Similar problems occur in a scientific context where there may be a continuous family of
-plausible data generation processes associated to the presence of systematic uncertainties. 
-Robust inference is possible if it is based on a
-    pivot -- a quantity whose distribution does not depend on the unknown value of the
-    nuisance parameters that parametrize this family of
-    data generation processes. In this work,  we introduce and derive theoretical results
-     for ta training procedure
-    based on adversarial networks for enforcing the pivotal property (or, equivalently, fairness with respect to continuous attributes) on a predictive model.
-    The method includes a hyperparameter, which one can tune a trade-off between accuracy and robustness. We
-    demonstrate the effectiveness of this approach with a toy example and an examples
-    from particle physics.
+
+Several techniques for domain adaptation have been proposed to account for
+differences in the distribution of the data used for training and testing. The
+majority of this work focuses on a binary domain label. Similar problems occur
+in a scientific context where there may be a continuous family of plausible data
+generation processes associated to the presence of systematic uncertainties.
+Robust inference is possible if it is based on a pivot -- a quantity whose
+distribution does not depend on the unknown value of the nuisance parameters
+that parametrize this family of data generation processes. In this work,  we
+introduce and derive theoretical results for a training procedure based on
+adversarial networks for enforcing the pivotal property (or, equivalently,
+fairness with respect to continuous attributes) on a predictive model. The
+method includes a hyperparameter to control the trade-off between accuracy and
+robustness. We demonstrate the effectiveness of this approach with a toy example
+and an examples from particle physics.
+
 \end{abstract}
@@
 observed random variables is desired, as in domain adaptation or fairness in
-classification (e.g. \citep{Shimmin:2017mfk}).
+classification (e.g. \cite{Shimmin:2017mfk}).
 % not sure this is a great sentence to end with, sounds like a sales pitch

### 2017-05-19 tex/nips/nips_2017.tex
hack for broken figure
fix figure
address last comments, small fixes

@@
 Robust inference is possible if it is based on a pivot -- a quantity whose
-distribution does not depend on the unknown value of the nuisance parameters
+distribution does not depend on the unknown values of the nuisance parameters
 that parametrize this family of data generation processes. In this work,  we
@@
 robustness. We demonstrate the effectiveness of this approach with a toy example
-and an examples from particle physics.
+and examples from particle physics.
 
@@
 Finally, we demonstrate the
-effectiveness of the approach with a toy example and an example from particle physics.
+effectiveness of the approach with a toy example and examples from particle physics.
 
@@
 
-    \node[above] at (12.75,-2) {$\gamma_1(f(X;\theta_f);\theta_r)$};
-    \path[-o, shorten >= 0pt] (11,-2.0) edge (14,-2.0);
-    \node[above] at (12.75,-3) {$\gamma_2(f(X;\theta_f);\theta_r)$};
-    \path[-o, shorten >= 0pt] (11,-3) edge (14,-3);
-    \node[above] at (12.75,-4) {$\dots$};
-    \path[-o, shorten >= 0pt] (11,-4) edge (14,-4);
+   \node[above] at (13.25,-2) {$\gamma_1(f(X;\theta_f);\theta_r)$};
+   \path[-o, shorten >= 0pt] (11,-2.0) edge (15,-2.0);
+   \node[above] at (13.25,-3) {$\gamma_2(f(X;\theta_f);\theta_r)$};
+   \path[-o, shorten >= 0pt] (11,-3) edge (15,-3);
+   \node[above] at (13.25,-4) {$\dots$};
+   \path[-o, shorten >= 0pt] (11,-4) edge (15,-4);
 
@@
 
-    \draw (14,-1.5) rectangle (17,-4.5);
-    \path[->, shorten >= 0pt] (15.5,-0.5) edge (15.5,-1.5);
-    \node[above] at (15.5,-0.5) {$Z$};
-    \path[->, shorten >= 0pt] (15.5,-4.5) edge (15.5,-5.5);
-    \node[below] at (15.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
-    \node at (15.5,-3) {${\cal P}(\gamma_1, \gamma_2, \dots)$};
+   \draw (15,-1.5) rectangle (18,-4.5);
+   \path[->, shorten >= 0pt] (16.5,-0.5) edge (16.5,-1.5);
+   \node[above] at (16.5,-0.5) {$Z$};
+   \path[->, shorten >= 0pt] (16.5,-4.5) edge (16.5,-5.5);
+   \node[below] at (16.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
+   \node at (16.5,-3) {${\cal P}(\gamma_1, \gamma_2, \dots)$};
 
-    \draw[dashed,-] (15.5,-5) -| (12.75,-6.5);
+   \draw[dashed,-] (16.5,-5) -| (12.75,-6.5);
    \node[below] at (12.75,-6.5) {${\cal L}_r(\theta_f, \theta_r)$};
@@
    \end{tikzpicture}
+
+% \includegraphics[width=\linewidth]{pivot-figure}
 
@@
 cross-entropy). In classification, values $\smash{s \in {\cal S} = \mathbb{R}^{|{\cal Y}|}}$
-correspond to scores used for mapping hard predictions $y \in {\cal Y}$, while ${\cal S} = {\cal Y}$ for regression.
+correspond to the classifier scores used for mapping hard predictions $y \in {\cal Y}$, while ${\cal S} = {\cal Y}$ for regression.
 
@@
 classifier minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not
-necessarily with $Y|X,Z$ (unless $Z$ is made explicit and is included among the
-input variables in $X$). Results hold for a nuisance parameters $Z$ taking either
+necessarily with $Y|X,Z$.
+% (unless $Z$ is made explicit and is included among the
+% input variables in $X$).
+Results hold for a nuisance parameters $Z$ taking either
 categorical or continuous values. By abuse of notation, $H(Z)$ denotes the
@@
 $\mathbb{E}_{s \sim f(X;\theta_f)} [ H({Z|f(X;\theta_f)=s}) ]$ of the conditional distribution of the nuisance parameters.
-This expectation is nothing else than the conditional entropy of the random variables
+This expectation corresponds to the conditional entropy of the random variables
 $Z$ and $f(X;\theta_f)$ and can be written as $H(Z|f(X;\theta_f))$.
@@
 
+In this section, we empirically demonstrate the effectiveness of the approach
+with a toy example and examples from particle physics. Notably,
+there are no other other approaches to compare to in the case of continuous
+nuisance parameters, as further explained in Sec.~\ref{sec:related}. In the case
+of binary parameters, we do not expect results to be much different from
+previous works.
+
 \subsection{A toy example with a continous nuisance parameter}
@@
     \end{center}
+
     \caption{Toy example.
@@
 will cause the two jets from its decay to merge into a single `$W$-jet'. These
-$W$-jets have a rich internal substructure (see e.g.
-\citep{Altheimer:2012mn,Altheimer:2013yza}). However,
+$W$-jets have a rich internal substructure.
+% (see e.g.\citep{Altheimer:2012mn,Altheimer:2013yza}).
+However,
 jets are also produced ubiquitously at high energy colliders through more
@@
 Recently, an independent group has used our approach to learn jet classifiers
-that are independent of the jet mass, which is a continuous attribute~\citep{Shimmin:2017mfk}.
+that are independent of the jet mass~\citep{Shimmin:2017mfk}, which is a continuous attribute.
 The results of their studies show that the adversarial training strategy works very well for
@@
 
-Learning a pivotal quantity can be related to the problem
+Learning to pivot can be related to the problem
 of domain adaptation
@@
 
-In terms of applications, the proposed solution can be used in any situation
+In terms of applications, our solution can be used in any situation
 where the training data may not be representative of the real data the
@@
 also extends to cases where independence of the predictive model with respect to
-observed random variables is desired, as in domain adaptation or fairness in
-classification (e.g. \cite{Shimmin:2017mfk}).
+observed random variables is desired, as in fairness for
+classification.
 % not sure this is a great sentence to end with, sounds like a sales pitch

### 2017-10-07 tex/paper.tex
camera ready version

@@
-\documentclass[twocolumn,superscriptaddress,aps]{revtex4-1}
+\documentclass{article}
 
-\usepackage[utf8]{inputenc}
+% if you need to pass options to natbib, use, e.g.:
+% \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading nips_2017
+%
+% to avoid loading the natbib package, add option nonatbib:
+% \usepackage[nonatbib]{nips_2017}
+
+%\usepackage{nips_2017}
+
+% to compile a camera-ready version, add the [final] option, e.g.:
+\usepackage[final]{nips_2017}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
 
@@
 \usepackage{amsthm}
+\usepackage{graphicx}
+\usepackage{subfigure}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 
-\usepackage{bm}
-\usepackage{cancel}
-\usepackage{bbold}
-\usepackage{slashed}
-\usepackage{graphicx}
+\usepackage[font=footnotesize]{caption}
+
 \usepackage{color}
-\usepackage{hyperref}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
+\newcommand{\glnote}[1]{\textcolor{red}{[GL: #1]}}
+
 \usepackage{tikz}
-
-\newcommand{\glnote}[1]{\textcolor{red}{[GL: #1]}}
-\newcommand{\kcnote}[1]{\textcolor{red}{[KC: #1]}}
-
 \theoremstyle{plain}
@@
 
+%\newcommand{\theHalgorithm}{\arabic{algorithm}}
+
+\title{Learning to Pivot with Adversarial Networks}
+
+% The \author macro works with any number of authors. There are two
+% commands used to separate the names and addresses of multiple
+% authors: \And and \AND.
+%
+% Using \And between authors leaves it to LaTeX to determine where to
+% break the lines. Using \AND forces a line break at that point. So,
+% if LaTeX puts 3 of 4 authors names on the first line, and the last
+% on the second line, try using \AND instead of \And before the third
+% author name.
+
+\author{
+  Gilles Louppe\\
+  New York University\\
+  \texttt{g.louppe@nyu.edu}\\
+  \And
+  Michael Kagan\\
+  SLAC National Accelerator Laboratory\\
+  \texttt{makagan@slac.stanford.edu}\\
+  \And
+  Kyle Cranmer\\
+  New York University\\
+  \texttt{kyle.cranmer@nyu.edu}
+}
+
 \begin{document}
+% \nipsfinalcopy is no longer used
 
-% ==============================================================================
-
-\title{\Large{Learning to Pivot with Adversarial Networks}}
-\vspace{1cm}
-\author{\small{\bf Gilles Louppe}\thanks{\texttt{g.louppe@nyu.edu}}}
-\affiliation{New York University}
-\author{\small{\bf Michael Kagan}\thanks{\texttt{makagan@slac.stanford.edu}}}
-\affiliation{SLAC National Accelerator Laboratory}
-\author{\small{\bf Kyle Cranmer}\thanks{\texttt{kyle.cranmer@nyu.edu}}}
-\affiliation{New York University}
+\maketitle
 
@@
 
-Many inference problems involve data generation processes that are not uniquely
-specified or are uncertain in some way. In a scientific context, the presence of
-several plausible data generation processes is often associated to the presence
-of systematic uncertainties. Robust inference is possible if it is based on a
-pivot -- a quantity whose distribution is invariant to the unknown value of the
-(categorical or continuous) nuisance parameters that parametrizes this family of
-generation processes. In this work,  we introduce a flexible training procedure
-based on adversarial networks for enforcing the pivotal property on a predictive
-model. We derive theoretical results showing that the proposed algorithm tends towards a minimax
-solution corresponding to a predictive model that is both optimal and
-independent of the nuisance parameters (if that models exists) or for which one
-can tune the trade-off between power and robustness. Finally, we demonstrate the
-effectiveness of this approach with a toy example and an example from particle
-physics.
+Several techniques for domain adaptation have been proposed to account for
+differences in the distribution of the data used for training and testing. The
+majority of this work focuses on a binary domain label. Similar problems occur
+in a scientific context where there may be a continuous family of plausible data
+generation processes associated to the presence of systematic uncertainties.
+Robust inference is possible if it is based on a pivot -- a quantity whose
+distribution does not depend on the unknown values of the nuisance parameters
+that parametrize this family of data generation processes. In this work,  we
+introduce and derive theoretical results for a training procedure based on
+adversarial networks for enforcing the pivotal property (or, equivalently,
+fairness with respect to continuous attributes) on a predictive model. The
+method includes a hyperparameter to control the trade-off between accuracy and
+robustness. We demonstrate the effectiveness of this approach with a toy example
+and examples from particle physics.
 
@@
 
-\maketitle
+%\begin{abstract}
+%    % Many inference problems involve data generation processes that are not uniquely
+%    % specified or are uncertain in some way.
+%    In a scientific context, the presence of
+%    several plausible data generation processes is often associated to systematic uncertainties. Robust inference is possible if it is based on a
+%    pivot -- a quantity whose distribution does not depend on the unknown value of the
+%    (typically continuous) nuisance parameters that parametrize this family of
+%    generation processes. In this work,  we introduce a training procedure
+%    based on adversarial networks for enforcing the pivotal property on a predictive model.
+%    We derive theoretical results showing that the procedure tends
+%    towards a minimax solution corresponding to a predictive model that is both
+%    optimal and independent of the nuisance parameters (if that model exists) or
+%    for which one can tune a trade-off between accuracy and robustness. We
+%    demonstrate the effectiveness of this approach with a toy example and an example
+%    from particle physics.
+%\end{abstract}
+
 % ==============================================================================
@@
 scientific process. One of the challenges of applying machine learning
-techniques to scientific problems is the need to incorporate systematic
-uncertainties. The presence of systematic uncertainties affects both the
-robustness of inference and the metrics used to evaluate the performance of a
-particular data analysis strategy.
+to scientific problems is the need to incorporate systematic
+uncertainties, which affect both the
+robustness of inference and the metrics used to evaluate a
+particular analysis strategy.
 
-In this work, we focus on supervised learning techniques where the presence of
-systematic uncertainty can be associated to the fact that the data generation
-process is not uniquely specified. In other words, the lack of systematic
+In this work, we focus on supervised learning techniques where
+systematic uncertainties can be associated to a data generation
+process that is not uniquely specified. In other words, the lack of systematic
 uncertainties corresponds to the (rare) case that the process that generates
@@
 real world data. By contrast, a common situation when systematic uncertainty is
-present is when the training data are not representative of the real data that a
-predictive model will be applied to in practice. Several techniques for domain
-adaptation have been developed to create predictive models that are more robust
-to this type of uncertainty. A more generic situation is that there are several
-plausible data generation processes, specified as a family of data generation
-processes parametrized by nuisance parameters, which can be categorical or
-continuous.
+present is when the training data are not representative of the real data.
+Several techniques for domain
+adaptation have been developed to create models that are more robust
+to this binary type of uncertainty.
+A more generic situation is that there are several
+plausible data generation processes, specified as a family
+parametrized by continuous nuisance parameters, as is typically found in scientific domains.
+In this broader context, statisticians have for long been working on robust inference techniques
+based on the concept of a pivot -- a quantity whose distribution is invariant
+with the nuisance parameters (see e.g., \citep{degroot1986probability}).
+% Pivotal quantities are e.g. fundamental to the construction
+% of test statistics and confidence intervals.
 
@@
 target labels, and $Z$ are the nuisance parameters, we consider the problem of
-learning a predictive model for $Y$ conditional on the observed values of $X$
-that is robust to uncertainty in the unknown value of $Z$. More specifically, we
-introduce a learning procedure for enforcing the pivotal property
-\citep{degroot1986probability}  on a
-predictive model by jointly training two neural networks in an adversarial
-fashion. We derive theoretical results proving that the proposed procedure
-converges towards a predictive model that is both optimal and independent of the
-nuisance parameters (if that models exists) or for which one can tune the
-trade-off between power and robustness. Finally, we demonstrate the
-effectiveness of this approach with a toy example and an example from particle physics.
-
-The remainder of the paper is structured as follows. In Sec.~\ref{sec:problem},
-we first formally define the problem of learning a pivotal, predictive model. In
-Sec. \ref{sec:method} and \ref{sec:theory}, we describe how adversarial
-networks can be trained to obtain predictive models that are independent of
-nuisance parameters and show that the proposed procedure converges
-towards a model that is both optimal and robust. The  effectiveness of the
-approach is then demonstrated empirically in Sec. \ref{sec:toy} and \ref{sec:hep}.
-In Sec.~\ref{sec:related}, we relate our work to
-close contributions from statistical inference, domain adaptation, and fairness
-in classification. Finally, we gather our concluding remarks in
-Sec.~\ref{sec:conclusions}.
-
-% ==============================================================================
+learning a predictive model $f(X)$ for $Y$ conditional on the observed values of $X$
+that is robust to uncertainty in the unknown value of $Z$. We
+introduce a flexible learning procedure based on adversarial networks~\citep{goodfellow2014generative}
+for enforcing that  $f(X)$ is a pivot with respect to $Z$.
+We derive theoretical results proving that the procedure
+converges towards a model that is both optimal and statistically independent of the
+nuisance parameters (if that model exists) or for which one can tune a
+trade-off between accuracy and robustness (e.g., as driven by a higher level objective).
+In particular, and to the best of our knowledge, our contribution is the first
+solution for imposing pivotal constraints on a predictive model,
+working regardless of the type of
+the nuisance parameter (discrete or continuous) or of its prior.
+Finally, we demonstrate the
+effectiveness of the approach with a toy example and examples from particle physics.
 
@@
 
-    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.85]
-    \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!30, line width=0.3mm, fill=white]
+   \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.65]
+   \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!20, line width=0.3mm, fill=white]
 
@@
        \foreach \dest in {1,...,5}
-            \path[black!20] (f-I-\source) edge (f-H1-\dest);
+           \path[black] (f-I-\source) edge (f-H1-\dest);
 
    \foreach \source in {1,...,5}
-        \path[black!30] (f-H2-\source) edge (f-O);
+       \path[black] (f-H2-\source) edge (f-O);
 
-    \node[black!30] at (1.5,-3) {...};
+   \node[black] at (1.5,-3) {...};
 
@@
        \foreach \dest in {1,...,5}
-            \path[black!20] (r-I-\source) edge (r-H1-\dest);
+           \path[black] (r-I-\source) edge (r-H1-\dest);
 
    \foreach \source in {1,...,5}
-        \path[black!30] (r-H2-\source) edge (r-O1);
+       \path[black] (r-H2-\source) edge (r-O1);
    \foreach \source in {1,...,5}
-        \path[black!30] (r-H2-\source) edge (r-O2);
+       \path[black] (r-H2-\source) edge (r-O2);
    \foreach \source in {1,...,5}
-        \path[black!30] (r-H2-\source) edge (r-O3);
+       \path[black] (r-H2-\source) edge (r-O3);
 
-    \node[black!30] at (9,-3) {...};
+   \node[black] at (9,-3) {...};
 
-    \draw (14,-1.5) rectangle (17,-4.5);
-    \path[->, shorten >= 0pt] (15.5,-0.5) edge (15.5,-1.5);
-    \node[above] at (15.5,-0.5) {$Z$};
-    \path[->, shorten >= 0pt] (15.5,-4.5) edge (15.5,-5.5);
-    \node[below] at (15.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
-    \node at (15.5,-3) {${\cal P}(\gamma_1, \gamma_2, \dots)$};
+   \draw (15,-1.5) rectangle (18,-4.5);
+   \path[->, shorten >= 0pt] (16.5,-0.5) edge (16.5,-1.5);
+   \node[above] at (16.5,-0.5) {$Z$};
+   \path[->, shorten >= 0pt] (16.5,-4.5) edge (16.5,-5.5);
+   \node[below] at (16.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
+   \node at (16.5,-3) {${\cal P}(\gamma_1, \gamma_2, \dots)$};
 
-    \draw[dashed,-] (15.5,-5) -| (12.75,-6.5);
+   \draw[dashed,-] (16.5,-5) -| (12.75,-6.5);
    \node[below] at (12.75,-6.5) {${\cal L}_r(\theta_f, \theta_r)$};
@@
 
+% \includegraphics[width=\linewidth]{pivot-figure}
+
     \caption{Architecture for the adversarial training of a binary classifier
@@
 \end{figure*}
+
+% % this adds nothing
+% The remainder of the paper is structured as follows. In Sec.~\ref{sec:problem},
+% we first formally define the problem of learning a pivotal, predictive model. In
+% Sec. \ref{sec:method} and \ref{sec:theory}, we describe how adversarial
+% networks can be trained to obtain predictive models that are independent of
+% nuisance parameters and show that the proposed procedure converges
+% towards a model that is both optimal and robust. The  effectiveness of the
+% approach is then demonstrated empirically in Sec. \ref{sec:toy} and \ref{sec:hep}.
+% In Sec.~\ref{sec:related}, we relate our work to
+% close contributions from statistical inference, domain adaptation, and fairness
+% in classification. Finally, we gather our concluding remarks in
+% Sec.~\ref{sec:conclusions}.
+
+% ==============================================================================
 
@@
 Let us assume that prior to incorporating the effect of uncertainty
-in $Z$, our goal is to learn a regression function $f : \mathcal{X} \mapsto
-{\cal S}$ of parameters $\theta_f$ (e.g., a neural network-based probabilistic
+in $Z$, our goal is to learn a regression function $f : \mathcal{X} \to
+{\cal S}$ with parameters $\theta_f$ (e.g., a neural network-based probabilistic
 classifier) that minimizes a loss ${\cal L}_f(\theta_f)$ (e.g., the
-cross-entropy).
+cross-entropy). In classification, values $\smash{s \in {\cal S} = \mathbb{R}^{|{\cal Y}|}}$
+correspond to the classifier scores used for mapping hard predictions $y \in {\cal Y}$, while ${\cal S} = {\cal Y}$ for regression.
 
@@
 In words, we are looking for a predictive function $f$
-which is a pivotal quantity \citep{degroot1986probability} with respect to the
+which is a pivotal quantity with respect to the
 nuisance parameters. This implies that  $f(X; \theta_f)$ and $Z$ are independent random variables.
@@
 
+    {\footnotesize
     \begin{flushleft}
-        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;\\
-        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$;\\
-        {\it Hyper-parameters:} Number $T$ of training iterations,
-                                Number $K$ of gradient steps to update $r$,
-                                Size $M$ of the minibatches.
+        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;
+        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$.
+        % {\it Hyper-parameters:} Number $T$ of iterations,
+        %                         Number $K$ of gradient steps to update $r$,
+        %                         Size $M$ of the minibatches.
     \end{flushleft}
@@
     \begin{algorithmic}[1]
-        \For{$t=1$ to $T$}
-            \For{$k=1$ to $K$} \Comment{Update $r$}
-                \State{Sample minibatch $\{x_m, z_m, s_m = f(x_m;\theta_f) \}_{m=1}^M$} of size $M$;
-                \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
+        \FOR{$t=1$ to $T$}
+            \FOR{$k=1$ to $K$} %\COMMENT{Update $r$}
+                \STATE{Sample minibatch $\{x_m, z_m, s_m = f(x_m;\theta_f) \}_{m=1}^M$ of size $M$;}
+                \STATE{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
                 $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|s_m)  ;$$}
-            \EndFor
-            \State{Sample minibatch $\{x_m, y_m, z_m, s_m = f(x_m;\theta_f)  \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
-            \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
+            \ENDFOR
+            \STATE{Sample minibatch $\{x_m, y_m, z_m, s_m = f(x_m;\theta_f)  \}_{m=1}^M$ of size $M$;} %\Comment{Update $f$}
+            \STATE{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
             $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|s_m)  \right],$$
             \indent where $p_{\theta_f}(y_m|x_m)$ denotes $\mathbf{1}(y_m=0)(1-s_m) + \mathbf{1}(y_m=1)s_m$;}
-        \EndFor
+        \ENDFOR
     \end{algorithmic}
+    }
     \end{algorithm}
@@
 predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. As
-illustrated in Fig.~\ref{fig:architecture}, we pit $f$ against an adversary
-model $r := p_{\theta_r}(z | f(X;\theta_f)=s)$ of parameters $\theta_r$ and
+illustrated in Fig.~\ref{fig:architecture}, we pit $f$ against an adversarial
+model $r := p_{\theta_r}(z | f(X;\theta_f)=s)$ with parameters $\theta_r$ and
 associated loss ${\cal L}_r(\theta_f, \theta_r)$. This model takes  as input
-realizations $f(X; \theta_f)$ and produces as output a function
+realizations $s$ of $f(X; \theta_f)$ and produces as output a function
 modeling the posterior probability density $p_{\theta_r}(z | f(X;\theta_f)=s)$.
@@
 
-If $Z$ takes discrete values, then $p_{\theta_r}$ can be represented e.g. as a
-probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal Z|}}$ whose
+If $Z$ takes discrete values, then $p_{\theta_r}$ can be represented as a
+probabilistic classifier $\mathbb{R} \to \mathbb{R}^{|{\cal Z|}}$ whose
 $j^\textrm{th}$ output (for $j=1, \dots, |{\cal Z}|$) is the estimated probability mass
@@
 
-\subsection{Toy example}
+In this section, we empirically demonstrate the effectiveness of the approach
+with a toy example and examples from particle physics. Notably,
+there are no other other approaches to compare to in the case of continuous
+nuisance parameters, as further explained in Sec.~\ref{sec:related}. In the case
+of binary parameters, we do not expect results to be much different from
+previous works. The source code to reproduce the experiments is available online~\footnote{\url{https://github.com/glouppe/paper-learning-to-pivot}}.
+
+\subsection{A toy example with a continous nuisance parameter}
 \label{sec:toy}
@@
     \end{center}
-    \vspace{-0.5cm}
+
     \caption{Toy example.
     (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
-       when $f$ is built without adversarial training. The resulting densities
-       are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
-    (Middle left) The corresponding decision surface, highlighting
+        without adversarial training. The resulting densities
+       are dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
+    (Middle left) The associated decision surface, highlighting
        the fact that samples are easier to classify for values of $Z$ above $\sigma$,
-       hence partially explaining the dependency.
+       hence explaining the dependency.
     (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
-       built with adversarial training, as outlined in Sec.~\ref{sec:method}.
+       built with adversarial training.
        The resulting densities are now almost identical to each other, indicating only a
        small dependency on $Z$.
-    (Right) The corresponding decision surface, illustrating how adversarial
+    (Right) The associated decision surface, illustrating how adversarial
        training bends the decision function vertically to erase the dependency on $Z$.
@@
 
-\subsection{High energy physics example}
+\begin{figure}
+\centering
+\begin{minipage}{.46\linewidth}
+    \centering
+        \includegraphics[width=\textwidth]{figures/training.pdf}\vspace{-1em}
+        \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
+                 and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
+                 Initialized with a pre-trained classifier $f$, adversarial training was performed for $200$ iterations, mini-batches of size $M=128$, $K=500$ and $\lambda=50$.}
+        \label{fig:toy-training}
+\end{minipage}
+\hspace{.05\linewidth}
+\begin{minipage}{.46\linewidth}
+    \centering
+        \includegraphics[width=\textwidth]{figures/ams.pdf}\vspace{-1em}
+        \caption{Physics example. Approximate median significance as a function of the decision threshold
+                 on the output of $f$. At $\lambda=10$, trading
+                 accuracy for independence to pileup
+                 results in a net benefit in terms of statistical significance.}
+        \label{fig:physics-ams}
+\end{minipage}
+\end{figure}
+
+\subsection{High energy physics examples}
 \label{sec:hep}
 
-Experiments at high energy colliders like the LHC \citep{LHCMachine} are
-searching for evidence of new particles beyond the ones described by the well
-established Standard Model (SM) of particle physics. A wide array of extended
+\paragraph{Binary Case} Experiments at high energy colliders like the LHC \citep{LHCMachine} are
+searching for evidence of new particles beyond those described by the
+Standard Model (SM) of particle physics. A wide array of
 theories predict the existence of new massive particles that would decay to
@@
 can decay to two quarks, each of which produce collimated sprays of particles
-known as \textit{jets} (collectively denoted $W\to jj$). If the exotic particle
+known as jets.
+%(collectively denoted $W\to jj$).
+If the exotic particle
 is heavy, then the $W$ boson will be moving very fast, and  relativistic effects
 will cause the two jets from its decay to merge into a single `$W$-jet'. These
-$W$-jets have a rich internal substructure (see e.g.
-\citep{Altheimer:2012mn,Altheimer:2013yza} and references therein). However,
+$W$-jets have a rich internal substructure.
+% (see e.g.\citep{Altheimer:2012mn,Altheimer:2013yza}).
+However,
 jets are also produced ubiquitously at high energy colliders through more
@@
 wish to impose the pivotal property on $Y=0$ events. To this end,
-we use as a higher level objective the Approximate Median Significance (AMS), which was used
-in the Higgs boson machine learning challenge and is a natural generalization
+we use as a higher level objective the Approximate Median Significance (AMS), which is a natural generalization
 of the power of a hypothesis test when systematic uncertainties are taken into account
-(see Eqn.~20 of \citep{adam2014higgs}).
+(see Eqn.~20 of \cite{adam2014higgs}).
 
@@
 
-\begin{figure}
-    \includegraphics[width=0.48\textwidth]{figures/ams.pdf}
-    \caption{Physics example. Approximate median significance as a function of the decision threshold
-             on the output of $f$.  As shown at $\lambda=10$, trading
-             classification accuracy for independence with respect to pileup
-             results in a net benefit in terms of statistical significance.}
-    \label{fig:physics-ams}
-\end{figure}
-
 As  Fig.~\ref{fig:physics-ams} illustrates, without adversarial training (at
@@
 is made stronger (for $\lambda > 0$) the AMS peak moves higher, with a maximum
-value around  $7.8$ for $\lambda=10$. In other words, trading classification
-accuracy for robustness to pileup results in a net
-benefit in terms of the power of the resulting hypothesis test. Setting $\lambda$ too high however
+value around  $7.8$ for $\lambda=10$. Trading classification
+accuracy for robustness to pileup thereby results in a net
+benefit in terms of the power of the hypothesis test. Setting $\lambda$ too high however
 (e.g. $\lambda=500$) results in a decrease of the maximum AMS, by
 focusing the capacity of $f$ too strongly on independence with $Z$, at the expense of
-classification accuracy. As demonstrated in this example, optimizing $\lambda$
-gives us a principled and effective approach to control the trade-off between
+accuracy. In effect, optimizing $\lambda$
+yields a principled and effective approach to control the trade-off between
 accuracy and robustness that ultimately maximizes the
 power of the enveloping hypothesis test.
+
+\paragraph{Continous Case}
+%Our method also applies to cases where one desires the predictive model to be independent of
+%some other observed random variables, as in domain adaptation or fairness in
+%classification. Importantly, our approach is able to enforce fairness over
+%continuous attributes.
+Recently, an independent group has used our approach to learn jet classifiers
+that are independent of the jet mass~\citep{Shimmin:2017mfk}, which is a continuous attribute.
+The results of their studies show that the adversarial training strategy works very well for
+real-world problems with continuous attributes, thus enhancing the sensitivity of searches for
+new physics at the LHC.
 
@@
 
-In machine learning, learning a pivotal quantity can be related to the problem
+Learning to pivot can be related to the problem
 of domain adaptation
@@
 closest to our work, in which domain invariance and fairness are enforced
-through an adversarial minimax setup composed of a classifier and an adversary
-discriminator. Following this line of work and without making assumptions on the
-prior $p(Z)$, our method can be regarded as a generalization that also supports
+through an adversarial minimax setup composed of a classifier and an adversarial
+discriminator. Following this line of work, our method can be regarded as a unified generalization that also supports
 a continuously parametrized family of domains or as enforcing fairness over
 continuous attributes.
+
+%While our contribution may first appear as a minor extension, most
+Most related
+work is based on the strong and limiting assumption that $Z$ is a binary
+random variable (e.g., $Z=0$ for the source domain, and
+$Z=1$ for the target domain).
+In particular, \citep{pan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,zemel2013learning,ganin2014unsupervised,ajakan2014domain,EdwardsS15,louizos2015variational}
+are all based on the minimization of some form of divergence between the two distributions of
+$f(X)|Z=0$ and $f(X)|Z=1$. For this reason,
+%and in contrast with our procedure,
+these works cannot directly be generalized to non-binary or
+continuous nuisance parameters, both from a practical and theoretical point of view.
+Notably, \cite{kamishima2012fairness} enforces fairness through
+a prejudice regularization term based on empirical estimates of $p(f(X)|Z)$.
+While this approach is in principle sufficient for handling non-binary
+nuisance parameters $Z$, it requires accurate
+empirical estimates of $p(f(X)|Z=z)$ for all values $z$, which quickly becomes
+impractical as the cardinality of $Z$ increases. By contrast, our approach models the conditional
+dependence through an adversarial network, which allows for generalization without
+necessarily requiring an exponentially growing number of training examples.
 
@@
 the most statistically powerful approach for incorporating the effect of
-systematics on the underlying classification task.  As argued in
-\citep{Neal:2007zz}, it is not obvious how such classifiers can be used on real
-data since the correct value $z$ of the nuisance often remains unknown. This is
-not an issue in the context of parameter
-inference~\citep{cranmer2015approximating}, where nuisance parameters are
-eliminated via optimization or marginalization. In practice, parametrized
+systematics on the underlying classification task.
+%As argued in
+%\citep{Neal:2007zz}, it is not obvious how such classifiers can be used on real
+%data since the correct value $z$ of the nuisance often remains unknown. This is
+%not an issue in the context of parameter
+%inference~\citep{cranmer2015approximating}, where nuisance parameters are
+%eliminated via optimization or marginalization.
+In practice, parametrized
 classifiers  are also computationally expensive to build and evaluate. In
@@
 $p(f(x,z)|y,z)$ as a continuous function of $z$, remains an open challenge. By
-contrast, constraining $f$ to be pivotal yields a classifier which may not be
-optimal with respect to $Y|X,Z$, as discussed in Sec.~\ref{sec:theory}, but that
-can otherwise be directly used in a wider range of applications, since the
-dependence on the nuisance parameter $Z$ has already been eliminated. Moreover,
-if $f$ is sufficiently close to being pivotal, then calibration only needs to be
-carried out once.
+contrast, constraining $f$ to be pivotal yields a classifier
+%which may not be
+%optimal with respect to $Y|X,Z$, as discussed in Sec.~\ref{sec:theory}, but
+that can
+%otherwise
+be directly used in a wider range of applications, since the
+dependence on the nuisance parameter $Z$ has already been eliminated.
+%Moreover, if $f$ is sufficiently close to being pivotal, then calibration only needs to be
+%carried out once.
 
@@
 
-Despite the theoretical characterization of the minimax solution of Eqn.~\ref{eqn:min_thetaf}
-that motivates this work,
-Algorithm~\ref{alg:adversarial-training} remains a heuristic in the sense
-that we do not provide formal guarantees of convergence of the procedure towards
-that solution. Empirical evidence gathered here and in similar minimax-based
- works (see e.g., \citep{goodfellow2014generative,EdwardsS15}) suggests
-that the alternating optimization algorithm works sufficiently well in practice,
-but further research studying its stability and convergence would certainly
-be beneficial for a large body of works.
+% Despite the theoretical characterization of the minimax solution of Eqn.~\ref{eqn:min_thetaf}
+% that motivates this work,
+% Algorithm~\ref{alg:adversarial-training} remains a heuristic in the sense
+% that we do not provide formal guarantees of convergence of the procedure towards
+% that solution. Empirical evidence gathered here and in similar minimax-based
+%  works suggests
+% that the alternating optimization algorithm works sufficiently well in practice,
+% but further research studying its stability and convergence would certainly
+% be beneficial for a large body of works.
 
-In terms of applications, the proposed solution can be used in any situation
+In terms of applications, our solution can be used in any situation
 where the training data may not be representative of the real data the
@@
 of data generation processes, and it would be worth revisiting those scientific problems
-that utilize machine learning in light of this technique. Moreover, the approach
+that utilize machine learning in light of this technique. The approach
 also extends to cases where independence of the predictive model with respect to
-observed random variables is desired, as in domain adaptation or fairness in
-classification. Again, trying our algorithm on these use cases is certainly
-worth investigating.
+observed random variables is desired, as in fairness for
+classification.
+% not sure this is a great sentence to end with, sounds like a sales pitch
+% Again, trying our algorithm on these use cases is certainly
+%worth investigating.
 
@@
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 
-    We would like to thank the authors of Ref.\citep{baldi2016jet} for sharing the data used in their
+We would like to thank the authors of \citep{baldi2016jet} for sharing the data used in their
 studies.
@@
 
-% ==============================================================================
-
-\bibliographystyle{acm}
-\bibliography{bibliography.bib}
+\bibliography{bibliography}
+\bibliographystyle{apalike}
 
